### 2012-09-13 chi2012.tex
Initial version.

\documentclass{sigchi}

% Use this command to override the default ACM copyright statement (e.g. for preprints). 
% Consult the conference website for the camera-ready copyright statement.
\toappear{
Permission to make digital or hard copies of all or part of this
work for personal or classroom use is granted without fee provided that 
copies are not made or distributed for profit or commercial advantage and
that copies bear this notice and the full citation on the first page. To
copy otherwise, or republish, to post on servers or to redistribute to 
lists, requires prior specific permission and/or a fee.\\
{\confname{CHI'12}}, May 5--10, 2012, Austin, Texas, USA.\\
Copyright 2012 ACM 978-1-4503-1015-4/12/05...\$10.00.
}

% Arabic page numbers for submission. 
% Remove this line to eliminate page numbers for the camera ready copy
%\pagenumbering{arabic}

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, 
% to give it a fighting chance of not being over-written, 
% since its job is to redefine many LaTeX commands.
\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={SIGCHI Conference Proceedings Format},
pdfauthor={LaTeX},
pdfkeywords={SIGCHI, proceedings, archival format},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

% create a shortcut to typeset table headings
\newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management}

\numberofauthors{3}
\author{
  \alignauthor 1st Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Address}\\
    \email{e-mail address}\\
    \affaddr{Optional phone number}
  \alignauthor 2nd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Address}\\
    \email{e-mail address}\\
    \affaddr{Optional phone number}    
  \alignauthor 3rd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Address}\\
    \email{e-mail address}\\
    \affaddr{Optional phone number}
}

% Teaser figure can go here
%\teaser{
%  \centering
%  \includegraphics{Figure1}
%  \caption{Teaser Image}
%  \label{fig:teaser}
%}

\maketitle

\begin{abstract}
\end{abstract}

\keywords{
	Guides; instructions; author's kit; conference publications;
	keywords should be separated by a semi-colon.
	\textcolor{red}{Mandatory section to be included in your final version.}
}

\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}

See: \url{http://www.acm.org/about/class/1998/}
for more information and the full list of ACM classifiers
and descriptors. 
\textcolor{red}{Mandatory section to be included in your
final version. On the submission page only the classifiers'
letter-number combination will need to be entered.}

\terms{
	Human Factors; Design; Measurement. 
	If you choose more than one ACM General Term, 
	separate the terms with a semi-colon.
}

See list of the limited ACM 16 terms in the
instructions and additional information:
\url{http://www.sheridanprinting.com/sigchi/generalterms.htm}.
\textcolor{red}{Optional section to be included in your final version.}

\section{Introduction}

In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web, through APIs, datasets, and data feeds.  This information includes access to data previously hidden behind services and applications, such as retailer product catalogues, information not previously directly released to the public, such as open government data or records of personal financial transactions, and new kinds of data generated by emerging kinds of data sources, such as wearable and smartphone-based biosensors.

While a primary goal of opening up such direct access to information is the hope that end-user citizens will make more informed decisions pertaining to their health, wealth, or well-being \cite{}, use of this data beyond specialised app developers, journalists and other “data specialists” remains minimal.  Our hypothesis is that the main impediment towards end-user access to this data is a mismatch between the capabilities of tools currently used to satisfy daily information needs, and those required to effectively browse, use and consume heterogeneous structured data from diverse sources.  For example, while much of this data represents a many and varied set of things, activities, and descriptions at various granularities, most commonly used structured personal information management (PIM) tools either can only manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{}.

If PIM tools were extended to enable end-users to make effective use of the emerging ecosystem of personal data, how would this impact personal information practice? This paper presents our initial results towards answering this question through a multi-method investigation of the personal data integration problem.  We start with a small interview study examining information tasks people perform that rely on multiple sources of information, and how such needs are currently satisfied.  Our results suggest that people prefer to rely upon multiple, diverse sources to singular integrated ones for a number of reasons, including coverage, reliability, and \_\_\_\_\_\_.  Second, we conduct a analysis of various popular personal data sources available today to identify barriers to effective unification of data from each.  We draw upon definitions and work from the database integration systems literature to characterise types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.

Based upon the data integration needs revealed by people in the interviews, and the kinds of heterogeneity observed in the data feeds, we embarked on a user-centric design exercise towards an interface to facilitate simple, "light-touch" integration tasks of diverse, heterogeneous data sources.  We describe our initial foray into this exercise, an interface we call DataPalette, which focuses on the most common types of heterogeneity to enable serendipitous "data mixing" - a form of simple integration sufficient to let people easily and effectively combine and compare information sources without the need to write code.  Our usability evaluation of Data Palette reveals that most users were comfortable with and that this interface is effectively expedites the most sophisticated of the typical kinds of "data integration" tasks people need to perform while planning personal,

\section{Related Work}

As the theoretical basis of data integration extends nearly a half-century in the database systems literature . As described by Halvey et al’s comprehensive textbook on data integration problems and techniques \cite{principlesofdataintegration} much of this focus of this community has focused on \emph{engineered} or \emph{automatic} approaches to data integration at a large scale,

In scientific laboratory settings,

We only briefly touch on the extensive literature covering the “deep end” of data integration - both automatic and “hand-coded” techniques to link multiple, typically large, data sources in consistent, correct and efficient ways, little work has pertained on the opposite end -- the “light”, “small” and “sketchy” end of data integration, such as arises when multiple heterogeneous sources of data about similar kinds of things are brought together spontaneously

Interfaces that let end-users find, consolidate, and work with heterogeneous data 
PIM 
	Amy Voida’s databases paper
User-subjective approach to information management

Information scraps
End-user Data Gathering and integration platforms
	Potluck
	Cards (M. Dontcheva)

End-user data gathering
	Piggy Bank
“data notebooks”	Hunter-gatherer
	Yahoo Pipes
PROMPT

A key assumption of this work is that ontologies are not going to converge.  (Studies about the diversity of web ontologies backing this claim up?)

Why end-user data integration is desirable
User-subjective approach to personal information management

Why is the data integration problem so hard?  
The problem of reconciling schema heterogeneity has been a subject of research for decades, but solutions are few.  In the words of leading data integration researcher Alon Halevy,
[The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs. Hence, even if they model
overlapping domains, they will model them in different ways. Differing structures are a byproduct of human nature ?people think differently from one another even when faced with  the same modeling goal. \cite{alonhalevy}

Exhibit - designed for personal structured curation
Yahoo Pipes! - designed for not-personal (public) curation
MySupermarket - http://mysupermarket.co.uk/
Demonstrates the power of joining across data sources: people can gather a grocery list, and compare how much the same shopping list would cost at each of the major UK supermarket chains (ASDA, Tesco, Waitrose, ocado) --

But what about the mom + pop groceries?
What if they’re missing data: allergies, suitable for -- available from somewhere else

Schema.org  - is it helping or not?

\section{Pre-study 1 - Understanding Heterogeneity for Everyday Tasks}
\subsection{Method}
Our pre-study investigated the types of tasks people perform online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.  We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions (see Appendix ?).  This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant’s experiences, while collecting the same general areas of information.  Each interview was recorded and notes taken, the interviewer was trained so that they were knowledgeable about the importance of the study and to minimise bias.
\subsection{Results and Discussion}
The participant demographic consisted of 8 people, 7 of which were male and 1 female.  They were split evenly into two age groups of 18-25 and 26-32.  All but one participant used social network sites very regularly, and they either logged in multiple times a day, left a website page open or used dedicated application for their social networking site’s updates.  Those that used social networking websites primarily use Facebook.  All of the participants used Twitter to listen to broadcasted tweets, and 6 out of 8 people regularly tweeted.

Question 1: Tasks
All participants had experience in using multiple websites to complete tasks.  They listed example tasks such as shopping (including food, consumer electronics, hotels and flights), searching for jobs, choosing recipes, study and work.  The biggest focus for shopping focused tasks was balancing price, quality and speed of delivery.  In general, people did not not trust a single websites for reviews because they said that they could be biased and did not always review the features they were interested in.  One participant said that video reviews of items were really important because they showed the aesthetics and scale of the item.  In general for initial research into organising a task it was typical to Google keywords to find related websites to their tasks.  They identified that they used different website for different jobs, such as manufacturers website for technical details, review aggregators for a range of opinions, and Google maps for location based decisions.  Social networks were also identified to have different functions, such as different groups of people belong to each and share different types of opinions.  One participant said that they shared their outlook and gmail calendars, but only for work events and did not record social activities on them.

Question 2: Large Social Event
All of the participants would organise a large social event by talking to their friends, about their ideas, preferences and recommendations.  The tools they would use to acquire this information included face-to-face, phone calls, email, Skype, Facebook events.  The method chosen depended on the time scale required to organise an event, in general if there was a short time scale people would speak in person or on the phone.  Otherwise, people would set up a FaceBook event page to discuss ideas with their friends.  If their friends were not on Facebook they would email them.  All but one participant said they would not use Doddle to organise day and time of an event, because they felt that people did not fill in the form and it was more suit to organising work event not social.

Most people felt that being assertive and posting their decisions about the date and time on Facebook meant that organising events were more successful than trying to gather a consensus.  They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.   They selected venues, locations, restaurants and activities based on their own knowledge, friend’s recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.  People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.  They considered how to get there and looked up train or bus times on the Web.  One participant organised carpools informally and at last minute via Facebook or text messages.

When organising an event the weather was not that important because it they did not trust its accuracy for events plan further in the future than one week.  Also, their friend’s preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.  Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
\subsection{Prestudy 1 Summary of Findings}
People said that they would like a website that could support the evolution of an event, that allowed them to post drafts of an event.  They felt that organising an event should be a process that changes over time and did not need rushing.  Three of the participants said that they would like a recommendation system for places to eat and activities. Another strong requirement was that it had to be ubiquitous so that everyone could use it because they wanted a single place to communicate with their friends.  While Facebook was the most popular social networking site for organising events, half of the participants felt that it was not the best solution and would prefer a collaborative environment, which could be wiki based.

\section{Prestudy 2 : Technical Challenges of Integrating the Data}

\subsection{Method}

\subsection{Results}

\subsection{Analysis}

\subsection{Prestudy 2 Summary of Findings}

\section{WebBox Interface - Design of the Data Palette}

Design Goals, 
	What are the problems observed needed to address
	What kind of interactions can be designed to solve them?

Interaction features 
	multiple paths selection
	sameas interactions
	visualisation of partial results

Design process

\section{Methodology}

Lab based usability study,

Design of Study
How we designed the tasks, subjective task that requires multi dimensions 
How we got data
How we recruited participants - International student list

Our aim was XXX, To accomplish thies XX participants were recruited to the lab for a 1 hour session

	Latin square dance!
	4 factors
		Interface or not
		Task (restaurant or uni)
		Dataset (glasgow / cambridge; history or sports)
		Participant

Hypotheses
Interaction-method specific hypotheses
	h1.1 - Do people understanding the problem of co-reference, and the need to reconcile coreference problems?
	h1.2. Does the ability to do drag + drop combination for co-reference reconciliation effectively solve this problem?
	h2.1- Do people understand the problem of structural differences in data?
h2.2 - Does multipath selection let people effectively deal with this -- and work with collections of heterogeneous items?

Are these interaction methods sufficient to perform common tasks involving heterogeneous data sources?  Does supporting these simple techniques facilitate task completion?

\section{Results}
How many native english speakers, only 8 were all enrolled at university all past proficiency test.
\section{Discussion}
What we found

What we didn’t find (evidence for)

What we might look at next -- follow-on work
		We focused on terminological heterogeneity -- didn’t address structural heterogeneity, semantic heterogeneity
		We focused on finite (small) sets Streaming data (how do select future data --) 
		Scale - small data, what are the facilites required for large data
		Collaborative - how might interaction support multiple users collaborating?

Limitations found from study... what is the take away from the limitations
\section{Conclusion} %Key findings of the whole paper and its implications

\section{Acknowledgments}

% Balancing columns in a ref list is a bit of a pain because you
% either use a hack like flushend or balance, or manually insert
% a column break.  http://www.tex.ac.uk/cgi-bin/texfaq2html?label=balance
% multicols doesn't work because we're already in two-column mode,
% and flushend isn't awesome, so I choose balance.  See this
% for more info: http://cs.brown.edu/system/software/latex/doc/balance.pdf
%
% Note that in a perfect world balance wants to be in the first
% column of the last page.
%
% If balance doesn't work for you, you can remove that and
% hard-code a column break into the bbl file right before you
% submit:
%
% http://stackoverflow.com/questions/2149854/how-to-manually-equalize-columns-
% in-an-ieee-paper-if-using-bibtex
%
% Or, just remove \balance and give up on balancing the last page.
%
\balance

% If you want to use smaller typesetting for the reference list,
% uncomment the following line:
% \small
\bibliographystyle{acm-sigchi}
\bibliography{sample}
\end{document}

### 2012-09-13 sample.tex
Initial version.

\documentclass{sigchi}

% Use this command to override the default ACM copyright statement (e.g. for preprints). 
% Consult the conference website for the camera-ready copyright statement.
\toappear{
Permission to make digital or hard copies of all or part of this
work for personal or classroom use is granted without fee provided that 
copies are not made or distributed for profit or commercial advantage and
that copies bear this notice and the full citation on the first page. To
copy otherwise, or republish, to post on servers or to redistribute to 
lists, requires prior specific permission and/or a fee.\\
{\confname{CHI'12}}, May 5--10, 2012, Austin, Texas, USA.\\
Copyright 2012 ACM 978-1-4503-1015-4/12/05...\$10.00.
}

% Arabic page numbers for submission. 
% Remove this line to eliminate page numbers for the camera ready copy
%\pagenumbering{arabic}

% Load basic packages
\usepackage{balance}  % to better equalize the last page
\usepackage{graphics} % for EPS, load graphicx instead
\usepackage{times}    % comment if you want LaTeX's default font
\usepackage{url}      % llt: nicely formatted URLs

% llt: Define a global style for URLs, rather that the default one
\makeatletter
\def\url@leostyle{%
  \@ifundefined{selectfont}{\def\UrlFont{\sf}}{\def\UrlFont{\small\bf\ttfamily}}}
\makeatother
\urlstyle{leo}

% To make various LaTeX processors do the right thing with page size.
\def\pprw{8.5in}
\def\pprh{11in}
\special{papersize=\pprw,\pprh}
\setlength{\paperwidth}{\pprw}
\setlength{\paperheight}{\pprh}
\setlength{\pdfpagewidth}{\pprw}
\setlength{\pdfpageheight}{\pprh}

% Make sure hyperref comes last of your loaded packages, 
% to give it a fighting chance of not being over-written, 
% since its job is to redefine many LaTeX commands.
\usepackage[pdftex]{hyperref}
\hypersetup{
pdftitle={SIGCHI Conference Proceedings Format},
pdfauthor={LaTeX},
pdfkeywords={SIGCHI, proceedings, archival format},
bookmarksnumbered,
pdfstartview={FitH},
colorlinks,
citecolor=black,
filecolor=black,
linkcolor=black,
urlcolor=black,
breaklinks=true,
}

% create a shortcut to typeset table headings
\newcommand\tabhead[1]{\small\textbf{#1}}

% End of preamble. Here it comes the document.
\begin{document}

\title{Carp\'{e} Data: Supporting Serendipitous Data Integration in Personal Information Management}

\numberofauthors{3}
\author{
  \alignauthor 1st Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Address}\\
    \email{e-mail address}\\
    \affaddr{Optional phone number}
  \alignauthor 2nd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Address}\\
    \email{e-mail address}\\
    \affaddr{Optional phone number}    
  \alignauthor 3rd Author Name\\
    \affaddr{Affiliation}\\
    \affaddr{Address}\\
    \email{e-mail address}\\
    \affaddr{Optional phone number}
}

% Teaser figure can go here
%\teaser{
%  \centering
%  \includegraphics{Figure1}
%  \caption{Teaser Image}
%  \label{fig:teaser}
%}

\maketitle

\begin{abstract}
In this paper we describe the formatting requirements for
SIGCHI Conference Proceedings, and this sample file
offers recommendations on writing for the worldwide
SIGCHI readership. Please review this document even if
you have submitted to SIGCHI conferences before, some
format details have changed relative to previous years.
\end{abstract}

\keywords{
	Guides; instructions; author's kit; conference publications;
	keywords should be separated by a semi-colon.
	\textcolor{red}{Mandatory section to be included in your final version.}
}

\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}

See: \url{http://www.acm.org/about/class/1998/}
for more information and the full list of ACM classifiers
and descriptors. 
\textcolor{red}{Mandatory section to be included in your
final version. On the submission page only the classifiers'
letter-number combination will need to be entered.}

\terms{
	Human Factors; Design; Measurement. 
	If you choose more than one ACM General Term, 
	separate the terms with a semi-colon.
}

See list of the limited ACM 16 terms in the
instructions and additional information:
\url{http://www.sheridanprinting.com/sigchi/generalterms.htm}.
\textcolor{red}{Optional section to be included in your final version.}

\section{Introduction}

This format is to be used for submissions that are
published in the conference proceedings.  We wish to give
this volume a consistent, high-quality appearance. We
therefore ask that authors follow some simple
guidelines. In essence, you should format your paper
exactly like this document. The easiest way to do this is
simply to download a template from the conference web
site, and replace the content with your own material.

\section{Page Size and Columns}

On each page your material (not including the page number) should fit
within a rectangle of 18 x 23.5 cm (7 x 9.25 in.), centered on a US
letter page, beginning 1.9 cm (.75 in.) from the top of the page, with
a .85 cm (.33 in.) space between two 8.4 cm (3.3 in.) columns.  Right
margins should be justified, not ragged. Beware, especially when using
this template on a Macintosh, Word can change these dimensions in
unexpected ways. Please be sure that your PDF is US letter and not
A4. If your PDF or paper are formatted for A4, the submission will be
returned to you to fix.

\section{Typeset Text}

Prepare your submissions on a word processor or typesetter.  Please
note that page layout may change slightly depending upon the printer
you have specified.  \LaTeX\ sometimes will create overfull lines
that extend into columns.  To attempt to combat this, the .cls
file has a command, {\textbackslash}sloppy, that essentially asks
\LaTeX\ to prefer underfull lines with extra whitespace.  For more
details on this, and info on how to control it more finely, check out
{\url{http://www.economics.utoronto.ca/osborne/latex/PMAKEUP.HTM}}.

\subsection{Title and Authors}

Your paper's title, authors and affiliations should run across the
full width of the page in a single column 17.8 cm (7 in.) wide.  The
title should be in Helvetica 18-point bold; use Arial if Helvetica is
not available.  Authors' names should be in Times Roman 12-point bold,
and affiliations in Times Roman 12-point.  For more than three authors,
you may have to place some address information in a footnote, or in a named
section at the end of your paper. Please use full international addresses and
telephone dialing prefixes.  Leave one 10-pt line of white space below the last
line of affiliations.

\subsection{Abstract and Keywords}

Every submission should begin with an abstract of about 150 words,
followed by a set of keywords. The abstract and keywords should be
placed in the left column of the first page under the left half of the
title. The abstract should be a concise statement of the problem,
approach and conclusions of the work described.  It should clearly
state the paper's contribution to the field of HCI.

The first set of keywords will be used to index the paper in the
proceedings. The second set are used to catalogue the paper in the ACM
Digital Library. The latter are entries from the ACM Classification
System~\cite{acm_categories}.  In general, it should only be necessary
to pick one or more of the H5 subcategories, see
\url{http://www.acm.org/class/1998/ccs98.html}

\subsection{Normal or Body Text}

Please use a 10-point Times Roman font or, if this is unavailable,
another proportional font with serifs, as close as possible in
appearance to Times Roman 10-point. The Press 10-point font available
to users of Script is a good substitute for Times Roman. If Times
Roman is not available, try the font named Computer Modern Roman. On a
Macintosh, use the font named Times and not Times New Roman. Please
use sans-serif or non-proportional fonts only for special purposes,
such as headings or source code text.

\subsection{First Page Copyright Notice}

Leave 3 cm (1.25 in.) of blank space for the copyright notice at the
bottom of the left column of the first page. In this template a
floating text box will automatically generate the required space. Note
however that the text box is anchored to the \textbf{ABSTRACT}
heading, so if that heading is deleted the text box will disappear as
well.  You can replace the default copyright notice by uncommenting
the {\textbackslash}toappear block at the beginning of the document
and inserting your own text, for example, for versions under review.

\subsection{Subsequent Pages}

On pages beyond the first, start at the top of the page and continue
in double-column format.  The two columns on the last page should be
of equal length.

\begin{figure}[!h]
\centering
\includegraphics[width=0.9\columnwidth]{Figure1}
\caption{With Caption Below, be sure to have a good resolution image
  (see item D within the preparation instructions).}
\label{fig:figure1}
\end{figure}

\subsection{References and Citations}

Use a numbered list of references at the end of the article, ordered
alphabetically by first author, and referenced by numbers in brackets
\cite{ethics,
  Klemmer:2002:WSC:503376.503378,
  Mather:2000:MUT,
  Zellweger:2001:FAO:504216.504224}. For
papers from conference proceedings, include the title of the paper and
an abbreviated name of the conference (e.g., for Interact 2003
proceedings, use \textit{Proc. Interact 2003}). Do not include the
location of the conference or the exact date; do include the page
numbers if available. See the examples of citations at the end of this
document. Within this template file, use the \texttt{References} style
for the text of your citation.

Your references should be published materials accessible to the
public.  Internal technical reports may be cited only if they are
easily accessible (i.e., you provide the address for obtaining the
report within your citation) and may be obtained by any reader for a
nominal fee.  Proprietary information may not be cited. Private
communications should be acknowledged in the main text, not referenced
(e.g., ``[Robertson, personal communication]'').

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|}
    \hline
    \tabhead{Objects} &
    \multicolumn{1}{|p{0.3\columnwidth}|}{\centering\tabhead{Caption --- pre-2002}} &
    \multicolumn{1}{|p{0.4\columnwidth}|}{\centering\tabhead{Caption --- 2003 and afterwards}} \\
    \hline
    Tables & Above & Below \\
    \hline
    Figures & Below & Below \\
    \hline
  \end{tabular}
  \caption{Table captions should be placed below the table.}
  \label{tab:table1}
\end{table}

\section{Sections}

The heading of a section should be in Helvetica 9-point bold, all in
capitals. Use Arial if Helvetica is not available. Sections should
not be numbered.

\subsection{Subsections}

Headings of subsections should be in Helvetica 9-point bold with
initial letters capitalized.  For
sub-sections and sub-subsections, a word like \emph{the} or \emph{of}
is not capitalized unless it is the first word of the heading.)

\subsubsection{Sub-subsections}

Headings for sub-subsections should be in Helvetica 9-point italic
with initial letters capitalized.  Standard {\textbackslash}section,
{\textbackslash}subsection, and {\textbackslash}subsubsection commands
will work fine.

\section{Figures/Captions}

Place figures and tables at the top or bottom of the appropriate
column or columns, on the same page as the relevant text
(see Figure~\ref{fig:figure1}). A figure or table may extend across both
columns to a maximum width of 17.78 cm (7 in.).

Captions should be Times New Roman 9-point bold.  They should be numbered (e.g.,
``Table~\ref{tab:table1}'' or ``Figure~\ref{fig:figure2}''), centered
and placed beneath the figure or table.  Please note that the words
``Figure'' and ``Table'' should be spelled out (e.g., ``Figure''
rather than ``Fig.'') wherever they occur.

Papers and notes may use color figures, which are included in the page
limit; the figures must be usable when printed in black and white in
the proceedings.  The paper may be accompanied by a short video figure
up to five minutes in length.  However, the paper should stand on its
own without the video figure, as the video may not be available to
everyone who reads the paper.

\section{Language, Style and Content}

The written and spoken language of SIGCHI is English. Spelling and
punctuation may use any dialect of English (e.g., British, Canadian,
US, etc.) provided this is done consistently. Hyphenation is
optional. To ensure suitability for an international audience, please
pay attention to the following:

\begin{itemize}
\item Write in a straightforward style.
\item Try to avoid long or complex sentence structures.
\item Briefly define or explain all technical terms that may be
  unfamiliar to readers.
\item Explain all acronyms the first time they are used in your text---e.g.,
``Digital Signal Processing (DSP)''.
\item Explain local references (e.g., not everyone knows all city
  names in a particular country).
\item Explain ``insider'' comments. Ensure that your whole audience
  understands any reference whose meaning you do not describe (e.g.,
  do not assume that everyone has used a Macintosh or a particular
  application).
\item Explain colloquial language and puns. Understanding phrases like
  ``red herring'' may require a local knowledge of English.  Humor and
  irony are difficult to translate.
\item Use unambiguous forms for culturally localized concepts, such as
  times, dates, currencies and numbers (e.g., ``1-5-97'' or ``5/1/97''
  may mean 5 January or 1 May, and ``seven o'clock'' may mean 7:00 am or
  19:00).  For currencies, indicate equivalences---e.g., ``Participants
  were paid 10,000 lire, or roughly \$5.''
\item Be careful with the use of gender-specific pronouns (he, she)
  and other gendered words (chairman, manpower, man-months). Use
  inclusive language that is gender-neutral (e.g., she or he, they,
  s/he, chair, staff, staff-hours,
  person-years). See~\cite{Schwartz:1995:GBF} for further advice and
  examples regarding gender and other personal attributes.
\item If possible, use the full (extended) alphabetic character set
  for names of persons, institutions, and places (e.g.,
  Gr{\o}nb{\ae}k, Lafreni\'ere, S\'anchez, Universit{\"a}t,
  Wei{\ss}enbach, Z{\"u}llighoven, \r{A}rhus, etc.).  These characters
  are already included in most versions of Times, Helvetica, and Arial
  fonts.
\end{itemize}

\section{Page Numbering, Headers and Footers}

Please submit your anonymous version for reviewing with page numbers
centered in the footer.  These must be removed in the final version of
accepted papers, as page numbers, headers, and footers will be added
by the conference printers.  Comment out the {\textbackslash}pagenumbering
command at the top of the document to remove page numbers.

\section{Producing and Testing PDF Files}

We recommend that you produce a PDF version of your submission well
before the final deadline.  Your PDF file must be ACM DL
Compliant. The requirements for an ACM Compliant PDF are available at:
{\url{http://www.sheridanprinting.com/typedept/ACM-distilling-settings.htm}}.

Test your PDF file by viewing or printing it with the same software we
will use when we receive it, Adobe Acrobat Reader Version 7. This is
widely available at no cost from~\cite{acrobat}.  Note that most
reviewers will use a North American/European version of Acrobat
reader, which cannot handle documents containing non-North American or
non-European fonts (e.g. Asian fonts).  Please therefore do not use
Asian fonts, and verify this by testing with a North American/European
Acrobat reader (obtainable as above). Something as minor as including
a space or punctuation character in a two-byte font can render a file
unreadable.

\section{Blind Review}

For archival submissions, CHI requires a ``blind review.'' To prepare
your submission for blind review, remove author and institutional
identities in the title and header areas of the paper. You may also
need to remove part or all of the Acknowledgments text.  Further
suppression of identity in the body of the paper and references is
left to the authors' discretion. For more details, see the submission
guidelines and checklist for your submission category.

\section{Conclusion}

It is important that you write for the SIGCHI audience.  Please read
previous years' Proceedings to understand the writing style and
conventions that successful authors have used.  It is particularly
important that you state clearly what you have done, not merely what
you plan to do, and explain how your work is different from previously
published work, i.e., what is the unique contribution that your work
makes to the field?  Please consider what the reader will learn from
your submission, and how they will find your work useful.  If you
write with these questions in mind, your work is more likely to be
successful, both in being accepted into the Conference, and in
influencing the work of our field.

\section{Acknowledgments}

We thank CHI, PDC and CSCW volunteers, and all publications support
and staff, who wrote and provided helpful comments on previous
versions of this document.  Some of the references cited in this paper
are included for illustrative purposes only.  \textbf{Don't forget
to acknowledge funding sources as well}, so you don't wind up
having to correct it later.

% Balancing columns in a ref list is a bit of a pain because you
% either use a hack like flushend or balance, or manually insert
% a column break.  http://www.tex.ac.uk/cgi-bin/texfaq2html?label=balance
% multicols doesn't work because we're already in two-column mode,
% and flushend isn't awesome, so I choose balance.  See this
% for more info: http://cs.brown.edu/system/software/latex/doc/balance.pdf
%
% Note that in a perfect world balance wants to be in the first
% column of the last page.
%
% If balance doesn't work for you, you can remove that and
% hard-code a column break into the bbl file right before you
% submit:
%
% http://stackoverflow.com/questions/2149854/how-to-manually-equalize-columns-
% in-an-ieee-paper-if-using-bibtex
%
% Or, just remove \balance and give up on balancing the last page.
%
\balance

% If you want to use smaller typesetting for the reference list,
% uncomment the following line:
% \small
\bibliographystyle{acm-sigchi}
\bibliography{sample}
\end{document}

### 2012-09-13 chi2012.tex
relwork and bibtex

@@
 % Consult the conference website for the camera-ready copyright statement.
-\toappear{
-Permission to make digital or hard copies of all or part of this
-work for personal or classroom use is granted without fee provided that 
-copies are not made or distributed for profit or commercial advantage and
-that copies bear this notice and the full citation on the first page. To
-copy otherwise, or republish, to post on servers or to redistribute to 
-lists, requires prior specific permission and/or a fee.\\
-{\confname{CHI'12}}, May 5--10, 2012, Austin, Texas, USA.\\
-Copyright 2012 ACM 978-1-4503-1015-4/12/05...\$10.00.
-}
+%% \toappear{
+%% Permission to make digital or hard copies of all or part of this
+%% work for personal or classroom use is granted without fee provided that 
+%% copies are not made or distributed for profit or commercial advantage and
+%% that copies bear this notice and the full citation on the first page. To
+%% copy otherwise, or republish, to post on servers or to redistribute to 
+%% lists, requires prior specific permission and/or a fee.\\
+%{\confname{CHI'13}}, May 5--10, 2012, Austin, Texas, USA.\\
+%Copyright 2012 ACM 978-1-4503-1015-4/12/05...\$10.00.
+%}
 
@@
 \begin{abstract}
+
 \end{abstract}
 
-\keywords{
-	Guides; instructions; author's kit; conference publications;
-	keywords should be separated by a semi-colon.
-	\textcolor{red}{Mandatory section to be included in your final version.}
-}
+\keywords{}
 
@@
 
-See: \url{http://www.acm.org/about/class/1998/}
-for more information and the full list of ACM classifiers
-and descriptors. 
-\textcolor{red}{Mandatory section to be included in your
-final version. On the submission page only the classifiers'
-letter-number combination will need to be entered.}
-
-\terms{
-	Human Factors; Design; Measurement. 
-	If you choose more than one ACM General Term, 
-	separate the terms with a semi-colon.
-}
-
-See list of the limited ACM 16 terms in the
-instructions and additional information:
-\url{http://www.sheridanprinting.com/sigchi/generalterms.htm}.
-\textcolor{red}{Optional section to be included in your final version.}
+\terms{Human Factors; Design; Measurement.}
 
@@
 We start with a small interview study examining information tasks people perform that rely on multiple sources of information, and how such needs are currently satisfied.
-Our results suggest that people prefer to rely upon multiple, diverse sources to singular integrated ones for a number of reasons, including coverage, reliability, and \_\_\_\_\_\_.
+Our results suggest that people prefer to rely upon multiple, diverse sources to singular integrated ones for a number of reasons, including coverage, reliability, and seeking out more suitable alternatives.
 Second, we conduct a analysis of various popular personal data sources available today to identify barriers to effective unification of data from each.
@@
 
-Based upon the data integration needs revealed by people in the interviews, and the kinds of heterogeneity observed in the data feeds, we embarked on a user-centric design exercise towards an interface to facilitate simple, "light-touch" integration tasks of diverse, heterogeneous data sources.
-We describe our initial foray into this exercise, an interface we call DataPalette, which focuses on the most common types of heterogeneity to enable serendipitous "data mixing" - a form of simple integration sufficient to let people easily and effectively combine and compare information sources without the need to write code.
-Our usability evaluation of Data Palette reveals that most users were comfortable with and that this interface is effectively expedites the most sophisticated of the typical kinds of "data integration" tasks people need to perform while planning personal,
+Based upon the data integration needs revealed by people in the interviews, and the kinds of heterogeneity observed in the data feeds, we embarked on a user-centric design exercise towards an interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
+We describe our initial foray into this exercise, an interface we call DataPalette, which focuses on the most common types of heterogeneity to enable serendipitous ``data mixing'' - a form of simple integration sufficient to let people easily and effectively combine and compare information sources without the need to write code.
+Our usability evaluation of Data Palette reveals that most users were comfortable with and that this interface is effectively expedites the most sophisticated of the typical kinds of ``data integration'' tasks people need to perform while planning personal,
 
@@
 
-As the theoretical basis of data integration extends nearly a half-century in the database systems literature .
-As described by Halvey et al’s comprehensive textbook on data integration problems and techniques \cite{principlesofdataintegration} much of this focus of this community has focused on \emph{engineered} or \emph{automatic} approaches to data integration at a large scale,
+In this section, we touch on motivating work for DataPalette from three fields: peronal information management, database integration, and end-user mashups and toolkits.
 
-In scientific laboratory settings,
+First,
 
-We only briefly touch on the extensive literature covering the “deep end” of data integration - both automatic and “hand-coded” techniques to link multiple, typically large, data sources in consistent, correct and efficient ways, little work has pertained on the opposite end -- the “light”, “small” and “sketchy” end of data integration, such as arises when multiple heterogeneous sources of data about similar kinds of things are brought together spontaneously
+Data integration has been a major recurring theme in the literature of personal information management.
+The problem of data fragmentation, for example, defined by Jones et al \cite{jones}, can be seen as as an inability failure of current PIM practices to support effective consolidation of data from multiple sources.
+In our own study, (reference anonymised) we observed the manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions)  devised to manage and updating data from external sources.
+Similar observations have been made in others' studies, such as Voida et al.'s findings concerning volunteer coordinators' ``homebrew databases'', manually-mainted information assemblages that were concocted to meet data integration needs in the absence of tools with suitable capabilities.
 
-Interfaces that let end-users find, consolidate, and work with heterogeneous data 
-PIM 
-	Amy Voida’s databases paper
-User-subjective approach to information management
-
-Information scraps
-End-user Data Gathering and integration platforms
-	Potluck
-	Cards (M. Dontcheva)
-
-End-user data gathering
-	Piggy Bank
-“data notebooks”	Hunter-gatherer
-	Yahoo Pipes
-PROMPT
-
-A key assumption of this work is that ontologies are not going to converge.  (Studies about the diversity of web ontologies backing this claim up?)
-
-Why end-user data integration is desirable
-User-subjective approach to personal information management
-
-Why is the data integration problem so hard?  
-The problem of reconciling schema heterogeneity has been a subject of research for decades, but solutions are few.  In the words of leading data integration researcher Alon Halevy,
-[The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs. Hence, even if they model
-overlapping domains, they will model them in different ways.
-Differing structures are a byproduct of human nature ?people think differently from one another even when faced with  the same modeling goal.
-\cite{alonhalevy}
-
-Exhibit - designed for personal structured curation
-Yahoo Pipes! - designed for not-personal (public) curation
-MySupermarket - http://mysupermarket.co.uk/
-Demonstrates the power of joining across data sources: people can gather a grocery list, and compare how much the same shopping list would cost at each of the major UK supermarket chains (ASDA, Tesco, Waitrose, ocado) --
-
-But what about the mom + pop groceries?
-What if they’re missing data: allergies, suitable for -- available from somewhere else
-
-Schema.org  - is it helping or not?
+A number of research PIM systems have been devised to solve this problem.
 
@@
 \bibliographystyle{acm-sigchi}
-\bibliography{sample}
+\bibliography{carpedata}
 \end{document}

### 2012-09-13 relwork-scratch.tex
relwork and bibtex

extraction/collection from pages
                      sifter
                      hunter-gatherer
                      cards
     TX2 - learning during collection

post-scraping structured data mashups
    pipes, popfly, QDEWiki

PIM platform (on the fly data integration)
    SEMEX - 
     Haystack -

mediating schema, pairwise mediationnz

scraperwiki

----

As the theoretical basis of data integration extends nearly a half-century in the database systems literature . As described by Halvey et al’s comprehensive textbook on data integration problems and techniques \cite{principlesofdataintegration} much of this focus of this community has focused on \emph{engineered} or \emph{automatic} approaches to data integration at a large scale,

In scientific laboratory settings,

We only briefly touch on the extensive literature covering the “deep end” of data integration - both automatic and “hand-coded” techniques to link multiple, typically large, data sources in consistent, correct and efficient ways, little work has pertained on the opposite end -- the `light', “small” and “sketchy” end of data integration, such as arises when multiple heterogeneous sources of data about similar kinds of things are brought together spontaneously

Interfaces that let end-users find, consolidate, and work with heterogeneous data 
PIM 
	Amy Voida’s databases paper
User-subjective approach to information management

Information scraps
End-user Data Gathering and integration platforms
	Potluck
	Cards (M. Dontcheva)

End-user data gathering
	Piggy Bank
“data notebooks”	Hunter-gatherer
	Yahoo Pipes
PROMPT

A key assumption of this work is that ontologies are not going to converge.  (Studies about the diversity of web ontologies backing this claim up?)

Why end-user data integration is desirable
User-subjective approach to personal information management

Why is the data integration problem so hard?  
The problem of reconciling schema heterogeneity has been a subject of research for decades, but solutions are few.  In the words of leading data integration researcher Alon Halevy,
[The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs. Hence, even if they model
overlapping domains, they will model them in different ways. Differing structures are a byproduct of human nature ?people think differently from one another even when faced with  the same modeling goal. \cite{alonhalevy}

Exhibit - designed for personal structured curation
Yahoo Pipes! - designed for not-personal (public) curation
MySupermarket - http://mysupermarket.co.uk/
Demonstrates the power of joining across data sources: people can gather a grocery list, and compare how much the same shopping list would cost at each of the major UK supermarket chains (ASDA, Tesco, Waitrose, ocado) --

But what about the mom + pop groceries?
What if they’re missing data: allergies, suitable for -- available from somewhere else

Schema.org  - is it helping or not?

### 2012-09-13 chi2012.tex
a bit more in rel work

@@
 
-In this section, we touch on motivating work for DataPalette from three fields: peronal information management, database integration, and end-user mashups and toolkits.
+In this section, we touch on motivating work for DataPalette from three fields: database integration, peronal information management, and end-user mashups and toolkits.
 
-First,
+The primary motivation for this work stems from field studies in personal information management, where data integration has been a major recurring theme.
+The problem of data fragmentation, for example, defined by Jones et al \cite{jones-fragmentation}, stems from a failure of current PIM practices to support effective consolidation of data from multiple sources.
+In our own study, (reference anonymised) we observed several manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) that people devised to manage data when they came from multiple sources -- be them online, offline, or from other people.
+Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-mainted information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate tehir activities \cite{voida-homebrew}.
+%% TODO MORE HERE
 
-Data integration has been a major recurring theme in the literature of personal information management.
-The problem of data fragmentation, for example, defined by Jones et al \cite{jones}, can be seen as as an inability failure of current PIM practices to support effective consolidation of data from multiple sources.
-In our own study, (reference anonymised) we observed the manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions)  devised to manage and updating data from external sources.
-Similar observations have been made in others' studies, such as Voida et al.'s findings concerning volunteer coordinators' ``homebrew databases'', manually-mainted information assemblages that were concocted to meet data integration needs in the absence of tools with suitable capabilities.
+In order for PIM tools to start to support integration of arbitrary external data, we consulted the database ``data integration'' a transient
+
+\begin{quote} 
+[The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs. Hence, even if they model
+overlapping domains, they will model them in different ways.
+Differing structures are a byproduct of human nature ?people think differently from one another even when faced with  the same modeling goal.
+\cite{alonhalevy}
+\end{quote}
 

### 2012-09-13 chi2012.tex
related work etc

@@
 We describe our initial foray into this exercise, an interface we call DataPalette, which focuses on the most common types of heterogeneity to enable serendipitous ``data mixing'' - a form of simple integration sufficient to let people easily and effectively combine and compare information sources without the need to write code.
-Our usability evaluation of Data Palette reveals that most users were comfortable with and that this interface is effectively expedites the most sophisticated of the typical kinds of ``data integration'' tasks people need to perform while planning personal,
+Our usability evaluation of DataPalette reveals that most users were comfortable with and that this interface is effectively expedites the most sophisticated of the typical kinds of ``data integration'' tasks people need to perform while planning personal,
 
-\section{Related Work}
+\section{Background}
 
@@
 
-The primary motivation for this work stems from field studies in personal information management, where data integration has been a major recurring theme.
-The problem of data fragmentation, for example, defined by Jones et al \cite{jones-fragmentation}, stems from a failure of current PIM practices to support effective consolidation of data from multiple sources.
-In our own study, (reference anonymised) we observed several manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) that people devised to manage data when they came from multiple sources -- be them online, offline, or from other people.
-Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-mainted information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate tehir activities \cite{voida-homebrew}.
+The primary motivation for this work stems from field studies in personal information management, where effectively consolidating and
+combining data from multiple sources has been a major recurring theme.
+The problem of data fragmentation (\cite{Jones05towardsa}, for example, arises, in part, from a failure of PIM tools to adequately support the effective consolidation of data across multiple applications, services and platforms (e.g., \cite{bergman},  \cite{boardmansasse}).
+In our own previous field study, (reference anonymised) we observed several different variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) that people devised to merge in data from external sources -- be them online, offline, or other people.
+Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-mainted information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
 %% TODO MORE HERE
 
-In order for PIM tools to start to support integration of arbitrary external data, we consulted the database ``data integration'' a transient
+%% don't like the beginning of this > 
+It became clear that effectively coping with data heterogeneity would be a central focus to letting people effectively draw upon multiple, arbitrary sources of data when consulting the theoretical data integration literature.
+Alon Halvey describes hetereogeneity as an inevitable by-product of the distributed processes that are used to author it:
 
 \begin{quote} 
-[The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs. Hence, even if they model
-overlapping domains, they will model them in different ways.
+[The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs.
+Hence, even if they model overlapping domains, they will model them in different ways.
 Differing structures are a byproduct of human nature ?people think differently from one another even when faced with  the same modeling goal.
@@
 
-A number of research PIM systems have been devised to solve this problem.
+In this paper, we take the position that there needs to be a difference in focus between the data integration approaches discussed in the databases literature and that of PIM, because of corresponding differences in requirements.
+In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one.
+This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted.
+The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined.
+In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would be
+fully consume the time and effort a person has allocated for their task.
+Moreover, while a fullly, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
+
+A number of prototype PIM systems have demonstrated the kinds of
+capabilities that effective management of data heterogeneity can provide for users. The SEMEX \cite{semex} and Haystack \cite{haystack} systems,
+
+The mash-up
 
@@
 
-\section{WebBox Interface - Design of the Data Palette}
+\section{WebBox Interface - Design of the DataPalette}
 

### 2012-09-13 chi2012.tex
Added 1st draft of study methodology

@@
 
-Lab based usability study,
+\subsection{Participants}
+We recruited 20 participants (10 female) from the University of Southampton’s international student mailing list.
+All of the participants spoke English, and for 12 of them, English was not their first language, however they had all passed the university English proficiency test.
+Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.
+All of the participants had previously encountered tasks requiring them to use and evaluate data from multiple websites.
 
-Design of Study
-How we designed the tasks, subjective task that requires multi dimensions 
-How we got data
-How we recruited participants - International student list
+\subsection{Task Design}
+We designed a semi-structured lab-based study around four possible tasks shown in Table \ref{tab:studyfactors}.
+The tasks were to select a restaurant for an event or university at which to study, out of six possible choices.
+We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choices.
+The shortlist of universities and restaurants were taken from the top six ranked from the Complete University Guide (for task 1 and 2), Tripadvisor (for task 3) and Yelp (for task 4).
+The top six were chosen from these websites because they provide a realistic shortlist and are comparable in terms of rankings and properties.
+The web sites used as data sources were chosen because they are the most popular and contain comparable statistics for universities and restaurants.
 
-Our aim was XXX, To accomplish thies XX participants were recruited to the lab for a 1 hour session
+\begin{table*}[t]
+\caption{Tasks, choices and data sources}
+\begin{center}
 
-	Latin square dance!
-	4 factors
-		Interface or not
-		Task (restaurant or uni)
-		Dataset (glasgow / cambridge; history or sports)
-		Participant
+\begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
+\hline
+Task	 &Choices	&Data Sources\\
+\hline
+1) Three universities they would apply to study Sport Science & Universities: 1) Loughborough,  2) Durham,  3) Exeter, 4) Edinburgh, 5) Birmingham, and 6) Bath & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
+\hline
+2) Three universities they would apply to study History & Universities: 1) Cambridge, 2) London School of Economics and Political Science, 3) Durham, 4) Oxford, 5) St Andrews, and 6) University College London & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
+\hline
+3) Restaurant they would book for 12 friends in Cambridge&Restaurants : 1) Alimentum, 2) Gardenia Restaurant, 3) The Cambridge Chop House, 4) Al Casbah, 5) Midsummer House Restaurant, and 6) Nando's Restaurant.
+& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
+\hline
+4) Restaurant they would book for 12 friends in Glagow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollopes Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia.
+& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
+\hline
+\end{tabular}
+\end{center}
+\label{tab:studyfactors}
+\end{table*}%
 
-Hypotheses
+Each task had an A and B condition.
+The A condition, the participant could only use WebBox to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or spreadsheets in Microsoft Excel containing the data taken from the data sources to make their choice.
+
+Each study consisted of two tasks, one A and one B condition.
+Each participant was given 10 minutes to complete a task.
+In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.
+In order to train the user on how to use our interface, they viewed a five-minute WebBox tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
+
+\subsection{Data Collection}
+Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study’s protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
+During the lab study, we recorded the audio and the participant’s actions on screen.
+We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.
+
+The studies lasted 40 minutes, on average.
+Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn’t influence the study.
+The studies were run over 4 days.
+At the end of each day the facilitators and observers met to discuss the studies’ data and to revisit protocols, if necessary.
+
+\subsection{Data Analysis}
+
+During the data collection we collated a spreadsheet summarising the participant’s gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
+After we had collected the study’s data, we performed our data analysis.
+
+TODO look further into types of data analysis....
+
+\subsection{Hypotheses}
 Interaction-method specific hypotheses
-	h1.1 - Do people understanding the problem of co-reference, and the need to reconcile coreference problems?
-	h1.2. Does the ability to do drag + drop combination for co-reference reconciliation effectively solve this problem?
-	h2.1- Do people understand the problem of structural differences in data?
-h2.2 - Does multipath selection let people effectively deal with this -- and work with collections of heterogeneous items?
-
+\begin{itemize}
+	\item h1.1 - Do people understanding the problem of co-reference, and the need to reconcile coreference problems?
+	\item h1.2. Does the ability to do drag + drop combination for co-reference reconciliation effectively solve this problem?
+	\item h2.1- Do people understand the problem of structural differences in data?
+	\item h2.2 - Does multipath selection let people effectively deal with this -- and work with collections of heterogeneous items?
+\end{itemize}
 Are these interaction methods sufficient to perform common tasks involving heterogeneous data sources?  Does supporting these simple techniques facilitate task completion?

### 2012-09-13 chi2012.tex
Added paragraphs to background section

@@
 The mash-up
+
+%the following needs integrating with the rest of the lit rev.
+The Semantic Web community use ontologies to formally describe vocabularies for specific domains by defining axioms describing entities, instances of entities and the relationships that hold among them \cite{borst1997construction}.
+The axioms in an ontology can be either assertional or terminological, the former describes the schema of the entities and the latter the instances of those entities.
+Typically, these ontologies are marked up using OWL (Web Ontology Language) and Resource Description Framework Schema (RDF(S)).
+ There has been an explosion of ontologies published online, many of which describe the same entities.
+ These entities may or may not share the same entity names or properties.
+ Therefore using data from heterogeneous ontologies can be difficult.
+The field of ontology matching can be split into two fields: schema and instance matching.
+ Instance matching differs from schema level matching, because it aims to detect instances referring to the same real world object, while schema matching aims to find a set of mappings between concepts and properties in different ontologies.
+
+For our purposes of allowing users to explore and compare data from different websites, we are particularly interested in aligning instances of the most common forms of data that people compare from website to website.
+ Instance matching within this scope is in its infancy, some of its problems have been addressed by the record linkage problem in the database field \cite{fellegi1969theory,winkler1999state,gu2003record}.
+ However, there are new problems to address in this field, which are strongly related to schema matching, including structural heterogeneity and logical heterogeneity.
+
+The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instances’ schema’s structural and logical definitions \cite{castano2006matching}.
+ In our case, this is problematic because information from the web does not always have a schema.
+ In real-world scenarios, semi-automatic ontology matching approaches are favoured over automatic because reconciling the differences between ontologies that were designed for different purposes is not always accurate.
+ Semi-automatic approaches use varying levels of support to aid a user, from prompting users to manually check flagged alignments, to suggesting possible alignments.
+ Allowing users to compare instances across heterogeneous data is powerful because users can combine and evaluate multiple sources.
+ In principle, automatic instance matching approaches would allow users to seamlessly browse this data, however for now and the foreseeable future such approaches cannot make use of personal knowledge and do not guarantee their alignments.
+ Therefore in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.
 

### 2012-09-13 chi2012.tex
working on related work

--- 2012-09-13 chi2012.tex
@@
 
-The primary motivation for this work stems from field studies in personal information management, where effectively consolidating and
-combining data from multiple sources has been a major recurring theme.
+The primary motivation for this work stems from field studies in personal information management, where letting users effectively consolidating and work with data from multiple sources has remained a challenge for over a decade.
 The problem of data fragmentation (\cite{Jones05towardsa}, for example, arises, in part, from a failure of PIM tools to adequately support the effective consolidation of data across multiple applications, services and platforms (e.g., \cite{bergman}, \cite{boardmansasse}).
@@
 fully consume the time and effort a person has allocated for their task.
-Moreover, while a fullly, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
+Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
 
-A number of prototype PIM systems have demonstrated the kinds of
-capabilities that effective management of data heterogeneity can provide for users. The SEMEX \cite{semex} and Haystack \cite{haystack} systems,
+A number of prototype PIM systems have dmonstrated data integration capacities.
+For example, SEMEX \cite{semex} was a PIM environment that provided users a uniform query interface over the contents of their personal information environments, across heterogeneous files, folders and repositories.
+Similarly, Haystack was a personal information manager that allowed users to manipulate and managing their data in a uniform, consolidated manner \cite{haystack}.
+Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
 
-The mash-up
+Systems that have focused solely on the end-user-driven, interactive reconciliation of data heterogeneity have come out of the web ``data mash-up'' community.
+Yahoo Pipes!
+is a visual programming language aimed at letting novice programmers easily create data-transformation workflows, but remained rather difficult to use and poorly suited for one-off information tasks.
+Mash-up makers such as Intel Mashmaker \cite{}, QDEWiki, and TX2
+
+DataPalette's approach is most directly inspired by ``lightweight'' visual, user-interactive data collection, consolidation and alignment tools such as Potluck \cite{citeulike:3875264} and Cards \cite{Dontcheva:2007:RCS:1294211.1294224}.
 
@@
 Each interview was recorded and notes taken, the interviewer was trained so that they were knowledgeable about the importance of the study and to minimise bias.
+
 \subsection{Results and Discussion}
@@
 
-Question 1: Tasks
+\subsubsection{Question 1: Tasks}
 All participants had experience in using multiple websites to complete tasks.
@@
 
-Question 2: Large Social Event
+\subsubsection{Question 2: Large Social Event}
 All of the participants would organise a large social event by talking to their friends, about their ideas, preferences and recommendations.

### 2012-09-13 chi2012.tex
Added qual

--- 2012-09-13 chi2012.tex
@@
 \section{Results}
-How many native english speakers, only 8 were all enrolled at university all past proficiency test.
+
+% behaviours in websites
+
+Users who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues. In
+particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of
+university teaching by the Times Higher Education. There were numerous usability issues in using websites, particularly
+in searching for university courses and course requirements. Similarly, participants found it difficult to understand the
+locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto
+the same map using the websites. Overall, the users that spent the most time using websites exhibited significant
+frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations),
+and all ended up falling back to using the supplied spreadsheets.
+
+% behaviours in excel
+One of the designs of the study was to produce a spreadsheet worksheet for each website, which meant that users
+had to switch between the worksheets to gather data from multiple sources. Different approaches to using the spreadsheets
+were observed. Some users were keen to collate all data onto a single sheet, so that they could compare the values
+all at once. However, this approach was error prone, for example participants noticing too late that the ordering of rows
+was different on each sheet, so that their pasted values were incorrectly mapped. Others used pen and paper to collate
+the values, which was often faster and resulted in less observed frustration. While the majority of users did not
+attempt anything complex with Excel (in fact only one users used charts), there was some advanced use of Excel observed.
+One user explained that they did not like to look at numbers, and so coloured the cells of the types of restaurant, and the
+friends' preferences, and matched the colours to make their decision. Another user calculated their own weighted metric
+of which university to choose by averaging different metrics supplied by the data sources.
+
+% webbox client
+
+The most commonly observed use of WebBox was to collate the data on-screen, and compare multiple values directly.
+About half of the observed users actually used the instance combination feature of WebBox, with the other half able
+to successfully make decisions without using it. Those who tried to use it were largely successful, and were able
+to therefore use the global brush-highlighting to simultaneously compare different statistics. Most users instantly opened
+up all data sources into individual boxes, in order to determine which sources contained specific statistics. The result
+of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.
+
+%% more from doc here
+
+% weird and wonderful
+A number of users admittedly introduced their own bias into the choices: "I prefer Indian food, so I will choose the Indian
+restaurant", and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating
+those that required bus transfers (the protocol did not mention routing/transfers at all).
+
 \section{Discussion}
-What we found
 
-What we didn’t find (evidence for)
+% processes
 
-What we might look at next -- follow-on work
-		We focused on terminological heterogeneity -- didn’t address structural heterogeneity, semantic heterogeneity
-		We focused on finite (small) sets Streaming data (how do select future data --) 
-		Scale - small data, what are the facilites required for large data
-		Collaborative - how might interaction support multiple users collaborating?
+In order to complete the tasks, the users followed a number of different processes. The most common process followed
+was {\it collation} of data, either on-screen or on paper. Some users were proficient enough to copy and paste data
+in the supplied spreadsheets from worksheet to worksheet, while others manually typed data value-by-value across worksheets.
+Others preferred to collate the data from screen to paper, in order to compare multiple data points. A similar process that
+some participants followed was to determine which values of a particular statistic were acceptable, and keep a paper tally
+of these next to the set of choices. They then made their choice based on the value of this tally.
 
-Limitations found from study... what is the take away from the limitations
+Two contrasting approaches that participants followed was to use the data to rule out candidates, or to use the data to
+directly compare candidates at once. Our pool of participants exhibited both behaviours evenly, however in the condition
+when using the WebBox interface, significantly more comparison (and visualisation of values) was performed, with fewer
+participants using the "ruling out" process. A consequence of this was that when asked to justify their choices, users were
+able to verify their values on-screen immediately, whereas those who had ruled-out had to memorise their reasons, with
+mixed success  -- many participants had to look through the data sources to re-find the reasons why they has ruled out some choices.
+Therefore, the uses that were using WebBox were more confident in their choices, and knew exactly why they had chosen them,
+often pointing to the screens to show and verify each choice as they explained them.
+
+%What we found 
+%
+%What we didn’t find (evidence for)
+%
+%What we might look at next -- follow-on work
+%		We focused on terminological heterogeneity -- didn’t address structural heterogeneity, semantic heterogeneity
+%		We focused on finite (small) sets Streaming data (how do select future data --) 
+%		Scale - small data, what are the facilites required for large data
+%		Collaborative - how might interaction support multiple users collaborating?
+%
+%Limitations found from study... what is the take away from the limitations
+
 \section{Conclusion} %Key findings of the whole paper and its implications

### 2012-09-13 chi2012.tex
got rid of Es

@@
 The primary motivation for this work stems from field studies in personal information management, where letting users effectively consolidating and work with data from multiple sources has remained a challenge for over a decade.
-The problem of data fragmentation (\cite{Jones05towardsa}, for example, arises, in part, from a failure of PIM tools to adequately support the effective consolidation of data across multiple applications, services and platforms (e.g., \cite{bergman}, \cite{boardmansasse}).
-In our own previous field study, (reference anonymised) we observed several different variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) that people devised to merge in data from external sources -- be them online, offline, or other people.
+The problem of data fragmentation (\cite{Jones05towardsa}, for example, derives from an inability to effectively consolidate data across multiple applications, services and platforms (e.g., \cite{bergman}, \cite{boardmansasse}).
+In our own previously conducted field study, (reference anonymised) we observed several different variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) that people devised to merge in data from external sources -- be them online, offline, or other people.
 Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-mainted information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
@@
 In principle, automatic instance matching approaches would allow users to seamlessly browse this data, however for now and the foreseeable future such approaches cannot make use of personal knowledge and do not guarantee their alignments.
- Therefore in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.
+Therefore, in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.
 

### 2012-09-13 chi2012.tex
a bit of massage

@@
 
-In this section, we touch on motivating work for DataPalette from three fields: database integration, peronal information management, and end-user mashups and toolkits.
+In this section, we touch on motivating work for DataPalette from three fields: peronal information management, data integration, and end-user mashups and toolkits.
 
-The primary motivation for this work stems from field studies in personal information management, where letting users effectively consolidating and work with data from multiple sources has remained a challenge for over a decade.
+The primary motivation for this work stems from field studies in personal information management, where letting users effectively consolidate and work with data from multiple sources has remained a challenge for over a decade.
 The problem of data fragmentation (\cite{Jones05towardsa}, for example, derives from an inability to effectively consolidate data across multiple applications, services and platforms (e.g., \cite{bergman}, \cite{boardmansasse}).
-In our own previously conducted field study, (reference anonymised) we observed several different variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) that people devised to merge in data from external sources -- be them online, offline, or other people.
+In our own previously-conducted field study, (reference anonymised) we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage heterogeneous data from disparate sources -- be them online, offline, or other people.
 Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-mainted information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
@@
 %% don't like the beginning of this > 
-It became clear that effectively coping with data heterogeneity would be a central focus to letting people effectively draw upon multiple, arbitrary sources of data when consulting the theoretical data integration literature.
-Alon Halvey describes hetereogeneity as an inevitable by-product of the distributed processes that are used to author it:
+Enabling PIM tools to supporting ad-hoc data integration has remained
+difficult because of the complex, theoretically difficult nature of data hetereogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
+As described by Alon Halvey:
 
@@
 Hence, even if they model overlapping domains, they will model them in different ways.
-Differing structures are a byproduct of human nature ?people think differently from one another even when faced with  the same modeling goal.
+Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modeling goal.
 \cite{alonhalevy}
 \end{quote}
+
+Yet, research prototypes of PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.
+For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
+Similarly, Haystack, a personal information manager, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
+Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
 
@@
 Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
-
-A number of prototype PIM systems have dmonstrated data integration capacities.
-For example, SEMEX \cite{semex} was a PIM environment that provided users a uniform query interface over the contents of their personal information environments, across heterogeneous files, folders and repositories.
-Similarly, Haystack was a personal information manager that allowed users to manipulate and managing their data in a uniform, consolidated manner \cite{haystack}.
-Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
 
@@
 
-The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instances’ schema’s structural and logical definitions \cite{castano2006matching}.
+The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instance schemas structural and logical definitions \cite{castano2006matching}.
 In our case, this is problematic because information from the web does not always have a schema.
@@
 
-\section{Pre-study 1 - Understanding Heterogeneity for Everyday Tasks}
-\subsection{Method}
+\section{Pre-studies: Understanding the use and shape of personal data}
+
+% little introductory ditty about why we're doing pre-studies 
+% actually we may wish to combine discussion of both (!)
+
+\subsection{1 - Understanding Heterogeneity in Everyday Tasks}
+\subsubsection{Method}
 Our pre-study investigated the types of tasks people perform online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.
@@
 
-\subsection{Results and Discussion}
+\subsubsection{Results}
 The participant demographic consisted of 8 people, 7 of which were male and 1 female.
@@
 Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
-\subsection{Prestudy 1 Summary of Findings}
+
+\subsubsection{Summary of Findings}
 People said that they would like a website that could support the evolution of an event, that allowed them to post drafts of an event.
@@
 
-\section{Prestudy 2 : Technical Challenges of Integrating the Data}
+\subsection{2 - Technical Challenges of Integrating the Data}
 
-\subsection{Method}
+The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.
 
-\subsection{Results}
+\subsubsection{Method}
+We sought out APIs and data feeds of popular Web services and sites that were aimed at developers targeting consumption of the data by end-users.
+Based on statistics collected We first selected 4 different categories
 
-\subsection{Analysis}
+\subsubsection{Results}
 
-\subsection{Prestudy 2 Summary of Findings}
+\subsubsection{Summary of Findings}
 

### 2012-09-14 chi2012.tex
integrated related work

@@
 %% don't like the beginning of this > 
-Enabling PIM tools to supporting ad-hoc data integration has remained
-difficult because of the complex, theoretically difficult nature of data hetereogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
+Enabling PIM tools to supporting ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data hetereogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
 As described by Alon Halvey:
@@
 
-Yet, research prototypes of PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.
+% i think we do not need to differentiate between semantic web and databse integration here, we're saying they're both bad, let's lump them together to keep the argument flowing and not kill our readers
+The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} research communities have primarily focused efforts on purely-automatic approaches to the problem of data integration.
+Such approaches typically take the form of either \emph{ontology} or \emph{schema-level matching}, or \emph{direct} or \emph{instance-level} matching.
+We presently focus  on \emph{instance-level} matching, as schemas may not always be available, nor easily accesssed by users.
+Approaches to automatic instance matching have included applied natural language processing (NLP) techniques for terminological matching (e.g., \cite{euzenat2004api}), machine-learning techniques that use both structural and terminological features for matching \cite{doan2003learning}, probabilistic approaches \cite{suchanek2011paris}, and instance schema comparisons using structural and logical definitions \cite{castano2006matching}.
+In real-world settings, however, manual and semi-automatic matching approaches are favoured over automatic because of the need for high degrees of accuracy.
+
+A different approach entirely has been taken in the web mash-up community; this community has largely focused on end-user interfaces that allow people to interactively reconcile data from multiple sources.
+Yahoo Pipes!
+is a visual programming language aimed at letting novice programmers easily create data-transformation workflows, but remained rather challenging to  use and effort-intensive for one-off information tasks.
+Mash-up makers such as Intel Mashmaker \cite{}, QDEWiki \cite{}, and TX2 \cite{}.
+DataPalette's approach is most directly inspired by ``lightweight'' visual, user-interactive data collection, consolidation and alignment tools such as Potluck \cite{citeulike:3875264} and Cards \cite{Dontcheva:2007:RCS:1294211.1294224}.
+
+% Semi-automatic approaches use varying levels of support to aid a user, from prompting users to manually check flagged alignments, to suggesting possible alignments.
+Allowing users to compare instances across heterogeneous data is powerful because users can combine and evaluate multiple sources.
+In principle, automatic instance matching approaches would allow users to seamlessly browse this data, however for now and the foreseeable future such approaches cannot make use of personal knowledge and do not guarantee their alignments.
+Therefore, in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.
+
+A number of research PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.
 For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
-Similarly, Haystack, a personal information manager, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
+Similarly, Haystack, a personal information platform, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
 Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
 
-In this paper, we take the position that there needs to be a difference in focus between the data integration approaches discussed in the databases literature and that of PIM, because of corresponding differences in requirements.
-In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one.
-This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted.
-The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined.
-In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would be
-fully consume the time and effort a person has allocated for their task.
-Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
+%the following needs integrating with the rest of the lit rev.
 
-Systems that have focused solely on the end-user-driven, interactive reconciliation of data heterogeneity have come out of the web ``data mash-up'' community.
-Yahoo Pipes!
-is a visual programming language aimed at letting novice programmers easily create data-transformation workflows, but remained rather difficult to use and poorly suited for one-off information tasks.
-Mash-up makers such as Intel Mashmaker \cite{}, QDEWiki, and TX2
+%% The Semantic Web research community has approached the problem of data integration in terms of \emph{ontology matching}, which can be split into two distinct strategies: \emph{schema} and \emph{instance} matching.
+Instance matching differs from schema level matching in that it aims to detect instances referring to the same real world object, while schema matching aims to find a set of mappings between concepts and properties in different ontologies.
+For our purposes of allowing users to explore and compare data from different websites, we are particularly interested in aligning instances of the most common forms of data that people compare from website to website.
 
-DataPalette's approach is most directly inspired by ``lightweight'' visual, user-interactive data collection, consolidation and alignment tools such as Potluck \cite{citeulike:3875264} and Cards \cite{Dontcheva:2007:RCS:1294211.1294224}.
+%% Instance matching within this scope is in its infancy, some of its problems have been addressed by the record linkage problem in the database field \cite{fellegi1969theory,winkler1999state,gu2003record}.
+However, there are new problems to address in this field, which are strongly related to schema matching, including structural heterogeneity and logical heterogeneity.
+The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet\footnote{\emph{Wordnet} - \url{http://wordnet.princeton.edu}} to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instance schemas structural and logical definitions \cite{castano2006matching}.
+In our case, this is problematic because information from the web does not always have a schema.
+In real-world scenarios, semi-automatic ontology matching approaches are favoured over automatic because reconciling the differences between ontologies that were designed for different purposes is not always accurate.
+Semi-automatic approaches use varying levels of support to aid a user, from prompting users to manually check flagged alignments, to suggesting possible alignments.
+Allowing users to compare instances across heterogeneous data is powerful because users can combine and evaluate multiple sources.
+In principle, automatic instance matching approaches would allow users to seamlessly browse this data, however for now and the foreseeable future such approaches cannot make use of personal knowledge and do not guarantee their alignments.
+Therefore, in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.
 
-%the following needs integrating with the rest of the lit rev.
-The Semantic Web community use ontologies to formally describe vocabularies for specific domains by defining axioms describing entities, instances of entities and the relationships that hold among them \cite{borst1997construction}.
+%% The Semantic Web community use ontologies to formally describe vocabularies for specific domains by defining axioms describing entities, instances of entities and the relationships that hold among them \cite{borst1997construction}.
 The axioms in an ontology can be either assertional or terminological, the former describes the schema of the entities and the latter the instances of those entities.
@@
 
-For our purposes of allowing users to explore and compare data from different websites, we are particularly interested in aligning instances of the most common forms of data that people compare from website to website.
+% For our purposes of allowing users to explore and compare data from different websites, we are particularly interested in aligning instances of the most common forms of data that people compare from website to website.
 Instance matching within this scope is in its infancy, some of its problems have been addressed by the record linkage problem in the database field \cite{fellegi1969theory,winkler1999state,gu2003record}.
@@
 
-The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instance schemas structural and logical definitions \cite{castano2006matching}.
+% The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instance schemas structural and logical definitions \cite{castano2006matching}.
 In our case, this is problematic because information from the web does not always have a schema.
@@
 
+In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
+In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one.
+This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted.
+The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined.
+In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.
+Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
+
 % little introductory ditty about why we're doing pre-studies 
@@
 
-\subsection{1 - Understanding Heterogeneity in Everyday Tasks}
+\subsection{PS.1 - Understanding Data Diversity in Everyday Tasks}
 \subsubsection{Method}
@@
 
-\subsection{2 - Technical Challenges of Integrating the Data}
+\subsection{PS.2 - Technical Challenges of Integrating the Data}
 

### 2012-09-14 chi2012.tex
prestudy introductory ditty in place

@@
 
-%the following needs integrating with the rest of the lit rev.
+% integrated into the above ::
 
@@
 For our purposes of allowing users to explore and compare data from different websites, we are particularly interested in aligning instances of the most common forms of data that people compare from website to website.
-
 %% Instance matching within this scope is in its infancy, some of its problems have been addressed by the record linkage problem in the database field \cite{fellegi1969theory,winkler1999state,gu2003record}.
@@
 Therefore, in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.
-
 %% The Semantic Web community use ontologies to formally describe vocabularies for specific domains by defining axioms describing entities, instances of entities and the relationships that hold among them \cite{borst1997construction}.
@@
 Instance matching differs from schema level matching, because it aims to detect instances referring to the same real world object, while schema matching aims to find a set of mappings between concepts and properties in different ontologies.
-
 % For our purposes of allowing users to explore and compare data from different websites, we are particularly interested in aligning instances of the most common forms of data that people compare from website to website.
@@
 However, there are new problems to address in this field, which are strongly related to schema matching, including structural heterogeneity and logical heterogeneity.
-
 % The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instance schemas structural and logical definitions \cite{castano2006matching}.
@@
 
-\section{Pre-studies: Understanding the use and shape of personal data}
+\section{Pre-studies: The Need and Challenge of Data Integration in PIM}
 
-In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
+%% not sure where this goes >> but i want to keep it
+
+Setting out to investigate the potential role for better support for data integration in PIM, we first sought an updated understanding of how and whether people still used multiple information sources in the course of everyday information-related activities.
+There were several reasons for this inquiry; first, we wondered whether people were increasingly relying only on the handful of the massive ``supersites'' (Facebook, Google, Wikipedia) for most of their activities than actually searching for distributed sources of information.
+If this were the case, much of the benefit could be gained by merely integrating PIM tools with the handful of sites, rather than the substantially more challenging problem of integrating data from arbitrary information sources on the Web.
+Second, if people did rely on multiple sites, we wished to know why, and the typical kinds of tasks people did that required diverse information collection - were these kinds of tasks rare?
+If so, this would suggest that perhaps specialised interfaces could be concocted for this purpose rather than it being a PIM need.
+
+The second study we performed an analysis on existing on data available from APIs and structured data feeds on the Web.
+The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.
+We describe our methodology below.
+
+%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
 In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one.
@@
 Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
-
-% little introductory ditty about why we're doing pre-studies 
-% actually we may wish to combine discussion of both (!)
 
@@
 They were split evenly into two age groups of 18-25 and 26-32.
-All but one participant used social network sites very regularly, and they either logged in multiple times a day, left a website page open or used dedicated application for their social networking site’s updates.
-Those that used social networking websites primarily use Facebook.
+All but one participant reported they used social network sites very regularly (many times a day); they said they either logged in multiple times a day, left a website page open or used dedicated application for their social networking site updates.
+Those that used social networking websites primarily used Facebook.
 All of the participants used Twitter to listen to broadcasted tweets, and 6 out of 8 people regularly tweeted.
@@
 \subsubsection{Question 1: Tasks}
-All participants had experience in using multiple websites to complete tasks.
+All participants reported having used multiple websites to accomplish a task.
 They listed example tasks such as shopping (including food, consumer electronics, hotels and flights), searching for jobs, choosing recipes, study and work.

### 2012-09-14 chi2012.tex
prestudy introductory ditty in place

@@
 %% not sure where this goes >> but i want to keep it
-
 Setting out to investigate the potential role for better support for data integration in PIM, we first sought an updated understanding of how and whether people still used multiple information sources in the course of everyday information-related activities.
@@
 \subsection{PS.1 - Understanding Data Diversity in Everyday Tasks}
-\subsubsection{Method}
+% \subsubsection{Method}
 Our pre-study investigated the types of tasks people perform online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.

### 2012-09-14 chi2012.tex
Added intro section to prestudies

--- 2012-09-14 chi2012.tex
@@
 %% don't like the beginning of this > 
-Enabling PIM tools to supporting ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data hetereogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
+Enabling PIM tools to supporting ad-hoc data integration has remained
+difficult because of the complex, theoretically difficult nature of data hetereogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
 As described by Alon Halvey:
@@
 
-% i think we do not need to differentiate between semantic web and databse integration here, we're saying they're both bad, let's lump them together to keep the argument flowing and not kill our readers
-The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} research communities have primarily focused efforts on purely-automatic approaches to the problem of data integration.
-Such approaches typically take the form of either \emph{ontology} or \emph{schema-level matching}, or \emph{direct} or \emph{instance-level} matching.
-We presently focus  on \emph{instance-level} matching, as schemas may not always be available, nor easily accesssed by users.
-Approaches to automatic instance matching have included applied natural language processing (NLP) techniques for terminological matching (e.g., \cite{euzenat2004api}), machine-learning techniques that use both structural and terminological features for matching \cite{doan2003learning}, probabilistic approaches \cite{suchanek2011paris}, and instance schema comparisons using structural and logical definitions \cite{castano2006matching}.
-In real-world settings, however, manual and semi-automatic matching approaches are favoured over automatic because of the need for high degrees of accuracy.
+Yet, research prototypes of PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.
+For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
+Similarly, Haystack, a personal information manager, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
+Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
 
-A different approach entirely has been taken in the web mash-up community; this community has largely focused on end-user interfaces that allow people to interactively reconcile data from multiple sources.
+In this paper, we take the position that there needs to be a difference in focus between the data integration approaches discussed in the databases literature and that of PIM, because of corresponding differences in requirements.
+In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one.
+This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted.
+The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined.
+In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would be
+fully consume the time and effort a person has allocated for their task.
+Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
+
+Systems that have focused solely on the end-user-driven, interactive reconciliation of data heterogeneity have come out of the web ``data mash-up'' community.
 Yahoo Pipes!
-is a visual programming language aimed at letting novice programmers easily create data-transformation workflows, but remained rather challenging to  use and effort-intensive for one-off information tasks.
-Mash-up makers such as Intel Mashmaker \cite{}, QDEWiki \cite{}, and TX2 \cite{}.
+is a visual programming language aimed at letting novice programmers easily create data-transformation workflows, but remained rather difficult to use and poorly suited for one-off information tasks.
+Mash-up makers such as Intel Mashmaker \cite{}, QDEWiki, and TX2
+
 DataPalette's approach is most directly inspired by ``lightweight'' visual, user-interactive data collection, consolidation and alignment tools such as Potluck \cite{citeulike:3875264} and Cards \cite{Dontcheva:2007:RCS:1294211.1294224}.
 
-% Semi-automatic approaches use varying levels of support to aid a user, from prompting users to manually check flagged alignments, to suggesting possible alignments.
-Allowing users to compare instances across heterogeneous data is powerful because users can combine and evaluate multiple sources.
-In principle, automatic instance matching approaches would allow users to seamlessly browse this data, however for now and the foreseeable future such approaches cannot make use of personal knowledge and do not guarantee their alignments.
-Therefore, in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.
-
-A number of research PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.
-For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
-Similarly, Haystack, a personal information platform, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
-Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
-
 %the following needs integrating with the rest of the lit rev.
-
-%% The Semantic Web research community has approached the problem of data integration in terms of \emph{ontology matching}, which can be split into two distinct strategies: \emph{schema} and \emph{instance} matching.
-Instance matching differs from schema level matching in that it aims to detect instances referring to the same real world object, while schema matching aims to find a set of mappings between concepts and properties in different ontologies.
-For our purposes of allowing users to explore and compare data from different websites, we are particularly interested in aligning instances of the most common forms of data that people compare from website to website.
-
-%% Instance matching within this scope is in its infancy, some of its problems have been addressed by the record linkage problem in the database field \cite{fellegi1969theory,winkler1999state,gu2003record}.
-However, there are new problems to address in this field, which are strongly related to schema matching, including structural heterogeneity and logical heterogeneity.
-The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet\footnote{\emph{Wordnet} - \url{http://wordnet.princeton.edu}} to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instance schemas structural and logical definitions \cite{castano2006matching}.
-In our case, this is problematic because information from the web does not always have a schema.
-In real-world scenarios, semi-automatic ontology matching approaches are favoured over automatic because reconciling the differences between ontologies that were designed for different purposes is not always accurate.
-Semi-automatic approaches use varying levels of support to aid a user, from prompting users to manually check flagged alignments, to suggesting possible alignments.
-Allowing users to compare instances across heterogeneous data is powerful because users can combine and evaluate multiple sources.
-In principle, automatic instance matching approaches would allow users to seamlessly browse this data, however for now and the foreseeable future such approaches cannot make use of personal knowledge and do not guarantee their alignments.
-Therefore, in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.
-
-%% The Semantic Web community use ontologies to formally describe vocabularies for specific domains by defining axioms describing entities, instances of entities and the relationships that hold among them \cite{borst1997construction}.
+The Semantic Web community use ontologies to formally describe vocabularies for specific domains by defining axioms describing entities, instances of entities and the relationships that hold among them \cite{borst1997construction}.
 The axioms in an ontology can be either assertional or terminological, the former describes the schema of the entities and the latter the instances of those entities.
@@
 
-% For our purposes of allowing users to explore and compare data from different websites, we are particularly interested in aligning instances of the most common forms of data that people compare from website to website.
+For our purposes of allowing users to explore and compare data from different websites, we are particularly interested in aligning instances of the most common forms of data that people compare from website to website.
 Instance matching within this scope is in its infancy, some of its problems have been addressed by the record linkage problem in the database field \cite{fellegi1969theory,winkler1999state,gu2003record}.
@@
 
-% The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instance schemas structural and logical definitions \cite{castano2006matching}.
+The approaches used to match instances typically use: natural language processing, using lexicons like Wordnet to identify synonyms, and word stems \cite{euzenat2004api}; machine learning techniques \cite{doan2003learning}; probabilistic approaches \cite{suchanek2011paris} and comparisons of instance schemas structural and logical definitions \cite{castano2006matching}.
 In our case, this is problematic because information from the web does not always have a schema.
@@
 
-In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
-In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one.
-This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted.
-The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined.
-In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.
-Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
+In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.
+For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a task requiring multiple websites (see Section \ref{sec:prestudy1}).
+We analysed the data collected from this study using a thematic approach.
+From our first pre-study's qualitative analysis, we then designed our second pre-study which investigates the data and data formats available for people to use from popular websites (see Section \ref{sec:prestudy2}).
+We performed a statistical analysis of the features of the available data.
+In the next two sections, we describe our two pre-studies.
 
-% little introductory ditty about why we're doing pre-studies 
-% actually we may wish to combine discussion of both (!)
-
-\subsection{PS.1 - Understanding Data Diversity in Everyday Tasks}
+\subsection{1 - Understanding Heterogeneity in Everyday Tasks}\label{sec:prestudy1}
 \subsubsection{Method}
@@
 
-\subsection{PS.2 - Technical Challenges of Integrating the Data}
+\subsection{2 - Technical Challenges of Integrating the Data}\label{sec:prestudy2}
 

### 2012-09-14 chi2012.tex
prestudy

--- 2012-09-14 chi2012.tex
@@
 \subsection{PS.1 - Understanding Data Diversity in Everyday Tasks}
-\subsubsection{Method}
 Our pre-study investigated the types of tasks people perform online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.
@@
 \subsection{PS.2 - Technical Challenges of Integrating the Data}
+Our investigation of integration challenges started with identifying a set of data sources to analyze.
+For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} and looked at the 50 most popular data feeds, based upon number of developer API keys issued for each.
+In order to get a variety of different kinds of sources represented, we feeds pertaining to 5 topics: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.
+Table The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.
 
-The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.
+We first sought out the APIs and data feeds of popular Web services and sites that were aimed at developers creating apps for end-users.
+Based on statistics collected We first selected 4 different categories
 
-\subsubsection{Method}
-We sought out APIs and data feeds of popular Web services and sites that were aimed at developers targeting consumption of the data by end-users.
-Based on statistics collected We first selected 4 different categories
+\begin{table*}
+\small
+\begin{tabular}{| p{2.0cm} | p{1.5cm} | p{3.6cm} | p{2cm} | p{2cm} | l | l | l |}
+\hline
+source category  & data type & sources & API type & avg fields/rec & avg depth & term. het & struc. het \\
+\hline
+social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
+event calendars & events & eventbrit, gcal  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
+retail (shopping)  & product profile & Amazon, shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
+music catalogues & songs and albums & echonest, rdio, soundcloud, spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
+weather & forecasts & wunderground, weatherbug, yahoo weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
+\hline
+\end{tabular}
+\caption{\emph{PS.2 Results} - Analysis of heterogeneity present across data feeds, by category of feed} \label{tbl_tps}
+\end{table*}
 

### 2012-09-14 chi2012.tex
prestudy

@@
 
-While a primary goal of opening up such direct access to information is the hope that end-user citizens will make more informed decisions pertaining to their health, wealth, or well-being \cite{}, use of this data beyond specialised app developers, journalists and other “data specialists” remains minimal.
+While a primary goal of opening up such direct access to information is the hope that end-user citizens will make more informed decisions pertaining to their health, wealth, or well-being \cite{}, use of this data beyond specialised app developers, journalists and other Òdata specialistsÓ remains minimal.
 Our hypothesis is that the main impediment towards end-user access to this data is a mismatch between the capabilities of tools currently used to satisfy daily information needs, and those required to effectively browse, use and consume heterogeneous structured data from diverse sources.
@@
 We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions (see Appendix ?).
-This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant’s experiences, while collecting the same general areas of information.
+This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participantÕs experiences, while collecting the same general areas of information.
 Each interview was recorded and notes taken, the interviewer was trained so that they were knowledgeable about the importance of the study and to minimise bias.
@@
 They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.
-They selected venues, locations, restaurants and activities based on their own knowledge, friend’s recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.
+They selected venues, locations, restaurants and activities based on their own knowledge, friendÕs recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.
 People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.
@@
 When organising an event the weather was not that important because it they did not trust its accuracy for events plan further in the future than one week.
-Also, their friend’s preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.
+Also, their friendÕs preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.
 Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
@@
 
+%% TODO THIS IS PLACEHOLDER DATA >> 
 \begin{table*}
@@
 \end{table*}
+%% <<
 
@@
 \subsection{Participants}
-We recruited 20 participants (10 female) from the University of Southampton’s international student mailing list.
+We recruited 20 participants (10 female) from the University of SouthamptonÕs international student mailing list.
 All of the participants spoke English, and for 12 of them, English was not their first language, however they had all passed the university English proficiency test.
@@
 \subsection{Data Collection}
-Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study’s protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
-During the lab study, we recorded the audio and the participant’s actions on screen.
+Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the studyÕs protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
+During the lab study, we recorded the audio and the participantÕs actions on screen.
 We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.
@@
 The studies lasted 40 minutes, on average.
-Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn’t influence the study.
+Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldnÕt influence the study.
 The studies were run over 4 days.
-At the end of each day the facilitators and observers met to discuss the studies’ data and to revisit protocols, if necessary.
+At the end of each day the facilitators and observers met to discuss the studiesÕ data and to revisit protocols, if necessary.
 
@@
 
-During the data collection we collated a spreadsheet summarising the participant’s gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
-After we had collected the study’s data, we performed our data analysis.
+During the data collection we collated a spreadsheet summarising the participantÕs gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
+After we had collected the studyÕs data, we performed our data analysis.
 
@@
 %
-%What we didn’t find (evidence for)
+%What we didnÕt find (evidence for)
 %
 %What we might look at next -- follow-on work
-%		We focused on terminological heterogeneity -- didn’t address structural heterogeneity, semantic heterogeneity
+%		We focused on terminological heterogeneity -- didnÕt address structural heterogeneity, semantic heterogeneity
 %		We focused on finite (small) sets Streaming data (how do select future data --) 

### 2012-09-14 chi2012.tex
prestudy
Fix apostrophes

@@
 We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions (see Appendix ?).
-This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participantÕs experiences, while collecting the same general areas of information.
+This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant's experiences, while collecting the same general areas of information.
 Each interview was recorded and notes taken, the interviewer was trained so that they were knowledgeable about the importance of the study and to minimise bias.
@@
 They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.
-They selected venues, locations, restaurants and activities based on their own knowledge, friendÕs recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.
+They selected venues, locations, restaurants and activities based on their own knowledge, friend's recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.
 People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.
@@
 When organising an event the weather was not that important because it they did not trust its accuracy for events plan further in the future than one week.
-Also, their friendÕs preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.
+Also, their friend's preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.
 Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
@@
 \subsection{Participants}
-We recruited 20 participants (10 female) from the University of SouthamptonÕs international student mailing list.
+We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.
 All of the participants spoke English, and for 12 of them, English was not their first language, however they had all passed the university English proficiency test.
@@
 \subsection{Data Collection}
-Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the studyÕs protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
-During the lab study, we recorded the audio and the participantÕs actions on screen.
+Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
+During the lab study, we recorded the audio and the participant's actions on screen.
 We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.
@@
 The studies lasted 40 minutes, on average.
-Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldnÕt influence the study.
+Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn't influence the study.
 The studies were run over 4 days.
-At the end of each day the facilitators and observers met to discuss the studiesÕ data and to revisit protocols, if necessary.
+At the end of each day the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.
 
@@
 
-During the data collection we collated a spreadsheet summarising the participantÕs gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
-After we had collected the studyÕs data, we performed our data analysis.
+During the data collection we collated a spreadsheet summarising the participant's gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
+After we had collected the study's data, we performed our data analysis.
 
@@
 %
-%What we didnÕt find (evidence for)
+%What we didn't find (evidence for)
 %
 %What we might look at next -- follow-on work
-%		We focused on terminological heterogeneity -- didnÕt address structural heterogeneity, semantic heterogeneity
+%		We focused on terminological heterogeneity -- didn't address structural heterogeneity, semantic heterogeneity
 %		We focused on finite (small) sets Streaming data (how do select future data --) 

### 2012-09-14 chi2012.tex
fix double quotes

@@
 % weird and wonderful
-A number of users admittedly introduced their own bias into the choices: "I prefer Indian food, so I will choose the Indian
-restaurant", and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating
+A number of users admittedly introduced their own bias into the choices: ``I prefer Indian food, so I will choose the Indian
+restaurant'', and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating
 those that required bus transfers (the protocol did not mention routing/transfers at all).
@@
 when using the WebBox interface, significantly more comparison (and visualisation of values) was performed, with fewer
-participants using the "ruling out" process. A consequence of this was that when asked to justify their choices, users were
+participants using the ``ruling out'' process. A consequence of this was that when asked to justify their choices, users were
 able to verify their values on-screen immediately, whereas those who had ruled-out had to memorise their reasons, with

### 2012-09-14 chi2012.tex

--- 2012-09-14 chi2012.tex
@@
 
-While a primary goal of opening up such direct access to information is the hope that end-user citizens will make more informed decisions pertaining to their health, wealth, or well-being \cite{}, use of this data beyond specialised app developers, journalists and other Òdata specialistsÓ remains minimal.
+While a primary goal of opening up such direct access to information is the hope that end-user citizens will make more informed decisions pertaining to their health, wealth, or well-being \cite{}, use of this data beyond specialised app developers, journalists and other “data specialists” remains minimal.
 Our hypothesis is that the main impediment towards end-user access to this data is a mismatch between the capabilities of tools currently used to satisfy daily information needs, and those required to effectively browse, use and consume heterogeneous structured data from diverse sources.
@@
 
+In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.
+For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a task requiring multiple websites (see Section \ref{sec:prestudy1}).
+We analysed the data collected from this study using a thematic approach.
+From our first pre-study's qualitative analysis, we then designed our second pre-study which investigates the data and data formats available for people to use from popular websites (see Section \ref{sec:prestudy2}).
+We performed a statistical analysis of the features of the available data.
+In the next two sections, we describe our two pre-studies.
 %% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
@@
 \subsection{PS.1 - Understanding Data Diversity in Everyday Tasks}
+% \subsubsection{Method}
 Our pre-study investigated the types of tasks people perform online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.
 We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions (see Appendix ?).
-This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant's experiences, while collecting the same general areas of information.
+This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant’s experiences, while collecting the same general areas of information.
 Each interview was recorded and notes taken, the interviewer was trained so that they were knowledgeable about the importance of the study and to minimise bias.
@@
 They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.
-They selected venues, locations, restaurants and activities based on their own knowledge, friend's recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.
+They selected venues, locations, restaurants and activities based on their own knowledge, friend’s recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.
 People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.
@@
 When organising an event the weather was not that important because it they did not trust its accuracy for events plan further in the future than one week.
-Also, their friend's preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.
+Also, their friend’s preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.
 Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
@@
 \subsection{PS.2 - Technical Challenges of Integrating the Data}
-Our investigation of integration challenges started with identifying a set of data sources to analyze.
-For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} and looked at the 50 most popular data feeds, based upon number of developer API keys issued for each.
-In order to get a variety of different kinds of sources represented, we feeds pertaining to 5 topics: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.
-Table The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.
 
-We first sought out the APIs and data feeds of popular Web services and sites that were aimed at developers creating apps for end-users.
+The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.
+
+\subsubsection{Method}
+We sought out APIs and data feeds of popular Web services and sites that were aimed at developers targeting consumption of the data by end-users.
 Based on statistics collected We first selected 4 different categories
-
-%% TODO THIS IS PLACEHOLDER DATA >> 
-\begin{table*}
-\small
-\begin{tabular}{| p{2.0cm} | p{1.5cm} | p{3.6cm} | p{2cm} | p{2cm} | l | l | l |}
-\hline
-source category  & data type & sources & API type & avg fields/rec & avg depth & term. het & struc. het \\
-\hline
-social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-event calendars & events & eventbrit, gcal  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-retail (shopping)  & product profile & Amazon, shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-music catalogues & songs and albums & echonest, rdio, soundcloud, spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-weather & forecasts & wunderground, weatherbug, yahoo weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-\hline
-\end{tabular}
-\caption{\emph{PS.2 Results} - Analysis of heterogeneity present across data feeds, by category of feed} \label{tbl_tps}
-\end{table*}
-%% <<
 
@@
 \subsection{Participants}
-We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.
+We recruited 20 participants (10 female) from the University of Southampton’s international student mailing list.
 All of the participants spoke English, and for 12 of them, English was not their first language, however they had all passed the university English proficiency test.
@@
 \subsection{Data Collection}
-Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
-During the lab study, we recorded the audio and the participant's actions on screen.
+Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study’s protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
+During the lab study, we recorded the audio and the participant’s actions on screen.
 We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.
@@
 The studies lasted 40 minutes, on average.
-Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn't influence the study.
+Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn’t influence the study.
 The studies were run over 4 days.
-At the end of each day the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.
+At the end of each day the facilitators and observers met to discuss the studies’ data and to revisit protocols, if necessary.
 
@@
 
-During the data collection we collated a spreadsheet summarising the participant's gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
-After we had collected the study's data, we performed our data analysis.
+During the data collection we collated a spreadsheet summarising the participant’s gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
+After we had collected the study’s data, we performed our data analysis.
 
@@
 %
-%What we didn't find (evidence for)
+%What we didn’t find (evidence for)
 %
 %What we might look at next -- follow-on work
-%		We focused on terminological heterogeneity -- didn't address structural heterogeneity, semantic heterogeneity
+%		We focused on terminological heterogeneity -- didn’t address structural heterogeneity, semantic heterogeneity
 %		We focused on finite (small) sets Streaming data (how do select future data --) 

### 2012-09-14 chi2012.tex
incremental updates
consolidation action

@@
 \begin{abstract}
-
+Exploratory sequential mixed method study
 \end{abstract}
@@
 If PIM tools were extended to enable end-users to make effective use of the emerging ecosystem of personal data, how would this impact personal information practice?
-This paper presents our initial results towards answering this question through a multi-method investigation of the personal data integration problem.
+This paper presents our initial results towards answering this question through a sequential exploratory mixed-method investigation of the personal data integration problem.
 We start with a small interview study examining information tasks people perform that rely on multiple sources of information, and how such needs are currently satisfied.
@@
 
-\section{Pre-studies: The Need and Challenge of Data Integration in PIM}
+\section{Pre-studies: The Need for and Challenges in realizing Ad-Hoc Data Integration in PIM Tools}
 
@@
 
-The second study we performed an analysis on existing on data available from APIs and structured data feeds on the Web.
-The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.
-We describe our methodology below.
-
-In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.
+% blargh. this needs to be integrated above! 
+In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed-method investigation to help us design our contribution.
 For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a task requiring multiple websites (see Section \ref{sec:prestudy1}).
@@
 In the next two sections, we describe our two pre-studies.
+
+The second study we performed an analysis on existing on data available from APIs and structured data feeds on the Web.
+The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.
+We describe our methodology below.
+
 %% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
@@
 The method chosen depended on the time scale required to organise an event, in general if there was a short time scale people would speak in person or on the phone.
-Otherwise, people would set up a FaceBook event page to discuss ideas with their friends.
+Otherwise, people would set up a Facebook event page to discuss ideas with their friends.
 If their friends were not on Facebook they would email them.
-All but one participant said they would not use Doddle to organise day and time of an event, because they felt that people did not fill in the form and it was more suit to organising work event not social.
+All but one participant said they would not use Doodle\footnote{Doodle - \emph{www.doodle.com}} to organise day and time of an event, because they felt that people did not fill in the form and it was more suit to organising work event not social.
 
@@
 They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.
-They selected venues, locations, restaurants and activities based on their own knowledge, friend's recommendations, Google maps, reviews through Googling and review sites such as tripadvisor.
+They selected venues, locations, restaurants and activities based on their own knowledge, friend's recommendations, Google maps, reviews obtained through web search and sites such as tripadvisor.
 People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.
@@
 While Facebook was the most popular social networking site for organising events, half of the participants felt that it was not the best solution and would prefer a collaborative environment, which could be wiki based.
-
-\subsection{PS.2 - Technical Challenges of Integrating the Data}
-Our investigation of integration challenges started with identifying a set of data sources to analyze.
-For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} and looked at the 50 most popular data feeds, based upon number of developer API keys issued for each.
-In order to get a variety of different kinds of sources represented, we feeds pertaining to 5 topics: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.
-Table The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.
-
-We first sought out the APIs and data feeds of popular Web services and sites that were aimed at developers creating apps for end-users.
-Based on statistics collected We first selected 4 different categories
 
@@
 %% <<
+
+\subsection{PS.2 - Technical Challenges of Integrating the Data}
+Our investigation of integration challenges started with identifying a set of data sources to analyze.
+For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} and looked at the 50 most popular data feeds, based upon number of developer API keys issued for each.
+In order to get a variety of different kinds of sources represented, based upon available choices, we first chose 5 feed \emph{categories}: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.
+Then, we populated each category with the most popular feeds from the list, and took samples of data records from each.
 

### 2012-09-14 chi2012.tex
Adding images into document

--- 2012-09-14 chi2012.tex
@@
 
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=18cm]{img/dragdrop}
+\caption{WebBox uses a Drag and Drop interface to add data into the workspace.}
+\label{fig:dragdrop}
+\end{center}
+\end{figure*}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=8cm]{img/multipathing}
+\caption{To see properties, users click on the name of the property. When instances do not have a particular property, additional property names can be added.}
+\label{fig:multipathing}
+\end{center}
+\end{figure}
+
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=18cm]{img/sameas}
+\caption{When two instances represent the same entity, users can drag one on top of the other, and they are combined.}
+\label{fig:sameas}
+\end{center}
+\end{figure*}
+
 \section{Methodology}

### 2012-09-14 chi2012.tex
table

--- 2012-09-14 chi2012.tex
@@
 social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-event calendars & events & eventbrit, gcal  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-retail (shopping)  & product profile & Amazon, shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-music catalogues & songs and albums & echonest, rdio, soundcloud, spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-weather & forecasts & wunderground, weatherbug, yahoo weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
+event calendars & events & Eventbrite, Google Calendar  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
+retail (shopping)  & product profile & Amazon, Shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
+music catalogues & songs and albums & Echonest, rdio, Soundcloud, Spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
+weather & forecasts & W.Underground, Weatherbug, Yahoo! Weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
 \hline
 \end{tabular}
+\label{tbl:prestudy2}
 \caption{\emph{PS.2 Results} - Analysis of heterogeneity present across data feeds, by category of feed} \label{tbl_tps}
@@
 
-\subsection{PS.2 - Technical Challenges of Integrating the Data}
-Our investigation of integration challenges started with identifying a set of data sources to analyze.
-For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} and looked at the 50 most popular data feeds, based upon number of developer API keys issued for each.
-In order to get a variety of different kinds of sources represented, based upon available choices, we first chose 5 feed \emph{categories}: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.
-Then, we populated each category with the most popular feeds from the list, and took samples of data records from each.
+\subsection{PS.2 - Technical Challenges Data Integration}
+Our investigation of integration challenges started with identifying a set of candidate data sources for analysis.
+For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} of 50 most popular data feeds measured by the number of developer API keys issued for each.
+In order to get a variety of different kinds of sources represented, we first categorised feeds into 5 topic categories: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.
+We then populated each category with 2-5 feeds each depending on feed availability, for a total of 15 total feeds.
+
+For each feed, 3-5 data records were obtained by querying the API directly for records of a that category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}; that is, while many of the APIs also provided access to other kinds of data; for example, \url{shopping.com} offers programmatic access to ``offers'' in addition to particular product descriptions, Amazon offers access to the shopper's profile, etc, but only product records were obtained for
 

### 2012-09-14 chi2012.tex
added more in results section

--- 2012-09-14 chi2012.tex
@@
 
-The second study we performed an analysis on existing on data available from APIs and structured data feeds on the Web.
+In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.
+For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a tasks requiring multiple websites (see Section \ref{sec:prestudy1}).
+We analysed the data collected from this study using a thematic approach.
+From our first pre-study's qualitative analysis, we designed and performed our second study which analyses existing on data available from APIs and structured data feeds on the Web (see Section \ref{sec:prestudy2}).
 The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.
-We describe our methodology below.
-
-In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.
-For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a task requiring multiple websites (see Section \ref{sec:prestudy1}).
-We analysed the data collected from this study using a thematic approach.
-From our first pre-study's qualitative analysis, we then designed our second pre-study which investigates the data and data formats available for people to use from popular websites (see Section \ref{sec:prestudy2}).
 We performed a statistical analysis of the features of the available data.
@@
 
-\begin{figure*}[htbp]
-\begin{center}
-\includegraphics[width=18cm]{img/dragdrop}
-\caption{WebBox uses a Drag and Drop interface to add data into the workspace.}
-\label{fig:dragdrop}
-\end{center}
-\end{figure*}
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=8cm]{img/multipathing}
-\caption{To see properties, users click on the name of the property. When instances do not have a particular property, additional property names can be added.}
-\label{fig:multipathing}
-\end{center}
-\end{figure}
-
-\begin{figure*}[htbp]
-\begin{center}
-\includegraphics[width=18cm]{img/sameas}
-\caption{When two instances represent the same entity, users can drag one on top of the other, and they are combined.}
-\label{fig:sameas}
-\end{center}
-\end{figure*}
-
 \section{Methodology}
@@
 \subsection{Participants}
-We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.
-All of the participants spoke English, and for 12 of them, English was not their first language, however they had all passed the university English proficiency test.
-Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.
-All of the participants had previously encountered tasks requiring them to use and evaluate data from multiple websites.
+We recruited participants by calling for participant on the University of Southampton's international student mailing list.
+There is no guarantee of the mailing list's recipients first language, however all of them passed the university English proficiency test.
+We required that all participants had encountered tasks that required them to use and evaluate data from multiple websites.
 
@@
 \begin{center}
-
+\small
 \begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
@@
 \label{tab:studyfactors}
+\normalsize
 \end{table*}%
@@
 Each task had an A and B condition.
-The A condition, the participant could only use WebBox to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or spreadsheets in Microsoft Excel containing the data taken from the data sources to make their choice.
+The A condition, the participant could only use WebBox to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or workbook in Microsoft Excel containing the data taken from the data sources to make their choice on individual spreadsheets.
 
@@
 We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.
+At the end of the study the participant completed an exit survey.
 
@@
 
-% behaviours in websites
+We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.
+All of the participants spoke English, and for 12 of them, English was not their first language.
+Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.
 
-Users who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues. In
-particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of
-university teaching by the Times Higher Education. There were numerous usability issues in using websites, particularly
-in searching for university courses and course requirements. Similarly, participants found it difficult to understand the
-locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto
-the same map using the websites. Overall, the users that spent the most time using websites exhibited significant
-frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations),
-and all ended up falling back to using the supplied spreadsheets.
+The information in Table \ref{} was captured during the studies, so that we could statistically analyses the features used by the participants.
 
-% behaviours in excel
-One of the designs of the study was to produce a spreadsheet worksheet for each website, which meant that users
-had to switch between the worksheets to gather data from multiple sources. Different approaches to using the spreadsheets
-were observed. Some users were keen to collate all data onto a single sheet, so that they could compare the values
-all at once. However, this approach was error prone, for example participants noticing too late that the ordering of rows
-was different on each sheet, so that their pasted values were incorrectly mapped. Others used pen and paper to collate
-the values, which was often faster and resulted in less observed frustration. While the majority of users did not
+\begin{table*}[htdp]
+\caption{default}
+\small
+\begin{center}
+\begin{tabular}{|r|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
+\hline
+Participants & 1 &2&3&4&5&6&7&8&9&10&11&12&13&14&15&16&17&18&19&20\\
+\hline\hline
+gender&f&m&f&f&f&f&f&m&f&f&m&f&m&f&m&m&m&m&m&m\\
+A condition first task &y&y&y&n&n&y&n&n&y&y&n&n&y&y&n&n&n&n&n&n\\
+\hline\hline
+A Condition: task &4&2&3&4&3&&2&1&4&1&4&1&3&2&3&2&1&4&3&2\\
+paper used&n&n&n&n&n&n&y&n&y&y&n&y&n&n&y&n&n&n&n&n\\
+charts used &1&2&2&0&2&4&5&0&0&0&2&3&2&0&0&0&3&3&4&1\\
+maps used &1&1&2&2&2&1&1&0&1&0&1&1&2&1&0&1&0&0&2&1\\ 
+same as &4&3&3&6&0&2&0&2&1&0&6&5&6&2&7&0&6&7&7&7\\
+number data sources  &4&3&4&4&4&3&4&3&4&4&3&3&5&4&5&4&4&5&5&4\\
+chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, 1.3, 1.2& 4.2& 1.4, 1.5, 1.3& 3.2& 2.1, 2.3, 2.6& 3.5, 3.3&2.6, 2.3, 2.5&1.6, 1.5, 1.3&4.1&3.3&2.4, 2.5, 2.6\\
+\hline\hline
+ B Condition:task &3&&4&3&4&2&1&2&2&3&2&3&1&4&1&4&2&2&1&3\\
+ paper used &n&n&n&n&n&n&y&n&y&y&n&y&n&y&y&n&n&y&n&n\\
+ spreadsheet/website&&&&&&&&&&&&&&&&&&&&\\
+ number of spreadsheets used&5/5&4/4&2/5&1.5&2/5&5/5&4/4&2/4&3/4&&2/4&4/5&2/4&4/5&2/4&2/5&4/4&4/4&0&5/5\\
+ number of websites used&1&1&4&3&3&0&2&3&2&&1&2&3&1&2&3&2&0&4&0\\
+ chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4,1.2,1.6&2.1&2.3,2.6,2.5&3.5&2.2,2.3,2.6&3.5&1.1&4.1&1.6&4.2&2.6,2.3, 2.5&2.3,2.5,2.6&1.1, 1.3, 1.4&3.5\\
+ \hline
+\end{tabular}
+\end{center}
+\label{default}
+\normalsize
+\end{table*}
+
+In the A condition, where participants used WebBox to complete their tasks they used it to collate the data on-screen, and compare multiple values directly.
+About half of the observed users actually used the instance combination feature of WebBox, with the other half able to successfully make decisions without using it.
+Those who tried to use it were largely successful, and were able to therefore use the global brush-highlighting to simultaneously compare different statistics.
+Most users instantly opened up all data sources into individual boxes, in order to determine which sources contained specific statistics.
+The result of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.
+
+In the B conditions, users were given the option of using websites and or a spreadsheet to make their choice.
+Participants who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues.
+In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education.
+There were numerous usability issues in using websites, particularly in searching for university courses and course requirements.
+Similarly, participants found it difficult to understand the locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto the same map using the websites.
+Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and all ended up falling back to using the supplied spreadsheets.
+
+Participants who opted to use the spreadsheet to support their decisions switched between worksheets to view data from different websites.
+They used different approaches to using the spreadsheets, some participants were keen to collate all data onto a single sheet, so that they could compare the values all at once.
+However, this approach was error prone, for example participants noticing too late that the ordering of rows was different on each sheet, so that their pasted values were incorrectly mapped.
+Others used pen and paper to collate the values, which was often faster and resulted in less observed frustration.
+While the majority of users did not
 attempt anything complex with Excel (in fact only one users used charts), there was some advanced use of Excel observed.
@@
 
-% webbox client
+Notably, a number of users admittedly introduced their own bias into the choices in both A and B conditions: ``I prefer Indian food, so I will choose the Indian restaurant'', and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating those that required bus transfers (the protocol did not mention routing/transfers at all).
 
-The most commonly observed use of WebBox was to collate the data on-screen, and compare multiple values directly.
-About half of the observed users actually used the instance combination feature of WebBox, with the other half able
-to successfully make decisions without using it. Those who tried to use it were largely successful, and were able
-to therefore use the global brush-highlighting to simultaneously compare different statistics. Most users instantly opened
-up all data sources into individual boxes, in order to determine which sources contained specific statistics. The result
-of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.
-
-%% more from doc here
-
-% weird and wonderful
-A number of users admittedly introduced their own bias into the choices: ``I prefer Indian food, so I will choose the Indian
-restaurant'', and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating
-those that required bus transfers (the protocol did not mention routing/transfers at all).
+In the exit survey, the users noted that they would use WebBox for finding a property to rent/buy, choosing a job and purchasing electronic devices, in the future.
+However, one participant said that they would not use it for things like deciding on a restaurant because ``it would cheapen the experience'' by removing the spontaneity, but would use it for making more important decisions.
+In order to improve our solutions the participants suggested making the origin of the information clearer, sorting instances, and explanations of ratings and scoring systems (however some of these are not transparent on websites).
 

### 2012-09-14 chi2012.tex

@@
 If so, this would suggest that perhaps specialised interfaces could be concocted for this purpose rather than it being a PIM need.
+
+% blargh. this needs to be integrated above! 
+In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed-method investigation to help us design our contribution.
+For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a task requiring multiple websites (see Section \ref{sec:prestudy1}).
+We analysed the data collected from this study using a thematic approach.
+From our first pre-study's qualitative analysis, we then designed our second pre-study which investigates the data and data formats available for people to use from popular websites (see Section \ref{sec:prestudy2}).
+We performed a statistical analysis of the features of the available data.
+In the next two sections, we describe our two pre-studies.
+
+The second study we performed an analysis on existing on data available from APIs and structured data feeds on the Web.
+The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.
+We describe our methodology below.
 
@@
 
-\subsection{PS.2 - Technical Challenges of Integrating the Data}
-Our investigation of integration challenges started with identifying a set of data sources to analyze.
-For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} and looked at the 50 most popular data feeds, based upon number of developer API keys issued for each.
-In order to get a variety of different kinds of sources represented, we feeds pertaining to 5 topics: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.
-Table The second took a look at the structured data available via popular APIs, data feeds and downloadable data sets in order to characterise and understand challenges of combining data from these sources.
-
-We first sought out the APIs and data feeds of popular Web services and sites that were aimed at developers creating apps for end-users.
-Based on statistics collected We first selected 4 different categories
-
 %% TODO THIS IS PLACEHOLDER DATA >> 
@@
 %% <<
+
+\subsection{PS.2 - Technical Challenges Data Integration}
+Our investigation of integration challenges started with identifying a set of candidate data sources for analysis.
+For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} of 50 most popular data feeds measured by the number of developer API keys issued for each.
+In order to get a variety of different kinds of sources represented, we first categorised feeds into 5 topic categories: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.
+We then populated each category with 2-5 feeds each depending on feed availability, for a total of 15 total feeds.
+
+For each feed, 3-5 data records were obtained by querying the API directly for records of a that category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}; that is, while many of the APIs also provided access to other kinds of data; for example, \url{shopping.com} offers programmatic access to ``offers'' in addition to particular product descriptions, Amazon offers access to the shopper's profile, etc, but only product records were obtained for
 
@@
 Design process
+
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=18cm]{img/dragdrop}
+\caption{WebBox uses a Drag and Drop interface to add data into the workspace.}
+\label{fig:dragdrop}
+\end{center}
+\end{figure*}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=8cm]{img/multipathing}
+\caption{To see properties, users click on the name of the property. When instances do not have a particular property, additional property names can be added.}
+\label{fig:multipathing}
+\end{center}
+\end{figure}
+
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=18cm]{img/sameas}
+\caption{When two instances represent the same entity, users can drag one on top of the other, and they are combined.}
+\label{fig:sameas}
+\end{center}
+\end{figure*}
 

### 2012-09-14 chi2012.tex
Fix conflict

@@
 % blargh. this needs to be integrated above! 
-In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed-method investigation to help us design our contribution.
-For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a task requiring multiple websites (see Section \ref{sec:prestudy1}).
+In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.
+For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a tasks requiring multiple websites (see Section \ref{sec:prestudy1}).
 We analysed the data collected from this study using a thematic approach.
-From our first pre-study's qualitative analysis, we then designed our second pre-study which investigates the data and data formats available for people to use from popular websites (see Section \ref{sec:prestudy2}).
+From our first pre-study's qualitative analysis, we designed and performed our second study which analyses existing on data available from APIs and structured data feeds on the Web (see Section \ref{sec:prestudy2}).
+The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.
 We performed a statistical analysis of the features of the available data.
@@
 
-In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.
-For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a tasks requiring multiple websites (see Section \ref{sec:prestudy1}).
-We analysed the data collected from this study using a thematic approach.
-From our first pre-study's qualitative analysis, we designed and performed our second study which analyses existing on data available from APIs and structured data feeds on the Web (see Section \ref{sec:prestudy2}).
-The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.
-We performed a statistical analysis of the features of the available data.
-In the next two sections, we describe our two pre-studies.
 %% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.

### 2012-09-14 chi2012.tex
p2 update

--- 2012-09-14 chi2012.tex
@@
 
-For each feed, 3-5 data records were obtained by querying the API directly for records of a that category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}; that is, while many of the APIs also provided access to other kinds of data; for example, \url{shopping.com} offers programmatic access to ``offers'' in addition to particular product descriptions, Amazon offers access to the shopper's profile, etc, but only product records were obtained for
+For each feed, 3-5 data records were obtained by querying the API directly for records of a that category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}.
+For example, while many of the APIs also provided access to other kinds of data, ranging from reviews to customer profilesonly product records were obtained for this category.
+Once data records were obtained, each set of record was analysed individually for the following metrics
+
+\begin{itemize}
+\item \emph{Average number of fields per record} - A simple metric of how many properties were typically associated with each field from each source
+\item \emph{Variance in number of fields per record} - 
+\item \emph{Average Depth of record} - Levels of data which 
+\end{itemize}
+
+Pairwise analysis was done to compute the following heterogeneity metrics
+\begin{itemize}
+\item \emph{Terminological consistency} - 
+\item \emph{Structural consistency} - 
+\end{itemize}
 
 \subsubsection{Results}
-
-\subsubsection{Summary of Findings}
-
 \section{WebBox Interface - Design of the DataPalette}
@@
 \subsection{Participants}
-We recruited participants by calling for participant on the University of Southampton's international student mailing list.
-There is no guarantee of the mailing list's recipients first language, however all of them passed the university English proficiency test.
-We required that all participants had encountered tasks that required them to use and evaluate data from multiple websites.
+We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.
+All of the participants spoke English, and for 12 of them, English was not their first language, however they had all passed the university English proficiency test.
+Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.
+All of the participants had previously encountered tasks requiring them to use and evaluate data from multiple websites.
 
@@
 \begin{center}
-\small
+
 \begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
@@
 \label{tab:studyfactors}
-\normalsize
 \end{table*}%
@@
 Each task had an A and B condition.
-The A condition, the participant could only use WebBox to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or workbook in Microsoft Excel containing the data taken from the data sources to make their choice on individual spreadsheets.
+The A condition, the participant could only use WebBox to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or spreadsheets in Microsoft Excel containing the data taken from the data sources to make their choice.
 
@@
 We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.
-At the end of the study the participant completed an exit survey.
 
@@
 
-We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.
-All of the participants spoke English, and for 12 of them, English was not their first language.
-Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.
+% behaviours in websites
 
-The information in Table \ref{} was captured during the studies, so that we could statistically analyses the features used by the participants.
+Users who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues. In
+particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of
+university teaching by the Times Higher Education. There were numerous usability issues in using websites, particularly
+in searching for university courses and course requirements. Similarly, participants found it difficult to understand the
+locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto
+the same map using the websites. Overall, the users that spent the most time using websites exhibited significant
+frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations),
+and all ended up falling back to using the supplied spreadsheets.
 
-\begin{table*}[htdp]
-\caption{default}
-\small
-\begin{center}
-\begin{tabular}{|r|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
-\hline
-Participants & 1 &2&3&4&5&6&7&8&9&10&11&12&13&14&15&16&17&18&19&20\\
-\hline\hline
-gender&f&m&f&f&f&f&f&m&f&f&m&f&m&f&m&m&m&m&m&m\\
-A condition first task &y&y&y&n&n&y&n&n&y&y&n&n&y&y&n&n&n&n&n&n\\
-\hline\hline
-A Condition: task &4&2&3&4&3&&2&1&4&1&4&1&3&2&3&2&1&4&3&2\\
-paper used&n&n&n&n&n&n&y&n&y&y&n&y&n&n&y&n&n&n&n&n\\
-charts used &1&2&2&0&2&4&5&0&0&0&2&3&2&0&0&0&3&3&4&1\\
-maps used &1&1&2&2&2&1&1&0&1&0&1&1&2&1&0&1&0&0&2&1\\ 
-same as &4&3&3&6&0&2&0&2&1&0&6&5&6&2&7&0&6&7&7&7\\
-number data sources  &4&3&4&4&4&3&4&3&4&4&3&3&5&4&5&4&4&5&5&4\\
-chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, 1.3, 1.2& 4.2& 1.4, 1.5, 1.3& 3.2& 2.1, 2.3, 2.6& 3.5, 3.3&2.6, 2.3, 2.5&1.6, 1.5, 1.3&4.1&3.3&2.4, 2.5, 2.6\\
-\hline\hline
- B Condition:task &3&&4&3&4&2&1&2&2&3&2&3&1&4&1&4&2&2&1&3\\
- paper used &n&n&n&n&n&n&y&n&y&y&n&y&n&y&y&n&n&y&n&n\\
- spreadsheet/website&&&&&&&&&&&&&&&&&&&&\\
- number of spreadsheets used&5/5&4/4&2/5&1.5&2/5&5/5&4/4&2/4&3/4&&2/4&4/5&2/4&4/5&2/4&2/5&4/4&4/4&0&5/5\\
- number of websites used&1&1&4&3&3&0&2&3&2&&1&2&3&1&2&3&2&0&4&0\\
- chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4,1.2,1.6&2.1&2.3,2.6,2.5&3.5&2.2,2.3,2.6&3.5&1.1&4.1&1.6&4.2&2.6,2.3, 2.5&2.3,2.5,2.6&1.1, 1.3, 1.4&3.5\\
- \hline
-\end{tabular}
-\end{center}
-\label{default}
-\normalsize
-\end{table*}
-
-In the A condition, where participants used WebBox to complete their tasks they used it to collate the data on-screen, and compare multiple values directly.
-About half of the observed users actually used the instance combination feature of WebBox, with the other half able to successfully make decisions without using it.
-Those who tried to use it were largely successful, and were able to therefore use the global brush-highlighting to simultaneously compare different statistics.
-Most users instantly opened up all data sources into individual boxes, in order to determine which sources contained specific statistics.
-The result of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.
-
-In the B conditions, users were given the option of using websites and or a spreadsheet to make their choice.
-Participants who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues.
-In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education.
-There were numerous usability issues in using websites, particularly in searching for university courses and course requirements.
-Similarly, participants found it difficult to understand the locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto the same map using the websites.
-Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and all ended up falling back to using the supplied spreadsheets.
-
-Participants who opted to use the spreadsheet to support their decisions switched between worksheets to view data from different websites.
-They used different approaches to using the spreadsheets, some participants were keen to collate all data onto a single sheet, so that they could compare the values all at once.
-However, this approach was error prone, for example participants noticing too late that the ordering of rows was different on each sheet, so that their pasted values were incorrectly mapped.
-Others used pen and paper to collate the values, which was often faster and resulted in less observed frustration.
-While the majority of users did not
+% behaviours in excel
+One of the designs of the study was to produce a spreadsheet worksheet for each website, which meant that users
+had to switch between the worksheets to gather data from multiple sources. Different approaches to using the spreadsheets
+were observed. Some users were keen to collate all data onto a single sheet, so that they could compare the values
+all at once. However, this approach was error prone, for example participants noticing too late that the ordering of rows
+was different on each sheet, so that their pasted values were incorrectly mapped. Others used pen and paper to collate
+the values, which was often faster and resulted in less observed frustration. While the majority of users did not
 attempt anything complex with Excel (in fact only one users used charts), there was some advanced use of Excel observed.
@@
 
-Notably, a number of users admittedly introduced their own bias into the choices in both A and B conditions: ``I prefer Indian food, so I will choose the Indian restaurant'', and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating those that required bus transfers (the protocol did not mention routing/transfers at all).
+% webbox client
 
-In the exit survey, the users noted that they would use WebBox for finding a property to rent/buy, choosing a job and purchasing electronic devices, in the future.
-However, one participant said that they would not use it for things like deciding on a restaurant because ``it would cheapen the experience'' by removing the spontaneity, but would use it for making more important decisions.
-In order to improve our solutions the participants suggested making the origin of the information clearer, sorting instances, and explanations of ratings and scoring systems (however some of these are not transparent on websites).
+The most commonly observed use of WebBox was to collate the data on-screen, and compare multiple values directly.
+About half of the observed users actually used the instance combination feature of WebBox, with the other half able
+to successfully make decisions without using it. Those who tried to use it were largely successful, and were able
+to therefore use the global brush-highlighting to simultaneously compare different statistics. Most users instantly opened
+up all data sources into individual boxes, in order to determine which sources contained specific statistics. The result
+of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.
+
+%% more from doc here
+
+% weird and wonderful
+A number of users admittedly introduced their own bias into the choices: ``I prefer Indian food, so I will choose the Indian
+restaurant'', and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating
+those that required bus transfers (the protocol did not mention routing/transfers at all).
 

### 2012-09-14 chi2012.tex
Fixed table refs

--- 2012-09-14 chi2012.tex
@@
 \end{tabular}
-\label{tbl:prestudy2}
-\caption{\emph{PS.2 Results} - Analysis of heterogeneity present across data feeds, by category of feed} \label{tbl_tps}
+\caption{\emph{PS.2 Results} - Analysis of heterogeneity present across data feeds, by category of feed}\label{tbl:prestudy2}
+
 \end{table*}
@@
 
-For each feed, 3-5 data records were obtained by querying the API directly for records of a that category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}; that is, while many of the APIs also provided access to other kinds of data; for example, \url{shopping.com} offers programmatic access to ``offers'' in addition to particular product descriptions, Amazon offers access to the shopper's profile, etc, but only product records were obtained for
+For each feed, 3-5 data records were obtained by querying the API directly for records of a that category's primary \emph{corresponding type}, listed in the second column of Table \ref{tbl:prestudy2}; that is, while many of the APIs also provided access to other kinds of data; for example, \url{shopping.com} offers programmatic access to ``offers'' in addition to particular product descriptions, Amazon offers access to the shopper's profile, etc, but only product records were obtained for
 
@@
 \begin{table*}[t]
-\caption{Tasks, choices and data sources}
 \begin{center}
@@
 \end{center}
-\label{tab:studyfactors}
+\caption{Tasks, choices and data sources} \label{tab:studyfactors}
 \normalsize
@@
 
-The information in Table \ref{} was captured during the studies, so that we could statistically analyses the features used by the participants.
+The information in Table \ref{tbl:data} was captured during the studies, so that we could statistically analyses the features used by the participants.
 
 \begin{table*}[htdp]
-\caption{default}
 \small
@@
 \end{center}
-\label{default}
+\caption{Participants data.} \label{tbl:data}
 \normalsize

### 2012-09-14 chi2012.tex
ps2 progress

@@
 \small
-\begin{tabular}{| p{2.0cm} | p{1.5cm} | p{3.6cm} | p{2cm} | p{2cm} | l | l | l |}
+\begin{tabular}{| p{2.0cm} | p{1.5cm} | p{3.6cm} | p{2cm} | p{2cm} | l | l | l | l |}
 \hline
-source category  & data type & sources & API type & avg fields/rec & avg depth & term. het & struc. het \\
+source category  & data type & sources & API type & avg fields/rec & avg depth & term. & struct & semantic \\
 \hline
-social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-event calendars & events & Eventbrite, Google Calendar  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-retail (shopping)  & product profile & Amazon, Shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-music catalogues & songs and albums & Echonest, rdio, Soundcloud, Spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
-weather & forecasts & W.Underground, Weatherbug, Yahoo! Weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 \\
+social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
+event calendars & events & Eventbrite, Google Calendar  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
+retail (shopping)  & product profile & Amazon, Shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
+music catalogues & songs and albums & Echonest, rdio, Soundcloud, Spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
+weather & forecasts & W.Underground, Weatherbug, Yahoo! Weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
 \hline
@@
 
-\subsection{PS.2 - Technical Challenges Data Integration}
+\subsection{PS.2 - Technical Challenges of Data Integration}
 Our investigation of integration challenges started with identifying a set of candidate data sources for analysis.
@@
 For example, while many of the APIs also provided access to other kinds of data, ranging from reviews to customer profilesonly product records were obtained for this category.
-Once data records were obtained, each set of record was analysed individually for the following metrics
+Once data records were obtained, simple complexity metrics were computed individually, followed by pairwise heterogeneity metrics.
 
+The per-source metrics were as follows:
 \begin{itemize}
-\item \emph{Average number of fields per record} - A simple metric of how many properties were typically associated with each field from each source
-\item \emph{Variance in number of fields per record} - 
-\item \emph{Average Depth of record} - Levels of data which 
+\item \emph{Average Width of record} - Average number of properties per record.
+\item \emph{Average Depth of record} - Average depth of data record (number of subproperties in self-contained record).
+In JSON/XML representations which have tree representations, this would reflect the number of levels in the tree; for ``infinite'' relational representations (such as RDF with external links) this is merely the number of embedded entity definitions returned by a top-level query.
 \end{itemize}
 
-Pairwise analysis was done to compute the following heterogeneity metrics
+Heterogeneity was measured through \emph{pairwise consistency metrics} along the 
 \begin{itemize}
@@
 \item \emph{Structural consistency} - 
+\item \emph{Semantic/Model consistency} - 
 \end{itemize}
@@
 \subsubsection{Results}
+
 \section{WebBox Interface - Design of the DataPalette}

### 2012-09-14 chi2012.tex
Results section and table

--- 2012-09-14 chi2012.tex
@@
 
-The information in Table \ref{tbl:data} was captured during the studies, so that we could statistically analyses the features used by the participants.
+The information in Table \ref{tbl:data} was captured during the studies, so that we could statistically analyse the features used by the participants.
 
@@
 \begin{center}
-\begin{tabular}{|r|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
+\begin{tabular}{|r|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|}
 \hline
@@
 number data sources  &4&3&4&4&4&3&4&3&4&4&3&3&5&4&5&4&4&5&5&4\\
-chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, 1.3, 1.2& 4.2& 1.4, 1.5, 1.3& 3.2& 2.1, 2.3, 2.6& 3.5, 3.3&2.6, 2.3, 2.5&1.6, 1.5, 1.3&4.1&3.3&2.4, 2.5, 2.6\\
+chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, \newline 1.3, \newline 1.2& 4.2& 1.4, \newline 1.5, \newline 1.3& 3.2& 2.1, \newline 2.3, \newline 2.6& 3.5, \newline 3.3&2.6, \newline 2.3, \newline 2.5&1.6, \newline 1.5, \newline 1.3&4.1&3.3&2.4, \newline 2.5, \newline 2.6\\
 \hline\hline
@@
  number of websites used&1&1&4&3&3&0&2&3&2&&1&2&3&1&2&3&2&0&4&0\\
- chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4,1.2,1.6&2.1&2.3,2.6,2.5&3.5&2.2,2.3,2.6&3.5&1.1&4.1&1.6&4.2&2.6,2.3, 2.5&2.3,2.5,2.6&1.1, 1.3, 1.4&3.5\\
+ chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4, \newline 1.2, \newline 1.6&2.1&2.3, \newline 2.6, \newline 2.5&3.5&2.2, \newline 2.3, \newline 2.6&3.5&1.1&4.1&1.6&4.2&2.6, \newline 2.3, \newline 2.5&2.3, \newline 2.5, \newline 2.6&1.1, \newline 1.3,  \newline 1.4&3.5\\
  \hline
@@
 
-In the B conditions, users were given the option of using websites and or a spreadsheet to make their choice.
+In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.
 Participants who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues.
@@
 While the majority of users did not
-attempt anything complex with Excel (in fact only one users used charts), there was some advanced use of Excel observed.
+attempt anything complex with Excel (in fact only one user used charts), there was some advanced use of Excel observed.
 One user explained that they did not like to look at numbers, and so coloured the cells of the types of restaurant, and the

### 2012-09-14 chi2012.tex
ps2 preogress

--- 2012-09-14 chi2012.tex
@@
 For each feed, 3-5 data records were obtained by querying the API directly for records of a that category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}.
-For example, while many of the APIs also provided access to other kinds of data, ranging from reviews to customer profilesonly product records were obtained for this category.
-Once data records were obtained, simple complexity metrics were computed individually, followed by pairwise heterogeneity metrics.
+For example, while many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type (product records) were obtained for the \emph{shopping} category.
+Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics.
+Since, due to optional fields,  individual records could vary in the number of fields.
+An alternative approach to using instance documents that we considered was to directly compare API schema definitions
 
@@
 
-Heterogeneity was measured through \emph{pairwise consistency metrics} along the 
+Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from the 
 \begin{itemize}

### 2012-09-14 chi2012.tex
Merged conflict

@@
 
-% blargh. this needs to be integrated above! 
-In order to support our primary motivation to allow users to effectively consolidate and work with data from multiple sources, we performed an exploratory sequential mixed method study to help us design our contribution.
-For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a tasks requiring multiple websites (see Section \ref{sec:prestudy1}).
+We performed an exploratory sequential mixed method study to help us design our contribution.
+For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a tasks requiring multiple websites.
 We analysed the data collected from this study using a thematic approach.
-From our first pre-study's qualitative analysis, we designed and performed our second study which analyses existing on data available from APIs and structured data feeds on the Web (see Section \ref{sec:prestudy2}).
+
+The second study we performed an analysis on existing on data available from APIs and structured data feeds on the Web, based from our first pre-study's qualitative analysis.
 The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.
@@
 In the next two sections, we describe our two pre-studies.
-
-The second study we performed an analysis on existing on data available from APIs and structured data feeds on the Web.
-The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.
-We describe our methodology below.
 

### 2012-09-14 chi2012.tex
added subsections to discussion

@@
 
-% processes
+\subsection{Better Supporting the Task}
+
+- are people more thorough
+- do they evaluate more sources
+
+\subsection{Causing Less Frustration}
+
+- Gave up with aspects of the tasks later / not at all
+- With websites, users gave up on websites and moved on early
+
+\subsection{Keeping an Open Mind}
+
+- Did users rule out choices less/later
+- Did that mean they chose "better" options
+
+\subsection{What method did people use?}
+
+- Processes
+- Organisation of data/screen
+- A vs B, how are they different?
+
+% DAN's  qual analysis of processes
 

### 2012-09-14 chi2012.tex
more ps2

--- 2012-09-14 chi2012.tex
@@
 
-For each feed, 3-5 data records were obtained by querying the API directly for records of a that category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}.
-For example, while many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type (product records) were obtained for the \emph{shopping} category.
+For each feed, 3-5 data records were obtained by querying the API directly for records using popular key terms for each category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}.
+While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type (product records) were obtained for the \emph{shopping} category.
 Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics.
 Since, due to optional fields,  individual records could vary in the number of fields.
-An alternative approach to using instance documents that we considered was to directly compare API schema definitions
+An alternative approach to using instance documents was to directly compare schema definitions directly; however, this approach was abandoned after an initial investigation revealed two problems with schema analysis.
+First, many of the sites' schema documentation was not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
+For example, the schema documentation for \emph{shopping.com} listed full product attribute properties, but none of the records actually returned for the products we queried contained such attributes.
 
@@
 \item \emph{Average Depth of record} - Average depth of data record (number of subproperties in self-contained record).
-In JSON/XML representations which have tree representations, this would reflect the number of levels in the tree; for ``infinite'' relational representations (such as RDF with external links) this is merely the number of embedded entity definitions returned by a top-level query.
+(In JSON/XML representations, this is the tree depth; for RDF  representations this is merely the number of embedded entity resource definitions included in the top-level query).
 \end{itemize}
 
-Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from the 
+Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from paper on data heterogeneity \cite{george2005understanding}, that reflected the kinds of differences present in each set.
+These metrics are measured in a pairwies manner between sources and averaged :
 \begin{itemize}
-\item \emph{Terminological consistency} - 
-\item \emph{Structural consistency} - 
-\item \emph{Semantic/Model consistency} - 
+\item \emph{Terminological consistency} - Percentage of property names that match among schemas
+\item \emph{Structural consistency} - Percent of properties that have the same structural shape
+\item \emph{Semantic/Model consistency} - Percentage of entities/concepts that match corerspondences 
+\item \emph{Value consistency} - Percentage of values represente in the same way 
 \end{itemize}

### 2012-09-15 chi2012.tex
ps2

@@
 %% not sure where this goes >> but i want to keep it
-Setting out to investigate the potential role for better support for data integration in PIM, we first sought an updated understanding of how and whether people still used multiple information sources in the course of everyday information-related activities.
-There were several reasons for this inquiry; first, we wondered whether people were increasingly relying only on the handful of the massive ``supersites'' (Facebook, Google, Wikipedia) for most of their activities than actually searching for distributed sources of information.
+
+We performed an exploratory sequential mixed-method study to help us design our contribution.
+For our first pre-study, we first sought an updated understanding of how and whether people still used multiple information sources regularly when they sought information on the Web.
+There were several reasons for this inquiry; first, we wondered whether people were increasingly relying only on the handful of the massive ``supersites'' (Facebook, Google, Wikipedia) for most of their activities, instead of  searching across distributed sources of information.
 If this were the case, much of the benefit could be gained by merely integrating PIM tools with the handful of sites, rather than the substantially more challenging problem of integrating data from arbitrary information sources on the Web.
@@
 
-We performed an exploratory sequential mixed method study to help us design our contribution.
-For our first pre-study we designed a semi-structured interview about which tasks and the processes people use to perform a tasks requiring multiple websites.
-We analysed the data collected from this study using a thematic approach.
-
-The second study we performed an analysis on existing on data available from APIs and structured data feeds on the Web, based from our first pre-study's qualitative analysis.
-The purpose of this study was to identify the degree of technical difficulty of integrating data from these sources.
-We performed a statistical analysis of the features of the available data.
-In the next two sections, we describe our two pre-studies.
+The second pre-study focused on the characteristics of the data available from the kinds of sites used by the first pre-study's participants for the tasks described.
+The purpose of this follow-on  was to identify not specific characteristics of particular sites in question, but to identify, in general, the degree of complexity of the data feeds currently available across a variety of domains, and typical integration problems that might arise when mixing this data.
+In the remaining part sof this section, we summarise our methodologies and results for each of these two pre-studies.
 
@@
 \subsection{PS.1 - Understanding Data Diversity in Everyday Tasks}
-Our pre-study investigated the types of tasks people perform online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.
+We designed a semi-structured interview about tasks people performed on-line that required the use of multiple websites, and the processes people used to perform such tasks.
+We analysed the data collected from this study using a thematic approach, as described below.
+
+Our pre-study investigated the types of tasks people performed online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.
 We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions (see Appendix ?).
@@
 For each feed, 3-5 data records were obtained by querying the API directly for records using popular key terms for each category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}.
-While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type (product records) were obtained for the \emph{shopping} category.
+While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type were obtained for each category; e.g., product records for the \emph{shopping} category.
 Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics.
@@
 
-Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from paper on data heterogeneity \cite{george2005understanding}, that reflected the kinds of differences present in each set.
-These metrics are measured in a pairwies manner between sources and averaged :
+Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from definitions in \cite{george2005understanding} and feature-metrics used in statistical instance alignment algorithms (such as summarised in \cite{bhattacharya2007collective}).
+Such metrics were chosen because they reflected, the differences among the data feeds at multiple levels - from the simple (labelling) differneces to differences in how data was structured, and modeled.
+Each such metric listed above is the result of averaging the pairwise score between sources in each category, separately and defined as follows:
+
 \begin{itemize}
-\item \emph{Terminological consistency} - Percentage of property names that match among schemas
-\item \emph{Structural consistency} - Percent of properties that have the same structural shape
-\item \emph{Semantic/Model consistency} - Percentage of entities/concepts that match corerspondences 
-\item \emph{Value consistency} - Percentage of values represente in the same way 
+\item \emph{Terminological consistency} - Percent of property names that match
+\item \emph{Structural consistency} - Percent of properties that have \emph{compatible structures}
+\item \emph{Semantic/Model consistency} - Percent of entities/concepts that have corresponding fields 
+\item \emph{Value consistency} - Percent of values that are compatible
 \end{itemize}
@@
 \subsubsection{Results}
+% TODO TODO TODO TODO 
+Our findings suggest that
 
-\section{WebBox Interface - Design of the DataPalette}
+\section{Supporting Design of the DataPalette}
 
-Design Goals, 
+\subsection{Design Goals}
+
+At a high level, the overall goal of DataPalette was to empower end-user citizens to mix arbitrary structured data feeds on-the-fly while in the process of doing their particular information task - be it sensemaking or decision-making related.
+Unlike Mashup makers, including Yahoo Pipes!, our hope was that this could be accomplished without need for a separate, explicit mapping, integration or manual integration step which would make the effort-overhead and its use so high as to impede its use as a PIM tool.
+
+Thus we wanted to focus on the most common
+
+The second prestudy highlighted the most common inconisistencies that occurred among data sources and allowed us to the focus on a smaller set of problems.
+The most common examples were terminological (labeling) inconsitencies, minor structural inconsistencies (pertaining to values at different levels of a field's record), and value range terminological An additional problem that needed to be addressed for effectively mixing data from multipl
+
+\emph{coreference}
+
 What are the problems observed needed to address
 	What kind of interactions can be designed to solve them?
+\subsection{Process}
+
+We followed a spiral model consisting of the phases of plan, design, mock-up, prototype, test, re-plan, and re-design to derive the design of DataPalette.
+This process was used because of the high degree of initial uncertainty associated with how to design easy to use
 

### 2012-09-15 chi2012.tex
design itnro done

@@
 
-At a high level, the overall goal of DataPalette was to empower end-user citizens to mix arbitrary structured data feeds on-the-fly while in the process of doing their particular information task - be it sensemaking or decision-making related.
-Unlike Mashup makers, including Yahoo Pipes!, our hope was that this could be accomplished without need for a separate, explicit mapping, integration or manual integration step which would make the effort-overhead and its use so high as to impede its use as a PIM tool.
+At a high level, the overall goal of DataPalette was to empower end-users to mix arbitrary structured data feeds on-the-fly while solving their particular information task(s) - be them sensemaking or decision-making related - in order to more effectively accomplish these tasks.
+Unlike Mashup makers, including Yahoo Pipes!, the goal was to facilitate such integration without a need for an explicit separate mapping or integration step, which would detract from the overall usability or usefulness of the system by making people do something that did not immediately contribute to their resolving their need.
 
-Thus we wanted to focus on the most common
+Towards our design goal, the two prestudies provided valuable insight that allowed us to constrain the design search space.
+The first prestudy confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
+The second prestudy revealed that the most common inconisistencies that occurred in the APIs and data feeds available through from such data sources consisted predominantly of the simpler kinds of heterogeneity - terminological and simpler forms of structural heterogeneity, rather than the delicate semantic/modeling variety, which were comparatively rare.
+This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
 
-The second prestudy highlighted the most common inconisistencies that occurred among data sources and allowed us to the focus on a smaller set of problems.
-The most common examples were terminological (labeling) inconsitencies, minor structural inconsistencies (pertaining to values at different levels of a field's record), and value range terminological An additional problem that needed to be addressed for effectively mixing data from multipl
+In the following sections we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.
 
-\emph{coreference}
-
-What are the problems observed needed to address
-	What kind of interactions can be designed to solve them?
 \subsection{Process}

### 2012-09-15 chi2012.tex
more design

@@
 
-\section{Supporting Design of the DataPalette}
-
-\subsection{Design Goals}
+\section{Designng DataPalette : An Interface for Serendipitous Data Mixing}
 
@@
 
-Towards our design goal, the two prestudies provided valuable insight that allowed us to constrain the design search space.
+The two prestudies provided valuable insight towards this goal that allowed us to guide our design process.
 The first prestudy confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
@@
 
-We followed a spiral model consisting of the phases of plan, design, mock-up, prototype, test, re-plan, and re-design to derive the design of DataPalette.
-This process was used because of the high degree of initial uncertainty associated with how to design easy to use
+We followed a five-phase spiral model consisting of planning, designing, mocking-up, prototyping, and testing to derive the design of DataPalette.
+This process was used because of the high degree of initial uncertainty associated with how to design the mechanisms for facilitating the various aspects of effective data mixing we describe in the next section.
+We coped with this uncertainty through alternative generation and elimination; multiple alternatives were designed, drawn up and evaluated by expert colleagues (and in later iterations, non-expert potential users), who provided critiques of the alternatives.
+The ease with which designs could be prototyped in HTML5 with modern web frameworks allowed us to perform this develop-test-iterate rapidly through seven iterations of this process.
+
+\subsection{Interaction mechanisms for data mixing}
+
+In order to familiar to new users, we borrowed the overall interaction and feel from MacOS Finder-style file managers,  with support for separate windows which can be freely moved, resized and titled.
+(As described in Results, we discovered that windows caused considerable clutter and frustration for some users in later stages of their tasks, a fact we did not detect during prototype testing.)
+
+As we refined our designs, we arrived at four simple mechanisms to support data mixing.  These were combined with a number of simple interaction mechanisms that were borrowed from other systems
+
+The ultimatel design of datapal
 

### 2012-09-15 chi2012.tex
updated design figure

--- 2012-09-15 chi2012.tex
@@
 
-\section{Supporting Design of the DataPalette}
-
-\subsection{Design Goals}
+\section{Designing DataPalette : An Interface for Serendipitous Data Mixing}
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=18cm]{img/screenshot}
+\caption{WebBox uses a Drag and Drop interface to add data into the workspace.}
+\label{fig:dragdrop}
+\end{center}
+\end{figure*}
 
@@
 \subsection{Process}
+We followed a five-phase spiral model consisting of planning, designing, mocking-up, prototyping, and testing to derive the design of DataPalette.
+This process was used because of the high degree of initial uncertainty associated with the mechanisms for facilitating the various aspects of effective data mixing that we describe in the next section.
+We approached this design challenge through alternative generation and elimination; alternatives were designed, drawn up and evaluated by expert colleagues (and in later iterations, non-expert potential users), who provided critiques of the alternatives.
+The ease with which designs could be prototyped in HTML5 with modern web frameworks allowed us to perform this develop-test-iterate rapidly through seven iterations of this process.
 
-We followed a spiral model consisting of the phases of plan, design, mock-up, prototype, test, re-plan, and re-design to derive the design of DataPalette.
-This process was used because of the high degree of initial uncertainty associated with how to design easy to use
+\subsection{Interaction mechanisms for data mixing}
+In order to familiar to new users, we borrowed the overall interaction and feel from MacOS Finder-style file managers, with support for separate windows which can be freely moved, resized and titled.
+(As described in Results, we discovered that windows caused considerable clutter and frustration in later stages of some tasks, a fact we did not detect during prototype testing.)  Instead of files, however, the element of manipulation was a single, discrete structured data item, or \emph{instance}.
+(In order to do this, non-RDF data had to be first ``chunked'' into instances, which was usually a well-defined process).
+Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
+Figure \
+
+As we refined our designs, we arrived at four simple mechanisms to support data mixing.  These were combined with a number of simple interaction mechanisms that were borrowed from other systems
+
+The ultimatel design of datapal
 
@@
 Design process
-
-\begin{figure*}[htbp]
-\begin{center}
-\includegraphics[width=18cm]{img/dragdrop}
-\caption{WebBox uses a Drag and Drop interface to add data into the workspace.}
-\label{fig:dragdrop}
-\end{center}
-\end{figure*}
 

### 2012-09-15 chi2012.tex
updated design figure

@@
 \includegraphics[width=18cm]{img/screenshot}
-\caption{WebBox uses a Drag and Drop interface to add data into the workspace.}
-\label{fig:dragdrop}
+\caption{DataPalette Workspace, a file manager-inspired Drag and Drop interface for structured data mixing.
+Collapsable displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a windoow (group) on the workspace.}
+\label{fig:workspace}
 \end{center}
@@
 In order to familiar to new users, we borrowed the overall interaction and feel from MacOS Finder-style file managers, with support for separate windows which can be freely moved, resized and titled.
-(As described in Results, we discovered that windows caused considerable clutter and frustration in later stages of some tasks, a fact we did not detect during prototype testing.)  Instead of files, however, the element of manipulation was a single, discrete structured data item, or \emph{instance}.
-(In order to do this, non-RDF data had to be first ``chunked'' into instances, which was usually a well-defined process).
+(We later discovered that windows caused considerable clutter at the later stages of some tasks, a fact we did not detect until the main study as we describe later.)  Instead of files, however, the element of manipulation was a single, discrete structured data item, or \emph{instance}.
+In order to define \emph{instances} for non-RDF data, such as JSON or XML, we defined all subtrees with more than 1 child as its an instance.
 Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.

### 2012-09-15 chi2012.tex
updated design figure

@@
 Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
-Figure \
+Figure \ref{fig:workspace} displays the entire UI overview.
+
+\begin{itemize}
+\item \emph{Group multi-path value selection} - Selecting any of the properties of an instance causes all of the rest of the instances in that box to be dereferenced along the same property, displaying the resulting value underneath the instance label.
+Selecting a further property of the resulting value further dereferences this value.
+\item \emph{Drag-and-drop coreference / value combination} - In order to specify that two instances correspond to the same entity, they can be simly dragged together
+\item \emph{Group-based Drag-and-drop matching} - In order to specify that two instances correspond to the same entity, they can be simly dragged together
+\item \emph{Universal model-based brushing} -
+\end{itemize}
 

### 2012-09-15 chi2012.tex
design

@@
 In order to familiar to new users, we borrowed the overall interaction and feel from MacOS Finder-style file managers, with support for separate windows which can be freely moved, resized and titled.
-(We later discovered that windows caused considerable clutter at the later stages of some tasks, a fact we did not detect until the main study as we describe later.)  Instead of files, however, the element of manipulation was a single, discrete structured data item, or \emph{instance}.
+(As described in results, we discovered that windows caused considerable clutter at the later stages of some tasks, a fact we did not detect until the main study.)  Instead of files, however, the element of manipulation was a single, discrete structured data item, or \emph{instance}.
 In order to define \emph{instances} for non-RDF data, such as JSON or XML, we defined all subtrees with more than 1 child as its an instance.

### 2012-09-15 chi2012.tex
revisions to introduction

--- 2012-09-15 chi2012.tex
@@
 In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web, through APIs, datasets, and data feeds.
-This information includes access to data previously hidden behind services and applications, such as retailer product catalogues, information not previously directly released to the public, such as open government data or records of personal financial transactions, and new kinds of data generated by emerging kinds of data sources, such as wearable and smartphone-based biosensors.
+This information includes: access to data previously hidden behind services and applications, such as retailer product catalogues; information not previously directly released to the public, such as open government data or records of personal financial transactions; and new kinds of data generated by emerging kinds of data sources, such as wearable and smartphone-based biosensors.
+While a primary goal of opening up such direct access to information is that end-user citizens will make more informed decisions pertaining to their personal information, for example health, wealth, and well-being \cite{}.
+This data is predominately used by  specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.
+For example, the most commonly used structured personal information management (PIM) tools either manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{}.
 
-While a primary goal of opening up such direct access to information is the hope that end-user citizens will make more informed decisions pertaining to their health, wealth, or well-being \cite{}, use of this data beyond specialised app developers, journalists and other Òdata specialistsÓ remains minimal.
-Our hypothesis is that the main impediment towards end-user access to this data is a mismatch between the capabilities of tools currently used to satisfy daily information needs, and those required to effectively browse, use and consume heterogeneous structured data from diverse sources.
-For example, while much of this data represents a many and varied set of things, activities, and descriptions at various granularities, most commonly used structured personal information management (PIM) tools either can only manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{}.
+We hypothesise that end-users would benefit from this rich data with tools allowing them to effectively browse, use and consume heterogeneous structured data from diverse sources.
+In this paper, we aim to investigate how extending PIM tools to effectively use emerging ecosystems of personal data would impact information practice.
+We use a three stage sequential exploratory mixed method approach: First, we performed a qualitative analysis, using a small semi-structured interview study to examine the types of tasks people perform requiring more than one source of information, and the processes that they would use to perform such a task.
+Our results suggest that people prefer to rely upon multiple, diverse sources to singular integrated ones for a number of reasons, including coverage, reliability, and seeking out more suitable alternatives.
 
-If PIM tools were extended to enable end-users to make effective use of the emerging ecosystem of personal data, how would this impact personal information practice?
-This paper presents our initial results towards answering this question through a sequential exploratory mixed-method investigation of the personal data integration problem.
-We start with a small interview study examining information tasks people perform that rely on multiple sources of information, and how such needs are currently satisfied.
-Our results suggest that people prefer to rely upon multiple, diverse sources to singular integrated ones for a number of reasons, including coverage, reliability, and seeking out more suitable alternatives.
-Second, we conduct a analysis of various popular personal data sources available today to identify barriers to effective unification of data from each.
-We draw upon definitions and work from the database integration systems literature to characterise types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
+Second, we conducted a quantitative analysis of various popular personal data sources available today to identify potential barriers to effective unification of data from each.
+We then draw upon definitions and work from the database integration systems literature to characterise types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
 
-Based upon the data integration needs revealed by people in the interviews, and the kinds of heterogeneity observed in the data feeds, we embarked on a user-centric design exercise towards an interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
-We describe our initial foray into this exercise, an interface we call DataPalette, which focuses on the most common types of heterogeneity to enable serendipitous ``data mixing'' - a form of simple integration sufficient to let people easily and effectively combine and compare information sources without the need to write code.
-Our usability evaluation of DataPalette reveals that most users were comfortable with and that this interface is effectively expedites the most sophisticated of the typical kinds of ``data integration'' tasks people need to perform while planning personal,
+Third, we developed DataPalette, which focuses on the most common types of heterogeneity to enable serendipitous ``data mixing'' - a form of simple integration sufficient to allow people to easily and effectively combine and compare information sources without the need to write code.
+It was principally designed based upon people's data integration needs and to support data heterogeneity observed in the previous studies.
+We embarked on a user-centric design exercise towards an interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
+We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with and that this interface is effectively expedites the most sophisticated of the typical kinds of ``data integration'' tasks people need to perform while planning personal.
 
@@
 
-\section{Designing DataPalette : An Interface for Serendipitous Data Mixing}
-\begin{figure*}[htbp]
-\begin{center}
-\includegraphics[width=18cm]{img/screenshot}
-\caption{DataPalette Workspace, a file manager-inspired Drag and Drop interface for structured data mixing.
-Collapsable displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a windoow (group) on the workspace.}
-\label{fig:workspace}
-\end{center}
-\end{figure*}
+\section{Supporting Design of the DataPalette}
 
-At a high level, the overall goal of DataPalette was to empower end-users to mix arbitrary structured data feeds on-the-fly while solving their particular information task(s) - be them sensemaking or decision-making related - in order to more effectively accomplish these tasks.
-Unlike Mashup makers, including Yahoo Pipes!, the goal was to facilitate such integration without a need for an explicit separate mapping or integration step, which would detract from the overall usability or usefulness of the system by making people do something that did not immediately contribute to their resolving their need.
+\subsection{Design Goals}
 
-The two prestudies provided valuable insight towards this goal that allowed us to guide our design process.
-The first prestudy confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
-The second prestudy revealed that the most common inconisistencies that occurred in the APIs and data feeds available through from such data sources consisted predominantly of the simpler kinds of heterogeneity - terminological and simpler forms of structural heterogeneity, rather than the delicate semantic/modeling variety, which were comparatively rare.
-This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
+At a high level, the overall goal of DataPalette was to empower end-user citizens to mix arbitrary structured data feeds on-the-fly while in the process of doing their particular information task - be it sensemaking or decision-making related.
+Unlike Mashup makers, including Yahoo Pipes!, our hope was that this could be accomplished without need for a separate, explicit mapping, integration or manual integration step which would make the effort-overhead and its use so high as to impede its use as a PIM tool.
 
-In the following sections we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.
+Thus we wanted to focus on the most common
 
+The second prestudy highlighted the most common inconisistencies that occurred among data sources and allowed us to the focus on a smaller set of problems.
+The most common examples were terminological (labeling) inconsitencies, minor structural inconsistencies (pertaining to values at different levels of a field's record), and value range terminological An additional problem that needed to be addressed for effectively mixing data from multipl
+
+\emph{coreference}
+
+What are the problems observed needed to address
+	What kind of interactions can be designed to solve them?
 \subsection{Process}
-We followed a five-phase spiral model consisting of planning, designing, mocking-up, prototyping, and testing to derive the design of DataPalette.
-This process was used because of the high degree of initial uncertainty associated with the mechanisms for facilitating the various aspects of effective data mixing that we describe in the next section.
-We approached this design challenge through alternative generation and elimination; alternatives were designed, drawn up and evaluated by expert colleagues (and in later iterations, non-expert potential users), who provided critiques of the alternatives.
-The ease with which designs could be prototyped in HTML5 with modern web frameworks allowed us to perform this develop-test-iterate rapidly through seven iterations of this process.
 
-\subsection{Interaction mechanisms for data mixing}
-In order to familiar to new users, we borrowed the overall interaction and feel from MacOS Finder-style file managers, with support for separate windows which can be freely moved, resized and titled.
-(We later discovered that windows caused considerable clutter at the later stages of some tasks, a fact we did not detect until the main study as we describe later.)  Instead of files, however, the element of manipulation was a single, discrete structured data item, or \emph{instance}.
-In order to define \emph{instances} for non-RDF data, such as JSON or XML, we defined all subtrees with more than 1 child as its an instance.
-Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
-Figure \
-
-As we refined our designs, we arrived at four simple mechanisms to support data mixing.  These were combined with a number of simple interaction mechanisms that were borrowed from other systems
-
-The ultimatel design of datapal
+We followed a spiral model consisting of the phases of plan, design, mock-up, prototype, test, re-plan, and re-design to derive the design of DataPalette.
+This process was used because of the high degree of initial uncertainty associated with how to design easy to use
 
@@
 Design process
+
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=18cm]{img/dragdrop}
+\caption{WebBox uses a Drag and Drop interface to add data into the workspace.}
+\label{fig:dragdrop}
+\end{center}
+\end{figure*}
 

### 2012-09-15 chi2012.tex

@@
 
-\section{Supporting Design of the DataPalette}
+\section{Designing DataPalette : An Interface for Serendipitous Data Mixing}
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=18cm]{img/screenshot}
+\caption{DataPalette Workspace, a file manager-inspired Drag and Drop interface for structured data mixing.
+Collapsable displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a windoow (group) on the workspace.}
+\label{fig:workspace}
+\end{center}
+\end{figure*}
 
-\subsection{Design Goals}
+At a high level, the overall goal of DataPalette was to empower end-users to mix arbitrary structured data feeds on-the-fly while solving their particular information task(s) - be them sensemaking or decision-making related - in order to more effectively accomplish these tasks.
+Unlike Mashup makers, including Yahoo Pipes!, the goal was to facilitate such integration without a need for an explicit separate mapping or integration step, which would detract from the overall usability or usefulness of the system by making people do something that did not immediately contribute to their resolving their need.
 
-At a high level, the overall goal of DataPalette was to empower end-user citizens to mix arbitrary structured data feeds on-the-fly while in the process of doing their particular information task - be it sensemaking or decision-making related.
-Unlike Mashup makers, including Yahoo Pipes!, our hope was that this could be accomplished without need for a separate, explicit mapping, integration or manual integration step which would make the effort-overhead and its use so high as to impede its use as a PIM tool.
+The two prestudies provided valuable insight towards this goal that allowed us to guide our design process.
+The first prestudy confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
+The second prestudy revealed that the most common inconisistencies that occurred in the APIs and data feeds available through from such data sources consisted predominantly of the simpler kinds of heterogeneity - terminological and simpler forms of structural heterogeneity, rather than the delicate semantic/modeling variety, which were comparatively rare.
+This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
 
-Thus we wanted to focus on the most common
+In the following sections we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.
 
-The second prestudy highlighted the most common inconisistencies that occurred among data sources and allowed us to the focus on a smaller set of problems.
-The most common examples were terminological (labeling) inconsitencies, minor structural inconsistencies (pertaining to values at different levels of a field's record), and value range terminological An additional problem that needed to be addressed for effectively mixing data from multipl
+\subsection{Process}
+We followed a five-phase spiral model consisting of planning, designing, mocking-up, prototyping, and testing to derive the design of DataPalette.
+This process was used because of the high degree of initial uncertainty associated with the mechanisms for facilitating the various aspects of effective data mixing that we describe in the next section.
+We approached this design challenge through alternative generation and elimination; alternatives were designed, drawn up and evaluated by expert colleagues (and in later iterations, non-expert potential users), who provided critiques of the alternatives.
+The ease with which designs could be prototyped in HTML5 with modern web frameworks allowed us to perform this develop-test-iterate rapidly through seven iterations of this process.
 
-\emph{coreference}
+\subsection{Interaction mechanisms for data mixing}
+In order to familiar to new users, we borrowed the overall interaction and feel from MacOS Finder-style file managers, with support for separate windows which can be freely moved, resized and titled.
+(As described in results, we discovered that windows caused considerable clutter at the later stages of some tasks, a fact we did not detect until the main study.)  Instead of files, however, the element of manipulation was a single, discrete structured data item, or \emph{instance}.
+In order to define \emph{instances} for non-RDF data, such as JSON or XML, we defined all subtrees with more than 1 child as its an instance.
+Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
+Figure \ref{fig:workspace} displays the entire UI overview.
 
-What are the problems observed needed to address
-	What kind of interactions can be designed to solve them?
-\subsection{Process}
+\begin{itemize}
+\item \emph{Group multi-path value selection} - Selecting any of the properties of an instance causes all of the rest of the instances in that box to be dereferenced along the same property, displaying the resulting value underneath the instance label.
+Selecting a further property of the resulting value further dereferences this value.
+\item \emph{Drag-and-drop coreference / value combination} - In order to specify that two instances correspond to the same entity, they can be simly dragged together
+\item \emph{Group-based Drag-and-drop matching} - In order to specify that two instances correspond to the same entity, they can be simly dragged together
+\item \emph{Universal model-based brushing} -
+\end{itemize}
 
-We followed a spiral model consisting of the phases of plan, design, mock-up, prototype, test, re-plan, and re-design to derive the design of DataPalette.
-This process was used because of the high degree of initial uncertainty associated with how to design easy to use
+As we refined our designs, we arrived at four simple mechanisms to support data mixing.  These were combined with a number of simple interaction mechanisms that were borrowed from other systems
+
+The ultimatel design of datapal
 

### 2012-09-15 chi2012.tex
tense wrestling

@@
 \begin{abstract}
-Exploratory sequential mixed method study
+Exploratory sequential mixed-method study
 \end{abstract}
@@
 This information includes: access to data previously hidden behind services and applications, such as retailer product catalogues; information not previously directly released to the public, such as open government data or records of personal financial transactions; and new kinds of data generated by emerging kinds of data sources, such as wearable and smartphone-based biosensors.
-While a primary goal of opening up such direct access to information is that end-user citizens will make more informed decisions pertaining to their personal information, for example health, wealth, and well-being \cite{}.
+A primary goal of opening up such direct access to information is that end-user citizens will make more informed decisions pertaining to their personal information, for example health, wealth, and well-being \cite{}.
 This data is predominately used by  specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.
@@
 We hypothesise that end-users would benefit from this rich data with tools allowing them to effectively browse, use and consume heterogeneous structured data from diverse sources.
-In this paper, we aim to investigate how extending PIM tools to effectively use emerging ecosystems of personal data would impact information practice.
-We use a three stage sequential exploratory mixed method approach: First, we performed a qualitative analysis, using a small semi-structured interview study to examine the types of tasks people perform requiring more than one source of information, and the processes that they would use to perform such a task.
-Our results suggest that people prefer to rely upon multiple, diverse sources to singular integrated ones for a number of reasons, including coverage, reliability, and seeking out more suitable alternatives.
+In this paper, we present our investigation of how extending PIM tools to effectively use emerging ecosystems of personal data impact information practice.
+We used a three stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of  information sources, and the processes that they relied upon to perform such tasks.
+Our results suggested that people preferred to rely upon multiple, diverse sources to singular, integrated ones for a number of reasons, including coverage, reliability, and seeking out more suitable alternatives.
 
-Second, we conducted a quantitative analysis of various popular personal data sources available today to identify potential barriers to effective unification of data from each.
-We then draw upon definitions and work from the database integration systems literature to characterise types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
+Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data from each.
+Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
 

### 2012-09-15 chi2012.tex
some edits to intro

@@
 We embarked on a user-centric design exercise towards an interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
-We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with and that this interface is effectively expedites the most sophisticated of the typical kinds of ``data integration'' tasks people need to perform while planning personal.
+We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision making tasks using multiple sources.
 

### 2012-09-15 chi2012.tex
some edits to intro

@@
 
-\subsection{Interaction mechanisms for data mixing}
-In order to familiar to new users, we borrowed the overall interaction and feel from MacOS Finder-style file managers, with support for separate windows which can be freely moved, resized and titled.
-(As described in results, we discovered that windows caused considerable clutter at the later stages of some tasks, a fact we did not detect until the main study.)  Instead of files, however, the element of manipulation was a single, discrete structured data item, or \emph{instance}.
-In order to define \emph{instances} for non-RDF data, such as JSON or XML, we defined all subtrees with more than 1 child as its an instance.
+\subsection{Interaction with DataPalette}
+In order to feel familiar to new users and focus our design efforts on integration-related aspects of the interface, we borrowed metaphors and interactions directly from MacOS Finder-style file managers.
+Specifically, in the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can arbitrarily drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
+(Although the design could be generalised to supporting nested groups, our prototype only supported a single level of grouping.)
+
+Unlike file managers, the basic unit of data is not a file, but an \emph{instance} - corresponding to an RDF instance or any small bit of structured data.
+In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data) into discrete instances with properties; for JSON or XML data feeds, we represented all subtrees with more than 1 child as its an instance.
+
+Each instance is displayed with its corresponding properties, and \emph{value selection}.
+
 Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
@@
 \begin{itemize}
-\item \emph{Group multi-path value selection} - Selecting any of the properties of an instance causes all of the rest of the instances in that box to be dereferenced along the same property, displaying the resulting value underneath the instance label.
+\item \emph{Group multi-path value selection} - Selecting any of the properties of an instance in a group (window) causes all of the rest of the instances that have that property in that group to be dereferenced along the same property, displaying the resulting value underneath the instance label.
 Selecting a further property of the resulting value further dereferences this value.

### 2012-09-15 chi2012.tex
interaction description

@@
 
+\subsubsection{Multipath selection: link-sliding for heterogeneous sets}
+When working with instances, users may be interested in a particular aspect of an instance rather than just the instance itself; for example, a user may be interested in the price of a particular product, its rating, manufacturer, and so on.
+These are represented as \emph{properties} of the item, listed under the name of the instance.
+Selecting any of such properties creates a \emph{path   selection} which dereferences all of the instances in that group that have that property, yielding their corresponding values.
+For example, selecting the \emph{manufacturer} property for an instance in a group will cause the manufacturer of that item to appear alongside its label, and will also cause the manufacturers of \emph{all other   instances in that group that have a manufacturer property} to be dereferenced (and displayed) as well.
+When such a path selection is applied, the properties displayed are updated to be the set of properties of the \emph{terminating value of the path}, so that such value can be further dereferenced.
+For example, once selecting a manufacturer, a user may want to know the product's manufacturer's overall reputation, so might select its ``reputation score'' property, thereby extending the path selection.
+
+Instances that cannot be dereferenced by a particular path selection  are left alone;  a second path selection can be created for such insances, which, like the first path selection, causes all like instances to be dereferenced.
+Each group can carry an arbitrary number of path selections, and each path selection can be extended arbitrarily deep by selecting successive properties of values.
+This achieves the ability to \emph{link slide} in the same way introduced by Parallax \cite{parallax}.
+
 Each instance is displayed with its corresponding properties, and \emph{value selection}.
@@
 Figure \ref{fig:workspace} displays the entire UI overview.
+
+\subsection{Summary of interaction features}
 

### 2012-09-15 chi2012.tex
interaction description

@@
 
-Each instance is displayed with its corresponding properties, and \emph{value selection}.
+The effect that multiple selection group-dereferencing of values are twofold: the fact that clicking on a property dereferences all corresponding values makes it easy to quickly compare all instances that have the same structure.
+Although we were concerned that users might be surprised that path selection applied to all objects within each group in parallel, users adapted very quickly and used this functionality to save them time.
 
-Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
+The second effect that this multiple paths is that supporting the selection of different paths by structure lets values be selected using separate different paths - making it possible to easily cope with properties with different names, and even at levels -- select the appropriate values even between instances that exhibit both  terminological and structurally heterogeneity.
+
+% Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
 Figure \ref{fig:workspace} displays the entire UI overview.

### 2012-09-15 chi2012.tex
interaction description

@@
 Selecting any of such properties creates a \emph{path   selection} which dereferences all of the instances in that group that have that property, yielding their corresponding values.
-For example, selecting the \emph{manufacturer} property for an instance in a group will cause the manufacturer of that item to appear alongside its label, and will also cause the manufacturers of \emph{all other   instances in that group that have a manufacturer property} to be dereferenced (and displayed) as well.
+For example, selecting the \emph{manufacturer} property for an instance in a group will cause the manufacturer of that item to appear alongside its label, and will also cause the manufacturers of \emph{all other   instances in that group that have a manufacturer property} to be dereferenced (and displayed next to their instances) as well.
 When such a path selection is applied, the properties displayed are updated to be the set of properties of the \emph{terminating value of the path}, so that such value can be further dereferenced.
-For example, once selecting a manufacturer, a user may want to know the product's manufacturer's overall reputation, so might select its ``reputation score'' property, thereby extending the path selection.
+For example, once selecting a manufacturer, a user may want to know the product's manufacturer's overall reputation, so might select its ``reputation score'' property, causing the previous path selection to be extended.
 
-Instances that cannot be dereferenced by a particular path selection  are left alone;  a second path selection can be created for such insances, which, like the first path selection, causes all like instances to be dereferenced.
+Instances that cannot be dereferenced by a particular path selection  are left alone; an additional path selection can be created for such insances, which, like the first path selection, causes all appropriate instances to be dereferenced.
 Each group can carry an arbitrary number of path selections, and each path selection can be extended arbitrarily deep by selecting successive properties of values.

### 2012-09-15 chi2012.tex
interaction description
minor edits

--- 2012-09-15 chi2012.tex
@@
 \begin{abstract}
-Exploratory sequential mixed-method study
+Exploratory sequential mixed method study
 \end{abstract}
@@
 This information includes: access to data previously hidden behind services and applications, such as retailer product catalogues; information not previously directly released to the public, such as open government data or records of personal financial transactions; and new kinds of data generated by emerging kinds of data sources, such as wearable and smartphone-based biosensors.
-A primary goal of opening up such direct access to information is that end-user citizens will make more informed decisions pertaining to their personal information, for example health, wealth, and well-being \cite{}.
+While a primary goal of opening up such direct access to information is that end-user citizens will make more informed decisions pertaining to their personal information, for example health, wealth, and well-being \cite{}.
 This data is predominately used by  specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.
@@
 We hypothesise that end-users would benefit from this rich data with tools allowing them to effectively browse, use and consume heterogeneous structured data from diverse sources.
-In this paper, we present our investigation of how extending PIM tools to effectively use emerging ecosystems of personal data impact information practice.
-We used a three stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of  information sources, and the processes that they relied upon to perform such tasks.
-Our results suggested that people preferred to rely upon multiple, diverse sources to singular, integrated ones for a number of reasons, including coverage, reliability, and seeking out more suitable alternatives.
+In this paper, we aim to investigate how extending PIM tools to effectively use emerging ecosystems of personal data would impact information practice.
+We use a three stage sequential exploratory mixed method approach: First, we performed a qualitative analysis, using a small semi-structured interview study to examine the types of tasks people perform requiring more than one source of information, and the processes that they would use to perform such a task.
+Our results suggest that people prefer to rely upon multiple, diverse sources to singular integrated ones for a number of reasons, including coverage, reliability, and seeking out more suitable alternatives.
 
-Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data from each.
-Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
+Second, we conducted a quantitative analysis of various popular personal data sources available today to identify potential barriers to effective unification of data from each.
+We then draw upon definitions and work from the database integration systems literature to characterise types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
 
@@
 We embarked on a user-centric design exercise towards an interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
-We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision making tasks using multiple sources.
+We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with and that this interface is effectively expedites the most sophisticated of the typical kinds of ``data integration'' tasks people need to perform while planning personal.
 
@@
 
-In this section, we touch on motivating work for DataPalette from three fields: peronal information management, data integration, and end-user mashups and toolkits.
+In this section, we touch on motivating work for DataPalette from three fields: peronsal information management, data integration, and end-user mashups and toolkits.
 
@@
 In our own previously-conducted field study, (reference anonymised) we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage heterogeneous data from disparate sources -- be them online, offline, or other people.
-Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-mainted information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
+Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-maintained information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
 %% TODO MORE HERE
@@
 \subsection{PS.1 - Understanding Data Diversity in Everyday Tasks}
-We designed a semi-structured interview about tasks people performed on-line that required the use of multiple websites, and the processes people used to perform such tasks.
+We designed a semi-structured interview about tasks people performed online that required the use of multiple websites, and the processes people used to perform such tasks.
 We analysed the data collected from this study using a thematic approach, as described below.
@@
 Our pre-study investigated the types of tasks people performed online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.
-We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions (see Appendix ?).
+We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions.
 This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant's experiences, while collecting the same general areas of information.
@@
 
-\subsubsection{Question 2: Large Social Event}
+\subsubsection{Question 2: Process Used to Organise a Large Social Event}
 All of the participants would organise a large social event by talking to their friends, about their ideas, preferences and recommendations.
@@
 
-\subsection{Interaction with DataPalette}
-In order to feel familiar to new users and focus our design efforts on integration-related aspects of the interface, we borrowed metaphors and interactions directly from MacOS Finder-style file managers.
-Specifically, in the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can arbitrarily drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
-(Although the design could be generalised to supporting nested groups, our prototype only supported a single level of grouping.)
-
-Unlike file managers, the basic unit of data is not a file, but an \emph{instance} - corresponding to an RDF instance or any small bit of structured data.
-In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data) into discrete instances with properties; for JSON or XML data feeds, we represented all subtrees with more than 1 child as its an instance.
-
-\subsubsection{Multipath selection: link-sliding for heterogeneous sets}
-When working with instances, users may be interested in a particular aspect of an instance rather than just the instance itself; for example, a user may be interested in the price of a particular product, its rating, manufacturer, and so on.
-These are represented as \emph{properties} of the item, listed under the name of the instance.
-Selecting any of such properties creates a \emph{path   selection} which dereferences all of the instances in that group that have that property, yielding their corresponding values.
-For example, selecting the \emph{manufacturer} property for an instance in a group will cause the manufacturer of that item to appear alongside its label, and will also cause the manufacturers of \emph{all other   instances in that group that have a manufacturer property} to be dereferenced (and displayed) as well.
-When such a path selection is applied, the properties displayed are updated to be the set of properties of the \emph{terminating value of the path}, so that such value can be further dereferenced.
-For example, once selecting a manufacturer, a user may want to know the product's manufacturer's overall reputation, so might select its ``reputation score'' property, thereby extending the path selection.
-
-Instances that cannot be dereferenced by a particular path selection  are left alone;  a second path selection can be created for such insances, which, like the first path selection, causes all like instances to be dereferenced.
-Each group can carry an arbitrary number of path selections, and each path selection can be extended arbitrarily deep by selecting successive properties of values.
-This achieves the ability to \emph{link slide} in the same way introduced by Parallax \cite{parallax}.
-
-Each instance is displayed with its corresponding properties, and \emph{value selection}.
-
+\subsection{Interaction mechanisms for data mixing}
+In order to familiar to new users, we borrowed the overall interaction and feel from MacOS Finder-style file managers, with support for separate windows which can be freely moved, resized and titled.
+(As described in results, we discovered that windows caused considerable clutter at the later stages of some tasks, a fact we did not detect until the main study.)  Instead of files, however, the element of manipulation was a single, discrete structured data item, or \emph{instance}.
+In order to define \emph{instances} for non-RDF data, such as JSON or XML, we defined all subtrees with more than 1 child as its an instance.
 Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
@@
 
-\subsection{Summary of interaction features}
-
 \begin{itemize}
-\item \emph{Group multi-path value selection} - Selecting any of the properties of an instance in a group (window) causes all of the rest of the instances that have that property in that group to be dereferenced along the same property, displaying the resulting value underneath the instance label.
+\item \emph{Group multi-path value selection} - Selecting any of the properties of an instance causes all of the rest of the instances in that box to be dereferenced along the same property, displaying the resulting value underneath the instance label.
 Selecting a further property of the resulting value further dereferences this value.

### 2012-09-15 chi2012.tex

--- 2012-09-15 chi2012.tex
@@
 Each group can carry an arbitrary number of path selections, and each path selection can be extended arbitrarily deep by selecting successive properties of values.
-This achieves the ability to \emph{link slide} in the same way introduced by Parallax \cite{parallax}.
+This achieves the ability to \emph{link slide} as introduced by Parallax, but across sets of heterogeneous items by supporting multple parallel dereferenced paths simultaneously \cite{parallax}.
 

### 2012-09-15 chi2012.tex
interaction description

--- 2012-09-15 chi2012.tex
@@
 The second effect that this multiple paths is that supporting the selection of different paths by structure lets values be selected using separate different paths - making it possible to easily cope with properties with different names, and even at levels -- select the appropriate values even between instances that exhibit both  terminological and structurally heterogeneity.
+
+\subsubsection{Coreference management: drag-and-drop sameas}
+
+A major problem with data integration is consolidating representations that co-refer to the same entity, e.g.
+person, place or thing.
+For example, effectively looking up properties of one's friends using Facebook and Google+ services, one might quickly want to combine the representations of our friends that have profiles on \emph{both} services to into single instances.
+
+Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances.  Thus,
+
+To make this as easy as possible,
+
+This was made \emph{optional}, so that people could opt to combine multiple representations co-referring to the same thing or not as it suits their purposes.
+Keeping representations from multiple sources separate,
 

### 2012-09-15 chi2012.tex
drag and drop

@@
 
-\subsubsection{Coreference management: drag-and-drop sameas}
+\subsubsection{Coreference management: Drag-and-Drop Same-As}
 
-A major problem with data integration is consolidating representations that co-refer to the same entity, e.g.
-person, place or thing.
-For example, effectively looking up properties of one's friends using Facebook and Google+ services, one might quickly want to combine the representations of our friends that have profiles on \emph{both} services to into single instances.
+A major challenge with integration from multiple sources is coreference consolidation, or combining representations that co-refer to the same entity into one.
+For example, consolidating one's friends from Facebook, Google+, and LinkedIn, one might quickly want to de-duplicate and combine friends that use multiple services so that they each have one representation.
 
-Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances.  Thus,
+Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, IDs used to identify particular individuals were system-specific.
+Thus, we sought to make instead make it use-driven, and easy to do manually.
 
-To make this as easy as possible,
+In order to remain consistent with the drag-and-drop interactions used throughout the interface, we decided that simply dragging an instance onto another instance was a likely intuitive way for people to specify that two things should be combined.
 
-This was made \emph{optional}, so that people could opt to combine multiple representations co-referring to the same thing or not as it suits their purposes.
-Keeping representations from multiple sources separate,
+Although general, this can be tedious for groups of items; to support more efficient combination, we extended this to allow entire groups to be dropped on other groups.
+When this is done, DataPalette attempts to identify matching pairs among items in the dragged and target groups by comparing each pairs' labels.
+Only exact matches (modulo white space/underscores) are considered matches and automatically combined.
+The rest are added to the target group as distinct elements, and may be explicitly combined by hand.
 
-% Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
+Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
 Figure \ref{fig:workspace} displays the entire UI overview.

### 2012-09-15 chi2012.tex

--- 2012-09-15 chi2012.tex
@@
 
-% Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
+\subsubsection{Coreference management: Drag-and-Drop Same-As}
+
+A major challenge with integration from multiple sources is coreference consolidation, or combining representations that co-refer to the same entity into one.
+For example, consolidating one's friends from Facebook, Google+, and LinkedIn, one might quickly want to de-duplicate and combine friends that use multiple services so that they each have one representation.
+
+Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, IDs used to identify particular individuals were system-specific.
+Thus, we sought to make instead make it use-driven, and easy to do manually.
+
+In order to remain consistent with the drag-and-drop interactions used throughout the interface, we decided that simply dragging an instance onto another instance was a likely intuitive way for people to specify that two things should be combined.
+
+Although general, this can be tedious for groups of items; to support more efficient combination, we extended this to allow entire groups to be dropped on other groups.
+When this is done, DataPalette attempts to identify matching pairs among items in the dragged and target groups by comparing each pairs' labels.
+Only exact matches (modulo white space/underscores) are considered matches and automatically combined.
+The rest are added to the target group as distinct elements, and may be explicitly combined by hand.
+
+Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
 Figure \ref{fig:workspace} displays the entire UI overview.

### 2012-09-15 chi2012.tex
Updated section on process

@@
 \subsubsection{Multipath selection: link-sliding for heterogeneous sets}
-When working with instances, users may be interested in a particular aspect of an instance rather than just the instance itself; for example, a user may be interested in the price of a particular product, its rating, manufacturer, and so on.
-These are represented as \emph{properties} of the item, listed under the name of the instance.
-Selecting any of such properties creates a \emph{path   selection} which dereferences all of the instances in that group that have that property, yielding their corresponding values.
-For example, selecting the \emph{manufacturer} property for an instance in a group will cause the manufacturer of that item to appear alongside its label, and will also cause the manufacturers of \emph{all other   instances in that group that have a manufacturer property} to be dereferenced (and displayed next to their instances) as well.
-When such a path selection is applied, the properties displayed are updated to be the set of properties of the \emph{terminating value of the path}, so that such value can be further dereferenced.
-For example, once selecting a manufacturer, a user may want to know the product's manufacturer's overall reputation, so might select its ``reputation score'' property, causing the previous path selection to be extended.
 
-Instances that cannot be dereferenced by a particular path selection  are left alone; an additional path selection can be created for such insances, which, like the first path selection, causes all appropriate instances to be dereferenced.
-Each group can carry an arbitrary number of path selections, and each path selection can be extended arbitrarily deep by selecting successive properties of values.
-This achieves the ability to \emph{link slide} as introduced by Parallax, but across sets of heterogeneous items by supporting multple parallel dereferenced paths simultaneously \cite{parallax}.
+Typically users are interested in different \emph{properties} of an instance, for example, a user is interested in a product and its price, rating, manufacturer and so on.
+An instance's properties are displayed beneath it; selecting a property creates a \emph{path selection}, which dereferences all of the instances in a group that contain the selected property and displays their corresponding values.
+For example, when a user selects the \emph{manufacturer} property for a product, all products in the same group with a manufacturer property will be dereferenced, causing the manufactueres to be displayed alongside their corresponding products.
+When such a path selection is applied, the properties displayed are updated to be the set of properties of the \emph{terminating value of the path}, so that this value can be dereferenced further.
+For example, once selecting a manufacturer, a user may want to know the product's manufacturer's reputation and selects their ``reputation score'' property, causing the previous path selection to be extended.
 
-The effect that multiple selection group-dereferencing of values are twofold: the fact that clicking on a property dereferences all corresponding values makes it easy to quickly compare all instances that have the same structure.
+When instances cannot be dereferenced by the chosen path selection, an additional path selection can be created.
+As with the first path selection, it causes all matching instances to be dereferenced.
+Each group can carry an arbitrary number of path selections, and each path selection can be extended continually by selecting successive properties of values.
+This achieves the ability to \emph{link slide} as introduced by Parallax~\cite{parallax}, but across sets of heterogeneous items by supporting multiple parallel dereferenced paths simultaneously.
+
+Multiple selection group-dereferencing results in two effects:
+
+\begin{enumerate}
+\item Clicking on a property dereferences all corresponding values makes it easy to quickly compare all instances that have the same structure.
 Although we were concerned that users might be surprised that path selection applied to all objects within each group in parallel, users adapted very quickly and used this functionality to save them time.
 
-The second effect that this multiple paths is that supporting the selection of different paths by structure lets values be selected using separate different paths - making it possible to easily cope with properties with different names, and even at levels -- select the appropriate values even between instances that exhibit both  terminological and structurally heterogeneity.
+\item That supporting the selection of different paths by structure lets values be selected using separate different paths - making it possible to easily cope with properties with different names, and even at levels -- select the appropriate values even between instances that exhibit both  terminological and structurally heterogeneity.
+\end{enumerate}
 
-\subsubsection{Coreference management: Drag-and-Drop Same-As}
-
-A major challenge with integration from multiple sources is coreference consolidation, or combining representations that co-refer to the same entity into one.
-For example, consolidating one's friends from Facebook, Google+, and LinkedIn, one might quickly want to de-duplicate and combine friends that use multiple services so that they each have one representation.
-
-Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, IDs used to identify particular individuals were system-specific.
-Thus, we sought to make instead make it use-driven, and easy to do manually.
-
-In order to remain consistent with the drag-and-drop interactions used throughout the interface, we decided that simply dragging an instance onto another instance was a likely intuitive way for people to specify that two things should be combined.
-
-Although general, this can be tedious for groups of items; to support more efficient combination, we extended this to allow entire groups to be dropped on other groups.
-When this is done, DataPalette attempts to identify matching pairs among items in the dragged and target groups by comparing each pairs' labels.
-Only exact matches (modulo white space/underscores) are considered matches and automatically combined.
-The rest are added to the target group as distinct elements, and may be explicitly combined by hand.
-
-Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
+% Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
 Figure \ref{fig:workspace} displays the entire UI overview.

### 2012-09-15 chi2012.tex
more dnd

--- 2012-09-15 chi2012.tex
@@
 
-Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, IDs used to identify particular individuals were system-specific.
-Thus, we sought to make instead make it use-driven, and easy to do manually.
+Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, ids used by each network to identify particular individuals were system-specific.
+Thus, we sought to make instead make it user-driven, and easy to do manually.
 
 In order to remain consistent with the drag-and-drop interactions used throughout the interface, we decided that simply dragging an instance onto another instance was a likely intuitive way for people to specify that two things should be combined.
-
 Although general, this can be tedious for groups of items; to support more efficient combination, we extended this to allow entire groups to be dropped on other groups.
@@
 The rest are added to the target group as distinct elements, and may be explicitly combined by hand.
+We are considering more sophisticated / less fragile approaches to matching in the future, which would make the system assistive.
 
-Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
+By having co-reference consolidation user-driven, we gave the user freedom to combine things when suited to his or her task.
+As we saw in the system evaluation, while in certain circumstances users preferred combining representations, in others people preferred to keep things separate.
+Thus, user-initiated consolidation afforded an additional flexibility that was ultimately deemed beneficial.
+
+% Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
 Figure \ref{fig:workspace} displays the entire UI overview.

### 2012-09-15 chi2012.tex
added reference to the figures int eraction

@@
 
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=8cm]{img/multipathing}
+\caption{To see properties, users click on the name of the property. When instances do not have a particular property, additional property names can be added.}
+\label{fig:multipathing}
+\end{center}
+\end{figure}
+
 Typically users are interested in different \emph{properties} of an instance, for example, a user is interested in a product and its price, rating, manufacturer and so on.
@@
 For example, once selecting a manufacturer, a user may want to know the product's manufacturer's reputation and selects their ``reputation score'' property, causing the previous path selection to be extended.
+This process is illustrated in Figure \ref{fig:multipathing}.
 
@@
 
-\item That supporting the selection of different paths by structure lets values be selected using separate different paths - making it possible to easily cope with properties with different names, and even at levels -- select the appropriate values even between instances that exhibit both  terminological and structurally heterogeneity.
+\item Supporting dereferencing along different paths \emph{by structure} allows instances with different property names, and even sub-structures to be quickly consolidated by separate dereferencing, thereby effectively coping with terminological and structural hetereogeneity.
 \end{enumerate}
@@
 \subsubsection{Coreference management: Drag-and-Drop Same-As}
+
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=18cm]{img/sameas}
+\caption{When two instances represent the same entity, users can drag one on top of the other, and they are combined.}
+\label{fig:sameas}
+\end{center}
+\end{figure*}
 
@@
 In order to remain consistent with the drag-and-drop interactions used throughout the interface, we decided that simply dragging an instance onto another instance was a likely intuitive way for people to specify that two things should be combined.
+An illustration of this interaction is provided in Figure \ref{fig:sameas}.
 Although general, this can be tedious for groups of items; to support more efficient combination, we extended this to allow entire groups to be dropped on other groups.
@@
 
+\subsubsection{Enumerated-type value consolidation}
+
+Related to the concept of co-reference, we extended the drag-and-drop sameas support to also be used to consolidate inconsistant enumerated values of properties, such as ``Casual dining'' vs ``Relaxed'' so that they can be uniformly compared.
+This allows the creation of meaningful, de-duplicated histograms and other visualisations.
+
+\subsubsection{Model-based brushing}
+
+Since it can get confusing having an instance represented in multiple groups,
+
 % Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
 Figure \ref{fig:workspace} displays the entire UI overview.
+
+\subsubsection{Adaptive visualisation modes}
 
@@
 
-As we refined our designs, we arrived at four simple mechanisms to support data mixing.  These were combined with a number of simple interaction mechanisms that were borrowed from other systems
+%% Interaction features 
+%% 	multiple paths selection
+%% 	sameas interactions
+%% 	visualisation of partial results
 
-The ultimatel design of datapal
-
-Interaction features 
-	multiple paths selection
-	sameas interactions
-	visualisation of partial results
-
-Design process
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=8cm]{img/multipathing}
-\caption{To see properties, users click on the name of the property. When instances do not have a particular property, additional property names can be added.}
-\label{fig:multipathing}
-\end{center}
-\end{figure}
-
-\begin{figure*}[htbp]
-\begin{center}
-\includegraphics[width=18cm]{img/sameas}
-\caption{When two instances represent the same entity, users can drag one on top of the other, and they are combined.}
-\label{fig:sameas}
-\end{center}
-\end{figure*}
+%% Design process
 

### 2012-09-15 chi2012.tex
webbox -> datapalette

@@
 Each task had an A and B condition.
-The A condition, the participant could only use WebBox to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or workbook in Microsoft Excel containing the data taken from the data sources to make their choice on individual spreadsheets.
+The A condition, the participant could only use DataPalette to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or workbook in Microsoft Excel containing the data taken from the data sources to make their choice on individual spreadsheets.
 
@@
 In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.
-In order to train the user on how to use our interface, they viewed a five-minute WebBox tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
+In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
 
@@
 
-In the A condition, where participants used WebBox to complete their tasks they used it to collate the data on-screen, and compare multiple values directly.
-About half of the observed users actually used the instance combination feature of WebBox, with the other half able to successfully make decisions without using it.
+In the A condition, where participants used DataPalette to complete their tasks they used it to collate the data on-screen, and compare multiple values directly.
+About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it.
 Those who tried to use it were largely successful, and were able to therefore use the global brush-highlighting to simultaneously compare different statistics.
@@
 
-In the exit survey, the users noted that they would use WebBox for finding a property to rent/buy, choosing a job and purchasing electronic devices, in the future.
+In the exit survey, the users noted that they would use DataPalette for finding a property to rent/buy, choosing a job and purchasing electronic devices, in the future.
 However, one participant said that they would not use it for things like deciding on a restaurant because ``it would cheapen the experience'' by removing the spontaneity, but would use it for making more important decisions.
@@
 directly compare candidates at once. Our pool of participants exhibited both behaviours evenly, however in the condition
-when using the WebBox interface, significantly more comparison (and visualisation of values) was performed, with fewer
+when using the DataPalette interface, significantly more comparison (and visualisation of values) was performed, with fewer
 participants using the ``ruling out'' process. A consequence of this was that when asked to justify their choices, users were
@@
 mixed success  -- many participants had to look through the data sources to re-find the reasons why they has ruled out some choices.
-Therefore, the uses that were using WebBox were more confident in their choices, and knew exactly why they had chosen them,
+Therefore, the uses that were using DataPalette were more confident in their choices, and knew exactly why they had chosen them,
 often pointing to the screens to show and verify each choice as they explained them.

### 2012-09-15 chi2012.tex
working on brushing

@@
 
-\subsubsection{Coreference management: Drag-and-Drop Same-As}
+\subsubsection{Coreference consolidation: Drag-and-Drop Same-As}
 
@@
 
-Related to the concept of co-reference, we extended the drag-and-drop sameas support to also be used to consolidate inconsistant enumerated values of properties, such as ``Casual dining'' vs ``Relaxed'' so that they can be uniformly compared.
-This allows the creation of meaningful, de-duplicated histograms and other visualisations.
+Related to the concept of co-reference, PS2 revealed that enumerated values of properties were often inconsistent, e.g.
+``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property.
+To reoncile such differences, so that they can be uniformly compared, we extended the drag and drop support of co-reference combination to support enumerated literal values as well,  This effectively allowed de-duplicated histograms and other visualisations to be made.
 

### 2012-09-15 chi2012.tex
working on brushing

@@
 
-Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, ids used by each network to identify particular individuals were system-specific.
+Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, identifiers used by each network were system-specific and inconsistent.
 Thus, we sought to make instead make it user-driven, and easy to do manually.
 
-In order to remain consistent with the drag-and-drop interactions used throughout the interface, we decided that simply dragging an instance onto another instance was a likely intuitive way for people to specify that two things should be combined.
+To remain consistent with the drag-and-drop interactions used throughout the interface, we introduced the ability to simply drag one instance on top of another to specify that two things should be combined.
 An illustration of this interaction is provided in Figure \ref{fig:sameas}.
@@
 
-Related to the concept of co-reference, PS2 revealed that enumerated values of properties were often inconsistent, e.g.
+Related to the concept of co-reference, PS.2 revealed that enumerated values of properties were often inconsistent, e.g.
 ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property.

### 2012-09-15 chi2012.tex
working on brushing

@@
 An illustration of this interaction is provided in Figure \ref{fig:sameas}.
-Although general, this can be tedious for groups of items; to support more efficient combination, we extended this to allow entire groups to be dropped on other groups.
+Although general, this can be tedious for groups of items.
+Thus, to support more efficient combination, we extended this to allow entire groups to be dropped on other groups.
 When this is done, DataPalette attempts to identify matching pairs among items in the dragged and target groups by comparing each pairs' labels.

### 2012-09-15 chi2012.tex
working on brushing

@@
 To remain consistent with the drag-and-drop interactions used throughout the interface, we introduced the ability to simply drag one instance on top of another to specify that two things should be combined.
+This action also caused the \emph{same-as display} in the lower left of the screen to display this new combination relationship.
+The display also acted as a mechanism to delete/undo combinations, which reverted representations of instances back to their original, separate forms.
 An illustration of this interaction is provided in Figure \ref{fig:sameas}.
@@
 
-By having co-reference consolidation user-driven, we gave the user freedom to combine things when suited to his or her task.
+By making such consolidation user-driven, we gave the user freedom to combine things when it suited him or her.
 As we saw in the system evaluation, while in certain circumstances users preferred combining representations, in others people preferred to keep things separate.

### 2012-09-15 chi2012.tex
done with interaction design section

@@
 
-\subsubsection{Coreference consolidation: Drag-and-Drop Same-As}
+\subsubsection{Coreference consolidation: Drag-and-Drop Same-as}
 
@@
 ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property.
-To reoncile such differences, so that they can be uniformly compared, we extended the drag and drop support of co-reference combination to support enumerated literal values as well,  This effectively allowed de-duplicated histograms and other visualisations to be made.
+To reoncile such differences, so that items from corresponding datasources can be effectively compared, we extended the drag and drop support of co-reference combination to support enumerated literal values as well.
+
+\subsubsection{``Do What I Mean'' Visualisation}
+
+To make it easier for people to quickly visually compare aspects of instances, we created a charting tool and integrated a mapping feature.
+To make these tools suitable for fast use, e.g.
+data exploration, we designed these tools to automatically configure display parameters instances or groups of were dropped on them, rather than making users manually specify them.
+In the current prototype, the DWIM behaviour of the chart automatically sets whether the plot is a histogram representation (counts of values) or numeric value display based upon the types of the dereferenced values being plotted; values that are non-numeric are counted, while numeric values are displayed.
+Similarly, the map was made to detect address strings and latitude/longitude pairs, geocoding them where appropriate, and determining the optimal bounding box to display all items at maximum zoom.
 
 \subsubsection{Model-based brushing}
-
-Since it can get confusing having an instance represented in multiple groups,
+Since it can get confusing having a single instance represented across several groups, we provided \emph{universal brushing} across the interface; hovering over any representation of any instance (an instance block, a map marker or a histogram bar), causes all other visible representations of that instance to lightly glow, so that the user can quickly identify all the places that a single instance is represented
 
@@
 
-\subsubsection{Adaptive visualisation modes}
+\subsection{Summary of interaction capabilities}
 
-\subsection{Summary of interaction features}
+%% TODO
 

### 2012-09-15 chi2012.tex
happy

@@
 
-In this section, we touch on motivating work for DataPalette from three fields: peronsal information management, data integration, and end-user mashups and toolkits.
+In this section, we touch on motivating work for DataPalette from three fields: personal information management, data integration, and end-user mashups and toolkits.
 
@@
 In the remaining part sof this section, we summarise our methodologies and results for each of these two pre-studies.
-
-%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
-In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one.
-This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted.
-The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined.
-In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.
-Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
 
@@
 This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
-
+[0
 In the following sections we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.
@@
 
-Typically users are interested in different \emph{properties} of an instance, for example, a user is interested in a product and its price, rating, manufacturer and so on.
+Typically, users are interested in viewing and comparing the \emph{properties} of instances; for example, a user might be keen to examine a product's price, rating, manufacturer and so on.
 An instance's properties are displayed beneath it; selecting a property creates a \emph{path selection}, which dereferences all of the instances in a group that contain the selected property and displays their corresponding values.
@@
 ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property.
-To reoncile such differences, so that items from corresponding datasources can be effectively compared, we extended the drag and drop support of co-reference combination to support enumerated literal values as well.
+To reconcile and consolidate corresponding values from diffeent  datasources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.
 
@@
 
-\subsection{Summary of interaction capabilities}
+%% 
+%% \subsection{Summary of interaction capabilities}
+%% \begin{itemize}
+%% \item \emph{Group multi-path value selection} - Selecting any of the %% properties of an instance in a group (window) causes all of the rest %% of the instances that have that property in that group to be %% dereferenced along the same property, displaying the resulting value %% underneath the instance label.
+Selecting a further property of the %% resulting value further dereferences this value.
+%% \item \emph{Drag-and-drop coreference / value combination} - In order %% to specify that two instances correspond to the same entity, they can %% be simly dragged together
+%% \item \emph{Group-based Drag-and-drop matching} - In order to specify %% that two instances correspond to the same entity, they can be simly %% dragged together
+%% \item \emph{Universal model-based brushing} -
+%% \end{itemize}
 
-%% TODO
+\section{Evaluation}
 
-\begin{itemize}
-\item \emph{Group multi-path value selection} - Selecting any of the properties of an instance in a group (window) causes all of the rest of the instances that have that property in that group to be dereferenced along the same property, displaying the resulting value underneath the instance label.
-Selecting a further property of the resulting value further dereferences this value.
-\item \emph{Drag-and-drop coreference / value combination} - In order to specify that two instances correspond to the same entity, they can be simly dragged together
-\item \emph{Group-based Drag-and-drop matching} - In order to specify that two instances correspond to the same entity, they can be simly dragged together
-\item \emph{Universal model-based brushing} -
-\end{itemize}
-
-%% Interaction features 
-%% 	multiple paths selection
-%% 	sameas interactions
-%% 	visualisation of partial results
-
-%% Design process
-
-\section{Methodology}
+In order to determine whether the interaction affordances introduced in DataPalette would effectively enable serendipitous data mixing by end-users, we devised a within-subjects A-B trial lab study which we describe next.
 
 \subsection{Participants}
-We recruited participants by calling for participant on the University of Southampton's international student mailing list.
-There is no guarantee of the mailing list's recipients first language, however all of them passed the university English proficiency test.
-We required that all participants had encountered tasks that required them to use and evaluate data from multiple websites.
+We recruited participants through adverts posted near the University campus and e-mails to student mailing lists.
+We screened participants to be at least 18 years of age, but did not filter or preferentially select particpants based on any other criteria.
+We took the first 20 that signed up and that successfully specified time slots that were available.
+As described below in results, we ended up getting a large number of international students, who had varying proficiency in English.
+Since all such participants had passed the requisite proficiency tests to enrol at University, we did not think this would substantially impact results.
 
@@
 
+\subsection{}
+
+- data integration literature demands 'correct' but perfect is the enemy of the good.
+
+%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
+In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one.
+This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted.
+The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined.
+In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.
+Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
+
 %What we found 

### 2012-09-16 chi2012.tex
meh method

@@
 \includegraphics[width=18cm]{img/sameas}
-\caption{When two instances represent the same entity, users can drag one on top of the other, and they are combined.}
+\caption{\emph{Coreference consolidation with drag-and-drop Same-As} - When two instances represent the same entity, users can drag one on top of the other, and they are combined - effectively displaying the union of properties of both source entities.
+This can be un-done by deleting the Same-As relaitonship in the view in the lower-left corner of the workspace.}
 \label{fig:sameas}
@@
 
-%% 
+%% it would be nice if we had space to summarize but i fear we don't
+
 %% \subsection{Summary of interaction capabilities}
@@
 
-In order to determine whether the interaction affordances introduced in DataPalette would effectively enable serendipitous data mixing by end-users, we devised a within-subjects A-B trial lab study which we describe next.
+In order to determine whether the interaction affordances introduced in DataPalette (henceforth DP) allow users to perform serendipitous data mixing on real data feeds by end-users, we devised a within-subjects A-B trial lab study which we describe next.
+
+\subsection{Methodology}
+
+We first describe our methodology, starting with the three hypotheses we set out to test.
+These hypothesis pertain to the usability of DP by end-users (h1), whether DP achieves its stated goal of enabling end-users to integrate data (h2), and finally if it has an effect on task performance (h3) - either in how well people perform a task, how quickly they perform it, or if it reduces task effort, as follows:
+
+\begin{itemize}
+\item \emph{h1 -} People understand how to use DP.
+\item \emph{h2 -} People can effectively mix heterogeneous data with DP.
+\item \emph{h3 -} Use of DP improves or facilitates task performance
+\end{itemize}
+
+%% Interaction-method specific hypotheses
+%% \begin{itemize}
+%% 	\item h1.1 - Do people understanding the problem of %% co-reference, and the need to reconcile coreference problems?
+%% 	\item h1.2. Does the ability to do drag + drop combination for %% co-reference reconciliation effectively solve this problem?
+%% 	\item h2.1- Do people understand the problem of structural %% differences in data?
+%% 	\item h2.2 - Does multipath selection let people effectively %% deal with this -- and work with collections of heterogeneous items?
+%% \end{itemize}
+%% Are these interaction methods sufficient to perform common tasks %% involving heterogeneous data sources?  Does supporting these simple %% techniques facilitate task completion?
+
+\subsubsection{Study design}
+To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two of four possible predefined tasks (depending on condition), with either DP (``condition A'') or a baseline interface (``condition B'') in 2 separate timed trials.
+Each participant was asked to use DP for one of the trials and the baseline interface for the other; conditions were fully counterbalanced among particiants to eliminate inadvertant ordering and task bias.
+
+The control (B) conditions were in a prepared Mac comprised of a web browser pre-loaded with a page with links to four live web sites, and an open instance of Excel with a workbook open containing spreadsheet versions of the same data as were loaded into DP for the A condition versions of the same tasks.
+Other Office applications were also available for use and participants were allowed to visit other sites than the prespecified ones.
+The purpose of this condition was to understand how people using standard desktop office software
+
+The tasks were to select a restaurant for an event or university at which to study, out of six possible choices  around four possible tasks shown in Table \ref{tab:studyfactors}.
+
+We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choices.
+The shortlist of universities and restaurants were taken from the top six ranked from the Complete University Guide (for task 1 and 2), Tripadvisor (for task 3) and Yelp (for task 4).
+The top six were chosen from these websites because they provide a realistic shortlist and are comparable in terms of rankings and properties.
+The web sites used as data sources were chosen because they are the most popular and contain comparable statistics for universities and restaurants.
+
+\subsection{Data Collection}
+Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
+During the lab study, we recorded the audio and the participant's actions on screen.
+We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.
+At the end of the study the participant completed an exit survey.
+
+In each trials, participants were given a paper notepad and pen to use if they wished.
+Screen capture software recorded all interactions with applications.
+Participants were asked to follow a \emph{think-aloud} protocol as they performed each task, and audio recordings (synchronized with the screen recordings) were taken.
+
+% This goes in results: The studies lasted 40 minutes, on average.  
+Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn't influence the study.
+The studies were run over 4 days.
+At the end of each day the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.
 
@@
 \subsection{Task Design}
-We designed a semi-structured lab-based study around four possible tasks shown in Table \ref{tab:studyfactors}.
-The tasks were to select a restaurant for an event or university at which to study, out of six possible choices.
-We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choices.
-The shortlist of universities and restaurants were taken from the top six ranked from the Complete University Guide (for task 1 and 2), Tripadvisor (for task 3) and Yelp (for task 4).
-The top six were chosen from these websites because they provide a realistic shortlist and are comparable in terms of rankings and properties.
-The web sites used as data sources were chosen because they are the most popular and contain comparable statistics for universities and restaurants.
 
@@
 
-\subsection{Data Collection}
-Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
-During the lab study, we recorded the audio and the participant's actions on screen.
-We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.
-At the end of the study the participant completed an exit survey.
-
-The studies lasted 40 minutes, on average.
-Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn't influence the study.
-The studies were run over 4 days.
-At the end of each day the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.
-
 \subsection{Data Analysis}
@@
 TODO look further into types of data analysis....
-
-\subsection{Hypotheses}
-Interaction-method specific hypotheses
-\begin{itemize}
-	\item h1.1 - Do people understanding the problem of co-reference, and the need to reconcile coreference problems?
-	\item h1.2. Does the ability to do drag + drop combination for co-reference reconciliation effectively solve this problem?
-	\item h2.1- Do people understand the problem of structural differences in data?
-	\item h2.2 - Does multipath selection let people effectively deal with this -- and work with collections of heterogeneous items?
-\end{itemize}
-Are these interaction methods sufficient to perform common tasks involving heterogeneous data sources?  Does supporting these simple techniques facilitate task completion?
 
@@
 mixed success  -- many participants had to look through the data sources to re-find the reasons why they has ruled out some choices.
+
 Therefore, the uses that were using DataPalette were more confident in their choices, and knew exactly why they had chosen them,
@@
 - data integration literature demands 'correct' but perfect is the enemy of the good.
-
-%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
-In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one.
-This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted.
-The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined.
-In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.
-Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
 
@@
 \section{Conclusion} %Key findings of the whole paper and its implications
+
+% ``perfect'' be the enemy of the good
+
+%% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
+In database systems, the overall objective is usually to allow multiple large databases to be accessed as if one.
+This requires essentially full mapping of all data types and objects and well-defined semantics concerning where and how data gets converted.
+The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined.
+In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.
+Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
+
+Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analyists the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{}} into the hands of the ordinary citizen \cite{datajournalismhandbook}.
+In order for this to happen, however, tools that allow people to effectively combine data more effectively than current structured PIM tools and generic spreadsheet tools will become increasingly important.
 

### 2012-09-16 chi2012.tex
some comments

@@
 
-\begin{itemize}
-\item \emph{h1 -} People understand how to use DP.
-\item \emph{h2 -} People can effectively mix heterogeneous data with DP.
-\item \emph{h3 -} Use of DP improves or facilitates task performance
-\end{itemize}
+% heather: instead of these low-level hypotheses, i've made higher level ones that we can explicitly address in the disucssion section - the more hypotheses we have, the more we have to come back to them; since we're short on space we might as well keep it simple!
 
@@
 
-\subsubsection{Study design}
+\begin{itemize}
+\item \emph{h1 -} People understand how to use DP.
+\item \emph{h2 -} People can effectively mix heterogeneous data with DP.
+\item \emph{h3 -} Use of DP improves or facilitates task performance
+\end{itemize}
+
+\subsubsection{Study and task design}
 To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two of four possible predefined tasks (depending on condition), with either DP (``condition A'') or a baseline interface (``condition B'') in 2 separate timed trials.
@@
 
-The control (B) conditions were in a prepared Mac comprised of a web browser pre-loaded with a page with links to four live web sites, and an open instance of Excel with a workbook open containing spreadsheet versions of the same data as were loaded into DP for the A condition versions of the same tasks.
-Other Office applications were also available for use and participants were allowed to visit other sites than the prespecified ones.
-The purpose of this condition was to understand how people using standard desktop office software
-
+% why? why were these tasks selected? can we drop the phrase ``multi-dimensional subjective choice task'' and refer back to PS.1 to justify the choice of tasks?
 The tasks were to select a restaurant for an event or university at which to study, out of six possible choices  around four possible tasks shown in Table \ref{tab:studyfactors}.
 
+\subsubsection{Task preparation}
+% how data was prepared for each of the tasks; exactly how the A condition was done (full screen DP in a web browser, no outside access to apps),  B condition and why
 We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choices.
@@
 
+% why was the B condition set up this way? What did we seek to achieve?
+The control (B) conditions were in a prepared Mac comprised of a web browser pre-loaded with a page with links to four live web sites, and an open instance of Excel with a workbook open containing spreadsheet versions of the same data as were loaded into DP for the A condition versions of the same tasks.
+Other Office applications were also available for use and participants were allowed to visit other sites than the prespecified ones.
+The purpose of this condition was to understand how people using standard desktop office software
+
 \subsection{Data Collection}
@@
 During the lab study, we recorded the audio and the participant's actions on screen.
-We asked the participants to speak out loud their thought process and what they were looking at, so that we could evaluate their reasoning behind their processes.
-At the end of the study the participant completed an exit survey.
+We asked the participants to follow a \emph{think-aloud} protocol as they worked, so that we could understand the reasoning behind their actions.
+At the end of the study, the participant completed an exit survey.
 
-In each trials, participants were given a paper notepad and pen to use if they wished.
+% In each trials, participants were given a paper notepad and pen to use if they wished.
 Screen capture software recorded all interactions with applications.
@@
 
-% This goes in results: The studies lasted 40 minutes, on average.  
+% This goes in results: 
+% The studies lasted 40 minutes, on average.  
 Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn't influence the study.
@@
 
-\subsection{Participants}
+\subsection{Participant recruitement}
 We recruited participants through adverts posted near the University campus and e-mails to student mailing lists.

### 2012-09-16 chi2012.tex
editted methodology section

@@
 All participants reported having used multiple websites to accomplish a task.
-They listed example tasks such as shopping (including food, consumer electronics, hotels and flights), searching for jobs, choosing recipes, study and work.
+They listed example tasks such as shopping (including food, consumer electronics, hotels and flights), choosing restaurants, searching for jobs, when choosing a university, choosing recipes, study and work.
 The biggest focus for shopping focused tasks was balancing price, quality and speed of delivery.
@@
 \item \emph{h2 -} People can effectively mix heterogeneous data with DP.
-\item \emph{h3 -} Use of DP improves or facilitates task performance
+\item \emph{h3 -} Use of DP improves or facilitates task performance.
 \end{itemize}
@@
 \subsubsection{Study and task design}
-To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two of four possible predefined tasks (depending on condition), with either DP (``condition A'') or a baseline interface (``condition B'') in 2 separate timed trials.
-Each participant was asked to use DP for one of the trials and the baseline interface for the other; conditions were fully counterbalanced among particiants to eliminate inadvertant ordering and task bias.
+To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials.
+The participants asked to use either our DP (``condition A'') or a baseline interface (``condition B'') to complete a task, and we used two possible dataset for each type of task.
+We used an A and B condition so that we could compare how users performed their tasks using our interface and the standard approach.
+We also used two datasets for each of the tasks, so that we evaluate whether the participant's performance of completing a task differed with different datasets.
+Each participant was asked to use DP for one of the trials and the baseline interface for the other; conditions were fully counterbalanced among participants to eliminate inadvertent ordering and task bias.
 
 % why? why were these tasks selected? can we drop the phrase ``multi-dimensional subjective choice task'' and refer back to PS.1 to justify the choice of tasks?
-The tasks were to select a restaurant for an event or university at which to study, out of six possible choices  around four possible tasks shown in Table \ref{tab:studyfactors}.
+We based our two tasks on these scenarios on choosing a university and picking a restaurant for a large social event, because in PS.1 the majority of people stated that they had used multiple websites for these tasks.
+We used different two different variables for both tasks, where the restaurant was located (Glasgow and Cambridge) and which subject people choose (sport science and history).
+The cities Glasgow and Cambridge were chosen because they are not geographically close to the participants, they are two heavily populated areas that have a choice of restaurants, and they contained detail data about the same restaurants on various review sites.
+When deciding on which cities to use, we found that different areas favoured different review websites and places such as Hull had sets of disjoint sets of restaurant reviews.
+We selected the top restaurants for on Tripadvisor and Yelp, because they were the top two websites people quoted for comparing reviews when travelling.
+We selected the two courses for the university task at random from a list taken form the complete university guide\footnote{}, and selected the top six university for that subject because they provide a realistic shortlist and are comparable in terms of rankings and properties.
+We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choicesex
 
@@
 % how data was prepared for each of the tasks; exactly how the A condition was done (full screen DP in a web browser, no outside access to apps),  B condition and why
-We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choices.
-The shortlist of universities and restaurants were taken from the top six ranked from the Complete University Guide (for task 1 and 2), Tripadvisor (for task 3) and Yelp (for task 4).
-The top six were chosen from these websites because they provide a realistic shortlist and are comparable in terms of rankings and properties.
-The web sites used as data sources were chosen because they are the most popular and contain comparable statistics for universities and restaurants.
+In order to prepare the data for our tasks, we identified websites that people commonly used for selecting universities and restaurants.
+Once we had compiled the list of websites, we select the top most used six for each task.
+We selected six because it enable us to scope our study, we found through two pilot studies that six data sources provided a sufficient challenge for users to explore the data and required them to use DP's main features.
+We manually collected the data from the nine selected websites into jason files, and converted them to Microsoft Excel worksheets (which 40\% of the participants in PS.1 said that they used to compare data in).
 
@@
 The control (B) conditions were in a prepared Mac comprised of a web browser pre-loaded with a page with links to four live web sites, and an open instance of Excel with a workbook open containing spreadsheet versions of the same data as were loaded into DP for the A condition versions of the same tasks.
-Other Office applications were also available for use and participants were allowed to visit other sites than the prespecified ones.
-The purpose of this condition was to understand how people using standard desktop office software
+Other Office applications were also available for use and participants were allowed to visit other sites than the pre-specified ones.
+The purpose of this condition was to understand how people using standard desktop office software.
 
@@
 Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
+Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn't influence the study.
+Explicitly both the facilitator and observer where trained in identifying processes that people used during a task, so that they could thematically categorise them.
+Initially, we identified the following themes: organisation of data, eliminating candidates, co-referencing, and visualisations.
+The studies were run over 4 days.
+At the end of each day the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.
 During the lab study, we recorded the audio and the participant's actions on screen.
@@
 
-% In each trials, participants were given a paper notepad and pen to use if they wished.
-Screen capture software recorded all interactions with applications.
-Participants were asked to follow a \emph{think-aloud} protocol as they performed each task, and audio recordings (synchronized with the screen recordings) were taken.
-
-% This goes in results: 
-% The studies lasted 40 minutes, on average.  
-Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn't influence the study.
-The studies were run over 4 days.
-At the end of each day the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.
-
-\subsection{Participant recruitement}
+\subsection{Participant Recruitement}
 We recruited participants through adverts posted near the University campus and e-mails to student mailing lists.
-We screened participants to be at least 18 years of age, but did not filter or preferentially select particpants based on any other criteria.
+We screened participants to be at least 18 years of age, but did not filter or preferentially select participants based on any other criteria.
 We took the first 20 that signed up and that successfully specified time slots that were available.
@@
 
-\subsection{Task Design}
+\subsection{Experimental Setup}
 
-\begin{table*}[t]
-\begin{center}
-\small
-\begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
-\hline
-Task	 &Choices	&Data Sources\\
-\hline
-1) Three universities they would apply to study Sport Science & Universities: 1) Loughborough,  2) Durham,  3) Exeter, 4) Edinburgh, 5) Birmingham, and 6) Bath & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
-\hline
-2) Three universities they would apply to study History & Universities: 1) Cambridge, 2) London School of Economics and Political Science, 3) Durham, 4) Oxford, 5) St Andrews, and 6) University College London & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
-\hline
-3) Restaurant they would book for 12 friends in Cambridge&Restaurants : 1) Alimentum, 2) Gardenia Restaurant, 3) The Cambridge Chop House, 4) Al Casbah, 5) Midsummer House Restaurant, and 6) Nando's Restaurant.
-& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
-\hline
-4) Restaurant they would book for 12 friends in Glagow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollopes Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia.
-& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
-\hline
-\end{tabular}
-\end{center}
-\caption{Tasks, choices and data sources} \label{tab:studyfactors}
-\normalsize
-\end{table*}%
-
-Each task had an A and B condition.
-The A condition, the participant could only use DataPalette to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or workbook in Microsoft Excel containing the data taken from the data sources to make their choice on individual spreadsheets.
-
-Each study consisted of two tasks, one A and one B condition.
+The participants were given a standard desktop PC running MacOS X on a 22" monitor with a standard keyboard and mouse.
+During the studies the Excel workbook and DP were initially set to full screen by default.
+Each participant was asked to complete two tasks, one A and one B condition.
 Each participant was given 10 minutes to complete a task.
@@
 In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
+
+%Each task had an A and B condition.
+The A condition, the participant could only use DataPalette to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or workbook in Microsoft Excel containing the data taken from the data sources to make their choice on individual spreadsheets.
+
+%Each study consisted of two tasks, one A and one B condition.
+Each participant was given 10 minutes to complete a task.
+In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.
+In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
+
+%\begin{table*}[t]
+%\begin{center}
+%\small
+%\begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
+%\hline
+%Task	 &Choices	&Data Sources\\
+%\hline
+%1) Three universities they would apply to study Sport Science & Universities: 1) Loughborough,  2) Durham,  3) Exeter, 4) Edinburgh, 5) Birmingham, and 6) Bath & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
+%\hline
+%2) Three universities they would apply to study History & Universities: 1) Cambridge, 2) London School of Economics and Political Science, 3) Durham, 4) Oxford, 5) St Andrews, and 6) University College London & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
+%\hline
+%3) Restaurant they would book for 12 friends in Cambridge&Restaurants : 1) Alimentum, 2) Gardenia Restaurant, 3) The Cambridge Chop House, 4) Al Casbah, 5) Midsummer House Restaurant, and 6) Nando's Restaurant.
+& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
+%\hline
+%4) Restaurant they would book for 12 friends in Glagow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollopes Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia.
+& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
+%\hline
+%\end{tabular}
+%\end{center}
+%\caption{Tasks, choices and data sources} \label{tab:studyfactors}
+%\normalsize
+%\end{table*}%
 
@@
 Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.
+On average each study took 40 mins to complete the tasks and exit surveys.
 

### 2012-09-16 chi2012.tex
strengthening first para of task design

@@
 \subsubsection{Study and task design}
-To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials.
-The participants asked to use either our DP (``condition A'') or a baseline interface (``condition B'') to complete a task, and we used two possible dataset for each type of task.
-We used an A and B condition so that we could compare how users performed their tasks using our interface and the standard approach.
-We also used two datasets for each of the tasks, so that we evaluate whether the participant's performance of completing a task differed with different datasets.
-Each participant was asked to use DP for one of the trials and the baseline interface for the other; conditions were fully counterbalanced among participants to eliminate inadvertent ordering and task bias.
+To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'' which we describe below).
+We used upon the within-subjects A-B format in order to make both quantiative measurements about an individual's performance with DP (relative to baseline), but also to make qualitative observations of the processes used by each participant as they solved each task.
+Each task was paired with one of two datasets, depending on condition, so that we could evaluate whether the participant's performance of completing a task differed with different datasets.
+Conditions and interface order were fully counterbalanced among participants to eliminate inadvertent ordering and task bias.
 
-% why? why were these tasks selected? can we drop the phrase ``multi-dimensional subjective choice task'' and refer back to PS.1 to justify the choice of tasks?
+% why? why were these tasks selected? can we drop in the phrase ``multi-dimensional subjective choice task'' and refer back to PS.1 to justify the choice of tasks?
 We based our two tasks on these scenarios on choosing a university and picking a restaurant for a large social event, because in PS.1 the majority of people stated that they had used multiple websites for these tasks.

### 2012-09-17 chi2012.tex
some rsults

@@
 Each task was paired with one of two datasets, depending on condition, so that we could evaluate whether the participant's performance of completing a task differed with different datasets.
-Conditions and interface order were fully counterbalanced among participants to eliminate inadvertent ordering and task bias.
+Conditions and interface order were fully counterbalanced among participants to eliminate ordering and task bias.
 
@@
 
-\subsection{Participant Recruitement}
+\subsection{Participant Recruitment}
 We recruited participants through adverts posted near the University campus and e-mails to student mailing lists.
@@
 
+Participants were offered a small gift certificate from an online retailer in exchange for their time.
+
 \subsection{Experimental Setup}
-
 The participants were given a standard desktop PC running MacOS X on a 22" monitor with a standard keyboard and mouse.
@@
 Each participant was asked to complete two tasks, one A and one B condition.
-Each participant was given 10 minutes to complete a task.
+Each participant was given a maximum of 10 minutes to complete a task, but were told that if they finished earlier they could tell us and move on to the next condition.
 In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.
@@
 
-We recruited 20 participants (10 female) from the University of Southampton's international student mailing list.
-All of the participants spoke English, and for 12 of them, English was not their first language.
-Predominately the participants were studying various levels of Computer Science, and 5 of which studied other subjects.
-On average each study took 40 mins to complete the tasks and exit surveys.
+Twenty-six (26) participants responded, from which we selected the first 20.
+Ten participants were female.
+All of the participants spoke English, but 12 of them were non-native English speakers.
+Seven were non-students; out of the remaining 13, eight studied computer science and 5 studied other subjects.
+On average, each study took 40 mins to complete.
 
-The information in Table \ref{tbl:data} was captured during the studies, so that we could statistically analyse the features used by the participants.
+The information in Table \ref{tbl:data} summarizes the metrics that were captured of the trials for each participant. We analyze this data next.
 
@@
 maps used &1&1&2&2&2&1&1&0&1&0&1&1&2&1&0&1&0&0&2&1\\ 
-same as &4&3&3&6&0&2&0&2&1&0&6&5&6&2&7&0&6&7&7&7\\
-number data sources  &4&3&4&4&4&3&4&3&4&4&3&3&5&4&5&4&4&5&5&4\\
+same as used &4&3&3&6&0&2&0&2&1&0&6&5&6&2&7&0&6&7&7&7\\
+\# data sources used &4&3&4&4&4&3&4&3&4&4&3&3&5&4&5&4&4&5&5&4\\
 chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, \newline 1.3, \newline 1.2& 4.2& 1.4, \newline 1.5, \newline 1.3& 3.2& 2.1, \newline 2.3, \newline 2.6& 3.5, \newline 3.3&2.6, \newline 2.3, \newline 2.5&1.6, \newline 1.5, \newline 1.3&4.1&3.3&2.4, \newline 2.5, \newline 2.6\\
+\# factors in choice &-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-\\
 \hline\hline
@@
  chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4, \newline 1.2, \newline 1.6&2.1&2.3, \newline 2.6, \newline 2.5&3.5&2.2, \newline 2.3, \newline 2.6&3.5&1.1&4.1&1.6&4.2&2.6, \newline 2.3, \newline 2.5&2.3, \newline 2.5, \newline 2.6&1.1, \newline 1.3,  \newline 1.4&3.5\\
+ \# factors in choice &-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-\\
  \hline
@@
 
+\subsection{Task performance}
+
+We used three metris to measure of task performance: time to completion, the number of factors used in the choice, and the number of candidates
+
+\subsubsection{Efficiency: Time taken}
+Participants spent an average of 8.9 minutes performing the tasks overall, with DP (condition A) taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B (the web browser/Excel interface) ($M=9.0$min; $SD=1.44$).
+An ANOVA test of time taken by interface condition blocking by participant did not find a interface use a significant predictor of task completion time.
+A comparison of times taken are visible in Figure \ref{fig:time}.
+
+Although across trials, restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor either.
+
+\subsubsection{Thoroughness: Factors weighed in final choice}
+A second metric was the number of factors the participant considered in making their final choice.
+To determine this, we asked participants to tell us precisely why their choice was made, and matched their justifications to properties in the data.
+This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95, p<0.10$); as visible in Figure \ref{fig:factors}, participants in the DP condition listed an average of 5.2 choices in their decision ($SD=1.7$) versus 4.5 ($SD=1.73$) in the standard condition.
+
+\subsubsection{Diversity: Data sources consulted}
+A third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
+
+\subsubsection{Effort: External cognition}
+Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much participants had to keep track of during the study.
+We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22$, $p\approx0.10$).
+
+\subsection{Condition A (DP)}
 In the A condition, where participants used DataPalette to complete their tasks they used it to collate the data on-screen, and compare multiple values directly.
@@
 
+\subsection{Condition B (control}
 In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.

### 2012-09-17 chi2012.tex
some rsults

@@
 To determine this, we asked participants to tell us precisely why their choice was made, and matched their justifications to properties in the data.
-This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95, p<0.10$); as visible in Figure \ref{fig:factors}, participants in the DP condition listed an average of 5.2 choices in their decision ($SD=1.7$) versus 4.5 ($SD=1.73$) in the standard condition.
+This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95, p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
 
@@
 A third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
+In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04  p\approx0.10$).
 

### 2012-09-17 chi2012.tex
Added hyphenation for datapalette
minor corrections

@@
 This information includes: access to data previously hidden behind services and applications, such as retailer product catalogues; information not previously directly released to the public, such as open government data or records of personal financial transactions; and new kinds of data generated by emerging kinds of data sources, such as wearable and smartphone-based biosensors.
-A primary goal of opening up such direct access to information is that end-user citizens will make more informed decisions pertaining to their personal information, for example health, wealth, and well-being \cite{}.
+A primary goal of opening up this data is that end-user citizens can make more informed decisions pertaining to their health, wealth, and well-being \cite{}.
 This data is predominately used by specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.
@@
 We hypothesise that end-users would benefit from this rich data with tools allowing them to effectively browse, use and consume heterogeneous structured data from diverse sources.
-In this paper, we present our investigation of how extending PIM tools to effectively use emerging ecosystems of personal data impact information practice.
+In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data impact information practice.
 We used a three stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of  information sources, and the processes that they relied upon to perform such tasks.
-Our results suggested that people preferred to rely upon multiple, diverse sources to singular, integrated ones for a number of reasons, including coverage, reliability, and seeking out more suitable alternatives.
+Our results suggested that people preferred to rely upon multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and alternative perspectives.
 
-Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data from each.
+Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data.
 Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
 
-Third, we developed DataPalette, which focuses on the most common types of heterogeneity to enable serendipitous ``data mixing'' - a form of simple integration sufficient to allow people to easily and effectively combine and compare information sources without the need to write code.
+Third, we developed DataPalette at interface that uses the most common types of heterogeneity to enable serendipitous ``data mixing'' - a form of simple integration sufficient to allow people to easily and effectively combine and compare information sources without the need to write code.
 It was principally designed based upon people's data integration needs and to support data heterogeneity observed in the previous studies.
@@
 
-In this section, we touch on motivating work for DataPalette from three fields: personal information management, data integration, and end-user mashups and toolkits.
+In this section, we motivate our work using three fields: PIM, data integration, and end-user mashups and toolkits.
 
-The primary motivation for this work stems from field studies in personal information management, where letting users effectively consolidate and work with data from multiple sources has remained a challenge for over a decade.
-The problem of data fragmentation (\cite{Jones05towardsa}, for example, derives from an inability to effectively consolidate data across multiple applications, services and platforms (e.g., \cite{bergman}, \cite{boardmansasse}).
+Our primary motivation stems from field studies in PIM, where effectively consolidating and working with data from multiple sources has remained a challenge for over a decade.
+The problem of data fragmentation (\cite{Jones05towardsa}, derives from an inability to effectively consolidate data across multiple applications, services and platforms (e.g., \cite{bergman}, \cite{boardmansasse}).
 In our own previously-conducted field study, (reference anonymised) we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage heterogeneous data from disparate sources -- be them online, offline, or other people.
@@
 %% don't like the beginning of this > 
-Enabling PIM tools to supporting ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data hetereogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
+Enabling PIM tools to supporting ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data heterogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
 As described by Alon Halvey:
@@
 \begin{quote} 
-  [The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs.
+  The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs.
 Hence, even if they model overlapping domains, they will model them in different ways.
-Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modeling goal.
+Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modelling goal.
 \cite{alonhalevy}
@@
 The purpose of this follow-on  was to identify not specific characteristics of particular sites in question, but to identify, in general, the degree of complexity of the data feeds currently available across a variety of domains, and typical integration problems that might arise when mixing this data.
-In the remaining part sof this section, we summarise our methodologies and results for each of these two pre-studies.
+In the following sections, we summarise our methodologies and results for each of these two pre-studies.
 
@@
 All of the participants used Twitter to listen to broadcasted tweets, and 6 out of 8 people regularly tweeted.
+The following two sections, describes the tasks people perform online with heterogeneous data and the processes they would use to organise a large social event.
 
@@
 Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from definitions in \cite{george2005understanding} and feature-metrics used in statistical instance alignment algorithms (such as summarised in \cite{bhattacharya2007collective}).
-Such metrics were chosen because they reflected, the differences among the data feeds at multiple levels - from the simple (labelling) differneces to differences in how data was structured, and modeled.
+Such metrics were chosen because they reflected, the differences among the data feeds at multiple levels - from the simple (labelling) differneces to differences in how data was structured, and modelled.
 Each such metric listed above is the result of averaging the pairwise score between sources in each category, separately and defined as follows:
@@
 
-At a high level, the overall goal of DataPalette was to empower end-users to mix arbitrary structured data feeds on-the-fly while solving their particular information task(s) - be them sensemaking or decision-making related - in order to more effectively accomplish these tasks.
+At a high level, the overall goal of DataPalette was to empower end-users to mix arbitrary structured data feeds on-the-fly while solving their particular information task(s) - be them sense-making or decision-making related - in order to more effectively accomplish these tasks.
 Unlike Mashup makers, including Yahoo Pipes!, the goal was to facilitate such integration without a need for an explicit separate mapping or integration step, which would detract from the overall usability or usefulness of the system by making people do something that did not immediately contribute to their resolving their need.

### 2012-09-17 chi2012.tex
happy times

--- 2012-09-17 chi2012.tex
@@
 Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from definitions in \cite{george2005understanding} and feature-metrics used in statistical instance alignment algorithms (such as summarised in \cite{bhattacharya2007collective}).
-Such metrics were chosen because they reflected, the differences among the data feeds at multiple levels - from the simple (labelling) differneces to differences in how data was structured, and modeled.
+Such metrics were chosen because they reflected the differences among data feeds at multiple levels -  the terminological (use of names/labels), structural level, and modelling level.
 Each such metric listed above is the result of averaging the pairwise score between sources in each category, separately and defined as follows:
@@
 To determine this, we asked participants to tell us precisely why their choice was made, and matched their justifications to properties in the data.
-This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95, p<0.10$); as visible in Figure \ref{fig:factors}, participants in the DP condition listed an average of 5.2 choices in their decision ($SD=1.7$) versus 4.5 ($SD=1.73$) in the standard condition.
+This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
 
@@
 A third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
+In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04$ $p\approx0.10$).
 
@@
 Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much participants had to keep track of during the study.
-We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22$, $p\approx0.10$).
+We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22; p\approx0.10$).
 

### 2012-09-17 chi2012.tex
blah

--- 2012-09-17 chi2012.tex
@@
 %\pagenumbering{arabic}
+
+\hyphenation{Data-Palette}
 
@@
 Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from definitions in \cite{george2005understanding} and feature-metrics used in statistical instance alignment algorithms (such as summarised in \cite{bhattacharya2007collective}).
-Such metrics were chosen because they reflected, the differences among the data feeds at multiple levels - from the simple (labelling) differneces to differences in how data was structured, and modelled.
+Such metrics were chosen because they reflected the differences among data feeds at multiple levels -  the terminological (use of names/labels), structural, and modelling levels.
 Each such metric listed above is the result of averaging the pairwise score between sources in each category, separately and defined as follows:
@@
 To determine this, we asked participants to tell us precisely why their choice was made, and matched their justifications to properties in the data.
-This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95, p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
+This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
 
@@
 A third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
-In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04  p\approx0.10$).
+In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04$ $p\approx0.10$).
 

### 2012-09-17 chi2012.tex
minor edits

--- 2012-09-17 chi2012.tex
@@
 
-The two prestudies provided valuable insight towards this goal that allowed us to guide our design process.
-The first prestudy confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
-The second prestudy revealed that the most common inconisistencies that occurred in the APIs and data feeds available through from such data sources consisted predominantly of the simpler kinds of heterogeneity - terminological and simpler forms of structural heterogeneity, rather than the delicate semantic/modeling variety, which were comparatively rare.
+The two pre-studies provided valuable insight towards this goal that allowed us to guide our design process.
+The first pre-study confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
+The second pre-study revealed that the most common inconsistencies that occurred in the APIs and data feeds available through from such data sources consisted predominantly of the simpler kinds of heterogeneity - terminological and simpler forms of structural heterogeneity, rather than the delicate semantic/modelling variety, which were comparatively rare.
 This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
@@
 
-\item Supporting dereferencing along different paths \emph{by structure} allows instances with different property names, and even sub-structures to be quickly consolidated by separate dereferencing, thereby effectively coping with terminological and structural hetereogeneity.
+\item Supporting dereferencing along different paths \emph{by structure} allows instances with different property names, and even sub-structures to be quickly consolidated by separate dereferencing, thereby effectively coping with terminological and structural heterogeneity.
 \end{enumerate}
@@
 \caption{\emph{Coreference consolidation with drag-and-drop Same-As} - When two instances represent the same entity, users can drag one on top of the other, and they are combined - effectively displaying the union of properties of both source entities.
-This can be un-done by deleting the Same-As relaitonship in the view in the lower-left corner of the workspace.}
+This can be un-done by deleting the Same-As relationship in the view in the lower-left corner of the workspace.}
 \label{fig:sameas}
@@
 ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property.
-To reconcile and consolidate corresponding values from diffeent  datasources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.
+To reconcile and consolidate corresponding values from different  datasources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.
 
@@
 To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'' which we describe below).
-We used upon the within-subjects A-B format in order to make both quantiative measurements about an individual's performance with DP (relative to baseline), but also to make qualitative observations of the processes used by each participant as they solved each task.
+We used upon the within-subjects A-B format in order to make both quantitative measurements about an individual's performance with DP (relative to baseline), but also to make qualitative observations of the processes used by each participant as they solved each task.
 Each task was paired with one of two datasets, depending on condition, so that we could evaluate whether the participant's performance of completing a task differed with different datasets.
@@
 
-\subsection{Condition B (control}
+\subsection{Condition B (control)}
 In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.

### 2012-09-17 chi2012.tex
Minor fixes

--- 2012-09-17 chi2012.tex
@@
 Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from definitions in \cite{george2005understanding} and feature-metrics used in statistical instance alignment algorithms (such as summarised in \cite{bhattacharya2007collective}).
-Such metrics were chosen because they reflected the differences among data feeds at multiple levels -  the terminological (use of names/labels), structural, and modelling levels.
+Such metrics were chosen because they reflected, the differences among the data feeds at multiple levels - from the simple (labelling) differneces to differences in how data was structured, and modelled.
 Each such metric listed above is the result of averaging the pairwise score between sources in each category, separately and defined as follows:
@@
 
-We used three metris to measure of task performance: time to completion, the number of factors used in the choice, and the number of candidates
+We used three metrics to measure task performance: time to completion; the number of factors used in the choice; and the number of candidates evaluated.
 
 \subsubsection{Efficiency: Time taken}
-Participants spent an average of 8.9 minutes performing the tasks overall, with DP (condition A) taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B (the web browser/Excel interface) ($M=9.0$min; $SD=1.44$).
-An ANOVA test of time taken by interface condition blocking by participant did not find a interface use a significant predictor of task completion time.
+Participants spent an average of 8.9 minutes performing the tasks overall, with condition A (DataPalette) taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B (the web browser/Excel interface) ($M=9.0$min; $SD=1.44$).
+An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time.
 A comparison of times taken are visible in Figure \ref{fig:time}.
 
-Although across trials, restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor either.
+Although across trials, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor.
 
 \subsubsection{Thoroughness: Factors weighed in final choice}
-A second metric was the number of factors the participant considered in making their final choice.
+The second metric was the number of factors the participant considered in making their final choice.
 To determine this, we asked participants to tell us precisely why their choice was made, and matched their justifications to properties in the data.
-This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
+This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95, p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
 
 \subsubsection{Diversity: Data sources consulted}
-A third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
-In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04$ $p\approx0.10$).
+The third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
+In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04  p\approx0.10$).
 
@@
 Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much participants had to keep track of during the study.
-We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22; p\approx0.10$).
+We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22$, $p\approx0.10$).
 
-\subsection{Condition A (DP)}
-In the A condition, where participants used DataPalette to complete their tasks they used it to collate the data on-screen, and compare multiple values directly.
+\subsection{Condition A (DataPalette)}
+In the A condition, where participants used DataPalette to complete their tasks, they used it to collate the data on-screen, and compare multiple values directly.
 About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it.

### 2012-09-17 chi2012.tex

--- 2012-09-17 chi2012.tex
@@
 
-We used three metris to measure of task performance: time to completion, the number of factors used in the choice, and the number of candidates
+We used three metrics to measure task performance: time to completion; the number of factors used in the choice; and the number of candidates evaluated.
 
 \subsubsection{Efficiency: Time taken}
-Participants spent an average of 8.9 minutes performing the tasks overall, with DP (condition A) taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B (the web browser/Excel interface) ($M=9.0$min; $SD=1.44$).
-An ANOVA test of time taken by interface condition blocking by participant did not find a interface use a significant predictor of task completion time.
+Participants spent an average of 8.9 minutes performing the tasks overall, with condition A (DataPalette) taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B (the web browser/Excel interface) ($M=9.0$min; $SD=1.44$).
+An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time.
 A comparison of times taken are visible in Figure \ref{fig:time}.
 
-Although across trials, restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor either.
+Although across trials, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor.
 
@@
 
-\subsection{Condition A (DP)}
-In the A condition, where participants used DataPalette to complete their tasks they used it to collate the data on-screen, and compare multiple values directly.
+\subsection{Condition A (DataPalette)}
+In the A condition, where participants used DataPalette to complete their tasks, they used it to collate the data on-screen, and compare multiple values directly.
 About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it.

### 2012-09-17 chi2012.tex
Minor fixes

@@
 \subsubsection{Thoroughness: Factors weighed in final choice}
-A second metric was the number of factors the participant considered in making their final choice.
+The second metric was the number of factors the participant considered in making their final choice.
 To determine this, we asked participants to tell us precisely why their choice was made, and matched their justifications to properties in the data.
@@
 \subsubsection{Diversity: Data sources consulted}
-A third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
+The third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
 In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04$ $p\approx0.10$).

### 2012-09-17 chi2012.tex
progress

--- 2012-09-17 chi2012.tex
@@
 % TODO TODO TODO TODO 
-Our findings suggest that
+Our findings suggest that -- FILL ME IN HERE PLEASE,
 
@@
 
-We used three metrics to measure task performance: time to completion; the number of factors used in the choice; and the number of candidates evaluated.
+We used three metris to measure of task performance: time to completion, the number of factors used in the choice, and the number of candidates
 
 \subsubsection{Efficiency: Time taken}
-Participants spent an average of 8.9 minutes performing the tasks overall, with condition A (DataPalette) taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B (the web browser/Excel interface) ($M=9.0$min; $SD=1.44$).
-An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time.
+Participants spent an average of 8.9 minutes performing the tasks overall, with DP (condition A) taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B (the web browser/Excel interface) ($M=9.0$min; $SD=1.44$).
+An ANOVA test of time taken by interface condition blocking by participant did not find a interface use a significant predictor of task completion time.
 A comparison of times taken are visible in Figure \ref{fig:time}.
 
-Although across trials, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor.
+Although across trials, restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor either.
 
 \subsubsection{Thoroughness: Factors weighed in final choice}
-The second metric was the number of factors the participant considered in making their final choice.
+A second metric was the number of factors the participant considered in making their final choice.
 To determine this, we asked participants to tell us precisely why their choice was made, and matched their justifications to properties in the data.
-This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95, p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
+This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
 
 \subsubsection{Diversity: Data sources consulted}
-The third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
-In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04  p\approx0.10$).
+A third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
+In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04$ $p\approx0.10$).
 
@@
 Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much participants had to keep track of during the study.
-We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22$, $p\approx0.10$).
+We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22; p\approx0.10$).
 
-\subsection{Condition A (DataPalette)}
-In the A condition, where participants used DataPalette to complete their tasks, they used it to collate the data on-screen, and compare multiple values directly.
+\subsection{Condition A (DP)}
+In the A condition, where participants used DataPalette to complete their tasks they used it to collate the data on-screen, and compare multiple values directly.
 About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it.

### 2012-09-17 chi2012.tex
Survey results

--- 2012-09-17 chi2012.tex
@@
 
-In the exit survey, the users noted that they would use DataPalette for finding a property to rent/buy, choosing a job and purchasing electronic devices, in the future.
+\subsection{Survey Results}
+
+%+ ÊWhat did people think about it? 
+% Ê Ê Ê ÊUsefulness,Ease,Confusion,etc
+
+In the exit survey, the users noted that they would use an interface like DataPalette in the future, for finding a property to rent/buy, choosing a job and purchasing electronic devices.
 However, one participant said that they would not use it for things like deciding on a restaurant because ``it would cheapen the experience'' by removing the spontaneity, but would use it for making more important decisions.
-In order to improve our solutions the participants suggested making the origin of the information clearer, sorting instances, and explanations of ratings and scoring systems (however some of these are not transparent on websites).
+In order to improve our solutions the participants suggested making the origin of the information clearer, particularly when combining multiple data sources.
+The participants also requested the ability to sort instances, as well as explanations of ratings and scoring systems (however some of these are not even transparent on websites).
+Participants also requested standard features not present in our prototype, such as window management options (specifically, arranging and minimising windows), and the ability to undo operations.
 

### 2012-09-17 chi2012.tex
strategy

--- 2012-09-17 chi2012.tex
@@
 
+%% TODO - emax doesn't like this paragraph - kill it? 
 During the data collection we collated a spreadsheet summarising the participant's gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
@@
 
-TODO look further into types of data analysis....
+%% TODO look further into types of data analysis....
 
@@
 We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22; p\approx0.10$).
+
+\subsubsection{Strategies: Successive Elimination vs Tallying}
+As described in Discussion later, we noticed several different strategies that people used to evalutate each candidate selection.
+One common strategy was \emph{successive elimination} - to look at a single dimension at a time, starting with the most important, ruling out candidates that did not meet the required minimum for that value.
+However, when there is no clear ``most important'' aspect, this  ``greedy-style'' can result in suboptimal decisions.
+By comparison, a different strategy, which we called \emph{tallying the pros and cons}, kept all candidates around, and counted up the advantages and disadvantages of each, potentially weighing each factor by its perceived importance.
+This strategy is less vulnerable to getting stuck in local maxima than the former.
+
+To understand whether interface influenced choice of strategy, we measured the number of candidates that the participant considered \emph{at final result}.
+This we found was a strong marker for use of each strategy; people using the successive elimination strategy arrived with only 1-2 candidates at decision time, while those that tallied still maintained the entire starting set.
+
+We found that the use of DP strongly influenced people to keep all candidates around and
 

### 2012-09-17 chi2012.tex
finished foo bar

@@
 
-We found that the use of DP strongly influenced people to keep all candidates around and
+We found that the use of DP strongly influenced people to keep all candidates around, while participants in the B condition eliminated choices early.
+An ANOVA test demonstrated a significant effect between condition and number of candidates maintained ($F(1,19)=5.323; p<0.05$); a post-hoc Tukey HSD test confirmed that  the number of candidates used in the final choice for participants in the A condition ($M=5.85;SD=0.45$) differed significantly from the B condition ($M=4.95; SD=2.26$).
 

### 2012-09-17 chi2012.tex
visualisation

@@
 
-\subsubsection{Strategies: Successive Elimination vs Tallying}
+\subsection{Strategies: Successive Elimination vs Tallying}
 As described in Discussion later, we noticed several different strategies that people used to evalutate each candidate selection.
@@
 An ANOVA test demonstrated a significant effect between condition and number of candidates maintained ($F(1,19)=5.323; p<0.05$); a post-hoc Tukey HSD test confirmed that  the number of candidates used in the final choice for participants in the A condition ($M=5.85;SD=0.45$) differed significantly from the B condition ($M=4.95; SD=2.26$).
+
+\subsection{Use of Data Integration Features}
+
+To determine whether participants used the data integration features of DP, we looked at use of the \emph{multipath selection} and \emph{Same-As} features of the system.
+Pertaining to path selection, all participants readily used the path selection process to select values.
+Five participants (25\%) deliberately used multipath selection, defined as selecting more than one active path per group at the same time to display common values across heterogeneous items.
+(We did not count accidental use of multipath, such as happened a few times when a participant had an already active path and wanted to switch to a separate path accidentally).
+
+Participants used the drag-and-drop \emph{Same-As} capability extensively.
+Sixteen (80\%) used \emph{Same-As} as least once; among these participants, they used \emph{Same-As} an average of 4.6 times per trial.
+We counted a single ``use'' as a single drag-and-drop of an individual item, or an entiredrag-and-drop of an entire group into another group.
+(We did not differentiate enumerated value consolidation from entity consolidation, as these operations were not discernably different from the interface perspective.)
+
+\subsection{Use of Visualisation Features}
+
+To determine to what extent visualisation tools were used in condition A, 13 (65\%) used charting tools at least once, while 15 (75\%) used the map.
+Determining the impact these had on performance, we discovered a significant interaction between the amount charting was used ($F(1,17)=3.26; p<0.10$) and maps were used ($F(1,17)=4.32; p<0.10$)  with time to complete the task.
+Examining the coefficients of our ANOVA regression ($Y_0=8.91;\beta_{charts}=0.49;\beta_{maps}=-1.05$) we find that as charting weakly positively influences time taken, while maps negatively influences it.
 

### 2012-09-17 chi2012.tex
typos and multiple regression

@@
 Sixteen (80\%) used \emph{Same-As} as least once; among these participants, they used \emph{Same-As} an average of 4.6 times per trial.
-We counted a single ``use'' as a single drag-and-drop of an individual item, or an entiredrag-and-drop of an entire group into another group.
+We counted a single ``use'' as a single drag-and-drop of an individual item, or a drag-and-drop operation of an entire group onto another.
 (We did not differentiate enumerated value consolidation from entity consolidation, as these operations were not discernably different from the interface perspective.)
@@
 To determine to what extent visualisation tools were used in condition A, 13 (65\%) used charting tools at least once, while 15 (75\%) used the map.
-Determining the impact these had on performance, we discovered a significant interaction between the amount charting was used ($F(1,17)=3.26; p<0.10$) and maps were used ($F(1,17)=4.32; p<0.10$)  with time to complete the task.
-Examining the coefficients of our ANOVA regression ($Y_0=8.91;\beta_{charts}=0.49;\beta_{maps}=-1.05$) we find that as charting weakly positively influences time taken, while maps negatively influences it.
+Determining the impact these had on performance, we performed a multiple regression on time taken with the number of uses of charting and mapping as variables.
+We found significant effects of both charting and mapping variables on time taken, ($R^2_{adj} = 0.04; F(2,17)=3.79; p<0.05$), with coefficients ($Y_0=8.91;\beta_{charts}=0.49;\beta_{maps}=-1.05$) demonstrating that charting positively influenced time taken, while the use of maps led to shorter times.
 

### 2012-09-17 chi2012.tex
Added new image

--- 2012-09-17 chi2012.tex
@@
 
-We first describe our methodology, starting with the three hypotheses we set out to test.
-These hypothesis pertain to the usability of DP by end-users (h1), whether DP achieves its stated goal of enabling end-users to integrate data (h2), and finally if it has an effect on task performance (h3) - either in how well people perform a task, how quickly they perform it, or if it reduces task effort, as follows:
+We first describe our methodology, starting with the three hypotheses we set out to test:
+%These hypothesis pertain to the usability of DP by end-users (h1), whether DP achieves its stated goal of enabling end-users to integrate data (h2), and finally if it has an effect on task performance (h3) - either in how well people perform a task, how quickly they perform it, or if it reduces task effort, as follows:
 
@@
 
-\begin{itemize}
-\item \emph{h1 -} People understand how to use DP.
-\item \emph{h2 -} People can effectively mix heterogeneous data with DP.
-\item \emph{h3 -} Use of DP improves or facilitates task performance.
-\end{itemize}
+\begin{description}
+\item [h1 - Usability] People understand how to use DP.
+\item [h2 - Data Integration] People can effectively mix heterogeneous data with DP.
+\item [h3 - Performance] Use of DP improves or facilitates task performance.
+\end{description}
 
@@
 
-%\begin{table*}[t]
-%\begin{center}
-%\small
-%\begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
-%\hline
-%Task	 &Choices	&Data Sources\\
-%\hline
-%1) Three universities they would apply to study Sport Science & Universities: 1) Loughborough,  2) Durham,  3) Exeter, 4) Edinburgh, 5) Birmingham, and 6) Bath & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
-%\hline
-%2) Three universities they would apply to study History & Universities: 1) Cambridge, 2) London School of Economics and Political Science, 3) Durham, 4) Oxford, 5) St Andrews, and 6) University College London & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
-%\hline
-%3) Restaurant they would book for 12 friends in Cambridge&Restaurants : 1) Alimentum, 2) Gardenia Restaurant, 3) The Cambridge Chop House, 4) Al Casbah, 5) Midsummer House Restaurant, and 6) Nando's Restaurant.
+\begin{table*}[t]
+\begin{center}
+\small
+\begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
+\hline
+Task	 &Choices	&Data Sources\\
+\hline
+1) Three universities they would apply to study Sport Science & Universities: 1) Loughborough,  2) Durham,  3) Exeter, 4) Edinburgh, 5) Birmingham, and 6) Bath & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
+\hline
+2) Three universities they would apply to study History & Universities: 1) Cambridge, 2) London School of Economics and Political Science, 3) Durham, 4) Oxford, 5) St Andrews, and 6) University College London & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
+\hline
+3) Restaurant they would book for 12 friends in Cambridge&Restaurants : 1) Alimentum, 2) Gardenia Restaurant, 3) The Cambridge Chop House, 4) Al Casbah, 5) Midsummer House Restaurant, and 6) Nando's Restaurant.
 & www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
-%\hline
-%4) Restaurant they would book for 12 friends in Glagow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollopes Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia.
+\hline
+4) Restaurant they would book for 12 friends in Glagow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollopes Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia.
 & www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
-%\hline
-%\end{tabular}
-%\end{center}
-%\caption{Tasks, choices and data sources} \label{tab:studyfactors}
-%\normalsize
-%\end{table*}%
+\hline
+\end{tabular}
+\end{center}
+\caption{Tasks, choices and data sources} \label{tab:studyfactors}
+\normalsize
+\end{table*}%
 
@@
 of which university to choose by averaging different metrics supplied by the data sources.
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=9cm]{img/results}
+\caption{Results, condition A and B shown.}
+\label{fig:results}
+\end{center}
+\end{figure}
 

### 2012-09-17 chi2012.tex

@@
 An ANOVA test demonstrated a significant effect between condition and number of candidates maintained ($F(1,19)=5.323; p<0.05$); a post-hoc Tukey HSD test confirmed that  the number of candidates used in the final choice for participants in the A condition ($M=5.85;SD=0.45$) differed significantly from the B condition ($M=4.95; SD=2.26$).
+
+\subsection{Use of Data Integration Features}
+
+To determine whether participants used the data integration features of DP, we looked at use of the \emph{multipath selection} and \emph{Same-As} features of the system.
+Pertaining to path selection, all participants readily used the path selection process to select values.
+Five participants (25\%) deliberately used multipath selection, defined as selecting more than one active path per group at the same time to display common values across heterogeneous items.
+(We did not count accidental use of multipath, such as happened a few times when a participant had an already active path and wanted to switch to a separate path accidentally).
+
+Participants used the drag-and-drop \emph{Same-As} capability extensively.
+Sixteen (80\%) used \emph{Same-As} as least once; among these participants, they used \emph{Same-As} an average of 4.6 times per trial.
+We counted a single ``use'' as a single drag-and-drop of an individual item, or a drag-and-drop operation of an entire group onto another.
+(We did not differentiate enumerated value consolidation from entity consolidation, as these operations were not discernably different from the interface perspective.)
+
+\subsection{Use of Visualisation Features}
+
+To determine to what extent visualisation tools were used in condition A, 13 (65\%) used charting tools at least once, while 15 (75\%) used the map.
+Determining the impact these had on performance, we performed a multiple regression on time taken with the number of uses of charting and mapping as variables.
+We found significant effects of both charting and mapping variables on time taken, ($R^2_{adj} = 0.04; F(2,17)=3.79; p<0.05$), with coefficients ($Y_0=8.91;\beta_{charts}=0.49;\beta_{maps}=-1.05$) demonstrating that charting positively influenced time taken, while the use of maps led to shorter times.
 

### 2012-09-17 chi2012.tex
Placed results graph
foo

--- 2012-09-17 chi2012.tex
@@
 
+% we examined use on performance but nothing significant was found :/
+
 \subsection{Use of Visualisation Features}
-
 To determine to what extent visualisation tools were used in condition A, 13 (65\%) used charting tools at least once, while 15 (75\%) used the map.
@@
 
+%% -> discussion and limitations
 \subsection{Condition A (DataPalette)}

### 2012-09-17 chi2012.tex

--- 2012-09-17 chi2012.tex
@@
 \begin{center}
-\includegraphics[width=9cm]{img/results}
-\caption{Results, condition A and B shown.}
+\includegraphics[width=8.5cm]{img/results}
+\caption{The chosen restaurant/university for each participant, showing the choice on the x axis and frequency on the y axis under conditions A and B.}
 \label{fig:results}

### 2012-09-17 chi2012.tex
moving things about

@@
 
-%% -> discussion and limitations
-\subsection{Condition A (DataPalette)}
-In the A condition, where participants used DataPalette to complete their tasks, they used it to collate the data on-screen, and compare multiple values directly.
-About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it.
-Those who tried to use it were largely successful, and were able to therefore use the global brush-highlighting to simultaneously compare different statistics.
-Most users instantly opened up all data sources into individual boxes, in order to determine which sources contained specific statistics.
-The result of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.
-
 \subsection{Condition B (control)}
@@
 \section{Discussion}
+
+%% -> discussion and limitations
+\subsection{DataPalette Strengths and Limitations}
+In the A condition, where participants used DataPalette to complete their tasks, they used it to collate the data on-screen, and compare multiple values directly.
+About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it.
+Those who tried to use it were largely successful, and were able to therefore use the global brush-highlighting to simultaneously compare different statistics.
+Most users instantly opened up all data sources into individual boxes, in order to determine which sources contained specific statistics.
+The result of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.
 

### 2012-09-17 chi2012.tex
minor edit

@@
 
-The information in Table \ref{tbl:data} summarizes the metrics that were captured of the trials for each participant. We analyze this data next.
+The information in Table \ref{tbl:data} summarises the metrics that were captured of the trials for each participant. We analyse this data next.
 
@@
 \includegraphics[width=8.5cm]{img/results}
-\caption{The chosen restaurant/university for each participant, showing the choice on the x axis and frequency on the y axis under conditions A and B.}
+\caption{The chosen restaurant/university for each participant, showing the choice on the x axis and frequency on the y axis under conditions A (blue) and B (red).}
 \label{fig:results}

### 2012-09-17 chi2012.tex
excel use

--- 2012-09-17 chi2012.tex
@@
 
-\subsection{Condition B (control)}
+\subsection{Condition B: Use of Websites and Excel}
 In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.
-Participants who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues.
+For B trials, we measured the fraction of the trial time spent in Excel versus looking at the sites in a rough, 5-value range (0,25,50,75,100\%).
+Six individuals used Excel entirely, while one avoided it altogether; most fell somewhere in between, with a median of 50\% use.
+
+Participants who used websites to make their decisions often encountered several difficulties and issues.
 In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education.

### 2012-09-17 chi2012.tex
delicious excel

--- 2012-09-17 chi2012.tex
@@
 
-\subsection{Condition B (control)}
+\subsection{Condition B: Observations and Issues (Websites + Excel)}
 In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.
-Participants who opted to use websites to make their decisions repeatedly encountered a number of difficulties and issues.
-In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education.
-There were numerous usability issues in using websites, particularly in searching for university courses and course requirements.
-Similarly, participants found it difficult to understand the locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto the same map using the websites.
-Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and all ended up falling back to using the supplied spreadsheets.
+For B trials, we measured the fraction of the trial time spent in Excel versus looking at the sites in a 5-value range (0,25,50,75,100\%).
+Six individuals used Excel entirely without consulting the web sites at all, while one avoided Excel altogether; most fell somewhere in between, with a median use of 75\% Excel.
+When asked to explain their choice post-hoc, participants who used Excel explained that it saved them the effort of having to individually seach and consolidate information from the websites (10 participants).
+Participants who looked at websites chose to do so when they needed to know more than what was provided in the spreadsheets (8 participants), when they mistrusted the information in them (1 participant), or if they were not familiar with Excel (1 participant).
+
+Of the Excel use, most participants made minor formatting and column sizing changes to facilitate reading and interpreting the spreadsheets, while four made extensive edits to the sheets to help them ``think through'' the task in various ways.
+These uses of Excel for external cognition varied from merely copying over candidate choices from one sheet to the next, sorting values by multiple columns (which participants had considerable trouble with), and colour-coding cells to allow easy counting of  ``pros'' and ``cons''; ``I can't deal with so many textual values; I need colours to help me see'' (p6).
 
@@
 of which university to choose by averaging different metrics supplied by the data sources.
+
+Participants who used websites to make their decisions often encountered several difficulties and issues.
+In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education.
+There were numerous usability issues in using websites, particularly in searching for university courses and course requirements.
+Similarly, participants found it difficult to understand the locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto the same map using the websites.
+Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and all ended up falling back to using the supplied spreadsheets.
 

### 2012-09-17 chi2012.tex
all but survey

@@
 Six individuals used Excel entirely without consulting the web sites at all, while one avoided Excel altogether; most fell somewhere in between, with a median use of 75\% Excel.
-When asked to explain their choice post-hoc, participants who used Excel explained that it saved them the effort of having to individually seach and consolidate information from the websites (10 participants).
+When asked to explain their choice post-hoc, participants who used Excel explained that it saved them the effort of having to individually search and consolidate information from the websites (10 participants).
 Participants who looked at websites chose to do so when they needed to know more than what was provided in the spreadsheets (8 participants), when they mistrusted the information in them (1 participant), or if they were not familiar with Excel (1 participant).
@@
 Of the Excel use, most participants made minor formatting and column sizing changes to facilitate reading and interpreting the spreadsheets, while four made extensive edits to the sheets to help them ``think through'' the task in various ways.
-These uses of Excel for external cognition varied from merely copying over candidate choices from one sheet to the next, sorting values by multiple columns (which participants had considerable trouble with), and colour-coding cells to allow easy counting of  ``pros'' and ``cons''; ``I can't deal with so many textual values; I need colours to help me see'' (p6).
+These uses of Excel for external cognition varied from merely copying over candidate choices from one sheet to the next to keep track of their previous selections, to sorting values by multiple columns (which participants had considerable trouble with due to the need to sort columns together), and colour-coding cells to allow easy counting of  ``pros'' and ``cons''; ``I can't deal with so many textual values; I need colours to help me see'' (p6).
+Another user calculated their own weighted metric of which university to choose by averaging different metrics supplied by the data sources and filling in their own chosen weights.
 
-Participants who opted to use the spreadsheet to support their decisions switched between worksheets to view data from different websites.
-They used different approaches to using the spreadsheets, some participants were keen to collate all data onto a single sheet, so that they could compare the values all at once.
-However, this approach was error prone, for example participants noticing too late that the ordering of rows was different on each sheet, so that their pasted values were incorrectly mapped.
-Others used pen and paper to collate the values, which was often faster and resulted in less observed frustration.
-While the majority of users did not
-attempt anything complex with Excel (in fact only one user used charts), there was some advanced use of Excel observed.
-One user explained that they did not like to look at numbers, and so coloured the cells of the types of restaurant, and the
-friends' preferences, and matched the colours to make their decision. Another user calculated their own weighted metric
-of which university to choose by averaging different metrics supplied by the data sources.
+Participants ran into a number of difficulties with using the Excel interface.
+Participants struggled with switching between different worksheets, and having to find the corresponding values in each.
+To cope with this fragmentation, some participants attempted to collate all data onto a single sheet, so that they could compare the values all at once as just described; however, this approach was often error prone.
+For example, two participants did not notice that the rows in two sheets were in different orders (because of a difference in the renderings of Unviersity names), which caused their pasted values to be incorrectly mapped.
+Four participants struggled with sorting columns, and only 1 attempted charting (p7); Excel's charting features caused her considerable difficulty and she eventually gave up trying to display multiple dimenions simultaneously and created simple bar charts instead.
+
+%%  was use of pen and paper to collate the values, which was often faster and resulted in less observed frustration.
+While the majority of users did not attempt anything complex with Excel (in fact only one user used charts), there was some advanced use of Excel observed.
 

### 2012-09-17 chi2012.tex
edits and changes

@@
 \subsection{Use of Visualisation Features}
-To determine to what extent visualisation tools were used in condition A, 13 (65\%) used charting tools at least once, while 15 (75\%) used the map.
+To determine to what extent visualisation tools were used in condition A, 13 (65\%) used charting tools at least once, while a majority (15, 75\%) used the map.
 Determining the impact these had on performance, we performed a multiple regression on time taken with the number of uses of charting and mapping as variables.
@@
 
-\subsection{Condition B: Observations and Issues (Websites + Excel)}
+\subsection{Condition B - Results and Observations}
 In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.
@@
 
-Participants ran into a number of difficulties with using the Excel interface.
-Participants struggled with switching between different worksheets, and having to find the corresponding values in each.
-To cope with this fragmentation, some participants attempted to collate all data onto a single sheet, so that they could compare the values all at once as just described; however, this approach was often error prone.
-For example, two participants did not notice that the rows in two sheets were in different orders (because of a difference in the renderings of Unviersity names), which caused their pasted values to be incorrectly mapped.
-Four participants struggled with sorting columns, and only 1 attempted charting (p7); Excel's charting features caused her considerable difficulty and she eventually gave up trying to display multiple dimenions simultaneously and created simple bar charts instead.
+Participants ran into a number of difficulties with the spreadsheet interface.
+Participants struggled with looking up values across different worksheets (data sources).
+Some participants coped by attempting to collate all data onto a single sheet, so that they could compare the values all at once as just described; however, this approach was often error prone.
+For example, two participants did not notice that the rows in two sheets were in different order, causing their pasted values to be incorrectly mapped.
+Four participants struggled with sorting columns, and only 1 attempted charting (p7).
+Although she attempted to plot multiple factors simultaneously, Excel's charting features caused her considerable difficulty and she eventually resorted at creating multiple simple bar-charts instead.
 
@@
 
-Participants who used websites to make their decisions often encountered several difficulties and issues.
-In particular, there was significant confusion over the meaning of data points, for example the percentage-based rankings of university teaching by the Times Higher Education.
-There were numerous usability issues in using websites, particularly in searching for university courses and course requirements.
-Similarly, participants found it difficult to understand the locations of restaurants in cities that they were not familiar, and could not find a way to plot the shortlist of restaurants onto the same map using the websites.
-Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and all ended up falling back to using the supplied spreadsheets.
+%% todo - look up to make sure it was really p7 here 
+Various other difficulties were encountered with the websites.
+There was significant confusion over various attributes, especially in the University selection task.
+For example, p15 could not tell how the teaching score from the Times Higher Education was calculated, in order to assess its relative importance.
+Several participants had difficulty finding information they needed because of inconsistencies in site layout, especially finding course requirements, information that ``should be obvious''.
+Similar issues were encountered with finding information in the restaurant condition; p7 said ``I wish they would put Cuisine Type in the same place for every restaurant - important information!'' Similarly, participants found it difficult to visualise the locations of restaurants in cities that they were not familiar, and could not find a way to easily map them.
+Overall, the users that spent the most time using websites exhibited significant frustration attempting to locate and compare different criteria (such as course entry requirements, or restaurant locations), and often ended up swapping back to using the supplied spreadsheets.
 
@@
 
-Notably, a number of users admittedly introduced their own bias into the choices in both A and B conditions: ``I prefer Indian food, so I will choose the Indian restaurant'', and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating those that required bus transfers (the protocol did not mention routing/transfers at all).
+% Notably, a number of users admittedly introduced their own bias into the choices in both A and B conditions: ``I prefer Indian food, so I will choose the Indian restaurant'', and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating those that required bus transfers (the protocol did not mention routing/transfers at all).
 

### 2012-09-17 chi2012.tex
edits and changes

@@
 
+\subsubsection{Candidate selection}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=8.5cm]{img/results}
+\caption{The chosen restaurant/university for each participant, showing the choice on the x axis and frequency on the y axis under conditions A (blue) and B (red).}
+\label{fig:chosen_results}
+\end{center}
+\end{figure}
+
+Since the tasks were subjective-choice, we could not evaluate the 'correctness' of people's answers.
+However, to determine whether there was a difference in variability of answers between interface conditions, we histogrammed responses for each task and dataset for both conditions, visible in  Figure \ref{fig:chosen_results}.
 \subsection{Strategies: Successive Elimination vs Tallying}
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=8.5cm]{img/results}
-\caption{The chosen restaurant/university for each participant, showing the choice on the x axis and frequency on the y axis under conditions A (blue) and B (red).}
-\label{fig:results}
-\end{center}
-\end{figure}
-
 % Notably, a number of users admittedly introduced their own bias into the choices in both A and B conditions: ``I prefer Indian food, so I will choose the Indian restaurant'', and one participant spent time on Google Maps looking at the routes to each restaurant, and eliminating those that required bus transfers (the protocol did not mention routing/transfers at all).
@@
 
-%+ ÊWhat did people think about it? 
-% Ê Ê Ê ÊUsefulness,Ease,Confusion,etc
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=8.5cm]{img/survey}
+\caption{The chosen restaurant/university for each participant, showing the choice on the x axis and frequency on the y axis under conditions A (blue) and B (red).}
+\label{fig:survey}
+\end{center}
+\end{figure}
+
+%+What did people think about it? 
+% Usefulness,Ease,Confusion,etc
 

### 2012-09-17 chi2012.tex
happy times git push

@@
 \section{Designing DataPalette : An Interface for Serendipitous Data Mixing}
-\begin{figure*}[htbp]
+\begin{figure*}[tbp]
 \begin{center}
@@
 \includegraphics[width=8.5cm]{img/results}
-\caption{The chosen restaurant/university for each participant, showing the choice on the x axis and frequency on the y axis under conditions A (blue) and B (red).}
+\caption{\emph{Choice picks per participant trial} - Histogram of number of times each restaurant/university was chosen in DP interface (A condition) in blue/left bars, compared to the Excel/website (B condition) in red/to the right.}
 \label{fig:chosen_results}
@@
 Since the tasks were subjective-choice, we could not evaluate the 'correctness' of people's answers.
-However, to determine whether there was a difference in variability of answers between interface conditions, we histogrammed responses for each task and dataset for both conditions, visible in  Figure \ref{fig:chosen_results}.
+However, to determine whether there was a difference in variability of answers between interface conditions, we histogrammed first choices for each trial, per task and dataset for both conditions, visible in  Figure \ref{fig:chosen_results}.
+As can be seen there is considerable variability and no discernable difference in variability between conditions.
+
 \subsection{Strategies: Successive Elimination vs Tallying}
@@
 Of the Excel use, most participants made minor formatting and column sizing changes to facilitate reading and interpreting the spreadsheets, while four made extensive edits to the sheets to help them ``think through'' the task in various ways.
-These uses of Excel for external cognition varied from merely copying over candidate choices from one sheet to the next to keep track of their previous selections, to sorting values by multiple columns (which participants had considerable trouble with due to the need to sort columns together), and colour-coding cells to allow easy counting of  ``pros'' and ``cons''; ``I can't deal with so many textual values; I need colours to help me see'' (p6).
+These uses of Excel for external cognition ranged from merely copying over candidate choices from one sheet to the next to keep track of their previous selections, to sorting values by multiple columns (which participants had considerable trouble with due to the need to sort columns together), and colour-coding cells to allow easy counting of  ``pros'' and ``cons''; ``I can't deal with so many textual values; I need colours to help me see'' (p6).
 Another user calculated their own weighted metric of which university to choose by averaging different metrics supplied by the data sources and filling in their own chosen weights.
@@
 Participants struggled with looking up values across different worksheets (data sources).
-Some participants coped by attempting to collate all data onto a single sheet, so that they could compare the values all at once as just described; however, this approach was often error prone.
+Some attempted to collate all data onto a single sheet, so that they could compare the values all at once as just described; however, this approach was often error prone.
 For example, two participants did not notice that the rows in two sheets were in different order, causing their pasted values to be incorrectly mapped.
@@
 
-\begin{figure}[htbp]
+\begin{figure}[tbp]
 \begin{center}
 \includegraphics[width=8.5cm]{img/survey}
-\caption{The chosen restaurant/university for each participant, showing the choice on the x axis and frequency on the y axis under conditions A (blue) and B (red).}
+\caption{\emph{Survey results} - Answers to questions on a 5-point Likert scale, from 1-\emph{strongly disagree} to 5-\emph{strongly agree}.}
 \label{fig:survey}
@@
 \end{figure}
+
+When asked to rate the DP interface, all participants responded that they felt they understood the system, with sixteen (80\%) responding that they understood it \emph{very strongly} (5).
+Eighteen (90\%) reported that they thought  DP was \emph{easy}, as well as useful, with one person being neutral on each.
+On the statement ``DP is confusing'',  11 strongly disagreed, 6 disagreed, and 3 agreed.
+Revisiting the time trials, we discovered that the 3 participants who rated DP as strongly confusing completed the trials statistically significantly slower than the rest of the group.
+Survey results were visualised in Figure \ref{fig:survey}.
 
@@
 In the exit survey, the users noted that they would use an interface like DataPalette in the future, for finding a property to rent/buy, choosing a job and purchasing electronic devices.
-However, one participant said that they would not use it for things like deciding on a restaurant because ``it would cheapen the experience'' by removing the spontaneity, but would use it for making more important decisions.
-In order to improve our solutions the participants suggested making the origin of the information clearer, particularly when combining multiple data sources.
+However, one participant said that they would not use it for things like deciding on a restaurant because ``it would cheapen the experience'' by removing the element of spontaneity of choice, but would use it for making more important decisions.
+Participants had a large number of suggestions for improvement, which In order to improve our solutions the participants suggested making the origin of the information clearer, particularly when combining multiple data sources.
 The participants also requested the ability to sort instances, as well as explanations of ratings and scoring systems (however some of these are not even transparent on websites).

### 2012-09-18 chi2012.tex
foo

@@
 \begin{description}
-\item [h1 - Usability] People understand how to use DP.
-\item [h2 - Data Integration] People can effectively mix heterogeneous data with DP.
-\item [h3 - Performance] Use of DP improves or facilitates task performance.
+\item [h1. Usability -] People understand how to use DP.
+\item [h2. Data Integration -] People can effectively mix heterogeneous data with DP.
+\item [h3. Task completion -] Use of DP improves people's ability to perform the task.
 \end{description}
@@
 For B trials, we measured the fraction of the trial time spent in Excel versus looking at the sites in a 5-value range (0,25,50,75,100\%).
-Six individuals used Excel entirely without consulting the web sites at all, while one avoided Excel altogether; most fell somewhere in between, with a median use of 75\% Excel.
-When asked to explain their choice post-hoc, participants who used Excel explained that it saved them the effort of having to individually search and consolidate information from the websites (10 participants).
-Participants who looked at websites chose to do so when they needed to know more than what was provided in the spreadsheets (8 participants), when they mistrusted the information in them (1 participant), or if they were not familiar with Excel (1 participant).
-
-Of the Excel use, most participants made minor formatting and column sizing changes to facilitate reading and interpreting the spreadsheets, while four made extensive edits to the sheets to help them ``think through'' the task in various ways.
-These uses of Excel for external cognition ranged from merely copying over candidate choices from one sheet to the next to keep track of their previous selections, to sorting values by multiple columns (which participants had considerable trouble with due to the need to sort columns together), and colour-coding cells to allow easy counting of  ``pros'' and ``cons''; ``I can't deal with so many textual values; I need colours to help me see'' (p6).
-Another user calculated their own weighted metric of which university to choose by averaging different metrics supplied by the data sources and filling in their own chosen weights.
+Six individuals used Excel entirely (100\%) without consulting the web sites, while one avoided Excel entirely; most fell somewhere in between, with median use at 75\%.
+When asked to explain their choice of tool post-hoc, participants who used Excel explained that it saved them time over individually searching and consolidating information from the websites (10 participants).
+Participants used websites when they needed to know more than what was provided in the spreadsheets (8 participants), when they mistrusted the information in them (1 participant), or if they were not familiar with Excel (1 participant).
 
@@
 Participants struggled with looking up values across different worksheets (data sources).
-Some attempted to collate all data onto a single sheet, so that they could compare the values all at once as just described; however, this approach was often error prone.
+Some attempted to collate all data onto a single sheet, so that they could compare the values all at once; however, this approach was often error prone.
 For example, two participants did not notice that the rows in two sheets were in different order, causing their pasted values to be incorrectly mapped.
@@
 Although she attempted to plot multiple factors simultaneously, Excel's charting features caused her considerable difficulty and she eventually resorted at creating multiple simple bar-charts instead.
+
+While most participants only made minor formatting and column sizing changes to the worksheets to facilitate reading and interpretation,  four made extensive edits to the sheets to help them ``think through'' the task in various ways.
+These uses of Excel for external cognition ranged from the collation behaviour just described to tallying annotations, to colour-coding cells to allow easy identification of  ``pros'' and ``cons''; ``I can't deal with so many textual values; I need colours to help me see'' (p6).
+Another user calculated their own weighted metric of which university to choose by averaging different metrics supplied by the data sources and filling in their own chosen weights.
 
@@
 %% todo - look up to make sure it was really p7 here 
-Various other difficulties were encountered with the websites.
+Various difficulties were also encountered with use of the websites.
 There was significant confusion over various attributes, especially in the University selection task.
@@
 \section{Discussion}
+
+In this section, we first re-visit our study hypotheses introduced in Methodology, using observations from results to support views on each.
+We follow this with a discussion of limitations, concerning the design of the interface, the state of the prototype, and the scope and execution of the study, including potential confounding factors, and how this impacts interpretability of results.
+Finally, we discuss our current and follow-on plans for continuing this line of research.
+
+\subsection{Revisiting the hypotheses: Does DesignPalette facilitate serendipitous data integration?}
+
+This study tested how
 

### 2012-09-18 chi2012.tex
discussion

@@
 \subsection{Experimental Setup}
-The participants were given a standard desktop PC running MacOS X on a 22" monitor with a standard keyboard and mouse.
+The participants were given a standard desktop PC running MacOS X on a 22-inch monitor with a standard keyboard and mouse.
 During the studies the Excel workbook and DP were initially set to full screen by default.
@@
 All of the participants spoke English, but 12 of them were non-native English speakers.
-Seven were non-students; out of the remaining 13, eight studied computer science and 5 studied other subjects.
+Seven were non-students, comprising University staff and alumni; out of the remaining 13, eight studied computer science and 5 studied other subjects.
 On average, each study took 40 mins to complete.
@@
 
-When asked to rate the DP interface, all participants responded that they felt they understood the system, with sixteen (80\%) responding that they understood it \emph{very strongly} (5).
-Eighteen (90\%) reported that they thought  DP was \emph{easy}, as well as useful, with one person being neutral on each.
+When asked to rate the DP interface, all participants responded that they felt they understood the system, with sixteen (80\%) responding that they agreed \emph{very strongly}.
+Eighteen (90\%) reported that they thought  DP was \emph{easy}, as well as \emph{useful}, with one person being neutral on each.
 On the statement ``DP is confusing'',  11 strongly disagreed, 6 disagreed, and 3 agreed.
-Revisiting the time trials, we discovered that the 3 participants who rated DP as strongly confusing completed the trials statistically significantly slower than the rest of the group.
+Revisiting the time trials, we discovered that the 3 participants who rated DP as confusing completed the trials statistically significantly slower than the rest of the group.
 Survey results were visualised in Figure \ref{fig:survey}.
@@
 
-\section{Discussion}
+\section{Discussion and Limitations}
 
 In this section, we first re-visit our study hypotheses introduced in Methodology, using observations from results to support views on each.
-We follow this with a discussion of limitations, concerning the design of the interface, the state of the prototype, and the scope and execution of the study, including potential confounding factors, and how this impacts interpretability of results.
+We follow this with a discussion of limitations regarding the design of the interface, the state of the prototype, and the design and execution of the study.
 Finally, we discuss our current and follow-on plans for continuing this line of research.
 
-\subsection{Revisiting the hypotheses: Does DesignPalette facilitate serendipitous data integration?}
+\subsection{Revisiting the hypotheses: Does DataPalette facilitate serendipitous data integration?}
 
-This study tested how
+We set out to answer three different questions with the DataPalette study.
+The first concerned whether people understood and could use the system (h1), whether the system enabled people to mix and integrate heterogeneous data (h2), and whether the interface facilitated task completion (h3).
+Using real data obtained from typical data sources identified during our prestudy (PS.1), we tested the DP prototype and its integration interaction mechanisms with ``real'' users, who were students, staff and alumni of our University who had varying backgrounds and levels of expertise with computers.
+
+We find substantial support for h1.
+All participants reported that they understood the tool, and that it was easy to use.
+All participants managed to use the interface to view, organize and collate data without running into major roadblocks or confusion, and all effectively were able to compare multiple attributes of heterogeneous data items directly.
+More than half of the participants used DP's visualisation features (charting or map -- often several times during the trial -- suggesting that these features were useful and usable as well.
+
+More convincing of participants' understanding of the system was their feedback, specifically feature requests and desired capabilities.
+Several participants requested search functionality, sort and filter functionality, the ability to display multiple properties for a single instance simultaneously, and visualise multiple attributes in a 2 or 3-dimensional visualisation.
+Perhaps the most interesting suggestion was also the most common -- the ability to selectively view the provenance of information after instances have been combined.
+For example, after his trial, P15 said:
+
+\begin{quote}
+Properties are a tricky issue.
+Sometimes you don’t care [which data source] a property is from, like ``address'' or ``phone number'' - for these the way it’s done now is fine.
+But in some cases you need context of where the properties come from in order to know what it really means - like for ``rating'', it makes a big difference whether you’re talking about ``Yelp rating'' or some random reviwer's rating.
+You also don’t know what typical ratings are, whether 5 stars are much better than 4, etc.
+\end{quote}
+
+Pertaining to h2, all participants were able to work with multiple sets of heterogeneous data effectively.
+With respect to integration specifically, a majority (80\%) of the particiapnts successfully and deliberately integrated data using drag-and-drop SameAs capabilities.
+While single-paths were used, multi-path features were not as widely used, only by 4 participants.
+We believe that participants may not have realized that creating multiple path selections per group was possible, instead assuming it to be similar to the single ``Current path'' state per window of most file managers.
+
+Ascertaining DP's impact on task performance (h3) was the most difficult to ascertain, given the small sample size of our study and large number of factors influencing each person's performance.
+While tasks were completed on average slightly faster than the control interface, these differences were not statistically significant.
+However, more data sources seemed to be consulted (we use ``seemed'' carefully because these findings approached significance) during trials with DP over the control interface, which meant that people were looking at more diverse information than the control conditions.
+We also found that particiapnts justified their choices with a greater number of factors in DP than the control interface, suggesting that decisions may have been made considering a greater number of factors.
+Participants entertained more possibilities for longer in the DP trial, while participants were more prone to eliminate candidates early in the baseline interface.
+Finally, people took fewer notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP, which strongly suggests that DP made the task easier to complete.
 
@@
 % DAN's  qual analysis of processes
-
 In order to complete the tasks, the users followed a number of different processes. The most common process followed

### 2012-09-18 chi2012.tex
foo bar

@@
 We hypothesise that end-users would benefit from this rich data with tools allowing them to effectively browse, use and consume heterogeneous structured data from diverse sources.
-In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data impact information practice.
-We used a three stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of  information sources, and the processes that they relied upon to perform such tasks.
+In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data.
+We used a three-stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of  information sources, and the processes that they relied upon to perform such tasks.
 Our results suggested that people preferred to rely upon multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and alternative perspectives.
@@
 Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data.
-Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in three popular domains: social networking, shopping, and dining, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
+Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: contacts, events, music, shopping (retail), social networking and weather, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
 
@@
 
-\subsection{Better Supporting the Task}
+\subsection{Supporting more sophisticated heterogeneity}
 
-- are people more thorough
-- do they evaluate more sources
+Supporting better structural heterogeneity:
+   Combining fields
+   Splitting fields
+   Rearranging fields (British dates American dates)
 
-\subsection{Causing Less Frustration}
+Modeling heterogeneity (?)
+   Specificity  
+     Education+Jobs vs ``Education'' + ``Jobs''
 
-- Gave up with aspects of the tasks later / not at all
-- With websites, users gave up on websites and moved on early
+   Unit conversion (kg -> stones)
+   Granularity
+     Crimes per area
 
-\subsection{Keeping an Open Mind}
+Working with data values
 
-- Did users rule out choices less/later
-- Did that mean they chose "better" options
+%% \subsection{Better Supporting the Task}
 
-\subsection{What method did people use?}
+%% - are people more thorough
+%% - do they evaluate more sources
 
-- Processes
-- Organisation of data/screen
-- A vs B, how are they different?
+%% \subsection{Causing Less Frustration}
+
+%% - Gave up with aspects of the tasks later / not at all
+%% - With websites, users gave up on websites and moved on early
+
+%% \subsection{Keeping an Open Mind}
+
+%% - Did users rule out choices less/later
+%% - Did that mean they chose "better" options
+
+%% \subsection{What method did people use?}
+
+%% - Processes
+%% - Organisation of data/screen
+%% - A vs B, how are they different?
 
 % DAN's  qual analysis of processes
-In order to complete the tasks, the users followed a number of different processes. The most common process followed
-was {\it collation} of data, either on-screen or on paper. Some users were proficient enough to copy and paste data
-in the supplied spreadsheets from worksheet to worksheet, while others manually typed data value-by-value across worksheets.
-Others preferred to collate the data from screen to paper, in order to compare multiple data points. A similar process that
-some participants followed was to determine which values of a particular statistic were acceptable, and keep a paper tally
-of these next to the set of choices. They then made their choice based on the value of this tally.
+%% In order to complete the tasks, the users followed a number of different processes. The most common process followed
+%% was {\it collation} of data, either on-screen or on paper. Some users were proficient enough to copy and paste data
+%% in the supplied spreadsheets from worksheet to worksheet, while others manually typed data value-by-value across worksheets.
+%% Others preferred to collate the data from screen to paper, in order to compare multiple data points. A similar process that
+%% some participants followed was to determine which values of a particular statistic were acceptable, and keep a paper tally
+%% of these next to the set of choices. They then made their choice based on the value of this tally.
 
-Two contrasting approaches that participants followed was to use the data to rule out candidates, or to use the data to
-directly compare candidates at once. Our pool of participants exhibited both behaviours evenly, however in the condition
-when using the DataPalette interface, significantly more comparison (and visualisation of values) was performed, with fewer
-participants using the ``ruling out'' process. A consequence of this was that when asked to justify their choices, users were
-able to verify their values on-screen immediately, whereas those who had ruled-out had to memorise their reasons, with
-mixed success  -- many participants had to look through the data sources to re-find the reasons why they has ruled out some choices.
+%% Two contrasting approaches that participants followed was to use the data to rule out candidates, or to use the data to
+%% directly compare candidates at once. Our pool of participants exhibited both behaviours evenly, however in the condition
+%% when using the DataPalette interface, significantly more comparison (and visualisation of values) was performed, with fewer
+%% participants using the ``ruling out'' process. A consequence of this was that when asked to justify their choices, users were
+%% able to verify their values on-screen immediately, whereas those who had ruled-out had to memorise their reasons, with
+%% mixed success  -- many participants had to look through the data sources to re-find the reasons why they has ruled out some choices.
 
-Therefore, the uses that were using DataPalette were more confident in their choices, and knew exactly why they had chosen them,
-often pointing to the screens to show and verify each choice as they explained them.
-
-\subsection{}
-
-- data integration literature demands 'correct' but perfect is the enemy of the good.
+%% Therefore, the uses that were using DataPalette were more confident in their choices, and knew exactly why they had chosen them,
+%% often pointing to the screens to show and verify each choice as they explained them.
 
@@
 
+% - data integration literature demands 'correct' but perfect is the enemy of the good.
 % ``perfect'' be the enemy of the good

### 2012-09-18 chi2012.tex
Design limits

@@
 
-More convincing of participants' understanding of the system was their feedback, specifically feature requests and desired capabilities.
+Additional evidence that participant had a solid understanding of the system was their feedback, specifically feature requests and desired capabilities.
 Several participants requested search functionality, sort and filter functionality, the ability to display multiple properties for a single instance simultaneously, and visualise multiple attributes in a 2 or 3-dimensional visualisation.
@@
 \begin{quote}
-Properties are a tricky issue.
+Properties are tricky.
 Sometimes you don’t care [which data source] a property is from, like ``address'' or ``phone number'' - for these the way it’s done now is fine.
-But in some cases you need context of where the properties come from in order to know what it really means - like for ``rating'', it makes a big difference whether you’re talking about ``Yelp rating'' or some random reviwer's rating.
+But in some cases you need context of where the properties come from in order to know what it really means - like for ``rating'', it makes a big difference whether you’re talking about ``Yelp rating'' or some random reviewer's rating.
 You also don’t know what typical ratings are, whether 5 stars are much better than 4, etc.
@@
 %% -> discussion and limitations
-\subsection{DataPalette Strengths and Limitations}
-In the A condition, where participants used DataPalette to complete their tasks, they used it to collate the data on-screen, and compare multiple values directly.
+\subsection{Design Limitations}
+
+In addition to the aforementioned feature requests, we observed a number of ``pain points'' in the design of the interface that we wish to address in the next iteration.
+One such pain point was window layout and screen real estate management; participants were prone to opening up a large number of windows, including maps and charts, and found that the screen filled up very quickly.
+When the screen got crowded, two problems were observed: people spent a lot of time arranging and moving windows about, and second, people tended to get lost in windows -- ``what was this again?''
+
+\subsection{More Integration Support}
+
+% In the A condition, where participants used DataPalette to complete their tasks, they used it to collate the data on-screen, and compare multiple values directly.
 About half of the observed users actually used the instance combination feature of DataPalette, with the other half able to successfully make decisions without using it.

### 2012-09-18 chi2012.tex
added data analysis methodology

--- 2012-09-18 chi2012.tex
@@
 
-\subsection{Data Collection}
+\subsection{Data Collection and Analysis}
 Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
@@
 
+We planned to analyse the qualitative and quantitative data using thematic and statistic analysis, respectively.
+In order to carry out our data analysis, we used the recorded transcripts to populate a spreadsheet summarising the quantitative metrics (see Table \ref{tbl:data} for specific fields).
+In our study, we aimed to identify process and strategies people used to complete their tasks, so that we could evaluate the effectiveness of DP's interface when aligning heterogeneous data for organisational tasks.
+We also categorised people's views voiced during the study and in their exit survey pertaining to DP's usability and interface, and clustered them to evaluate common feedback.
+
 \subsection{Participant Recruitment}
@@
 Since all such participants had passed the requisite proficiency tests to enrol at University, we did not think this would substantially impact results.
-
 Participants were offered a small gift certificate from an online retailer in exchange for their time.
@@
 
-\subsection{Data Analysis}
+%\subsection{Data Analysis}
 
 %% TODO - emax doesn't like this paragraph - kill it? 
-During the data collection we collated a spreadsheet summarising the participant's gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
+%During the data collection we collated a spreadsheet summarising the participant's gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
 After we had collected the study's data, we performed our data analysis.

### 2012-09-18 chi2012.tex
minor changes

@@
 At a high level, the overall goal of DataPalette was to empower end-users to mix arbitrary structured data feeds on-the-fly while solving their particular information task(s) - be them sense-making or decision-making related - in order to more effectively accomplish these tasks.
-Unlike Mashup makers, including Yahoo Pipes!, the goal was to facilitate such integration without a need for an explicit separate mapping or integration step, which would detract from the overall usability or usefulness of the system by making people do something that did not immediately contribute to their resolving their need.
+Unlike Mashup makers, including Yahoo Pipes!, the goal was to facilitate such integration without a need for an explicit separate mapping or integration step.
+This would detract from the overall usability and usefulness of the system by requiring that users perform tasks that did not immediately contribute to their resolving their need.
 
@@
 % why? why were these tasks selected? can we drop in the phrase ``multi-dimensional subjective choice task'' and refer back to PS.1 to justify the choice of tasks?
-We based our two tasks on these scenarios on choosing a university and picking a restaurant for a large social event, because in PS.1 the majority of people stated that they had used multiple websites for these tasks.
+We based our two tasks on choosing a university and picking a restaurant for a large social event, because in PS.1 the majority of people stated that they had used multiple websites for these tasks.
 We used different two different variables for both tasks, where the restaurant was located (Glasgow and Cambridge) and which subject people choose (sport science and history).
@@
 We selected the top restaurants for on Tripadvisor and Yelp, because they were the top two websites people quoted for comparing reviews when travelling.
-We selected the two courses for the university task at random from a list taken form the complete university guide\footnote{}, and selected the top six university for that subject because they provide a realistic shortlist and are comparable in terms of rankings and properties.
-We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choicesex
+We selected the two university courses at random from a list taken form the complete university guide\footnote{}, and selected the top six university for that subject because they provide a realistic shortlist and are comparable in terms of rankings and properties.
+We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choice.
 
@@
 We selected six because it enable us to scope our study, we found through two pilot studies that six data sources provided a sufficient challenge for users to explore the data and required them to use DP's main features.
-We manually collected the data from the nine selected websites into jason files, and converted them to Microsoft Excel worksheets (which 40\% of the participants in PS.1 said that they used to compare data in).
+We manually collected the data from the nine selected websites into jason files, and converted them to Microsoft Excel worksheets (which 40\% of the participants in PS.1 said that had previously used to compare data).
 
@@
 Since all such participants had passed the requisite proficiency tests to enrol at University, we did not think this would substantially impact results.
-Participants were offered a small gift certificate from an online retailer in exchange for their time.
+Participants were offered a small gift certificate from an online retailer for their time.
 
@@
 The participants were given a standard desktop PC running MacOS X on a 22-inch monitor with a standard keyboard and mouse.
-During the studies the Excel workbook and DP were initially set to full screen by default.
+During the studies the Excel workbook and DP were initially set to full screen.
 Each participant was asked to complete two tasks, one A and one B condition.
@@
 On average, each study took 40 mins to complete.
-
-The information in Table \ref{tbl:data} summarises the metrics that were captured of the trials for each participant. We analyse this data next.
+The information in Table \ref{tbl:data} summarises the metrics that were captured of the trials for each participant.
+We analyse this data next.
 
@@
 
-We used three metrics to measure task performance: time to completion; the number of factors used in the choice; and the number of candidates evaluated.
+The following four sections describes the metrics we used to measure task performance.
+%We used three metrics to measure task performance: time to completion; the number of factors used in the choice; and the number of candidates evaluated.
 
 \subsubsection{Efficiency: Time taken}
-Participants spent an average of 8.9 minutes performing the tasks overall, with condition A (DataPalette) taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B (the web browser/Excel interface) ($M=9.0$min; $SD=1.44$).
-An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time.
-A comparison of times taken are visible in Figure \ref{fig:time}.
-
+Participants spent an average of 8.9 minutes performing the tasks, with our interface, condition A, taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).
+An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time (shown in Figure \ref{fig:time}).
 Although across trials, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor.
@@
 Since the tasks were subjective-choice, we could not evaluate the 'correctness' of people's answers.
-However, to determine whether there was a difference in variability of answers between interface conditions, we histogrammed first choices for each trial, per task and dataset for both conditions, visible in  Figure \ref{fig:chosen_results}.
-As can be seen there is considerable variability and no discernable difference in variability between conditions.
+However, to determine whether there was a difference in variability of answers between interface conditions, we plotted their choice on a histogram for each trial, per task and dataset for both conditions, visible in Figure \ref{fig:chosen_results}.
+As can be seen there is considerable variability and no discernible difference in variability between conditions.
 

### 2012-09-18 chi2012.tex
merging

--- 2012-09-18 chi2012.tex
@@
 We then populated each category with 2-5 feeds each depending on feed availability, for a total of 15 total feeds.
-
-For each feed, 3-5 data records were obtained by querying the API directly for records using popular key terms for each category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}.
+For each feed, 3-5 data records were obtained by randomly selecting product IDs by first querying for popular key terms for products in each category, which are listed in the second colum of Table \ref{tbl:prestudy2}.
 While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type were obtained for each category; e.g., product records for the \emph{shopping} category.
-Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics.
-Since, due to optional fields,  individual records could vary in the number of fields.
-An alternative approach to using instance documents was to directly compare schema definitions directly; however, this approach was abandoned after an initial investigation revealed two problems with schema analysis.
-First, many of the sites' schema documentation was not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
-For example, the schema documentation for \emph{shopping.com} listed full product attribute properties, but none of the records actually returned for the products we queried contained such attributes.
+Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics described below.
+An alternative approach to using instance documents was to directly compare schema definitions directly; however, many of the sites' schemas were not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
+Thus we decided to use an instance document random sampling technique instead.
 
@@
 
-\subsection{Data Collection and Analysis}
+\subsection{Data Collection}
 Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
@@
 
-We planned to analyse the qualitative and quantitative data using thematic and statistic analysis, respectively.
-In order to carry out our data analysis, we used the recorded transcripts to populate a spreadsheet summarising the quantitative metrics (see Table \ref{tbl:data} for specific fields).
-In our study, we aimed to identify process and strategies people used to complete their tasks, so that we could evaluate the effectiveness of DP's interface when aligning heterogeneous data for organisational tasks.
-We also categorised people's views voiced during the study and in their exit survey pertaining to DP's usability and interface, and clustered them to evaluate common feedback.
-
 \subsection{Participant Recruitment}
@@
 Since all such participants had passed the requisite proficiency tests to enrol at University, we did not think this would substantially impact results.
+
 Participants were offered a small gift certificate from an online retailer in exchange for their time.
@@
 
-%\subsection{Data Analysis}
+\subsection{Data Analysis}
 
 %% TODO - emax doesn't like this paragraph - kill it? 
-%During the data collection we collated a spreadsheet summarising the participant's gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
+During the data collection we collated a spreadsheet summarising the participant's gender, typical daily computer usage, the tasks allocated and basic actions performed during the task.
 After we had collected the study's data, we performed our data analysis.

### 2012-09-18 chi2012.tex
Made sameas image smaller
prestudy work

@@
 \usepackage{url}      % llt: nicely formatted URLs
+\usepackage{multirow} % multirow tables
 
@@
 Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data.
-Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: contacts, events, music, shopping (retail), social networking and weather, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
+Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: address books (contact info), event calendars (events), music (songs/tracks), shopping (product info) social networking (profiles) and weather (forecasts), finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
 
@@
 \caption{\emph{PS.2 Results} - Analysis of heterogeneity present across data feeds, by category of feed}\label{tbl:prestudy2}
-
 \end{table*}
@@
 \subsection{PS.2 - Technical Challenges of Data Integration}
-Our investigation of integration challenges started with identifying a set of candidate data sources for analysis.
+Our investigation of integration challenges of available structured data feeds started with identifying a set of candidate sources to examine.
 For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} of 50 most popular data feeds measured by the number of developer API keys issued for each.
-In order to get a variety of different kinds of sources represented, we first categorised feeds into 5 topic categories: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.
-We then populated each category with 2-5 feeds each depending on feed availability, for a total of 15 total feeds.
-For each feed, 3-5 data records were obtained by randomly selecting product IDs by first querying for popular key terms for products in each category, which are listed in the second colum of Table \ref{tbl:prestudy2}.
+To increase the variety of sources, we first categorised sources into 5 topic categories: social network services, retailers, event sources (e.g., online calendars and event sites) music, and weather, and selected 2-5 sources from each 2-5 feeds each for a total of 15 total feeds for analysis.
+
+For each feed, 3-5 representative data records were obtained; this was done either by finding record IDs by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query otherwise.
+listed in the second colum of Table \ref{tbl:prestudy2}.
 While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type were obtained for each category; e.g., product records for the \emph{shopping} category.

### 2012-09-18 chi2012.tex
minor edits

--- 2012-09-18 chi2012.tex
@@
 
-We hypothesise that end-users would benefit from this rich data with tools allowing them to effectively browse, use and consume heterogeneous structured data from diverse sources.
+We hypothesise that end-users will benefit from tools that allow them effectively browse, use and consume heterogeneous structured data from diverse sources.
 In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data.
 We used a three-stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of information sources, and the processes that they relied upon to perform such tasks.
-Our results suggested that people preferred to rely upon multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and alternative perspectives.
+Our results suggested that people rely on multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and for alternative perspectives.
 
-Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data.
-Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: contacts, events, music, shopping (retail), social networking and weather, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
+Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to unifying data.
+Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: contacts, events, music, shopping, social networking and weather, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
 
-Third, we developed DataPalette at interface that uses the most common types of heterogeneity to enable serendipitous ``data mixing'' - a form of simple integration sufficient to allow people to easily and effectively combine and compare information sources without the need to write code.
-It was principally designed based upon people's data integration needs and to support data heterogeneity observed in the previous studies.
-We embarked on a user-centric design exercise towards an interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
+Third, we developed DataPalette, an interface that enables serendipitous ``data mixing,'' negating the need for people to write bespoke code to effectively combine and compare heterogenous data.
+It was principally designed to support people's data integration needs for data heterogeneity, using observations made from the previous studies.
+In particular, we designed a user-centric interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
 We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision making tasks using multiple sources.
@@
 In this section, we motivate our work using three fields: PIM, data integration, and end-user mashups and toolkits.
-
-Our primary motivation stems from field studies in PIM, where effectively consolidating and working with data from multiple sources has remained a challenge for over a decade.
-The problem of data fragmentation (\cite{Jones05towardsa}, derives from an inability to effectively consolidate data across multiple applications, services and platforms (e.g., \cite{bergman}, \cite{boardmansasse}).
-In our own previously-conducted field study, (reference anonymised) we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage heterogeneous data from disparate sources -- be them online, offline, or other people.
-Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-maintained information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
-%% TODO MORE HERE
-
-%% don't like the beginning of this > 
+Our primary motivation stems from field studies in PIM, where consolidating and working with data from multiple sources has remained a challenge for over a decade.
 Enabling PIM tools to supporting ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data heterogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
@@
 \end{quote}
+
+Recently, users are experiencing data fragmentation~\cite{Jones05towardsa} across multiple web-based applications and services, and struggle to consolidate across them~\cite{bergman,boardmansasse}.
+In our own previously-conducted field study, (reference anonymised) we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage both online and offline data.
+Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-maintained information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
+
+%In this section, we motivate our work using three fields: PIM, data integration, and end-user mashups and toolkits.
+Our primary motivation stems from field studies in PIM, where consolidating and working with data from multiple sources has remained a challenge for over a decade.
+Recently however, users are experiencing data fragmentation~\cite{Jones05towardsa} across multiple web-based applications and services, and struggle to consolidate across them~\cite{bergman,boardmansasse}.
+In our own previously-conducted field study, (reference anonymised) we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage both online and offline data.
+Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-maintained information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
+%% TODO MORE HERE
+
+%% don't like the beginning of this > 
+%Enabling PIM tools to supporting ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data heterogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
+As described by Alon Halvey:
+
+%\begin{quote} 
+  %The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs.
+Hence, even if they model overlapping domains, they will model them in different ways.
+Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modelling goal.
+\cite{alonhalevy}
+%\end{quote}
 
@@
 We then populated each category with 2-5 feeds each depending on feed availability, for a total of 15 total feeds.
-For each feed, 3-5 data records were obtained by randomly selecting product IDs by first querying for popular key terms for products in each category, which are listed in the second colum of Table \ref{tbl:prestudy2}.
+
+For each feed, 3-5 data records were obtained by querying the API directly for records using popular key terms for each category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}.
 While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type were obtained for each category; e.g., product records for the \emph{shopping} category.
-Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics described below.
-An alternative approach to using instance documents was to directly compare schema definitions directly; however, many of the sites' schemas were not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
-Thus we decided to use an instance document random sampling technique instead.
+Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics.
+Since, due to optional fields,  individual records could vary in the number of fields.
+An alternative approach to using instance documents was to directly compare schema definitions directly; however, this approach was abandoned after an initial investigation revealed two problems with schema analysis.
+First, many of the sites' schema documentation was not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
+For example, the schema documentation for \emph{shopping.com} listed full product attribute properties, but none of the records actually returned for the products we queried contained such attributes.
 

### 2012-09-18 chi2012.tex
updated

--- 2012-09-18 chi2012.tex
@@
 For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} of 50 most popular data feeds measured by the number of developer API keys issued for each.
-To increase the variety of sources, we first categorised sources into 5 topic categories: social network services, retailers, event sources (e.g., online calendars and event sites) music, and weather, and selected 2-5 sources from each 2-5 feeds each for a total of 15 total feeds for analysis.
+To increase the variety of sources, we first categorised sources into 5 topic categories: social network services, retailers, event sources (e.g., online calendars and event sites) music, and weather, and selected 2-5 sources from each 2-5 feeds each for a total of 20 total feeds for analysis.
 
-For each feed, 3-5 representative data records were obtained; this was done either by finding record IDs by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query otherwise.
+For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, , this was user profiles, this was done either by finding record IDs by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query otherwise.
 listed in the second colum of Table \ref{tbl:prestudy2}.
@@
 
-\begin{figure*}[htbp]
+\begin{figure}[htbp]
 \begin{center}
-\includegraphics[width=18cm]{img/sameas}
+\includegraphics[width=8.5cm]{img/sameas}
 \caption{\emph{Coreference consolidation with drag-and-drop Same-As} - When two instances represent the same entity, users can drag one on top of the other, and they are combined - effectively displaying the union of properties of both source entities.
@@
 \end{center}
-\end{figure*}
+\end{figure}
 

### 2012-09-18 chi2012.tex

--- 2012-09-18 chi2012.tex
@@
 
-Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to unifying data.
+% Second, we conducted a quantitative, structural analysis of various popular personal data sources available today to identify potential barriers to effective unification of data.
+Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: address books (contact info), event calendars (events), music (songs/tracks), shopping (product info) social networking (profiles) and weather (forecasts), finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
+
+Second, we conducted a structural analysis of various popular personal data sources available today to identify potential barriers to unifying data.
 Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: contacts, events, music, shopping, social networking and weather, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
@@
 \subsection{PS.2 - Technical Challenges of Data Integration}
-Our investigation of integration challenges started with identifying a set of candidate data sources for analysis.
+Our investigation of integration challenges of available structured data feeds started with identifying a set of candidate sources to examine.
 For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} of 50 most popular data feeds measured by the number of developer API keys issued for each.
-In order to get a variety of different kinds of sources represented, we first categorised feeds into 5 topic categories: social networks (people), retail (shopping), events (calendar events, happenings, etc), music, and weather.
-We then populated each category with 2-5 feeds each depending on feed availability, for a total of 15 total feeds.
+To increase the variety of sources, we first categorised sources into 5 topic categories: social network services, retailers, event sources (e.g., online calendars and event sites) music, and weather, and selected 2-5 sources from each 2-5 feeds each for a total of 20 total feeds for analysis.
 
-For each feed, 3-5 data records were obtained by querying the API directly for records using popular key terms for each category's primary \emph{corresponding type}, listed in the second colum of Table \ref{tbl:prestudy2}.
+For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, , this was user profiles, this was done either by finding record IDs by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query otherwise.
+listed in the second colum of Table \ref{tbl:prestudy2}.
 While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type were obtained for each category; e.g., product records for the \emph{shopping} category.
-Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics.
-Since, due to optional fields,  individual records could vary in the number of fields.
-An alternative approach to using instance documents was to directly compare schema definitions directly; however, this approach was abandoned after an initial investigation revealed two problems with schema analysis.
-First, many of the sites' schema documentation was not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
-For example, the schema documentation for \emph{shopping.com} listed full product attribute properties, but none of the records actually returned for the products we queried contained such attributes.
+Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics described below.
+An alternative approach to using instance documents was to directly compare schema definitions directly; however, many of the sites' schemas were not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
+Thus we decided to use an instance document random sampling technique instead.
 

### 2012-09-18 chi2012.tex
looseness added

--- 2012-09-18 chi2012.tex
@@
 Hence, even if they model overlapping domains, they will model them in different ways.
-Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modelling goal.
-\cite{alonhalevy}
+Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modelling goal~\cite{alonhalevy}.\looseness=-1
 \end{quote}

### 2012-09-18 chi2012.tex
Moved titles of survey results and resized:wq
Shrunk multipathing image
Table made for PS2

@@
 
-%% TODO THIS IS PLACEHOLDER DATA >> 
 \begin{table*}
 \small
-\begin{tabular}{| p{2.0cm} | p{1.5cm} | p{3.6cm} | p{2cm} | p{2cm} | l | l | l | l |}
+\begin{tabular}{| p{2.0cm} | p{1.8cm} | p{1.8cm} | p{1.8cm} | p{1.8cm} | p{1.8cm} | p{1.8cm} | p{1.8cm} |}
 \hline
-source category  & data type & sources & API type & avg fields/rec & avg depth & term. & struct & semantic \\
+				& {\bf source}		& {\bf average props/record}	& {\bf average depth/record}	& {\bf equivalent to}	& {\bf prop name inconsistent}	& {\bf structural inconsistency} 	& {\bf model inconsistency} \\
 \hline
-social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
-event calendars & events & Eventbrite, Google Calendar  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
-retail (shopping)  & product profile & Amazon, Shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
-music catalogues & songs and albums & Echonest, rdio, Soundcloud, Spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
-weather & forecasts & W.Underground, Weatherbug, Yahoo! Weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
+social networking	& google+			& 12						& 1.52					& tw:5, li:4, fb:4		& tw:3, li:2, fb:2				& tw:0, li:1, fb:1				& tw:0, li:0, fb:0 \\
+(profiles)			& twitter			& 39						& 1.37					& li:4, fb:4 			& li:2, fb:1					& li:0, fb:0					& li:0, fb:0 \\
+				& linkedin			& 18						& 2.42					& fb:7			& fb:5					& fb:3					& fb:0 \\
+				& facebook		& 22						& 1.5						&				&						&						& \\
+\hline
+calendars			& google calendar	& 15						& 1.46					& eb:9			& eb:6					& eb:1					& eb:1 \\
+(events)			& eventbrite		& 32						& 1.43					&				&						& 						& \\
+\hline
+retail				& milo			& 5						& 1.2						& z:2, s:3, e:1, a:3	& z:2, s:2, e:1, a:3			& z:0, s:0, e:0, a:2			& z:0, s:0, e:0, a:0 \\
+(products)			& zappos			& 6						& 1						& s:2, e:2, a:3		& s:2, e:2, a:2				& s:0, e:0, a:0				& s:0, e:0, a:0 \\
+				& shopping.com	& 11						& 3.74					& e:2, a:8			& e:2, a:8					& e:0, a:0					& e:0, a:0 \\
+				& etsy			& 31						& 1.13					& a:11			& a:10					& a:1						& a:2 \\
+				& amazon			& 108					&						&				& 						&						& \\
+\hline
+contacts			& google contacts	& 7						& 1.46					& yp:2			& yp:2					& yp: 0					& yp: 0 \\
+				& yellowpages		& 25						& 1						&				&						&						& \\
+\hline
+music			& echonest		& 4						& 1.83					& so:3, r:3, sp:3		& so:3, r:3, sp:3				& so:0, r:0, sp:0				& so:0, r:0, sp:0 \\
+(songs/			& soundcloud		& 42						& 2.14					& r:12, sp:7		& r:11, sp:6 				& r:1, sp:1					& r:1, sp:0 \\
+tracks)			& rdio			& 27						& 2						& sp: 7			& sp: 6					& sp: 1					& sp: 1 \\
+				& spotify			& 10						& 2.44					&				&						&						& \\
+\hline
+weather			& weatherbug		& 8						& 2						& wu: 5, y: 5		& wu: 5, y: 5				& wu: 5, y: 4				& wu: 0, y: 0 \\
+(forecasts)		& wundergound	& 52						& 1.38					& y:12			& y:12					& y:7						& y:3 \\
+				& yahoo weather	& 22						& 1.58					&				&						&						& \\
 \hline
@@
 \end{table*}
-%% <<
 
@@
 \begin{center}
-\includegraphics[width=8cm]{img/multipathing}
+\includegraphics[width=8.5cm]{img/multipathing}
 \caption{To see properties, users click on the name of the property. When instances do not have a particular property, additional property names can be added.}
@@
 \begin{center}
-\includegraphics[width=8.5cm]{img/survey}
+\includegraphics[width=7.5cm]{img/survey}
 \caption{\emph{Survey results} - Answers to questions on a 5-point Likert scale, from 1-\emph{strongly disagree} to 5-\emph{strongly agree}.}

### 2012-09-18 chi2012.tex
ps2 description

--- 2012-09-18 chi2012.tex
@@
 
-For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, , this was user profiles, this was done either by finding record IDs by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query otherwise.
-listed in the second colum of Table \ref{tbl:prestudy2}.
-While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type were obtained for each category; e.g., product records for the \emph{shopping} category.
-Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics described below.
+For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for music sites this was song info, etc.
+Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.
+For each obtained data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.
+For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record has few attributes about the actual product compared to its ItemAttributes sub-property, which was used instead.
+Thus, all comparisons were made to maximize comparative similarity.
+Depth was computed as average from root
+
+In addition to the complexity metrics, we mapped out the amount of overlap among the sources in each category by first manually determiningr \emph{equivalent properties}.
+This was done manually by creating a pairwise alignment table between every two data sources in each category.
+Each row of the alignment table comprised a property in some schema, and all its closest matches in the other schemas.
+We used both property names and attribute descriptions (in documentation, if available) to make matching decisions; two properties  were considered equivalent if they intended to model the same piece of information;
+
+, followed by pairwise heterogeneity metrics described below.
 An alternative approach to using instance documents was to directly compare schema definitions directly; however, many of the sites' schemas were not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
@@
 \subsubsection{Enumerated-type value consolidation}
-
 Related to the concept of co-reference, PS.2 revealed that enumerated values of properties were often inconsistent, e.g.
@@
 
-Ascertaining DP's impact on task performance (h3) was the most difficult to ascertain, given the small sample size of our study and large number of factors influencing each person's performance.
+Assessing whether DP's improved task performance (h3) was the most difficult to ascertain, given the small sample size of our study and large number of factors influencing each person's performance.
 While tasks were completed on average slightly faster than the control interface, these differences were not statistically significant.

### 2012-09-18 chi2012.tex
no lines

@@
 \small
-\begin{tabular}{| p{2.0cm} | p{1.8cm} | p{1.8cm} | p{1.8cm} | p{1.8cm} | p{1.8cm} | p{1.8cm} | p{1.8cm} |}
-\hline
+\begin{tabular}{p{2.0cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}}
 				& {\bf source}		& {\bf average props/record}	& {\bf average depth/record}	& {\bf equivalent to}	& {\bf prop name inconsistent}	& {\bf structural inconsistency} 	& {\bf model inconsistency} \\
@@
 				& yahoo weather	& 22						& 1.58					&				&						&						& \\
-\hline
 \end{tabular}

### 2012-09-18 chi2012.tex
Fixed table alignment

@@
 \begin{tabular}{p{2.0cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}}
-				& {\bf source}		& {\bf average props/record}	& {\bf average depth/record}	& {\bf equivalent to}	& {\bf prop name inconsistent}	& {\bf structural inconsistency} 	& {\bf model inconsistency} \\
+				&				& {\bf average}				& {\bf average }				& 				& {\bf prop name}			& {\bf structural} 			& {\bf model } \\
+				& {\bf source}		& {\bf props/record}			& {\bf depth/record}			& {\bf equivalent to}	& {\bf inconsistent}			& {\bf inconsistency} 			& {\bf inconsistency} \\
 \hline

### 2012-09-18 chi2012.tex
Annotated screenshot
work on ps2

--- 2012-09-18 chi2012.tex
@@
 \begin{tabular}{p{2.0cm} p{1.8cm} p{1.8cm} p{1.8cm} p{1.8cm} p{1.8cm} p{1.8cm} p{1.8cm}}
-				& {\bf source}		& {\bf average props/record}	& {\bf average depth/record}	& {\bf equivalent to}	& {\bf prop name inconsistent}	& {\bf structural inconsistency} 	& {\bf model inconsistency} \\
+				& {\bf source}		& {\bf avg props/record}	& {\bf avg depth/record}	& {\bf equivalent props}	& {\bf prop name inconsistent}	& {\bf structural inconsistency} 	& {\bf model inconsistency} \\
 \hline
@@
 For each obtained data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.
-For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record has few attributes about the actual product compared to its ItemAttributes sub-property, which was used instead.
+For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record had few attributes about the actual product compared to its ``ItemAttributes'' sub-property, which was used instead.
 Thus, all comparisons were made to maximize comparative similarity.
-Depth was computed as average from root
+Depth was computed as the average number of levels from record root to leaves across all properties, and was measured, with width, to convey the overall size/complexity of each record.
 
-In addition to the complexity metrics, we mapped out the amount of overlap among the sources in each category by first manually determiningr \emph{equivalent properties}.
-This was done manually by creating a pairwise alignment table between every two data sources in each category.
-Each row of the alignment table comprised a property in some schema, and all its closest matches in the other schemas.
-We used both property names and attribute descriptions (in documentation, if available) to make matching decisions; two properties  were considered equivalent if they intended to model the same piece of information;
+The third metric measured the amount of overlap among the sources in each category.
+This was done by determining \emph{equivalent properties} between schemas, which we did manually by creating an alignment table for all properties across all data sources in each category.
+Each row of the alignment table comprised a property of a record from one data source, and all its closest matching properties from the others.
+We used both property names and attribute descriptions (in documentation, if available) to make matching decisions; two properties were aligned if they seemed to model the same piece of information.
 
-, followed by pairwise heterogeneity metrics described below.
+Once the equivalent properties were established, we look at how many of these properties were terminologically consistent, that is, identified using the same property name.
+If the names were different, we scored these as , followed by pairwise heterogeneity metrics described below.
 An alternative approach to using instance documents was to directly compare schema definitions directly; however, many of the sites' schemas were not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.

### 2012-09-18 chi2012.tex
Fix spello

--- 2012-09-18 chi2012.tex
@@
 \begin{center}
-\includegraphics[width=18cm]{img/screenshot}
+\includegraphics[width=14cm]{img/screenshot}
 \caption{DataPalette Workspace, a file manager-inspired Drag and Drop interface for structured data mixing.
-Collapsable displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a windoow (group) on the workspace.}
+Collapsible displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a windoow (group) on the workspace.}
 \label{fig:workspace}
@@
 
-% Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
+% Data sources were represented as collapsible panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
 Figure \ref{fig:workspace} displays the entire UI overview.

### 2012-09-18 chi2012.tex
better annotated screenshot :)
minor edits

--- 2012-09-18 chi2012.tex
@@
 For our first pre-study, we first sought an updated understanding of how and whether people still used multiple information sources regularly when they sought information on the Web.
-There were several reasons for this inquiry; first, we wondered whether people were increasingly relying only on the handful of the massive ``supersites'' (Facebook, Google, Wikipedia) for most of their activities, instead of  searching across distributed sources of information.
+There were two reasons for this inquiry; first, whether people were increasingly either relying only on the handful of the massive ``supersites'' (Facebook, Google, Wikipedia) for most of their activities, or searching across distributed sources of information.
 If this were the case, much of the benefit could be gained by merely integrating PIM tools with the handful of sites, rather than the substantially more challenging problem of integrating data from arbitrary information sources on the Web.
-Second, if people did rely on multiple sites, we wished to know why, and the typical kinds of tasks people did that required diverse information collection - were these kinds of tasks rare?
+Second, whether people relied on multiple sites and why.
+If they did what kinds of tasks did people required diverse information collection and were these kinds of tasks rare?
 If so, this would suggest that perhaps specialised interfaces could be concocted for this purpose rather than it being a PIM need.
@@
 The purpose of this follow-on  was to identify not specific characteristics of particular sites in question, but to identify, in general, the degree of complexity of the data feeds currently available across a variety of domains, and typical integration problems that might arise when mixing this data.
-In the following sections, we summarise our methodologies and results for each of these two pre-studies.
+The next two sections, we present our methodologies and results for our pre-studies.
 
@@
 We designed a semi-structured interview about tasks people performed online that required the use of multiple websites, and the processes people used to perform such tasks.
-We analysed the data collected from this study using a thematic approach, as described below.
-
-Our pre-study investigated the types of tasks people performed online that required heterogeneous data sources, and which tools they would use to organise a large social event involving 15 or more people.
-We interviewed 8 participants, using a general interview guide approach based loosely on a set of questions.
-This approach allowed both the participant and interviewer a degree of freedom in conversation and each interview could adapt to the participant's experiences, while collecting the same general areas of information.
-Each interview was recorded and notes taken, the interviewer was trained so that they were knowledgeable about the importance of the study and to minimise bias.
+We analysed the data collected from this study using a thematic approach.
+In our pre-study, we investigated the types of tasks people performed online that required heterogeneous data sources, and which tools they would use to organise a social event involving around 15 people.
+We planned to interview 8 participants, using a general interview guide approach based on a set of questions.
+This approach allowed the interviewer to adapt the interview to the participant's experiences, while collecting the same general areas of information.
+Each interview was recorded and notes taken, the interviewer was trained in the purpose of the study and to minimise bias.
 
 \subsubsection{Results}
-The participant demographic consisted of 8 people, 7 of which were male and 1 female.
+The participant demographic consisted of 8 people (7 male and 1 female).
 They were split evenly into two age groups of 18-25 and 26-32.
@@
 All of the participants used Twitter to listen to broadcasted tweets, and 6 out of 8 people regularly tweeted.
-The following two sections, describes the tasks people perform online with heterogeneous data and the processes they would use to organise a large social event.
+The following two sections, describe the tasks people perform online with heterogeneous data and how they would use to organise a large social event.
 
-\subsubsection{Question 1: Tasks}
+\subsubsection{Question 1: Heterogenous Data Tasks}
 All participants reported having used multiple websites to accomplish a task.
@@
 If their friends were not on Facebook they would email them.
-All but one participant said they would not use Doodle\footnote{Doodle - \emph{www.doodle.com}} to organise day and time of an event, because they felt that people did not fill in the form and it was more suit to organising work event not social.
+All but one participant said they would not use Doodle\footnote{Doodle - \emph{www.doodle.com}} to the organise day and time of an event, because they felt that people did not fill in the form and it was more suit to organising work events.
 
@@
 They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.
-They selected venues, locations, restaurants and activities based on their own knowledge, friend's recommendations, Google maps, reviews obtained through web search and sites such as tripadvisor.
+They selected venues, locations, restaurants and activities based on their own knowledge, friend's recommendations, Google maps, reviews obtained through web search and sites such as TripAdvisor.
 People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.
-They considered how to get there and looked up train or bus times on the Web.
-One participant organised carpools informally and at last minute via Facebook or text messages.
-
-When organising an event the weather was not that important because it they did not trust its accuracy for events plan further in the future than one week.
+They considered how to get there and looked up train or bus times on the Web, or organised carpools.
+When organising an event the weather was not that important because they did not trust the accuracy of predictions.
 Also, their friend's preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.
@@
 People said that they would like a website that could support the evolution of an event, that allowed them to post drafts of an event.
-They felt that organising an event should be a process that changes over time and did not need rushing.
-Three of the participants said that they would like a recommendation system for places to eat and activities.
+They felt that organising an event should be a process that changes over time.
+Three participants said that they would like a recommendation system for places to eat and activities.
 Another strong requirement was that it had to be ubiquitous so that everyone could use it because they wanted a single place to communicate with their friends.
-While Facebook was the most popular social networking site for organising events, half of the participants felt that it was not the best solution and would prefer a collaborative environment, which could be wiki based.
+While Facebook was the most popular social networking site for organising events, half of the participants felt that it was not the best solution and would prefer a collaborative environment.
 
+%% TODO THIS IS PLACEHOLDER DATA >> 
 \begin{table*}
 \small
-\begin{tabular}{p{2.0cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}}
-				&				& {\bf average}				& {\bf average }				& 				& {\bf prop name}			& {\bf structural} 			& {\bf model } \\
-				& {\bf source}		& {\bf props/record}			& {\bf depth/record}			& {\bf equivalent to}	& {\bf inconsistent}			& {\bf inconsistency} 			& {\bf inconsistency} \\
+\begin{tabular}{| p{2.0cm} | p{1.5cm} | p{3.6cm} | p{2cm} | p{2cm} | l | l | l | l |}
 \hline
-social networking	& google+			& 12						& 1.52					& tw:5, li:4, fb:4		& tw:3, li:2, fb:2				& tw:0, li:1, fb:1				& tw:0, li:0, fb:0 \\
-(profiles)			& twitter			& 39						& 1.37					& li:4, fb:4 			& li:2, fb:1					& li:0, fb:0					& li:0, fb:0 \\
-				& linkedin			& 18						& 2.42					& fb:7			& fb:5					& fb:3					& fb:0 \\
-				& facebook		& 22						& 1.5						&				&						&						& \\
+source category  & data type & sources & API type & avg fields/rec & avg depth & term. & struct & semantic \\
 \hline
-calendars			& google calendar	& 15						& 1.46					& eb:9			& eb:6					& eb:1					& eb:1 \\
-(events)			& eventbrite		& 32						& 1.43					&				&						& 						& \\
+social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
+event calendars & events & Eventbrite, Google Calendar  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
+retail (shopping)  & product profile & Amazon, Shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
+music catalogues & songs and albums & Echonest, rdio, Soundcloud, Spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
+weather & forecasts & W.Underground, Weatherbug, Yahoo! Weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
 \hline
-retail				& milo			& 5						& 1.2						& z:2, s:3, e:1, a:3	& z:2, s:2, e:1, a:3			& z:0, s:0, e:0, a:2			& z:0, s:0, e:0, a:0 \\
-(products)			& zappos			& 6						& 1						& s:2, e:2, a:3		& s:2, e:2, a:2				& s:0, e:0, a:0				& s:0, e:0, a:0 \\
-				& shopping.com	& 11						& 3.74					& e:2, a:8			& e:2, a:8					& e:0, a:0					& e:0, a:0 \\
-				& etsy			& 31						& 1.13					& a:11			& a:10					& a:1						& a:2 \\
-				& amazon			& 108					&						&				& 						&						& \\
-\hline
-contacts			& google contacts	& 7						& 1.46					& yp:2			& yp:2					& yp: 0					& yp: 0 \\
-				& yellowpages		& 25						& 1						&				&						&						& \\
-\hline
-music			& echonest		& 4						& 1.83					& so:3, r:3, sp:3		& so:3, r:3, sp:3				& so:0, r:0, sp:0				& so:0, r:0, sp:0 \\
-(songs/			& soundcloud		& 42						& 2.14					& r:12, sp:7		& r:11, sp:6 				& r:1, sp:1					& r:1, sp:0 \\
-tracks)			& rdio			& 27						& 2						& sp: 7			& sp: 6					& sp: 1					& sp: 1 \\
-				& spotify			& 10						& 2.44					&				&						&						& \\
-\hline
-weather			& weatherbug		& 8						& 2						& wu: 5, y: 5		& wu: 5, y: 5				& wu: 5, y: 4				& wu: 0, y: 0 \\
-(forecasts)		& wundergound	& 52						& 1.38					& y:12			& y:12					& y:7						& y:3 \\
-				& yahoo weather	& 22						& 1.58					&				&						&						& \\
 \end{tabular}
@@
 
-For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for music sites this was song info, etc.
-Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.
-For each obtained data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.
-For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record has few attributes about the actual product compared to its ItemAttributes sub-property, which was used instead.
-Thus, all comparisons were made to maximize comparative similarity.
-Depth was computed as average from root
-
-In addition to the complexity metrics, we mapped out the amount of overlap among the sources in each category by first manually determiningr \emph{equivalent properties}.
-This was done manually by creating a pairwise alignment table between every two data sources in each category.
-Each row of the alignment table comprised a property in some schema, and all its closest matches in the other schemas.
-We used both property names and attribute descriptions (in documentation, if available) to make matching decisions; two properties  were considered equivalent if they intended to model the same piece of information;
-
-, followed by pairwise heterogeneity metrics described below.
+For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, , this was user profiles, this was done either by finding record IDs by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query otherwise.
+listed in the second column of Table \ref{tbl:prestudy2}.
+While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type were obtained for each category; e.g., product records for the \emph{shopping} category.
+Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics described below.
 An alternative approach to using instance documents was to directly compare schema definitions directly; however, many of the sites' schemas were not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
@@
 \item \emph{Average Width of record} - Average number of properties per record.
-\item \emph{Average Depth of record} - Average depth of data record (number of subproperties in self-contained record).
+\item \emph{Average Depth of record} - Average depth of data record (number of sub-properties in self-contained record).
 (In JSON/XML representations, this is the tree depth; for RDF  representations this is merely the number of embedded entity resource definitions included in the top-level query).
@@
 
+%\section{Designing the DataPalette Interface}
 \section{Designing DataPalette : An Interface for Serendipitous Data Mixing}
@@
 \caption{DataPalette Workspace, a file manager-inspired Drag and Drop interface for structured data mixing.
-Collapsable displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a windoow (group) on the workspace.}
+Collapsable displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a window (group) on the workspace.}
 \label{fig:workspace}
@@
 This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
-[0
 In the following sections we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.
@@
 \subsubsection{Enumerated-type value consolidation}
+
 Related to the concept of co-reference, PS.2 revealed that enumerated values of properties were often inconsistent, e.g.
@@
 The second metric was the number of factors the participant considered in making their final choice.
-To determine this, we asked participants to tell us precisely why their choice was made, and matched their justifications to properties in the data.
+To determine this, we asked participants explain precisely why they made their choice, and matched their justifications to properties in the data.
 This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
@@
 \subsection{Strategies: Successive Elimination vs Tallying}
-As described in Discussion later, we noticed several different strategies that people used to evalutate each candidate selection.
+As described in later this section, we noticed several different strategies people used to evaluate each candidate selection.
 One common strategy was \emph{successive elimination} - to look at a single dimension at a time, starting with the most important, ruling out candidates that did not meet the required minimum for that value.
 However, when there is no clear ``most important'' aspect, this  ``greedy-style'' can result in suboptimal decisions.
-By comparison, a different strategy, which we called \emph{tallying the pros and cons}, kept all candidates around, and counted up the advantages and disadvantages of each, potentially weighing each factor by its perceived importance.
+Another strategy, which we called \emph{tallying the pros and cons}, kept all candidates around, and counted up the advantages and disadvantages of each, potentially weighing each factor by its perceived importance.
 This strategy is less vulnerable to getting stuck in local maxima than the former.
 
-To understand whether interface influenced choice of strategy, we measured the number of candidates that the participant considered \emph{at final result}.
+To understand whether the interface influenced choice of strategy, we measured the number of candidates that the participant considered \emph{at final result}.
 This we found was a strong marker for use of each strategy; people using the successive elimination strategy arrived with only 1-2 candidates at decision time, while those that tallied still maintained the entire starting set.
@@
 However, one participant said that they would not use it for things like deciding on a restaurant because ``it would cheapen the experience'' by removing the element of spontaneity of choice, but would use it for making more important decisions.
-Participants had a large number of suggestions for improvement, which In order to improve our solutions the participants suggested making the origin of the information clearer, particularly when combining multiple data sources.
+In order to improve our solutions the participants suggested making the origin of the information clearer, particularly when combining multiple data sources.
 The participants also requested the ability to sort instances, as well as explanations of ratings and scoring systems (however some of these are not even transparent on websites).
@@
 
-Assessing whether DP's improved task performance (h3) was the most difficult to ascertain, given the small sample size of our study and large number of factors influencing each person's performance.
+Ascertaining DP's impact on task performance (h3) was the most difficult to ascertain, given the small sample size of our study and large number of factors influencing each person's performance.
 While tasks were completed on average slightly faster than the control interface, these differences were not statistically significant.
 However, more data sources seemed to be consulted (we use ``seemed'' carefully because these findings approached significance) during trials with DP over the control interface, which meant that people were looking at more diverse information than the control conditions.
-We also found that particiapnts justified their choices with a greater number of factors in DP than the control interface, suggesting that decisions may have been made considering a greater number of factors.
+We also found that participants justified their choices with a greater number of factors in DP than the control interface, suggesting that decisions may have been made considering a greater number of factors.
 Participants entertained more possibilities for longer in the DP trial, while participants were more prone to eliminate candidates early in the baseline interface.
-Finally, people took fewer notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP, which strongly suggests that DP made the task easier to complete.
+Finally, people took fewer notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP, which strongly indicates that DP made the task easier to complete.
 

### 2012-09-18 chi2012.tex

@@
 
-%% TODO THIS IS PLACEHOLDER DATA >> 
 \begin{table*}
 \small
-\begin{tabular}{| p{2.0cm} | p{1.5cm} | p{3.6cm} | p{2cm} | p{2cm} | l | l | l | l |}
+\begin{tabular}{p{2.0cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}}
+				&				& {\bf average}				& {\bf average }				& 				& {\bf prop name}			& {\bf structural} 			& {\bf model } \\
+				& {\bf source}		& {\bf props/record}			& {\bf depth/record}			& {\bf equivalent to}	& {\bf inconsistent}			& {\bf inconsistency} 			& {\bf inconsistency} \\
 \hline
-source category  & data type & sources & API type & avg fields/rec & avg depth & term. & struct & semantic \\
+social networking	& google+			& 12						& 1.52					& tw:5, li:4, fb:4		& tw:3, li:2, fb:2				& tw:0, li:1, fb:1				& tw:0, li:0, fb:0 \\
+(profiles)			& twitter			& 39						& 1.37					& li:4, fb:4 			& li:2, fb:1					& li:0, fb:0					& li:0, fb:0 \\
+				& linkedin			& 18						& 2.42					& fb:7			& fb:5					& fb:3					& fb:0 \\
+				& facebook		& 22						& 1.5						&				&						&						& \\
 \hline
-social networking  & person's profile & Facebook, LinkedIn, Google+ & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
-event calendars & events & Eventbrite, Google Calendar  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
-retail (shopping)  & product profile & Amazon, Shopping.com, Etsy, Groupon, Zappos  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
-music catalogues & songs and albums & Echonest, rdio, Soundcloud, Spotify  & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
-weather & forecasts & W.Underground, Weatherbug, Yahoo! Weather & REST (JSON/XML) & 10 & 1.2 & 0.90 & 0.10 & 1.0 \\
+calendars			& google calendar	& 15						& 1.46					& eb:9			& eb:6					& eb:1					& eb:1 \\
+(events)			& eventbrite		& 32						& 1.43					&				&						& 						& \\
 \hline
+retail				& milo			& 5						& 1.2						& z:2, s:3, e:1, a:3	& z:2, s:2, e:1, a:3			& z:0, s:0, e:0, a:2			& z:0, s:0, e:0, a:0 \\
+(products)			& zappos			& 6						& 1						& s:2, e:2, a:3		& s:2, e:2, a:2				& s:0, e:0, a:0				& s:0, e:0, a:0 \\
+				& shopping.com	& 11						& 3.74					& e:2, a:8			& e:2, a:8					& e:0, a:0					& e:0, a:0 \\
+				& etsy			& 31						& 1.13					& a:11			& a:10					& a:1						& a:2 \\
+				& amazon			& 108					&						&				& 						&						& \\
+\hline
+contacts			& google contacts	& 7						& 1.46					& yp:2			& yp:2					& yp: 0					& yp: 0 \\
+				& yellowpages		& 25						& 1						&				&						&						& \\
+\hline
+music			& echonest		& 4						& 1.83					& so:3, r:3, sp:3		& so:3, r:3, sp:3				& so:0, r:0, sp:0				& so:0, r:0, sp:0 \\
+(songs/			& soundcloud		& 42						& 2.14					& r:12, sp:7		& r:11, sp:6 				& r:1, sp:1					& r:1, sp:0 \\
+tracks)			& rdio			& 27						& 2						& sp: 7			& sp: 6					& sp: 1					& sp: 1 \\
+				& spotify			& 10						& 2.44					&				&						&						& \\
+\hline
+weather			& weatherbug		& 8						& 2						& wu: 5, y: 5		& wu: 5, y: 5				& wu: 5, y: 4				& wu: 0, y: 0 \\
+(forecasts)		& wundergound	& 52						& 1.38					& y:12			& y:12					& y:7						& y:3 \\
+				& yahoo weather	& 22						& 1.58					&				&						&						& \\
 \end{tabular}
@@
 
-For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, , this was user profiles, this was done either by finding record IDs by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query otherwise.
-listed in the second column of Table \ref{tbl:prestudy2}.
-While many of the APIs also provided access to other kinds of data, such as customer reviews, only the primary type were obtained for each category; e.g., product records for the \emph{shopping} category.
-Once data records were obtained, two complexity metrics were computed for each, followed by pairwise heterogeneity metrics described below.
+For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for music sites this was song info, etc.
+Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.
+For each obtained data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.
+For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record has few attributes about the actual product compared to its ItemAttributes sub-property, which was used instead.
+Thus, all comparisons were made to maximize comparative similarity.
+Depth was computed as average from root
+
+In addition to the complexity metrics, we mapped out the amount of overlap among the sources in each category by first manually determiningr \emph{equivalent properties}.
+This was done manually by creating a pairwise alignment table between every two data sources in each category.
+Each row of the alignment table comprised a property in some schema, and all its closest matches in the other schemas.
+We used both property names and attribute descriptions (in documentation, if available) to make matching decisions; two properties  were considered equivalent if they intended to model the same piece of information;
+
+, followed by pairwise heterogeneity metrics described below.
 An alternative approach to using instance documents was to directly compare schema definitions directly; however, many of the sites' schemas were not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
@@
 %\section{Designing the DataPalette Interface}
+
 \section{Designing DataPalette : An Interface for Serendipitous Data Mixing}
@@
 \includegraphics[width=18cm]{img/screenshot}
+
 \caption{DataPalette Workspace, a file manager-inspired Drag and Drop interface for structured data mixing.
-Collapsable displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a window (group) on the workspace.}
+Collapsible displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a window (group) on the workspace.}
+
 \label{fig:workspace}

### 2012-09-18 chi2012.tex

--- 2012-09-18 chi2012.tex
@@
 For each obtained data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.
-For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record has few attributes about the actual product compared to its ItemAttributes sub-property, which was used instead.
+For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record had few attributes about the actual product compared to its ``ItemAttributes'' sub-property, which was used instead.
 Thus, all comparisons were made to maximize comparative similarity.
-Depth was computed as average from root
+Depth was computed as the average number of levels from record root to leaves across all properties, and was measured, with width, to convey the overall size/complexity of each record.
 
-In addition to the complexity metrics, we mapped out the amount of overlap among the sources in each category by first manually determiningr \emph{equivalent properties}.
-This was done manually by creating a pairwise alignment table between every two data sources in each category.
-Each row of the alignment table comprised a property in some schema, and all its closest matches in the other schemas.
-We used both property names and attribute descriptions (in documentation, if available) to make matching decisions; two properties  were considered equivalent if they intended to model the same piece of information;
+The third metric measured the amount of overlap among the sources in each category.
+This was done by determining \emph{equivalent properties} between schemas, which we did manually by creating an alignment table for all properties across all data sources in each category.
+Each row of the alignment table comprised a property of a record from one data source, and all its closest matching properties from the others.
+We used both property names and attribute descriptions (in documentation, if available) to make matching decisions; two properties were aligned if they seemed to model the same piece of information.
 
-, followed by pairwise heterogeneity metrics described below.
-An alternative approach to using instance documents was to directly compare schema definitions directly; however, many of the sites' schemas were not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
-Thus we decided to use an instance document random sampling technique instead.
+Once the equivalent properties were established, we look at how many of these properties were terminologically inconsistent, that is, identified using different property names.
+Then, ignoring property name, we looked to see if its values were structurally comparable.
+If two fields were of the same data type and, if complex, had equivalent subfields, they were considered comparable; otherwise they were considered \emph{structurally inconsistent}.
+An example of such a structural inconsistency were``birthday'' fields in Facebook and LinkedIn; Facebook represented it as a single string formatted ``MM/DD/YYYY'', while LinkedIn represented it as a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.
 
-The per-source metrics were as follows:
-\begin{itemize}
-\item \emph{Average Width of record} - Average number of properties per record.
-\item \emph{Average Depth of record} - Average depth of data record (number of subproperties in self-contained record).
-(In JSON/XML representations, this is the tree depth; for RDF  representations this is merely the number of embedded entity resource definitions included in the top-level query).
-\end{itemize}
-
-Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from definitions in \cite{george2005understanding} and feature-metrics used in statistical instance alignment algorithms (such as summarised in \cite{bhattacharya2007collective}).
-Such metrics were chosen because they reflected the differences among data feeds at multiple levels -  the terminological (use of names/labels), structural, and modelling levels.
-Each such metric listed above is the result of averaging the pairwise score between sources in each category, separately and defined as follows:
-
-\begin{itemize}
-\item \emph{Terminological consistency} - Percent of property names that match
-\item \emph{Structural consistency} - Percent of properties that have \emph{compatible structures}
-\item \emph{Semantic/Model consistency} - Percent of entities/concepts that have corresponding fields 
-\item \emph{Value consistency} - Percent of values that are compatible
-\end{itemize}
+Finally, we looked at fields that were equivalent, yet which could not be compared for reasons other than the terminological and structural reasons above.
+These \emph{modeling differences} were caused by a variety of reasons including, but not limited to, the following: measurement unit inconsistencies, how data was measured (e.g., dewpoint vs relative humidity), differences in scale or granularity, and mismatches in what was being modeled (``offer'' versus ``listing'').
 
@@
 \caption{DataPalette Workspace, a file manager-inspired Drag and Drop interface for structured data mixing.
-Collapsible displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a windoow (group) on the workspace.}
+Collapsable displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a windoow (group) on the workspace.}
 \label{fig:workspace}
@@
 
-% Data sources were represented as collapsible panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
+% Data sources were represented as collapsable panels in a sidebar on the left of the workspace, from which instances or entire data sources could be ``pulled off'' onto the workspace.
 Figure \ref{fig:workspace} displays the entire UI overview.

### 2012-09-18 chi2012.tex
ps2 results

--- 2012-09-18 chi2012.tex
@@
 For each obtained data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.
-For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record has few attributes about the actual product compared to its ItemAttributes sub-property, which was used instead.
+For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record had few attributes about the actual product compared to its ``ItemAttributes'' sub-property, which was used instead.
 Thus, all comparisons were made to maximize comparative similarity.
-Depth was computed as average from root
+Depth was computed as the average number of levels from record root to leaves across all properties, and was measured, with width, to convey the overall size/complexity of each record.
 
-In addition to the complexity metrics, we mapped out the amount of overlap among the sources in each category by first manually determiningr \emph{equivalent properties}.
-This was done manually by creating a pairwise alignment table between every two data sources in each category.
-Each row of the alignment table comprised a property in some schema, and all its closest matches in the other schemas.
-We used both property names and attribute descriptions (in documentation, if available) to make matching decisions; two properties  were considered equivalent if they intended to model the same piece of information;
+The third metric measured the amount of overlap among the sources in each category.
+This was done by determining \emph{equivalent properties} between schemas, which we did manually by creating an alignment table for all properties across all data sources in each category.
+Each row of the alignment table comprised a property of a record from one data source, and all its closest matching properties from the others.
+We used both property names and attribute descriptions (in documentation, if available) to make matching decisions; two properties were aligned if they seemed to model the same piece of information.
 
-, followed by pairwise heterogeneity metrics described below.
-An alternative approach to using instance documents was to directly compare schema definitions directly; however, many of the sites' schemas were not actually publicly available; second, that many of the schemas were not actually representative of typical data returned by the service.
-Thus we decided to use an instance document random sampling technique instead.
+Once the equivalent properties were established, we look at how many of these properties were terminologically inconsistent, that is, identified using different property names.
+Then, ignoring property name, we looked to see if its values were structurally comparable.
+If two fields were of the same data type and, if complex, had equivalent subfields, they were considered comparable; otherwise they were considered \emph{structurally inconsistent}.
+An example of such a structural inconsistency were``birthday'' fields in Facebook and LinkedIn; Facebook represented it as a single string formatted ``MM/DD/YYYY'', while LinkedIn represented it as a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.
 
-The per-source metrics were as follows:
-\begin{itemize}
-\item \emph{Average Width of record} - Average number of properties per record.
-\item \emph{Average Depth of record} - Average depth of data record (number of sub-properties in self-contained record).
-(In JSON/XML representations, this is the tree depth; for RDF  representations this is merely the number of embedded entity resource definitions included in the top-level query).
-\end{itemize}
-
-Heterogeneity was measured through \emph{pairwise consistency metrics} we derived from definitions in \cite{george2005understanding} and feature-metrics used in statistical instance alignment algorithms (such as summarised in \cite{bhattacharya2007collective}).
-Such metrics were chosen because they reflected the differences among data feeds at multiple levels -  the terminological (use of names/labels), structural, and modelling levels.
-Each such metric listed above is the result of averaging the pairwise score between sources in each category, separately and defined as follows:
-
-\begin{itemize}
-\item \emph{Terminological consistency} - Percent of property names that match
-\item \emph{Structural consistency} - Percent of properties that have \emph{compatible structures}
-\item \emph{Semantic/Model consistency} - Percent of entities/concepts that have corresponding fields 
-\item \emph{Value consistency} - Percent of values that are compatible
-\end{itemize}
+Finally, we looked at fields that were equivalent, yet which could not be compared for reasons other than the terminological and structural reasons above.
+These \emph{modelling differences} were caused by a variety of reasons including, but not limited to, the following: measurement unit inconsistencies, how data was measured (e.g., dewpoint vs relative humidity), differences in scale or granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').
 
 \subsubsection{Results}
-% TODO TODO TODO TODO 
-Our findings suggest that -- FILL ME IN HERE PLEASE,
+Table \ref{tbl:prestudy2} displays the raw results of the process just described.
+A number of features were quite striking about the data.
+First, while most of the data records were small and relatively simple, nearly one in each category produced substantially more complex records than the rest; in particular, amazon, soundcloud, Twitter and Weather Underground stood out in their respective categories.
+
+Looking at overlap, we notice that, except for the smallest records, there was little overlap among records; that is, most of the data records had a large number of unique properties unaccounted for in the others.
+While some of these 'unique records' were primarily for internal, service-specific use (such as service-speicifc IDs) others were useful properties one would expect to find in a record that were simply not present.
+Only one of the music records, Spotify, included information about the album(s) that a song was featured on.
+Such omissions show us that these data APIs reflect very specifically each service provider's capabilities and internal data representations, and that integrating data such records from multiple sources would allow properties to be aggregated to form more complete representations.
+
+The second thing we find is that equivalent properties were sometimes, though only rarely, given the same name.
+Thus, terminological heterogeneity was prevalent across data records we observed.
+Structural heterogeneity, however, was comparatively rare, with the most common case (comprising a majority of the cases) occurring
 
 %\section{Designing the DataPalette Interface}
-
 \section{Designing DataPalette : An Interface for Serendipitous Data Mixing}
@@
 
-Modeling heterogeneity (?)
+Modelling heterogeneity (?)
    Specificity  

### 2012-09-18 chi2012.tex
a  few more

@@
 A number of features were quite striking about the data.
-First, while most of the data records were small and relatively simple, nearly one in each category produced substantially more complex records than the rest; in particular, amazon, soundcloud, Twitter and Weather Underground stood out in their respective categories.
+First, while most of the data records were small and relatively simple, nearly one in each category produced substantially more complex records than the rest; in particular, Amazon, Soundcloud, Twitter and Weather Underground stood out in their respective categories.
 
@@
 
-The second thing we find is that equivalent properties were sometimes, though only rarely, given the same name.
+The second finding is that equivalent properties were sometimes, though only rarely, given the same name.
 Thus, terminological heterogeneity was prevalent across data records we observed.
-Structural heterogeneity, however, was comparatively rare, with the most common case (comprising a majority of the cases) occurring
+Structural heterogeneity, however, was comparatively rare, with the most common case being that a leaf property on one record was modeled as a subtree in a target schema.
 

### 2012-09-18 chi2012.tex
a  few more

@@
 
-Looking at overlap, we notice that, except for the smallest records, there was little overlap among records; that is, most of the data records had a large number of unique properties unaccounted for in the others.
+Looking at overlap, we notice that, there was, overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones that overlapped with any of the other records.
 While some of these 'unique records' were primarily for internal, service-specific use (such as service-speicifc IDs) others were useful properties one would expect to find in a record that were simply not present.
@@
 Thus, terminological heterogeneity was prevalent across data records we observed.
-Structural heterogeneity, however, was comparatively rare, with the most common case being that a leaf property on one record was modeled as a subtree in a target schema.
+Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modeled as a subtree in the other.
+We also encountered, particularly in the Weather category, examples of fields which had data embedded in textual description fields
 

### 2012-09-18 chi2012.tex
a  few more

@@
 
-Looking at overlap, we notice that, there was, overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones that overlapped with any of the other records.
+Looking at overlap, we notice overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones that overlapped with any of the other records.
 While some of these 'unique records' were primarily for internal, service-specific use (such as service-speicifc IDs) others were useful properties one would expect to find in a record that were simply not present.

### 2012-09-18 chi2012.tex
a  few more

@@
 
-Looking at overlap, we notice overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones that overlapped with any of the other records.
+Looking at overlap, we notice overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
 While some of these 'unique records' were primarily for internal, service-specific use (such as service-speicifc IDs) others were useful properties one would expect to find in a record that were simply not present.

### 2012-09-18 chi2012.tex
a  few more

@@
 Looking at overlap, we notice overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
-While some of these 'unique records' were primarily for internal, service-specific use (such as service-speicifc IDs) others were useful properties one would expect to find in a record that were simply not present.
-Only one of the music records, Spotify, included information about the album(s) that a song was featured on.
-Such omissions show us that these data APIs reflect very specifically each service provider's capabilities and internal data representations, and that integrating data such records from multiple sources would allow properties to be aggregated to form more complete representations.
+While some of these 'unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties one would expect to find in a record that were simply not present.
+Only one of the music data providers, Spotify, included information about the album(s) that featured a particular song.
+Such omissions show us that these data APIs reflect each service provider's own capabilities and internal data representations very specifically, and that integrating data such records from multiple sources could allow more complete and useful representations to be assembled.
 

### 2012-09-18 chi2012.tex
a  few more

@@
 Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modeled as a subtree in the other.
-We also encountered, particularly in the Weather category, examples of fields which had data embedded in textual description fields
+We also encountered, particularly in the Weather category, examples of fields which had data embedded in textual description fields, which hints their intended use in RSS and other feed readers than as stuctured data feeds.
+
+Modelling
 

### 2012-09-18 chi2012.tex
a  few more

@@
 
-Looking at overlap, we notice overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
+Looking at equivalent properties, we foun overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
 While some of these 'unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties one would expect to find in a record that were simply not present.
@@
 
-Modelling
+Modelling heterogeneity was also comparatively rare,
+
+Although not included in the table above, we also examined enumerated values for a number of the data sources above.  Field even as simple as ``Gender'' often disagreed in the
 

### 2012-09-18 chi2012.tex
a  few more

@@
 
-Although not included in the table above, we also examined enumerated values for a number of the data sources above.  Field even as simple as ``Gender'' often disagreed in the
+Although not included in the table above, we also examined the consistency in enumerated type ranges across data sources above.
+Ranges of enumerated types for properties as seemingly simple as ``Gender'' disagreed among social networks; while Facebook allowed only ``Male'' and ``Female'' for values, Google+ included a third option,``other''.
+``Relationship status'' was another for which a majority of the values had correspondences between platforms but there wer a number of unique values not present in the rest.
 

### 2012-09-19 chi2012.tex
a  few more

@@
 Looking at equivalent properties, we foun overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
-While some of these 'unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties one would expect to find in a record that were simply not present.
+While some of these 'unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply not present.
 Only one of the music data providers, Spotify, included information about the album(s) that featured a particular song.
@@
 Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modeled as a subtree in the other.
-We also encountered, particularly in the Weather category, examples of fields which had data embedded in textual description fields, which hints their intended use in RSS and other feed readers than as stuctured data feeds.
+We also encountered, particularly in the Weather category, examples of fields which had data nested in textual description fields, which hints their intended use in RSS and other feed readers than as stuctured data feeds.
 
-Modelling heterogeneity was also comparatively rare,
+Examples of modelling/semantic heterogeneity instances were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site; ``Promotions'' versus ``Offers'',
 

### 2012-09-19 chi2012.tex
a  few more

@@
 Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modeled as a subtree in the other.
-We also encountered, particularly in the Weather category, examples of fields which had data nested in textual description fields, which hints their intended use in RSS and other feed readers than as stuctured data feeds.
+We also encountered only 2 examples of fields which had data nested in textual description fields, both from records in the Weather category, which hints their intended use in RSS and other feed readers than as stuctured data feeds.
 
@@
 
-Although not included in the table above, we also examined the consistency in enumerated type ranges across data sources above.
-Ranges of enumerated types for properties as seemingly simple as ``Gender'' disagreed among social networks; while Facebook allowed only ``Male'' and ``Female'' for values, Google+ included a third option,``other''.
-``Relationship status'' was another for which a majority of the values had correspondences between platforms but there wer a number of unique values not present in the rest.
+Although not included in the table above, we also examined the ranges of enumerated type values across data sources.
+There was substantial terminological heterogeneity among such values, but also many cases where one source had additional values not supported in the range of the other.
+For example, the property ``Gender''  on Facebook allowed only ``Male'' and ``Female'' for values, while Google+ included a third option,``other''.
+``Relationship status'' was another which was similarly complex.
+Amazon's
 

### 2012-09-19 chi2012.tex
a  few more

@@
 \end{tabular}
-\caption{\emph{PS.2 Results} - Analysis of heterogeneity present across data feeds, by category of feed}\label{tbl:prestudy2}
+\caption{\emph{PS.2 Analysis of Data Feeds} - Structural statistics about data sampled from each API organized by category.
+The first column lists the category and type of data record that was examined.
+Cols 3 and 4 list the average number of properties per data record, and number of levels from root to leaf.
+The remaining fields show results from a pairwise analysis of data records from each of the data sources with those of the others in each category; values are prefixed with the first letter(s) of the target schema the row schema was compared against.
+For example ``Tw:5'' in the 5th column of Google+ row indicates that Facebook and Twitter records had 5 equivalent fields.
+}\label{tbl:prestudy2}
 \end{table*}

### 2012-09-19 chi2012.tex
intro tweaks and foo bar

@@
 
-In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web, through APIs, datasets, and data feeds.
-This information includes: access to data previously hidden behind services and applications, such as retailer product catalogues; information not previously directly released to the public, such as open government data or records of personal financial transactions; and new kinds of data generated by emerging kinds of data sources, such as wearable and smartphone-based biosensors.
-A primary goal of opening up this data is that end-user citizens can make more informed decisions pertaining to their health, wealth, and well-being \cite{}.
+In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web.
+This information has come by way of many different channels, from the various APIs now offered by most of the major online retailers, apps, sharing and service platforms, the massive data sets published by governmental and private institutions, to new streams being produced by various kinds of emerging data sources, such as wearable and smartphone-based biosensors.
+A primary goal of making much of this data available has been to give end-user citizens the ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
 This data is predominately used by specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.
@@
 Thus, all comparisons were made to maximize comparative similarity.
-Depth was computed as the average number of levels from record root to leaves across all properties, and was measured, with width, to convey the overall size/complexity of each record.
+% Depth was computed as the average number of levels from record root to leaves across all properties, and was measured, with width, to convey the overall size/complexity of each record.
 
@@
 
-Once the equivalent properties were established, we look at how many of these properties were terminologically inconsistent, that is, identified using different property names.
+Once the equivalent properties were established, we counted the properties that were terminologically inconsistent, that is, named differently between records.
 Then, ignoring property name, we looked to see if its values were structurally comparable.
@@
 
-Finally, we looked at fields that were equivalent, yet which could not be compared for reasons other than the terminological and structural reasons above.
-These \emph{modelling differences} were caused by a variety of reasons including, but not limited to, the following: measurement unit inconsistencies, how data was measured (e.g., dewpoint vs relative humidity), differences in scale or granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').
+Finally, we looked at fields that were equivalent, yet could not be compared for reasons other than the two reasons above.
+These \emph{modelling differences} were had a variety of causes, including, but not limited to, the following: measurement unit inconsistencies, measurement method (e.g., dewpoint vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').
 
@@
 Looking at equivalent properties, we foun overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
-While some of these 'unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply not present.
-Only one of the music data providers, Spotify, included information about the album(s) that featured a particular song.
-Such omissions show us that these data APIs reflect each service provider's own capabilities and internal data representations very specifically, and that integrating data such records from multiple sources could allow more complete and useful representations to be assembled.
+While some of these 'unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.
+Only one Spotify, among music providers, included information about the album(s) featuring a given song.
+Such omissions show us that these data APIs reflect each provider's own specific capabilities and needs, and suggests that integrating information from multiple sources may allow more complete and useful representations to be assembled.
 
-The second finding is that equivalent properties were sometimes, though only rarely, given the same name.
+The second finding is that equivalent properties were only rarely given the same name.
 Thus, terminological heterogeneity was prevalent across data records we observed.
 Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modeled as a subtree in the other.
-We also encountered only 2 examples of fields which had data nested in textual description fields, both from records in the Weather category, which hints their intended use in RSS and other feed readers than as stuctured data feeds.
+We also encountered only 2 examples where data was nested in textual description fields, both from records in the Weather category, which hints their intended use in RSS feed readers rather than in stuctured database applications.
+Examples of modelling/semantic heterogeneity instances were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and mismatches in units of measurement (commonly found in the Weather forecast measurements).
 
-Examples of modelling/semantic heterogeneity instances were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site; ``Promotions'' versus ``Offers'',
+Although not included in the table above, we also examined the source schemas of many of hte data sources for property enumerated type values.
+There was substantial terminological inconsistency among such values, but also cases where some sources had values not represented in the other source's sets.
+For example, the property ``Gender''  on Facebook allowed only supported ``Male'' and ``Female'' for values, while Google+ included a third option,``other''.
+``Relationship status'' was another which was similarly complex.
+None of the string representations of enumerated value sets matched across data sources perfectly.
 
-Although not included in the table above, we also examined the ranges of enumerated type values across data sources.
-There was substantial terminological heterogeneity among such values, but also many cases where one source had additional values not supported in the range of the other.
-For example, the property ``Gender''  on Facebook allowed only ``Male'' and ``Female'' for values, while Google+ included a third option,``other''.
-``Relationship status'' was another which was similarly complex.
-Amazon's
+In summary, while data records ranged in complexity, no data source subsumed the rest; all contained properties that other sources did not have.
+Moreover, the most common types of heterogeneity exhibited were terminological and value-related, although simpler kinds of structural hetereogeneity (pertaining to representing substructures of values) were also common.
 

### 2012-09-19 chi2012.tex
starting to cut text

@@
 
-\subsubsection{Results}
-The participant demographic consisted of 8 people (7 male and 1 female).
-They were split evenly into two age groups of 18-25 and 26-32.
-All but one participant reported they used social network sites very regularly (many times a day); they said they either logged in multiple times a day, left a website page open or used dedicated application for their social networking site updates.
+\subsubsection{Results - Heterogeneous Data Tasks}
+The participant demographic consisted of 8 people (7 male and 1 female). They were split evenly into two age groups of 18-25 and 26-32.
+
+% All but one participant reported they used social network sites very regularly (many times a day); they said they either logged in multiple times a day, left a website page open or used dedicated application for their social networking site updates.
 Those that used social networking websites primarily used Facebook.
@@
 The following two sections, describe the tasks people perform online with heterogeneous data and how they would use to organise a large social event.
-
-\subsubsection{Question 1: Heterogenous Data Tasks}
-All participants reported having used multiple websites to accomplish a task.
-They listed example tasks such as shopping (including food, consumer electronics, hotels and flights), choosing restaurants, searching for jobs, when choosing a university, choosing recipes, study and work.
+% \subsubsection{Question 1: Heterogenous Data Tasks}
+All participants regularly using multiple websites to accomplish tasks.
+Example of such tasks included shopping (including food, consumer electronics, hotels and flights), choosing restaurants, searching for jobs, selecting a university, seeking recipes, and finding answers to technical problems.
 The biggest focus for shopping focused tasks was balancing price, quality and speed of delivery.
-In general, people did not not trust a single websites for reviews because they said that they could be biased and did not always review the features they were interested in.
+In general, people did not not trust single websites for reviews because they said that they could be biased and did not always review the features they were interested in.
 One participant said that video reviews of items were really important because they showed the aesthetics and scale of the item.
-In general for initial research into organising a task it was typical to Google keywords to find related websites to their tasks.
-They identified that they used different website for different jobs, such as manufacturers website for technical details, review aggregators for a range of opinions, and Google maps for location based decisions.
+In general, for initial research into organising a task, it was typical to use Google to search for several related websites.
+They identified that they used different websites for different jobs, such as manufacturers website for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
 Social networks were also identified to have different functions, such as different groups of people belong to each and share different types of opinions.
@@
 \subsubsection{Enumerated-type value consolidation}
-Related to the concept of co-reference, PS.2 revealed that enumerated values of properties were often inconsistent, e.g.
+PS.2 revealed that enumerated values of properties were often inconsistent, e.g.
 ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property.
@@
 We selected the top restaurants for on Tripadvisor and Yelp, because they were the top two websites people quoted for comparing reviews when travelling.
-We selected the two university courses at random from a list taken form the complete university guide\footnote{}, and selected the top six university for that subject because they provide a realistic shortlist and are comparable in terms of rankings and properties.
+We selected the two university courses at random from a list taken form the complete university guide\footnote{The Complete University Guide - \url{www.thecompleteuniversityguide.co.uk}}, and selected the top six university for that subject because they provide a realistic shortlist and are comparable in terms of rankings and properties.
 We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choice.
@@
 In addition to the aforementioned feature requests, we observed a number of ``pain points'' in the design of the interface that we wish to address in the next iteration.
-One such pain point was window layout and screen real estate management; participants were prone to opening up a large number of windows, including maps and charts, and found that the screen filled up very quickly.
-When the screen got crowded, two problems were observed: people spent a lot of time arranging and moving windows about, and second, people tended to get lost in windows -- ``what was this again?''
+One such pain point was window layout and screen real estate management; participants were prone to opening up a large number of windows, including maps and charts.
+When their screen started to fill up, we found that participants spent a lot of time arranging windows, and second, that people tended to forget what each window represented -- ``what was this again?''.
+We plan to address these issue in the next version of DataPalette.
 
-\subsection{More Integration Support}
+\subsection{Integration Support}
+
+A key design decision we made at the outset was to target the most common, simpler forms of heterogeneity we observed in PS.2.
+Given the use of theinterface mechanisms we observed, we wish to next broaden the scope of capabilities of DP so that it can become more useful in real settings.
+Among the integration capabilities we are considering are, first, field combination and splitting (by delimiters), in order to address the kinds of structural heterogeneity observed in several cases of PS.2, along with a framework for converting and reconciling units of measure, which would address several of the modelling issues.
+Extending the Do What I Mean capabilities of the histogram and bar-charting tools to be unit-aware would be a tremendous asset, we believe, for users who want to quickly compare things without paying close attention to the units used.
+A core problem to realising this, however, is that few data sources (besides the Weather category) explicitly annotated values with units used.
+
+\subsection{Study limitations}
+
+% \subsection{Integration Limitations}
 
@@
 The result of this technique is that the users instantly filled their screen, and ran out of space for plots and maps.
-
-\subsection{Supporting more sophisticated heterogeneity}
-
-Supporting better structural heterogeneity:
-   Combining fields
-   Splitting fields
-   Rearranging fields (British dates American dates)
-
-Modelling heterogeneity (?)
-   Specificity  
-     Education+Jobs vs ``Education'' + ``Jobs''
-
-   Unit conversion (kg -> stones)
-   Granularity
-     Crimes per area
-
-Working with data values
 

### 2012-09-19 chi2012.tex
trimming and cutting

@@
 
-\numberofauthors{3}
-\author{
-  \alignauthor 1st Author Name\\
-    \affaddr{Affiliation}\\
-    \affaddr{Address}\\
-    \email{e-mail address}\\
-    \affaddr{Optional phone number}
-  \alignauthor 2nd Author Name\\
-    \affaddr{Affiliation}\\
-    \affaddr{Address}\\
-    \email{e-mail address}\\
-    \affaddr{Optional phone number}    
-  \alignauthor 3rd Author Name\\
-    \affaddr{Affiliation}\\
-    \affaddr{Address}\\
-    \email{e-mail address}\\
-    \affaddr{Optional phone number}
-}
+%% \numberofauthors{0}
+%% \author{
+%%   \alignauthor 1st Author Name\\
+%%     \affaddr{Affiliation}\\
+%%     \affaddr{Address}\\
+%%     \email{e-mail address}\\
+%%     \affaddr{Optional phone number}
+%%   \alignauthor 2nd Author Name\\
+%%     \affaddr{Affiliation}\\
+%%     \affaddr{Address}\\
+%%     \email{e-mail address}\\
+%%     \affaddr{Optional phone number}    
+%%   \alignauthor 3rd Author Name\\
+%%     \affaddr{Affiliation}\\
+%%     \affaddr{Address}\\
+%%     \email{e-mail address}\\
+%%     \affaddr{Optional phone number}
+%% }
 
@@
 \begin{quote} 
-  The problem stems from the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs.
+  The problem stems from [the fact that] we are trying to integrate data systems that were developed for slightly (or vastly) different business needs.
 Hence, even if they model overlapping domains, they will model them in different ways.
@@
 \subsubsection{Results - Heterogeneous Data Tasks}
-The participant demographic consisted of 8 people (7 male and 1 female). They were split evenly into two age groups of 18-25 and 26-32.
+The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18-25 and 26-32.   
+% mention # that were computer scientists here?
 
+% not so relevant:
 % All but one participant reported they used social network sites very regularly (many times a day); they said they either logged in multiple times a day, left a website page open or used dedicated application for their social networking site updates.
@@
 \subsubsection{Question 2: Process Used to Organise a Large Social Event}
-All of the participants would organise a large social event by talking to their friends, about their ideas, preferences and recommendations.
-The tools they would use to acquire this information included face-to-face, phone calls, email, Skype, Facebook events.
-The method chosen depended on the time scale required to organise an event, in general if there was a short time scale people would speak in person or on the phone.
+Participants said they would organise a large social event by first conferring with friends about their ideas, preferences and recommendations.
+They selected venues, locations, restaurants and activities through a number of sources: their own prior experience, friends' recommendations, searching Google maps, dedicated reviews sites (such as Yelp for restaurants), and through web search.
+People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.
+They considered how to get there and looked up train or bus times on the Web, or organised carpools.
+When organising an event, the weather was not that important because they did not trust the accuracy of predictions.
+Also, friends' preferences in price and food choices were not seen a priority; they would try to be inclusive but only if it meant small changes to the plan.
+Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
+
+%% Most people felt that being assertive and posting their decisions about the date and time on Facebook meant that organising events were more successful than trying to gather a consensus.
+They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.
+They selected venues, locations, restaurants and activities based on their own knowledge, friend's recommendations, Google maps, reviews obtained through web search and sites such as TripAdvisor.
+People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.
+They considered how to get there and looked up train or bus times on the Web, or organised carpools.
+When organising an event the weather was not that important because they did not trust the accuracy of predictions.
+Also, their friend's preferences in price and food choices were not a priority; they would try to be inclusive but only if it meant small changes to the plan.
+Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
+% previous
+% All of the participants would organise a large social event by first conferring with friends about their ideas, preferences and recommendations.
+The medium they would use included face-to-face, phone calls, email, Skype, Facebook events, depending on the time scale required to organise an event in general if there was a short time scale people would speak in person or on the phone.
 Otherwise, people would set up a Facebook event page to discuss ideas with their friends.
@@
 
-Most people felt that being assertive and posting their decisions about the date and time on Facebook meant that organising events were more successful than trying to gather a consensus.
+%% Most people felt that being assertive and posting their decisions about the date and time on Facebook meant that organising events were more successful than trying to gather a consensus.
 They expected that their friends would voice any objections if they could not attend or like the restaurant or activity selected.
@@
 \subsubsection{Summary of Findings}
+% TODO: how does this summary support the thesis of the paper?
 People said that they would like a website that could support the evolution of an event, that allowed them to post drafts of an event.
@@
 \subsection{PS.2 - Technical Challenges of Data Integration}
-Our investigation of integration challenges of available structured data feeds started with identifying a set of candidate sources to examine.
-For this, we consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} of 50 most popular data feeds measured by the number of developer API keys issued for each.
-To increase the variety of sources, we first categorised sources into 5 topic categories: social network services, retailers, event sources (e.g., online calendars and event sites) music, and weather, and selected 2-5 sources from each 2-5 feeds each for a total of 20 total feeds for analysis.
+Our analysis of structured data feeds started with identifying a set of candidate sources to examine.
+We consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} list of most popular data feeds for this purpose.
+To increase the variety of sources, we selected from 5 categories: social network services, retailers, online event calendars, music sites, and weather, selecting 2-5 sources from each, for a total of 20.
 
 For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for music sites this was song info, etc.  
-Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.
-For each obtained data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.
-For determining the width, we first examined the record to find the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record had few attributes about the actual product compared to its ``ItemAttributes'' sub-property, which was used instead.
-Thus, all comparisons were made to maximize comparative similarity.
+%Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.  
+For each such data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.
+Determining the width required first finding the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record had few attributes about the actual product compared to its ``ItemAttributes'' sub-property, which was used instead.
+% This step was done manually, with an effort to select most similar components for comparison.  
 % Depth was computed as the average number of levels from record root to leaves across all properties, and was measured, with width, to convey the overall size/complexity of each record.
@@
 
-Once the equivalent properties were established, we counted the properties that were terminologically inconsistent, that is, named differently between records.
+Once the equivalent properties were established, equivalent properties were examined to see if they were named identically.
 Then, ignoring property name, we looked to see if its values were structurally comparable.
@@
 
-Finally, we looked at fields that were equivalent, yet could not be compared for reasons other than the two reasons above.
-These \emph{modelling differences} were had a variety of causes, including, but not limited to, the following: measurement unit inconsistencies, measurement method (e.g., dewpoint vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').
+Finally, we examined fields that were equivalent but not be compared for reasons other than the structural or terminological reasons just described.
+These \emph{modelling differences} arose from a variety of causes, including, but not limited to, the following: measurement unit inconsistencies, measurement method (e.g., dewpoint vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').
 

### 2012-09-19 chi2012.tex
massaging that sentence

@@
 In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web.
-This information has come by way of many different channels, from the various APIs now offered by most of the major online retailers, apps, sharing and service platforms, the massive data sets published by governmental and private institutions, to new streams being produced by various kinds of emerging data sources, such as wearable and smartphone-based biosensors.
+This information has emerged from many channels, from the APIs now offered by most of the major retailers, apps, and social media sites, to the massive publishing of datasets by governmental and other institutions, to even the emergence of new kinds of data sources, such as wearable and smartphone-based biosensors.
 A primary goal of making much of this data available has been to give end-user citizens the ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.

### 2012-09-19 chi2012.tex
foo
Shortening and polishing PS2 section.

@@
 \begin{abstract}
-Exploratory sequential mixed-method study
+%% While people often draw upon diverse sources of information in the course of common tasks, tool support for integrating information from diverse information sources remains scarce.
+This paper presents an exploratory, sequential mixed-method study of user-driven data integration methods for PIM tools...
+
+exploratory, sequential mixed-method study of 
 \end{abstract}
@@
 
-For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for music sites this was song info, etc.  
+For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for retail sites this was product info, etc.
 %Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.  
-For each such data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.
+For each such data record, two simple complexity metrics were computed: the \emph{average width} and \emph{average depth}.
 Determining the width required first finding the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record had few attributes about the actual product compared to its ``ItemAttributes'' sub-property, which was used instead.
+The average depth was simply the mean depth to each leaf node from the chosen comparison level.
 % This step was done manually, with an effort to select most similar components for comparison.  
@@
 
-Once the equivalent properties were established, equivalent properties were examined to see if they were named identically.
-Then, ignoring property name, we looked to see if its values were structurally comparable.
-If two fields were of the same data type and, if complex, had equivalent subfields, they were considered comparable; otherwise they were considered \emph{structurally inconsistent}.
-An example of such a structural inconsistency were``birthday'' fields in Facebook and LinkedIn; Facebook represented it as a single string formatted ``MM/DD/YYYY'', while LinkedIn represented it as a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.
+Once established, equivalent properties were examined firstly for naming inconsistancies.
+Then, ignoring property name, we looked at whether the values were structurally comparable; if two fields were of the same data type and, if complex, had structurally equivalent subfields, they were considered comparable, otherwise they were considered \emph{structurally inconsistent}.
+The ``birthday'' fields in Facebook and LinkedIn demonstrate structural inconcistancy; Facebook represents birthdays as a single string formatted ``MM/DD/YYYY'', while the LinkedIn representation is a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.
 
-Finally, we examined fields that were equivalent but not be compared for reasons other than the structural or terminological reasons just described.
-These \emph{modelling differences} arose from a variety of causes, including, but not limited to, the following: measurement unit inconsistencies, measurement method (e.g., dewpoint vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').
+Finally, we examined equivalent fields which were incomparabe for reasons other than structural or terminological differences.
+These \emph{modelling differences} arose from a variety of causes, including, but not limited to: measurement unit inconsistencies, measurement method (e.g., dewpoint vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').
 

### 2012-09-19 chi2012.tex
Polished and shortened first two paragraphs in prestudies section.

@@
 We performed an exploratory sequential mixed-method study to help us design our contribution.
-For our first pre-study, we first sought an updated understanding of how and whether people still used multiple information sources regularly when they sought information on the Web.
-There were two reasons for this inquiry; first, whether people were increasingly either relying only on the handful of the massive ``supersites'' (Facebook, Google, Wikipedia) for most of their activities, or searching across distributed sources of information.
-If this were the case, much of the benefit could be gained by merely integrating PIM tools with the handful of sites, rather than the substantially more challenging problem of integrating data from arbitrary information sources on the Web.
-Second, whether people relied on multiple sites and why.
-If they did what kinds of tasks did people required diverse information collection and were these kinds of tasks rare?
+For our first pre-study, we sought an updated understanding on the use of multiple information sources in information gathering on the Web.
+There were two driving questions behind this inquiry; first, were people increasingly relying only on a handful of ``supersites'' (Facebook, Wikipedia), or searching across distributed sources of information?
+If the former were true, merely integrating PIM tools with the small numbr of supersites would provide greater benefit than tackling the substantially more challenging problem of integrating data from arbitrary sources on the Web.
+Second, why did people choose a single or multiple sources for information gathering?
+If a number of websites were used, why was this preferred over a single source?
+
+% If they did what kinds of tasks did people required diverse information collection and were these kinds of tasks rare?
 If so, this would suggest that perhaps specialised interfaces could be concocted for this purpose rather than it being a PIM need.
@@
 The second pre-study focused on the characteristics of the data available from the kinds of sites used by the first pre-study's participants for the tasks described.
-The purpose of this follow-on  was to identify not specific characteristics of particular sites in question, but to identify, in general, the degree of complexity of the data feeds currently available across a variety of domains, and typical integration problems that might arise when mixing this data.
-The next two sections, we present our methodologies and results for our pre-studies.
+The purpose of this follow-on was not to identify specific characteristics of particular sites in question, but to identify, in general, the degree of complexity of the data feeds currently available across a variety of domains, and typical integration problems that might arise when mixing this data.
+In the next two sections, we present our methodologies and results for our pre-studies.
 

### 2012-09-19 chi2012.tex
Rewrote first paragraph in PS1.

@@
 \subsection{PS.1 - Understanding Data Diversity in Everyday Tasks}
-We designed a semi-structured interview about tasks people performed online that required the use of multiple websites, and the processes people used to perform such tasks.
+% Rewritten with paragraph below; feel free to revert, or delete this if you're happy, or whatever.
+% We designed a semi-structured interview about tasks people performed online that required the use of multiple websites, and the processes people used to perform such tasks.
 We analysed the data collected from this study using a thematic approach.
-In our pre-study, we investigated the types of tasks people performed online that required heterogeneous data sources, and which tools they would use to organise a social event involving around 15 people.
+In our first pre-study, we investigated the types of tasks people performed online that required heterogeneous data sources, and which tools they would use to organise a social event involving around 15 people.
 We planned to interview 8 participants, using a general interview guide approach based on a set of questions.
@@
 Each interview was recorded and notes taken, the interviewer was trained in the purpose of the study and to minimise bias.
+
+For the first prestudy, we designed and ran a semi-structured interview to discover the processes and tools people would use when combining data from multiple websites.
+We interviewed 8 participants and asked them a range of questions regarding their use of multiple websites in gathering information.
+We then inquired as to how each participant would go about planning various aspects of a large party, such as date, location and entertainment.
 

### 2012-09-19 chi2012.tex
edits

--- 2012-09-19 chi2012.tex
@@
 This paper presents an exploratory, sequential mixed-method study of user-driven data integration methods for PIM tools...
-
-exploratory, sequential mixed-method study of 
+%% exploratory, sequential mixed-method study of 
 \end{abstract}
@@
 
-For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for retail sites this was product info, etc.
+For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for music sites this was song info, etc.  
 %Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.  
-For each such data record, two simple complexity metrics were computed: the \emph{average width} and \emph{average depth}.
+For each such data record, two simple complexity metrics were computed: an \emph{average width} and \emph{average depth} of record.
 Determining the width required first finding the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record had few attributes about the actual product compared to its ``ItemAttributes'' sub-property, which was used instead.
-The average depth was simply the mean depth to each leaf node from the chosen comparison level.
 % This step was done manually, with an effort to select most similar components for comparison.  
@@
 
-Once established, equivalent properties were examined firstly for naming inconsistancies.
-Then, ignoring property name, we looked at whether the values were structurally comparable; if two fields were of the same data type and, if complex, had structurally equivalent subfields, they were considered comparable, otherwise they were considered \emph{structurally inconsistent}.
-The ``birthday'' fields in Facebook and LinkedIn demonstrate structural inconcistancy; Facebook represents birthdays as a single string formatted ``MM/DD/YYYY'', while the LinkedIn representation is a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.
+Once the equivalent properties were established, equivalent properties were examined to see if they were named identically.
+Then, ignoring property name, we looked to see if its values were structurally comparable.
+If two fields were of the same data type and, if complex, had equivalent subfields, they were considered comparable; otherwise they were considered \emph{structurally inconsistent}.
+An example of such a structural inconsistency were``birthday'' fields in Facebook and LinkedIn; Facebook represented it as a single string formatted ``MM/DD/YYYY'', while LinkedIn represented it as a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.
 
-Finally, we examined equivalent fields which were incomparabe for reasons other than structural or terminological differences.
-These \emph{modelling differences} arose from a variety of causes, including, but not limited to: measurement unit inconsistencies, measurement method (e.g., dewpoint vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').
+Finally, we examined fields that were equivalent but not be compared for reasons other than the structural or terminological reasons just described.
+These \emph{modelling differences} arose from a variety of causes, including, but not limited to, the following: measurement unit inconsistencies, measurement method (e.g., dewpoint vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').
 
@@
 
-\subsection{Revisiting the hypotheses: Does DataPalette facilitate serendipitous data integration?}
+\subsection{Does DataPalette facilitate serendipitous data integration?}
 
-We set out to answer three different questions with the DataPalette study.
-The first concerned whether people understood and could use the system (h1), whether the system enabled people to mix and integrate heterogeneous data (h2), and whether the interface facilitated task completion (h3).
+We set out to test three hypotheses with the DataPalette study.
+The first was that end-user citizens would understand the interface mechanisms of DP (h1).
+The second, h2, was that DP enables people to mix and integrate heterogeneous data in the process of performing their tasks.
+The third, h3, was that using DP facilitatedp task completion (h3).
 Using real data obtained from typical data sources identified during our prestudy (PS.1), we tested the DP prototype and its integration interaction mechanisms with ``real'' users, who were students, staff and alumni of our University who had varying backgrounds and levels of expertise with computers.

### 2012-09-19 chi2012.tex
yes

@@
 
-For the first prestudy, we designed and ran a semi-structured interview to discover the processes and tools people would use when combining data from multiple websites.
-We interviewed 8 participants and asked them a range of questions regarding their use of multiple websites in gathering information.
+For the first prestudy, we conducted a semi-structured interview to understand the reasons and ways that people drew information from diverse information sources.
+We interviewed 8 participants, asking them to identify tasks that performed recently that required use of multiple information sources, the kinds of sources used, and the tools used to manage the resulting information.
 We then inquired as to how each participant would go about planning various aspects of a large party, such as date, location and entertainment.

### 2012-09-19 chi2012.tex
abstract yo

@@
 \begin{abstract}
-%% While people often draw upon diverse sources of information in the course of common tasks, tool support for integrating information from diverse information sources remains scarce.
+While people naturally draw upon diverse sources of information while problem solving and making decisions, tool support for serendipitous data mixing in Personal Information Management tools remains mostly absent.
 This paper presents an exploratory, sequential mixed-method study of user-driven data integration methods for PIM tools...
+%% We find that the ``light-touch'' data integration mechanisms introduced allow people to effectively work with and combine information from multiple sources, and more thoroughly justify their choices
 %% exploratory, sequential mixed-method study of 

### 2012-09-19 chi2012.tex
Refactored: \subsubsection{Question 1: Heterogenous Data Tasks}

--- 2012-09-19 chi2012.tex
@@
 This paper presents an exploratory, sequential mixed-method study of user-driven data integration methods for PIM tools...
-%% exploratory, sequential mixed-method study of 
+
+exploratory, sequential mixed-method study of 
 \end{abstract}
@@
 \subsubsection{Results - Heterogeneous Data Tasks}
+
 The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18-25 and 26-32.   
@@
 
-% not so relevant:
-% All but one participant reported they used social network sites very regularly (many times a day); they said they either logged in multiple times a day, left a website page open or used dedicated application for their social networking site updates.
-Those that used social networking websites primarily used Facebook.
-All of the participants used Twitter to listen to broadcasted tweets, and 6 out of 8 people regularly tweeted.
-The following two sections, describe the tasks people perform online with heterogeneous data and how they would use to organise a large social event.
-% \subsubsection{Question 1: Heterogenous Data Tasks}
-All participants regularly using multiple websites to accomplish tasks.
-Example of such tasks included shopping (including food, consumer electronics, hotels and flights), choosing restaurants, searching for jobs, selecting a university, seeking recipes, and finding answers to technical problems.
-The biggest focus for shopping focused tasks was balancing price, quality and speed of delivery.
-In general, people did not not trust single websites for reviews because they said that they could be biased and did not always review the features they were interested in.
-One participant said that video reviews of items were really important because they showed the aesthetics and scale of the item.
-In general, for initial research into organising a task, it was typical to use Google to search for several related websites.
-They identified that they used different websites for different jobs, such as manufacturers website for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
-Social networks were also identified to have different functions, such as different groups of people belong to each and share different types of opinions.
+\subsubsection{Question 1: Heterogenous Data Tasks}
+
+All participants reported regular use of multiple websites to accomplish tasks, examples of which including: shopping (including food, consumer electronics, hotels and flights), choosing restaurants, job-seeking, selecting a university, seeking recipes, and finding answers to technical problems.
+When beginning a task, it was common for participants to use Google to discover several related websites, which would then be used in different ways depending on the task at hand.
+For shopping-oriented tasks, balancing price, quality and speed of delivery was the main aim.
+For product reviews, single sources were not trusted due to possible bias and incomplete feature coverage.
+% Jim - Not sure if this is relevant. If it is, put it back in.
+% one participant said that video reviews of items were particularly helpful, as the aesthetics and scale of the item could be seen.
+% Social networks were identified to have different functions, such as different groups of people belong to each and share different types of opinions.
 One participant said that they shared their outlook and gmail calendars, but only for work events and did not record social activities on them.
+To provide a full coverage of information, websites of different types were used, such as manufacturers website for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
 
@@
 
-\subsection{Does DataPalette facilitate serendipitous data integration?}
+\subsection{Revisiting the hypotheses: Does DataPalette facilitate serendipitous data integration?}
 
-We set out to test three hypotheses with the DataPalette study.
-The first was that end-user citizens would understand the interface mechanisms of DP (h1).
-The second, h2, was that DP enables people to mix and integrate heterogeneous data in the process of performing their tasks.
-The third, h3, was that using DP facilitatedp task completion (h3).
+We set out to answer three different questions with the DataPalette study.
+The first concerned whether people understood and could use the system (h1), whether the system enabled people to mix and integrate heterogeneous data (h2), and whether the interface facilitated task completion (h3).
 Using real data obtained from typical data sources identified during our prestudy (PS.1), we tested the DP prototype and its integration interaction mechanisms with ``real'' users, who were students, staff and alumni of our University who had varying backgrounds and levels of expertise with computers.

### 2012-09-19 chi2012.tex
updated ps2

--- 2012-09-19 chi2012.tex
@@
 \begin{abstract}
-While people naturally draw upon diverse sources of information while problem solving and making decisions, tool support for serendipitous data mixing in Personal Information Management tools remains mostly absent.
+%% While people often draw upon diverse sources of information in the course of common tasks, tool support for integrating information from diverse information sources remains scarce.
 This paper presents an exploratory, sequential mixed-method study of user-driven data integration methods for PIM tools...
-%% We find that the ``light-touch'' data integration mechanisms introduced allow people to effectively work with and combine information from multiple sources, and more thoroughly justify their choices
 %% exploratory, sequential mixed-method study of 
@@
 
-For the first prestudy, we conducted a semi-structured interview to understand the reasons and ways that people drew information from diverse information sources.
+For the first pre-study, we conducted a semi-structured interview to understand the reasons and ways that people drew information from diverse information sources.
 We interviewed 8 participants, asking them to identify tasks that performed recently that required use of multiple information sources, the kinds of sources used, and the tools used to manage the resulting information.
@@
 \subsubsection{Results - Heterogeneous Data Tasks}
-The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18-25 and 26-32.   
+The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
 % mention # that were computer scientists here?
@@
 % \subsubsection{Question 1: Heterogenous Data Tasks}
-All participants regularly using multiple websites to accomplish tasks.
+All participants regularly use multiple websites to accomplish tasks.
 Example of such tasks included shopping (including food, consumer electronics, hotels and flights), choosing restaurants, searching for jobs, selecting a university, seeking recipes, and finding answers to technical problems.
-The biggest focus for shopping focused tasks was balancing price, quality and speed of delivery.
+The biggest focus for shopping-focused tasks was balancing price, quality and speed of delivery.
 In general, people did not not trust single websites for reviews because they said that they could be biased and did not always review the features they were interested in.
@@
 In general, for initial research into organising a task, it was typical to use Google to search for several related websites.
-They identified that they used different websites for different jobs, such as manufacturers website for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
-Social networks were also identified to have different functions, such as different groups of people belong to each and share different types of opinions.
+They identified that they used different websites for different jobs, such as manufacturers websites for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
+%  Social networks were also identified to have different functions, such as different groups of people belong to each and share different types of opinions.
 One participant said that they shared their outlook and gmail calendars, but only for work events and did not record social activities on them.
 
-\subsubsection{Question 2: Process Used to Organise a Large Social Event}
+%\subsubsection{Question 2: Process Used to Organise a Large Social Event}
 Participants said they would organise a large social event by first conferring with friends about their ideas, preferences and recommendations.
@@
 When organising an event, the weather was not that important because they did not trust the accuracy of predictions.
-Also, friends' preferences in price and food choices were not seen a priority; they would try to be inclusive but only if it meant small changes to the plan.
+Also, friends' preferences in price and food choices were not seen as a priority; they would try to be inclusive but only if it meant small changes to the plan.
 Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.

### 2012-09-19 chi2012.tex

--- 2012-09-19 chi2012.tex
@@
 \subsubsection{Results - Heterogeneous Data Tasks}
-
-The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18-25 and 26-32.   
+The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
 % mention # that were computer scientists here?
 
-\subsubsection{Question 1: Heterogenous Data Tasks}
-
+%\subsubsection{Question 1: Heterogenous Data Tasks}
 All participants reported regular use of multiple websites to accomplish tasks, examples of which including: shopping (including food, consumer electronics, hotels and flights), choosing restaurants, job-seeking, selecting a university, seeking recipes, and finding answers to technical problems.
@@
 
-\subsubsection{Question 2: Process Used to Organise a Large Social Event}
+%\subsubsection{Question 2: Process Used to Organise a Large Social Event}
 Participants said they would organise a large social event by first conferring with friends about their ideas, preferences and recommendations.

### 2012-09-19 chi2012.tex
Couple of little grammatical/spelling fixes before working on a paragraph.

--- 2012-09-19 chi2012.tex
@@
 There were two driving questions behind this inquiry; first, were people increasingly relying only on a handful of ``supersites'' (Facebook, Wikipedia), or searching across distributed sources of information?
-If the former were true, merely integrating PIM tools with the small numbr of supersites would provide greater benefit than tackling the substantially more challenging problem of integrating data from arbitrary sources on the Web.
+If the former were true, merely integrating PIM tools with the small number of supersites would provide greater benefit than tackling the substantially more challenging problem of integrating data from arbitrary sources on the Web.
 Second, why did people choose a single or multiple sources for information gathering?
@@
 For the first prestudy, we conducted a semi-structured interview to understand the reasons and ways that people drew information from diverse information sources.
-We interviewed 8 participants, asking them to identify tasks that performed recently that required use of multiple information sources, the kinds of sources used, and the tools used to manage the resulting information.
+We interviewed 8 participants, asking them to identify tasks they performed recently that required use of multiple information sources, the kinds of sources used, and the tools used to manage the resulting information.
 We then inquired as to how each participant would go about planning various aspects of a large party, such as date, location and entertainment.
@@
 
-The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18-25 and 26-32.   
+The 8 participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18-25 and 26-32.   
 % mention # that were computer scientists here?

### 2012-09-19 chi2012.tex
abstract

--- 2012-09-19 chi2012.tex
@@
 \begin{abstract}
-%% While people often draw upon diverse sources of information in the course of common tasks, tool support for integrating information from diverse information sources remains scarce.
-This paper presents an exploratory, sequential mixed-method study of user-driven data integration methods for PIM tools...
+Although people naturally rely upon multiple, diverse information sources in the course of routine problem-solving and decision-making tasks, support for data integration in Personal Information Management (PIM) tools remains poor, if largely nonexistent.
+This paper presents an exploratory, sequential, mixed-method study of ways to extend PIM tools to support user-driven data integration while performing their primary tasks.
+We present results of two pre-study in which requirements, we designed and evaluated DataPalette.
+% We find that the ``light-touch'' data integration mechanisms introduced allow people to effectively work with and combine information from multiple sources, and more thoroughly justify their choices
 %% exploratory, sequential mixed-method study of 
@@
 Hence, even if they model overlapping domains, they will model them in different ways.
-Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modelling goal~\cite{alonhalevy}.\looseness=-1
+Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modelling goal~\cite{halevy2006data}.\looseness=-1
 \end{quote}
@@
 
-For the first pre-study, we conducted a semi-structured interview to understand the reasons and ways that people drew information from diverse information sources.
+For the first prestudy, we conducted a semi-structured interview to understand the reasons and ways that people drew information from diverse information sources.
 We interviewed 8 participants, asking them to identify tasks that performed recently that required use of multiple information sources, the kinds of sources used, and the tools used to manage the resulting information.
@@
 \subsubsection{Results - Heterogeneous Data Tasks}
-The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
+The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18-25 and 26-32.   
 % mention # that were computer scientists here?
@@
 % \subsubsection{Question 1: Heterogenous Data Tasks}
-All participants regularly use multiple websites to accomplish tasks.
+All participants regularly using multiple websites to accomplish tasks.
 Example of such tasks included shopping (including food, consumer electronics, hotels and flights), choosing restaurants, searching for jobs, selecting a university, seeking recipes, and finding answers to technical problems.
-The biggest focus for shopping-focused tasks was balancing price, quality and speed of delivery.
+The biggest focus for shopping focused tasks was balancing price, quality and speed of delivery.
 In general, people did not not trust single websites for reviews because they said that they could be biased and did not always review the features they were interested in.
@@
 In general, for initial research into organising a task, it was typical to use Google to search for several related websites.
-They identified that they used different websites for different jobs, such as manufacturers websites for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
-%  Social networks were also identified to have different functions, such as different groups of people belong to each and share different types of opinions.
+They identified that they used different websites for different jobs, such as manufacturers website for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
+Social networks were also identified to have different functions, such as different groups of people belong to each and share different types of opinions.
 One participant said that they shared their outlook and gmail calendars, but only for work events and did not record social activities on them.
 
-%\subsubsection{Question 2: Process Used to Organise a Large Social Event}
+\subsubsection{Question 2: Process Used to Organise a Large Social Event}
 Participants said they would organise a large social event by first conferring with friends about their ideas, preferences and recommendations.
@@
 When organising an event, the weather was not that important because they did not trust the accuracy of predictions.
-Also, friends' preferences in price and food choices were not seen as a priority; they would try to be inclusive but only if it meant small changes to the plan.
+Also, friends' preferences in price and food choices were not seen a priority; they would try to be inclusive but only if it meant small changes to the plan.
 Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
@@
 
-\subsection{Does DataPalette facilitate serendipitous data integration?}
+\subsection{Does DataPalette enable data integration?}
 

### 2012-09-19 chi2012.tex
refs in intro

@@
 This data is predominately used by specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.
-For example, the most commonly used structured personal information management (PIM) tools either manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{}.
+For example, the most commonly used structured personal information management (PIM) tools either manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{infoscraps}.
 

### 2012-09-19 chi2012.tex
Quotes, spelling errors.

--- 2012-09-19 chi2012.tex
@@
 \begin{abstract}
-Although people naturally rely upon multiple, diverse information sources in the course of routine problem-solving and decision-making tasks, support for data integration in Personal Information Management (PIM) tools remains poor, if largely nonexistent.
-This paper presents an exploratory, sequential, mixed-method study of ways to extend PIM tools to support user-driven data integration while performing their primary tasks.
-We present results of two pre-study in which requirements, we designed and evaluated DataPalette.
-% We find that the ``light-touch'' data integration mechanisms introduced allow people to effectively work with and combine information from multiple sources, and more thoroughly justify their choices
+%% While people often draw upon diverse sources of information in the course of common tasks, tool support for integrating information from diverse information sources remains scarce.
+This paper presents an exploratory, sequential mixed-method study of user-driven data integration methods for PIM tools...
 %% exploratory, sequential mixed-method study of 
@@
 Hence, even if they model overlapping domains, they will model them in different ways.
-Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modelling goal~\cite{halevy2006data}.\looseness=-1
+Differing structures are a byproduct of human nature --- people think differently from one another even when faced with the same modelling goal~\cite{alonhalevy}.\looseness=-1
 \end{quote}
@@
 \end{tabular}
-\caption{\emph{PS.2 Analysis of Data Feeds} - Structural statistics about data sampled from each API organized by category.
+\caption{\emph{PS.2 Analysis of Data Feeds} - Structural statistics about data sampled from each API organised by category.
 The first column lists the category and type of data record that was examined.
@@
 
-Once established, equivalent properties were examined firstly for naming inconsistancies.
+Once established, equivalent properties were examined firstly for naming inconsistencies.
 Then, ignoring property name, we looked at whether the values were structurally comparable; if two fields were of the same data type and, if complex, had structurally equivalent subfields, they were considered comparable, otherwise they were considered \emph{structurally inconsistent}.
-The ``birthday'' fields in Facebook and LinkedIn demonstrate structural inconcistancy; Facebook represents birthdays as a single string formatted ``MM/DD/YYYY'', while the LinkedIn representation is a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.
+The ``birthday'' fields in Facebook and LinkedIn demonstrate structural inconstancy; Facebook represents birthdays as a single string formatted ``MM/DD/YYYY'', while the LinkedIn representation is a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.
 
-Finally, we examined equivalent fields which were incomparabe for reasons other than structural or terminological differences.
-These \emph{modelling differences} arose from a variety of causes, including, but not limited to: measurement unit inconsistencies, measurement method (e.g., dewpoint vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').
+Finally, we examined equivalent fields which were incomparable for reasons other than structural or terminological differences.
+These \emph{modelling differences} arose from a variety of causes, including, but not limited to: measurement unit inconsistencies, measurement method (e.g., dew point vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').\looseness=-1
 
@@
 
-Looking at equivalent properties, we foun overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
-While some of these 'unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.
+Looking at equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
+While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.
 Only one Spotify, among music providers, included information about the album(s) featuring a given song.
@@
 Thus, terminological heterogeneity was prevalent across data records we observed.
-Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modeled as a subtree in the other.
-We also encountered only 2 examples where data was nested in textual description fields, both from records in the Weather category, which hints their intended use in RSS feed readers rather than in stuctured database applications.
+Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modelled as a subtree in the other.
+We also encountered only 2 examples where data was nested in textual description fields, both from records in the Weather category, which hints their intended use in RSS feed readers rather than in structured database applications.
 Examples of modelling/semantic heterogeneity instances were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and mismatches in units of measurement (commonly found in the Weather forecast measurements).
 
-Although not included in the table above, we also examined the source schemas of many of hte data sources for property enumerated type values.
+Although not included in the table above, we also examined the source schemas of many of the data sources for property enumerated type values.
 There was substantial terminological inconsistency among such values, but also cases where some sources had values not represented in the other source's sets.
@@
 In summary, while data records ranged in complexity, no data source subsumed the rest; all contained properties that other sources did not have.
-Moreover, the most common types of heterogeneity exhibited were terminological and value-related, although simpler kinds of structural hetereogeneity (pertaining to representing substructures of values) were also common.
+Moreover, the most common types of heterogeneity exhibited were terminological and value-related, although simpler kinds of structural heterogeneity (pertaining to representing substructures of values) were also common.
 
@@
 
-\subsubsection{Multipath selection: link-sliding for heterogeneous sets}
+\subsubsection{Multi-path selection: link-sliding for heterogeneous sets}
 
@@
 An instance's properties are displayed beneath it; selecting a property creates a \emph{path selection}, which dereferences all of the instances in a group that contain the selected property and displays their corresponding values.
-For example, when a user selects the \emph{manufacturer} property for a product, all products in the same group with a manufacturer property will be dereferenced, causing the manufactueres to be displayed alongside their corresponding products.
+For example, when a user selects the \emph{manufacturer} property for a product, all products in the same group with a manufacturer property will be dereferenced, causing the manufactures to be displayed alongside their corresponding products.
 When such a path selection is applied, the properties displayed are updated to be the set of properties of the \emph{terminating value of the path}, so that this value can be dereferenced further.
@@
 We selected six because it enable us to scope our study, we found through two pilot studies that six data sources provided a sufficient challenge for users to explore the data and required them to use DP's main features.
-We manually collected the data from the nine selected websites into jason files, and converted them to Microsoft Excel worksheets (which 40\% of the participants in PS.1 said that had previously used to compare data).
+We manually collected the data from the nine selected websites into JSON files, and converted them to Microsoft Excel worksheets (which 40\% of the participants in PS.1 said that had previously used to compare data).
 
@@
 \hline
-4) Restaurant they would book for 12 friends in Glagow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollopes Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia.
+4) Restaurant they would book for 12 friends in Glasgow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollope's Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia.
 & www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
@@
 
-\begin{table*}[htdp]
+\begin{table*}[htbp]
 \small
@@
 
-Since the tasks were subjective-choice, we could not evaluate the 'correctness' of people's answers.
+Since the tasks were subjective-choice, we could not evaluate the `correctness' of people's answers.
 However, to determine whether there was a difference in variability of answers between interface conditions, we plotted their choice on a histogram for each trial, per task and dataset for both conditions, visible in Figure \ref{fig:chosen_results}.
@@
 
-To determine whether participants used the data integration features of DP, we looked at use of the \emph{multipath selection} and \emph{Same-As} features of the system.
+To determine whether participants used the data integration features of DP, we looked at use of the \emph{multi-path selection} and \emph{Same-As} features of the system.
 Pertaining to path selection, all participants readily used the path selection process to select values.
-Five participants (25\%) deliberately used multipath selection, defined as selecting more than one active path per group at the same time to display common values across heterogeneous items.
-(We did not count accidental use of multipath, such as happened a few times when a participant had an already active path and wanted to switch to a separate path accidentally).
+Five participants (25\%) deliberately used multi-path selection, defined as selecting more than one active path per group at the same time to display common values across heterogeneous items.
+(We did not count accidental use of multi-path, such as happened a few times when a participant had an already active path and wanted to switch to a separate path accidentally).
 
@@
 We counted a single ``use'' as a single drag-and-drop of an individual item, or a drag-and-drop operation of an entire group onto another.
-(We did not differentiate enumerated value consolidation from entity consolidation, as these operations were not discernably different from the interface perspective.)
+(We did not differentiate enumerated value consolidation from entity consolidation, as these operations were not discernibly different from the interface perspective.)
 
@@
 
-\subsection{Does DataPalette enable data integration?}
+\subsection{Does DataPalette facilitate serendipitous data integration?}
 
@@
 The second, h2, was that DP enables people to mix and integrate heterogeneous data in the process of performing their tasks.
-The third, h3, was that using DP facilitatedp task completion (h3).
-Using real data obtained from typical data sources identified during our prestudy (PS.1), we tested the DP prototype and its integration interaction mechanisms with ``real'' users, who were students, staff and alumni of our University who had varying backgrounds and levels of expertise with computers.
+The third, h3, was that using DP facilitated task completion (h3).
+Using real data obtained from typical data sources identified during our pre-study (PS.1), we tested the DP prototype and its integration interaction mechanisms with ``real'' users, who were students, staff and alumni of our University who had varying backgrounds and levels of expertise with computers.
 
@@
 All participants reported that they understood the tool, and that it was easy to use.
-All participants managed to use the interface to view, organize and collate data without running into major roadblocks or confusion, and all effectively were able to compare multiple attributes of heterogeneous data items directly.
+All participants managed to use the interface to view, organise and collate data without running into major roadblocks or confusion, and all effectively were able to compare multiple attributes of heterogeneous data items directly.
 More than half of the participants used DP's visualisation features (charting or map -- often several times during the trial -- suggesting that these features were useful and usable as well.
@@
 Properties are tricky.
-Sometimes you don’t care [which data source] a property is from, like ``address'' or ``phone number'' - for these the way it’s done now is fine.
-But in some cases you need context of where the properties come from in order to know what it really means - like for ``rating'', it makes a big difference whether you’re talking about ``Yelp rating'' or some random reviewer's rating.
-You also don’t know what typical ratings are, whether 5 stars are much better than 4, etc.
+Sometimes you don't care [which data source] a property is from, like ``address'' or ``phone number'' - for these the way it's done now is fine.
+But in some cases you need context of where the properties come from in order to know what it really means - like for ``rating'', it makes a big difference whether you're talking about ``Yelp rating'' or some random reviewer's rating.
+You also don't know what typical ratings are, whether 5 stars are much better than 4, etc.
 \end{quote}
@@
 Pertaining to h2, all participants were able to work with multiple sets of heterogeneous data effectively.
-With respect to integration specifically, a majority (80\%) of the particiapnts successfully and deliberately integrated data using drag-and-drop SameAs capabilities.
+With respect to integration specifically, a majority (80\%) of the participants successfully and deliberately integrated data using drag-and-drop SameAs capabilities.
 While single-paths were used, multi-path features were not as widely used, only by 4 participants.
-We believe that participants may not have realized that creating multiple path selections per group was possible, instead assuming it to be similar to the single ``Current path'' state per window of most file managers.
+We believe that participants may not have realised that creating multiple path selections per group was possible, instead assuming it to be similar to the single ``Current path'' state per window of most file managers.
 
@@
 A key design decision we made at the outset was to target the most common, simpler forms of heterogeneity we observed in PS.2.
-Given the use of theinterface mechanisms we observed, we wish to next broaden the scope of capabilities of DP so that it can become more useful in real settings.
+Given the use of the interface mechanisms we observed, we wish to next broaden the scope of capabilities of DP so that it can become more useful in real settings.
 Among the integration capabilities we are considering are, first, field combination and splitting (by delimiters), in order to address the kinds of structural heterogeneity observed in several cases of PS.2, along with a framework for converting and reconciling units of measure, which would address several of the modelling issues.
@@
 
-% - data integration literature demands 'correct' but perfect is the enemy of the good.
+% - data integration literature demands `correct' but perfect is the enemy of the good.
 % ``perfect'' be the enemy of the good
@@
 
-Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analyists the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{}} into the hands of the ordinary citizen \cite{datajournalismhandbook}.
+Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \cite{datajournalismhandbook}.
 In order for this to happen, however, tools that allow people to effectively combine data more effectively than current structured PIM tools and generic spreadsheet tools will become increasingly important.

### 2012-09-19 chi2012.tex

--- 2012-09-19 chi2012.tex
@@
 \subsubsection{Results - Heterogeneous Data Tasks}
-The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
+
+The 8 participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18-25 and 26-32.   
 % mention # that were computer scientists here?

### 2012-09-19 chi2012.tex
Fixing minor errors in PS2 results section.

@@
 \subsubsection{Results}
-Table \ref{tbl:prestudy2} displays the raw results of the process just described.
-A number of features were quite striking about the data.
-First, while most of the data records were small and relatively simple, nearly one in each category produced substantially more complex records than the rest; in particular, Amazon, Soundcloud, Twitter and Weather Underground stood out in their respective categories.
+Table \ref{tbl:prestudy2} shows the raw results of the process described.
+A number of features were quite striking about the data; First, while most of the data records were small and relatively simple, nearly one in each category produced substantially more complex records than the rest.
+In particular, Amazon, Soundcloud, Twitter and Weather Underground stood out in their respective categories.
 
@@
 While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.
-Only one Spotify, among music providers, included information about the album(s) featuring a given song.
+Only Spotify, out of the 4 music providers, included information about the album(s) featuring a given song.
 Such omissions show us that these data APIs reflect each provider's own specific capabilities and needs, and suggests that integrating information from multiple sources may allow more complete and useful representations to be assembled.
@@
 There was substantial terminological inconsistency among such values, but also cases where some sources had values not represented in the other source's sets.
-For example, the property ``Gender''  on Facebook allowed only supported ``Male'' and ``Female'' for values, while Google+ included a third option,``other''.
-``Relationship status'' was another which was similarly complex.
-None of the string representations of enumerated value sets matched across data sources perfectly.
+For example, the property ``gender''  on Facebook allowed only supported ``male'' and ``female'' for values, while Google+ included a third option,``other''.
+``Relationship status'' was another which was similarly complex, as none of the string representations of enumerated value sets matched across data sources perfectly.
 

### 2012-09-19 chi2012.tex
blahdy blah

--- 2012-09-19 chi2012.tex
@@
 \begin{abstract}
-Although people naturally rely upon multiple, diverse information sources in the course of routine problem-solving and decision-making tasks, support for data integration in Personal Information Management (PIM) tools remains poor, if largely nonexistent.
-This paper presents an exploratory, sequential, mixed-method study of ways to extend PIM tools to support user-driven data integration while performing their primary tasks.
-We present results of two pre-study in which requirements, we designed and evaluated DataPalette.
+While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
+In this paper, we present an exploratory initial investigation of ways to support user-driven, ad-hoc data integration in PIM tools, towards opening up for use the emerging ecosystem of structured data now available via APIs and structured data feeds on the Web.
+We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on blah and blah, finding that blah ad blah.
+These
+
+designed and evaluated DataPalette.  
 % We find that the ``light-touch'' data integration mechanisms introduced allow people to effectively work with and combine information from multiple sources, and more thoroughly justify their choices
@@
 There were two driving questions behind this inquiry; first, were people increasingly relying only on a handful of ``supersites'' (Facebook, Wikipedia), or searching across distributed sources of information?
-If the former were true, merely integrating PIM tools with the small number of supersites would provide greater benefit than tackling the substantially more challenging problem of integrating data from arbitrary sources on the Web.
+If the former were true, merely integrating PIM tools with the small numbr of supersites would provide greater benefit than tackling the substantially more challenging problem of integrating data from arbitrary sources on the Web.
 Second, why did people choose a single or multiple sources for information gathering?
@@
 \subsubsection{Results - Heterogeneous Data Tasks}
-
-The 8 participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18-25 and 26-32.   
+The participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
 % mention # that were computer scientists here?

### 2012-09-19 chi2012.tex
Ran the entire thing through ispell with a british englosh dictionary.

--- 2012-09-19 chi2012.tex
@@
 % Load basic packages
-\usepackage{balance}  % to better equalize the last page
+\usepackage{balance}  % to better equalise the last page
 \usepackage{graphics} % for EPS, load graphicx instead
@@
 In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web.
-This information has emerged from many channels, from the APIs now offered by most of the major retailers, apps, and social media sites, to the massive publishing of datasets by governmental and other institutions, to even the emergence of new kinds of data sources, such as wearable and smartphone-based biosensors.
+This information has emerged from many channels, from the APIs now offered by most of the major retailers, apps, and social media sites, to the massive publishing of datasets by governmental and other institutions, to even the emergence of new kinds of data sources, such as wearable and smartphone-based bio-sensors.
 A primary goal of making much of this data available has been to give end-user citizens the ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
@@
 
-% i think we do not need to differentiate between semantic web and databse integration here, we're saying they're both bad, let's lump them together to keep the argument flowing and not kill our readers
+% i think we do not need to differentiate between semantic web and database integration here, we're saying they're both bad, let's lump them together to keep the argument flowing and not kill our readers
 The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} research communities have primarily focused efforts on purely-automatic approaches to the problem of data integration.
 Such approaches typically take the form of either \emph{ontology} or \emph{schema-level matching}, or \emph{direct} or \emph{instance-level} matching.
-We presently focus  on \emph{instance-level} matching, as schemas may not always be available, nor easily accesssed by users.
+We presently focus  on \emph{instance-level} matching, as schemas may not always be available, nor easily accessed by users.
 Approaches to automatic instance matching have included applied natural language processing (NLP) techniques for terminological matching (e.g., \cite{euzenat2004api}), machine-learning techniques that use both structural and terminological features for matching \cite{doan2003learning}, probabilistic approaches \cite{suchanek2011paris}, and instance schema comparisons using structural and logical definitions \cite{castano2006matching}.
@@
 
-\section{Pre-studies: The Need for and Challenges in realizing Ad-Hoc Data Integration in PIM Tools}
+\section{Pre-studies: The Need for and Challenges in realising Ad-Hoc Data Integration in PIM Tools}
 
@@
 ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property.
-To reconcile and consolidate corresponding values from different  datasources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.
+To reconcile and consolidate corresponding values from different  data sources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.
 
@@
 
-%% it would be nice if we had space to summarize but i fear we don't
+%% it would be nice if we had space to summarise but i fear we don't
 
@@
 Selecting a further property of the %% resulting value further dereferences this value.
-%% \item \emph{Drag-and-drop coreference / value combination} - In order %% to specify that two instances correspond to the same entity, they can %% be simly dragged together
-%% \item \emph{Group-based Drag-and-drop matching} - In order to specify %% that two instances correspond to the same entity, they can be simly %% dragged together
+%% \item \emph{Drag-and-drop coreference / value combination} - In order %% to specify that two instances correspond to the same entity, they can %% be simply dragged together
+%% \item \emph{Group-based Drag-and-drop matching} - In order to specify %% that two instances correspond to the same entity, they can be simply %% dragged together
 %% \item \emph{Universal model-based brushing} -
@@
 
-% heather: instead of these low-level hypotheses, i've made higher level ones that we can explicitly address in the disucssion section - the more hypotheses we have, the more we have to come back to them; since we're short on space we might as well keep it simple!
+% heather: instead of these low-level hypotheses, I've made higher level ones that we can explicitly address in the discussion section - the more hypotheses we have, the more we have to come back to them; since we're short on space we might as well keep it simple!
 
@@
 Each participant was given a maximum of 10 minutes to complete a task, but were told that if they finished earlier they could tell us and move on to the next condition.
-In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.
+In order to select which tasks the participant was given, we used a Latin square to permutate the tasks and conditions.
 In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
@@
 Each participant was given 10 minutes to complete a task.
-In order to select which tasks the participant was given, we used a latin square to permutate the tasks and conditions.
+In order to select which tasks the participant was given, we used a Latin square to permutate the tasks and conditions.
 In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
@@
 Seven were non-students, comprising University staff and alumni; out of the remaining 13, eight studied computer science and 5 studied other subjects.
-On average, each study took 40 mins to complete.
+On average, each study took 40 minutes to complete.
 The information in Table \ref{tbl:data} summarises the metrics that were captured of the trials for each participant.
@@
 %		We focused on finite (small) sets Streaming data (how do select future data --) 
-%		Scale - small data, what are the facilites required for large data
+%		Scale - small data, what are the facilities required for large data
 %		Collaborative - how might interaction support multiple users collaborating?
@@
 The upfront costs of such complete integration are typically high, but are typically justified in the long-term, high-volume use cases imagined.
-In typical sensemaking scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.
+In typical sense-making scenarios of PIM, however, people might need to combine information from multiple sources for a one-off task; meaning that any upfront cost would directly impact the time and effort a person has allocated for their task.
 Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.

### 2012-09-19 chi2012.tex
blahdyblah'

--- 2012-09-19 chi2012.tex
@@
 In this paper, we present an exploratory initial investigation of ways to support user-driven, ad-hoc data integration in PIM tools, towards opening up for use the emerging ecosystem of structured data now available via APIs and structured data feeds on the Web.
-We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on blah and blah, finding that blah ad blah.
-These
+We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on blah and blah, finding that blahdyblah.
+These observations led to the design of DataPalette, an interface for ``light-touch'' data integration that supported simple drag-and-drop and path-selection mechanisms for managing terminological and structurally heterogeneous data items.
+Our lab study of DataPalette showed that users understood the interaction mechanisms introduced and the problems they were designed to solve, and that DesignPalette allowed them to make more carefully informed decisions and draw from more data sources while performing multidimensional subjective choice-based tasks.
 
-designed and evaluated DataPalette.  
 % We find that the ``light-touch'' data integration mechanisms introduced allow people to effectively work with and combine information from multiple sources, and more thoroughly justify their choices

### 2012-09-19 chi2012.tex
Study limitiations added

--- 2012-09-19 chi2012.tex
@@
 In this paper, we present an exploratory initial investigation of ways to support user-driven, ad-hoc data integration in PIM tools, towards opening up for use the emerging ecosystem of structured data now available via APIs and structured data feeds on the Web.
-We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on blah and blah, finding that blahdyblah.
-These observations led to the design of DataPalette, an interface for ``light-touch'' data integration that supported simple drag-and-drop and path-selection mechanisms for managing terminological and structurally heterogeneous data items.
-Our lab study of DataPalette showed that users understood the interaction mechanisms introduced and the problems they were designed to solve, and that DesignPalette allowed them to make more carefully informed decisions and draw from more data sources while performing multidimensional subjective choice-based tasks.
+We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on blah and blah, finding that blah ad blah.
+These
 
+designed and evaluated DataPalette.  
 % We find that the ``light-touch'' data integration mechanisms introduced allow people to effectively work with and combine information from multiple sources, and more thoroughly justify their choices
@@
 
-\subsection{Study limitations}
+\subsection{Study Limitations}
+
+Our study had several limitations.
+First, our study consisted of 20 participants which meant that we could not show that some of our data was statistically significant.
+Second, we gave the participants an artificially limited amount of time (10 minutes) to make their decisions.
+Given additional time, the participants choices may have changed.
+Third, the data sources were fixed, because we were not testing the users' capabilities to search and retrieve data from the web.
+Given unlimited access to data sources online, users would be able to compare different aspects, which may also alter their choices.
 

### 2012-09-19 chi2012.tex
abstract

--- 2012-09-19 chi2012.tex
@@
 While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
-In this paper, we present an exploratory initial investigation of ways to support user-driven, ad-hoc data integration in PIM tools, towards opening up for use the emerging ecosystem of structured data now available via APIs and structured data feeds on the Web.
-We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on blah and blah, finding that blahdyblah.
-These observations led to the design of DataPalette, an interface for ``light-touch'' data integration that supported simple drag-and-drop and path-selection mechanisms for managing terminological and structurally heterogeneous data items.
-Our lab study of DataPalette showed that users understood the interaction mechanisms introduced and the problems they were designed to solve, and that DesignPalette allowed them to make more carefully informed decisions and draw from more data sources while performing multidimensional subjective choice-based tasks.
+In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of emerging ecosystem of structured data available on the Web.
+We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
+These observations led to the design of DataPalette, an interface for ``light-touch'' data integration that introduced drag-and-drop and path-selection mechanisms for managing collections of terminologically and structurally- heterogeneous data items.
+Our lab study of DataPalette with actual data obtained from a selection of Web sites  showed that users understood the interaction mechanisms introduced and the problems they were designed to address, and that DesignPalette allowed them to make more carefully informed decisions and draw from more data sources while performing multidimensional subjective choice-based tasks.
 

### 2012-09-19 chi2012.tex
abstract

@@
 While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
-In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of emerging ecosystem of structured data available on the Web.
+In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data available on the Web.
 We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.

### 2012-09-19 chi2012.tex
Split and minor changes

--- 2012-09-19 chi2012.tex
@@
 We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
-These observations led to the design of DataPalette, an interface for ``light-touch'' data integration that introduced drag-and-drop and path-selection mechanisms for managing collections of terminologically and structurally- heterogeneous data items.
+These observations led to the design of DataPalette, an interface for ``light-touch'' data integration that introduced drag-and-drop and path-selection mechanisms for managing collections of terminologically and structurally-heterogeneous data.
 Our lab study of DataPalette with actual data obtained from a selection of Web sites  showed that users understood the interaction mechanisms introduced and the problems they were designed to address, and that DesignPalette allowed them to make more carefully informed decisions and draw from more data sources while performing multidimensional subjective choice-based tasks.

### 2012-09-19 chi2012.tex

@@
 While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
-In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of emerging ecosystem of structured data available on the Web.
+In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of emerging ecosystems of structured data available on the Web.
 We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.

### 2012-09-19 chi2012.tex
Fix of abstract I

@@
 These observations led to the design of DataPalette, an interface for ``light-touch'' data integration that introduced drag-and-drop and path-selection mechanisms for managing collections of terminologically and structurally-heterogeneous data.
-Our lab study of DataPalette with actual data obtained from a selection of Web sites  showed that users understood the interaction mechanisms introduced and the problems they were designed to address, and that DesignPalette allowed them to make more carefully informed decisions and draw from more data sources while performing multidimensional subjective choice-based tasks.
+Our DataPalette lab study showed that users understood the interaction mechanisms and the problems they were designed to address, and that it allowed them to make more carefully informed decisions and draw from more data sources while making multidimensional subjective decisions.
 

### 2012-09-19 chi2012.tex
Fix abstract II

@@
 We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
-These observations led to the design of DataPalette, an interface for ``light-touch'' data integration that introduced drag-and-drop and path-selection mechanisms for managing collections of terminologically and structurally-heterogeneous data.
+These observations led to the design of DataPalette, an interface for ``light-touch'' data integration that introduced drag-and-drop and path-selection mechanisms for managing collections of terminologically and structurally heterogeneous data.
 Our DataPalette lab study showed that users understood the interaction mechanisms and the problems they were designed to address, and that it allowed them to make more carefully informed decisions and draw from more data sources while making multidimensional subjective decisions.

### 2012-09-19 chi2012.tex
tweakies

@@
 We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
-These observations led to the design of DataPalette, an interface for ``light-touch'' data integration that introduced drag-and-drop and path-selection mechanisms for managing collections of terminologically and structurally heterogeneous data.
+These observations led to DataPalette, an interface that introduced drag-and-drop and path-selection mechanisms for working with terminologically and structurally heterogeneous data.
 Our DataPalette lab study showed that users understood the interaction mechanisms and the problems they were designed to address, and that it allowed them to make more carefully informed decisions and draw from more data sources while making multidimensional subjective decisions.

### 2012-09-19 chi2012.tex
tweakies

@@
 We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
-These observations led to DataPalette, an interface that introduced drag-and-drop and path-selection mechanisms for working with terminologically and structurally heterogeneous data.
-Our DataPalette lab study showed that users understood the interaction mechanisms and the problems they were designed to address, and that it allowed them to make more carefully informed decisions and draw from more data sources while making multidimensional subjective decisions.
+These observations led to our design of DataPalette, an interface that introduced drag-and-drop and path-selection mechanisms for working with terminologically and structurally heterogeneous data.  
+A lab study showed that users understood the interaction mechanisms and the problems they were designed to address, and that it allowed them to make more carefully informed decisions and draw from more data sources while making multidimensional subjective decisions.
 

### 2012-09-19 chi2012.tex
147

@@
 \begin{abstract}
+
 While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
@@
 These observations led to our design of DataPalette, an interface that introduced drag-and-drop and path-selection mechanisms for working with terminologically and structurally heterogeneous data.  
-A lab study showed that users understood the interaction mechanisms and the problems they were designed to address, and that it allowed them to make more carefully informed decisions and draw from more data sources while making multidimensional subjective decisions.
+Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions, drawing upon more data sources with DataPalette than a control condition, during multidimensional subjective choice decisions.
+
+%% While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
+%% In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of emerging ecosystems of structured data available on the Web.
+%% We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
+%% These observations led to our design of DataPalette, an interface that introduced drag-and-drop and path-selection mechanisms for working with terminologically and structurally heterogeneous data.
+%% A lab study showed that users understood the interaction mechanisms introduced, and that DataPalette allowed them to make more carefully-informed decisions, drawing from more data sources while making multidimensional subjective decisions.
 

### 2012-09-19 chi2012.tex
147

@@
 While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
-In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of emerging ecosystems of structured data available on the Web.
+In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data available on the Web.
 We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.

### 2012-09-19 chi2012.tex
author

--- 2012-09-19 chi2012.tex
@@
 While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
-In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of emerging ecosystems of structured data available on the Web.
-We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
-These observations led to our design of DataPalette, an interface that introduced drag-and-drop and path-selection mechanisms for working with terminologically and structurally heterogeneous data.  
-Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions, drawing upon more data sources with DataPalette than a control condition, during multidimensional subjective choice decisions.
+In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data on the Web.
+We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies examining the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
+These observations led to our design of DataPalette, an interface that introduced simple drag-and-drop and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
+Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions, drawing upon more data sources with DataPalette than a control condition during multidimensional, subjective-choice decisions.
 
@@
 
-\keywords{}
+\keywords{Personal Information Management, Personal Data Management, User-driven data integration}
 

### 2012-09-19 chi2012.tex
Proof reading and minor fixes to results section.

--- 2012-09-19 chi2012.tex
@@
 \begin{abstract}
+While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
+In this paper, we present an exploratory initial investigation of ways to support user-driven, ad-hoc data integration in PIM tools, towards opening up for use the emerging ecosystem of structured data now available via APIs and structured data feeds on the Web.
+We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on blah and blah, finding that blah ad blah.
+These
 
-While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
-In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of emerging ecosystems of structured data available on the Web.
-We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
-These observations led to our design of DataPalette, an interface that introduced drag-and-drop and path-selection mechanisms for working with terminologically and structurally heterogeneous data.  
-Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions, drawing upon more data sources with DataPalette than a control condition, during multidimensional subjective choice decisions.
-
-%% While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
-%% In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of emerging ecosystems of structured data available on the Web.
-%% We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies focusing on the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
-%% These observations led to our design of DataPalette, an interface that introduced drag-and-drop and path-selection mechanisms for working with terminologically and structurally heterogeneous data.
-%% A lab study showed that users understood the interaction mechanisms introduced, and that DataPalette allowed them to make more carefully-informed decisions, drawing from more data sources while making multidimensional subjective decisions.
-
+designed and evaluated DataPalette.  
 % We find that the ``light-touch'' data integration mechanisms introduced allow people to effectively work with and combine information from multiple sources, and more thoroughly justify their choices
@@
 
-Twenty-six (26) participants responded, from which we selected the first 20.
-Ten participants were female.
-All of the participants spoke English, but 12 of them were non-native English speakers.
+Twenty-six (26) participants responded, from which we selected the first 20, ten of which were female.
+All participants spoke English, but 12 of them were non-native English speakers.
 Seven were non-students, comprising University staff and alumni; out of the remaining 13, eight studied computer science and 5 studied other subjects.
@@
 
-The following four sections describes the metrics we used to measure task performance.
+The following four sections describe the metrics we used to measure task performance.
 %We used three metrics to measure task performance: time to completion; the number of factors used in the choice; and the number of candidates evaluated.
@@
 \subsubsection{Efficiency: Time taken}
-Participants spent an average of 8.9 minutes performing the tasks, with our interface, condition A, taking slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).
+Participants spent an average of 8.9 minutes performing the tasks.
+Our interface, condition A, took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).
 An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time (shown in Figure \ref{fig:time}).
@@
 The second metric was the number of factors the participant considered in making their final choice.
-To determine this, we asked participants explain precisely why they made their choice, and matched their justifications to properties in the data.
-This analysis revealed a significant effect of interface on number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
+To determine this, we asked participants to explain precisely why they made their choice, and matched their justifications to properties in the data.
+This analysis revealed the interface used made a significant impact on the number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
 
@@
 \subsection{Condition B - Results and Observations}
-In the B conditions, participants were given the option of using websites and or a spreadsheet to make their choice.
+In the B conditions, participants were given the option of using websites and/or a spreadsheet to make their choice.
 For B trials, we measured the fraction of the trial time spent in Excel versus looking at the sites in a 5-value range (0,25,50,75,100\%).
@@
 All participants managed to use the interface to view, organise and collate data without running into major roadblocks or confusion, and all effectively were able to compare multiple attributes of heterogeneous data items directly.
-More than half of the participants used DP's visualisation features (charting or map -- often several times during the trial -- suggesting that these features were useful and usable as well.
+More than half of the participants used DP's visualisation features (charting or map) -- often several times during the trial -- suggesting that these features were useful and usable as well.
 
-Additional evidence that participant had a solid understanding of the system was their feedback, specifically feature requests and desired capabilities.
+Additional evidence that participants had a solid understanding of the system was their feedback, specifically feature requests and desired capabilities.
 Several participants requested search functionality, sort and filter functionality, the ability to display multiple properties for a single instance simultaneously, and visualise multiple attributes in a 2 or 3-dimensional visualisation.
@@
 
-Assessing whether DP's improved task performance (h3) was the most difficult to ascertain, given the small sample size of our study and large number of factors influencing each person's performance.
+Assessing whether DP improved task performance (h3) was difficult given the small sample size of our study and large number of factors influencing each person's performance.
 While tasks were completed on average slightly faster than the control interface, these differences were not statistically significant.
@@
 
-In addition to the aforementioned feature requests, we observed a number of ``pain points'' in the design of the interface that we wish to address in the next iteration.
+In addition to the aforementioned feature requests, we observed a number of ``pain points'' in the design of the DP interface.
 One such pain point was window layout and screen real estate management; participants were prone to opening up a large number of windows, including maps and charts.
@@
 
-\subsection{Study Limitations}
-
-Our study had several limitations.
-First, our study consisted of 20 participants which meant that we could not show that some of our data was statistically significant.
-Second, we gave the participants an artificially limited amount of time (10 minutes) to make their decisions.
-Given additional time, the participants choices may have changed.
-Third, the data sources were fixed, because we were not testing the users' capabilities to search and retrieve data from the web.
-Given unlimited access to data sources online, users would be able to compare different aspects, which may also alter their choices.
+\subsection{Study limitations}
 

### 2012-09-19 chi2012.tex

--- 2012-09-19 chi2012.tex
@@
 All participants managed to use the interface to view, organise and collate data without running into major roadblocks or confusion, and all effectively were able to compare multiple attributes of heterogeneous data items directly.
-More than half of the participants used DP's visualisation features (charting or map -- often several times during the trial -- suggesting that these features were useful and usable as well.
-
-Additional evidence that participant had a solid understanding of the system was their feedback, specifically feature requests and desired capabilities.
+More than half of the participants used DP's visualisation features (charting or map) -- often several times during the trial -- suggesting that these features were useful and usable as well.
+
+Additional evidence that participants had a solid understanding of the system was their feedback, specifically feature requests and desired capabilities.
 Several participants requested search functionality, sort and filter functionality, the ability to display multiple properties for a single instance simultaneously, and visualise multiple attributes in a 2 or 3-dimensional visualisation.

### 2012-09-19 chi2012.tex
moved fig2
Minor fixes to Introduction.

@@
 
-We hypothesise that end-users will benefit from tools that allow them effectively browse, use and consume heterogeneous structured data from diverse sources.
+We hypothesise that end-users will benefit from tools that allow them to effectively browse, use and consume heterogeneous structured data from diverse sources.
 In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data.
 We used a three-stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of information sources, and the processes that they relied upon to perform such tasks.
-Our results suggested that people rely on multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and for alternative perspectives.
+Our results suggested that people rely on multiple and diverse sources over singular, integrated ones for a number of reasons, including coverage, reliability, and alternative perspectives.
 
@@
 
-Third, we developed DataPalette, an interface that enables serendipitous ``data mixing,'' negating the need for people to write bespoke code to effectively combine and compare heterogenous data.
+Third, we developed DataPalette, an interface that enables serendipitous ``data mixing'', negating the need for people to write bespoke code to effectively combine and compare heterogenous data.
 It was principally designed to support people's data integration needs for data heterogeneity, using observations made from the previous studies.
 In particular, we designed a user-centric interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
-We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision making tasks using multiple sources.
+We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision-making tasks using multiple sources.
 
@@
 Similarly, Haystack, a personal information platform, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
-Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
+Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user manually specify such a mapping translation themselves.
 

### 2012-09-19 chi2012.tex

@@
 
-\begin{table*}[t]
+\begin{table}[t]
 \begin{center}
 \small
-\begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
-\hline
-Task	 &Choices	&Data Sources\\
-\hline
-1) Three universities they would apply to study Sport Science & Universities: 1) Loughborough,  2) Durham,  3) Exeter, 4) Edinburgh, 5) Birmingham, and 6) Bath & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
-\hline
-2) Three universities they would apply to study History & Universities: 1) Cambridge, 2) London School of Economics and Political Science, 3) Durham, 4) Oxford, 5) St Andrews, and 6) University College London & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
-\hline
-3) Restaurant they would book for 12 friends in Cambridge&Restaurants : 1) Alimentum, 2) Gardenia Restaurant, 3) The Cambridge Chop House, 4) Al Casbah, 5) Midsummer House Restaurant, and 6) Nando's Restaurant.
-& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
-\hline
-4) Restaurant they would book for 12 friends in Glasgow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollope's Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia.
-& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
+\begin{tabular}{|p{3cm}|p{5cm}|}
+\hline
+Task		&Data Sources\\
+\hline
+Three universities they would apply to study 1)Sport Science and 2) History & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
+\hline
+Restaurant they would book for 12 friends in 3) Cambridge and 4) Glasgow& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
 \hline
@@
 \normalsize
-\end{table*}%
+\end{table}%
+
+%\begin{table*}[t]
+%\begin{center}
+%\small
+%\begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
+%\hline
+%Task	 &Choices	&Data Sources\\
+%\hline
+%1) Three universities they would apply to study Sport Science & Universities: 1) Loughborough,  2) Durham,  3) Exeter, 4) Edinburgh, 5) Birmingham, and 6) Bath & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
+%\hline
+%2) Three universities they would apply to study History & Universities: 1) Cambridge, 2) London School of Economics and Political Science, 3) Durham, 4) Oxford, 5) St Andrews, and 6) University College London & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
+%\hline
+%3) Restaurant they would book for 12 friends in Cambridge&Restaurants : 1) Alimentum, 2) Gardenia Restaurant, 3) The Cambridge Chop House, 4) Al Casbah, 5) Midsummer House Restaurant, and 6) Nando's Restaurant.
+& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
+%\hline
+%4) Restaurant they would book for 12 friends in Glasgow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollope's Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia.
+& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
+%\hline
+%\end{tabular}
+%\end{center}
+%\caption{Tasks, choices and data sources} \label{tab:studyfactors}
+%\normalsize
+%\end{table*}%
 

### 2012-09-19 chi2012.tex

@@
 
-We hypothesise that end-users will benefit from tools that allow them to effectively browse, use and consume heterogeneous structured data from diverse sources.
+We hypothesise that end-users will benefit from tools that allow them effectively browse, use and consume heterogeneous structured data from diverse sources.
 In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data.
 We used a three-stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of information sources, and the processes that they relied upon to perform such tasks.
-Our results suggested that people rely on multiple and diverse sources over singular, integrated ones for a number of reasons, including coverage, reliability, and alternative perspectives.
+Our results suggested that people rely on multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and for alternative perspectives.
 
@@
 
-Third, we developed DataPalette, an interface that enables serendipitous ``data mixing'', negating the need for people to write bespoke code to effectively combine and compare heterogenous data.
+Third, we developed DataPalette, an interface that enables serendipitous ``data mixing,'' negating the need for people to write bespoke code to effectively combine and compare heterogenous data.
 It was principally designed to support people's data integration needs for data heterogeneity, using observations made from the previous studies.
 In particular, we designed a user-centric interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
-We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision-making tasks using multiple sources.
+We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision making tasks using multiple sources.
 
@@
 Similarly, Haystack, a personal information platform, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
-Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user manually specify such a mapping translation themselves.
+Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
 

### 2012-09-19 chi2012.tex
Figs fixed

--- 2012-09-19 chi2012.tex
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=8.5cm]{img/multipathing}
-\caption{To see properties, users click on the name of the property. When instances do not have a particular property, additional property names can be added.}
-\label{fig:multipathing}
-\end{center}
-\end{figure}
-
 Typically, users are interested in viewing and comparing the \emph{properties} of instances; for example, a user might be keen to examine a product's price, rating, manufacturer and so on.
@@
 
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=8.7cm]{img/multipathing}
+\caption{To see properties, users click on the name of the property. When instances do not have a particular property, additional property names can be added.}
+\label{fig:multipathing}
+\end{center}
+\end{figure}
+
 When instances cannot be dereferenced by the chosen path selection, an additional path selection can be created.
@@
 \begin{center}
-\includegraphics[width=8.5cm]{img/sameas}
+\includegraphics[width=8.7cm]{img/sameas}
 \caption{\emph{Coreference consolidation with drag-and-drop Same-As} - When two instances represent the same entity, users can drag one on top of the other, and they are combined - effectively displaying the union of properties of both source entities.

### 2012-09-19 chi2012.tex
Minor fixes in background section.

--- 2012-09-19 chi2012.tex
@@
 Our primary motivation stems from field studies in PIM, where consolidating and working with data from multiple sources has remained a challenge for over a decade.
-Enabling PIM tools to supporting ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data heterogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
+Enabling PIM tools to support ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data heterogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
 As described by Alon Halvey:

### 2012-09-19 chi2012.tex

--- 2012-09-19 chi2012.tex
@@
 We based our two tasks on choosing a university and picking a restaurant for a large social event, because in PS.1 the majority of people stated that they had used multiple websites for these tasks.
-We used different two different variables for both tasks, where the restaurant was located (Glasgow and Cambridge) and which subject people choose (sport science and history).
+We used different two different variables for both tasks, where the restaurant was located (Glasgow and Cambridge) and which subject people chose (sport science and history).
 The cities Glasgow and Cambridge were chosen because they are not geographically close to the participants, they are two heavily populated areas that have a choice of restaurants, and they contained detail data about the same restaurants on various review sites.

### 2012-09-19 chi2012.tex
Change as discussed

--- 2012-09-19 chi2012.tex
@@
 Our primary motivation stems from field studies in PIM, where consolidating and working with data from multiple sources has remained a challenge for over a decade.
-Enabling PIM tools to support ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data heterogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
+Enabling PIM tools to supporting ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data heterogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
 As described by Alon Halvey:
@@
 Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-maintained information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
+
+A number of research PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.
+For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
+Similarly, Haystack, a personal information platform, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
+Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
+
+% i think we do not need to differentiate between semantic web and database integration here, we're saying they're both bad, let's lump them together to keep the argument flowing and not kill our readers
+The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} research communities have primarily focused efforts on purely-automatic approaches to the problem of data integration.
+Such approaches typically take the form of either \emph{ontology} or \emph{schema-level matching}, or \emph{direct} or \emph{instance-level} matching.
+We presently focus  on \emph{instance-level} matching, as schemas may not always be available, nor easily accessed by users.
+Approaches to automatic instance matching have included applied natural language processing (NLP) techniques for terminological matching (e.g., \cite{euzenat2004api}), machine-learning techniques that use both structural and terminological features for matching \cite{doan2003learning}, probabilistic approaches \cite{suchanek2011paris}, and instance schema comparisons using structural and logical definitions \cite{castano2006matching}.
+In real-world settings, however, manual and semi-automatic matching approaches are favoured over automatic because of the need for high degrees of accuracy.
 
@@
 
-% i think we do not need to differentiate between semantic web and database integration here, we're saying they're both bad, let's lump them together to keep the argument flowing and not kill our readers
-The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} research communities have primarily focused efforts on purely-automatic approaches to the problem of data integration.
-Such approaches typically take the form of either \emph{ontology} or \emph{schema-level matching}, or \emph{direct} or \emph{instance-level} matching.
-We presently focus  on \emph{instance-level} matching, as schemas may not always be available, nor easily accessed by users.
-Approaches to automatic instance matching have included applied natural language processing (NLP) techniques for terminological matching (e.g., \cite{euzenat2004api}), machine-learning techniques that use both structural and terminological features for matching \cite{doan2003learning}, probabilistic approaches \cite{suchanek2011paris}, and instance schema comparisons using structural and logical definitions \cite{castano2006matching}.
-In real-world settings, however, manual and semi-automatic matching approaches are favoured over automatic because of the need for high degrees of accuracy.
-
-A different approach entirely has been taken in the web mash-up community; this community has largely focused on end-user interfaces that allow people to interactively reconcile data from multiple sources.
+%A different approach entirely has been taken in the web mash-up community; this community has largely focused on end-user interfaces that allow people to interactively reconcile data from multiple sources.
 Yahoo Pipes!
@@
 Therefore, in our work, we adopt a semi-automated matching approach, which will support the user to make alignments between web data and their personal data.
-
-A number of research PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.
-For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
-Similarly, Haystack, a personal information platform, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
-Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user manually specify such a mapping translation themselves.
 
@@
 \section{Designing DataPalette : An Interface for Serendipitous Data Mixing}
+
 \begin{figure*}[tbp]
@@
 \includegraphics[width=18cm]{img/screenshot}
-
 \caption{DataPalette Workspace, a file manager-inspired Drag and Drop interface for structured data mixing.
 Collapsible displays of data sources are listed on the left, from which instances or entire sets of items can be dragged to form a window (group) on the workspace.}
-
 \label{fig:workspace}
@@
 
-At a high level, the overall goal of DataPalette was to empower end-users to mix arbitrary structured data feeds on-the-fly while solving their particular information task(s) - be them sense-making or decision-making related - in order to more effectively accomplish these tasks.
+At a high level, the overall goal of DataPalette was to empower end-users to mix arbitrary structured data feeds on-the-fly while solving their particular information task(s) --- be them sense-making or decision-making related --- in order to more effectively accomplish these tasks.
 Unlike Mashup makers, including Yahoo Pipes!, the goal was to facilitate such integration without a need for an explicit separate mapping or integration step.
-This would detract from the overall usability and usefulness of the system by requiring that users perform tasks that did not immediately contribute to their resolving their need.
+Instead, DataPalette's approach is most directly inspired by ``lightweight'' visual, user-interactive data collection, consolidation and alignment tools such as Potluck \cite{citeulike:3875264} and Cards \cite{Dontcheva:2007:RCS:1294211.1294224}.
 
@@
 \subsubsection{Coreference consolidation: Drag-and-Drop Same-as}
+
+A major challenge with integration from multiple sources is coreference consolidation, or combining representations that co-refer to the same entity into one.
+For example, consolidating one's friends from Facebook, Google+, and LinkedIn, one might quickly want to de-duplicate and combine friends that use multiple services so that they each have one representation.
+Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, identifiers used by each network were system-specific and inconsistent.
+Thus, we sought to make instead make it user-driven, and easy to do manually.
 
@@
 \end{figure}
-
-A major challenge with integration from multiple sources is coreference consolidation, or combining representations that co-refer to the same entity into one.
-For example, consolidating one's friends from Facebook, Google+, and LinkedIn, one might quickly want to de-duplicate and combine friends that use multiple services so that they each have one representation.
-
-Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, identifiers used by each network were system-specific and inconsistent.
-Thus, we sought to make instead make it user-driven, and easy to do manually.
 
@@
 \hline
-Task		&Data Sources\\
-\hline
-Three universities they would apply to study 1)Sport Science and 2) History & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
-\hline
-Restaurant they would book for 12 friends in 3) Cambridge and 4) Glasgow& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
+Task	 & Data Sources\\
+\hline
+Three universities they would apply to study 1)~Sport Science, and 2)~History & www.thecompleteuniversityguide.co.uk, www.ucas.ac.uk, unistats.direct.gov.uk, www.timeshighereducation.co.uk/world-university-rankings\\
+\hline
+Restaurant they would book for 12 friends in 3)~Cambridge, and 4)~Glasgow & www.yelp.co.uk, www.tripadvisor.co.uk, ratings.food.gov.uk, facebook and google plus.\\
 \hline
@@
 \end{center}
-\caption{Tasks, choices and data sources} \label{tab:studyfactors}
+\caption{For each type of task we used two variables, which affected which data we used from the relevant data sources.} \label{tab:studyfactors}
 \normalsize

### 2012-09-19 chi2012.tex

@@
 
-\begin{table}[t]
+\begin{table}[htbp]
 \begin{center}
 \small
-\begin{tabular}{|p{3cm}|p{5cm}|}
+\begin{tabular}{|p{3cm}|p{4.7cm}|}
 \hline
@@
 \normalsize
-\end{table}%
-
-%\begin{table*}[t]
-%\begin{center}
-%\small
-%\begin{tabular}{|p{4cm}|p{6cm}|p{6cm}|}
-%\hline
-%Task	 &Choices	&Data Sources\\
-%\hline
-%1) Three universities they would apply to study Sport Science & Universities: 1) Loughborough,  2) Durham,  3) Exeter, 4) Edinburgh, 5) Birmingham, and 6) Bath & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
-%\hline
-%2) Three universities they would apply to study History & Universities: 1) Cambridge, 2) London School of Economics and Political Science, 3) Durham, 4) Oxford, 5) St Andrews, and 6) University College London & http://www.thecompleteuniversityguide.co.uk/, http://www.ucas.ac.uk/, http://unistats.direct.gov.uk/, http://www.timeshighereducation.co.uk/world-university-rankings/ \\
-%\hline
-%3) Restaurant they would book for 12 friends in Cambridge&Restaurants : 1) Alimentum, 2) Gardenia Restaurant, 3) The Cambridge Chop House, 4) Al Casbah, 5) Midsummer House Restaurant, and 6) Nando's Restaurant.
-& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
-%\hline
-%4) Restaurant they would book for 12 friends in Glasgow & Restaurants :1) Black Sheep Bistro, 2) Charcoals, 3) Number 16 Family Restaurant, 4) Fanny Trollope's Cafe, 5) Two Fat Ladies At The Buttery, and 6) Sapori D'Italia.
-& www.yelp.co.uk, www.tripadvisor.co.uk, http://ratings.food.gov.uk/, facebook and google plus.\\
-%\hline
-%\end{tabular}
-%\end{center}
-%\caption{Tasks, choices and data sources} \label{tab:studyfactors}
-%\normalsize
-%\end{table*}%
+\end{table}
 

### 2012-09-19 chi2012.tex
Added intel mashmaker reference.

--- 2012-09-19 chi2012.tex
@@
 is a visual programming language aimed at letting novice programmers easily create data-transformation workflows, but remained rather challenging to  use and effort-intensive for one-off information tasks.
-Mash-up makers such as Intel Mashmaker \cite{}, QDEWiki \cite{}, and TX2 \cite{}.
+Mash-up makers such as Intel Mashmaker \cite{intel_mashmaker}, QDEWiki \cite{}, and TX2 \cite{}.
 DataPalette's approach is most directly inspired by ``lightweight'' visual, user-interactive data collection, consolidation and alignment tools such as Potluck \cite{citeulike:3875264} and Cards \cite{Dontcheva:2007:RCS:1294211.1294224}.

### 2012-09-19 chi2012.tex
Fixed first paragraph of rw

--- 2012-09-19 chi2012.tex
@@
 In this section, we motivate our work using three fields: PIM, data integration, and end-user mashups and toolkits.
-Our primary motivation stems from field studies in PIM, where consolidating and working with data from multiple sources has remained a challenge for over a decade.
-Enabling PIM tools to support ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data heterogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
-As described by Alon Halvey:
+Our primary motivation stems from field studies in PIM, which foregrounded work to consolidate data from multiple sources and this still remains a challenge.
+Specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
+As described by Alon Halevy:
 
@@
 
-Recently, users are experiencing data fragmentation~\cite{Jones05towardsa} across multiple web-based applications and services, and struggle to consolidate across them~\cite{bergman,boardmansasse}.
+A result of this problem is that users can experience data fragmentation~\cite{Jones05towardsa} across multiple web-based applications and services, and struggle to consolidate across them~\cite{bergman,boardmansasse}.
 In our own previously-conducted field study, (reference anonymised) we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage both online and offline data.
-Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-maintained information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
+Similar findings from other studies observed volunteers' ``homebrew databases'': manually-maintained information collated to handle the heterogeneous data requirements of groups needing to coordinate their activities~\cite{Voida:2011:HDC:1978942.1979078}.\looseness=-1
 
-A number of research PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.
+A number of research PIM systems have demonstrated numerous ways that users cab benefit from better data integration support.
 For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.

### 2012-09-19 chi2012.tex
happy

--- 2012-09-19 chi2012.tex
@@
 is a visual programming language aimed at letting novice programmers easily create data-transformation workflows, but remained rather challenging to  use and effort-intensive for one-off information tasks.
-Mash-up makers such as Intel Mashmaker \cite{intel_mashmaker}, QDEWiki \cite{}, and TX2 \cite{}.
+Mash-up makers such as Intel Mashmaker \cite{}, QDEWiki \cite{}, and TX2 \cite{}.
 DataPalette's approach is most directly inspired by ``lightweight'' visual, user-interactive data collection, consolidation and alignment tools such as Potluck \cite{citeulike:3875264} and Cards \cite{Dontcheva:2007:RCS:1294211.1294224}.
@@
 Our study had several limitations.
-First, our study consisted of 20 participants which meant that we could not show that some of our data was statistically significant.
-Second, we gave the participants an artificially limited amount of time (10 minutes) to make their decisions.
-Given additional time, the participants choices may have changed.
-Third, the data sources were fixed, because we were not testing the users' capabilities to search and retrieve data from the web.
-Given unlimited access to data sources online, users would be able to compare different aspects, which may also alter their choices.
-
+First, the small sample size of our study (20 participants) meant that we could not confirm some of the effects observed with statistical significance.
+Second, giving participants a time limit of 10 minutes on trials may have affected the strategies they used to make their choices.
+For this reason, in our follow-up study, we will try to give some participants unlimited time on tasks of similar complexity.
+A second departure from a realistic task was the way in which we pre-determined the data sources for participants to use.
+We did this in order to focus participants' efforts on looking at and working with their choices, rather than retrieving it.
+However, as supporting effective retrieval and serendipitous discovery will be important for DP to be useful in practice, we plan to look at supporting these stages in the future.
 % \subsection{Integration Limitations}

### 2012-09-19 chi2012.tex

--- 2012-09-19 chi2012.tex
@@
 In this section, we motivate our work using three fields: PIM, data integration, and end-user mashups and toolkits.
-Our primary motivation stems from field studies in PIM, which foregrounded work to consolidate data from multiple sources and this still remains a challenge.
-Specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
-As described by Alon Halevy:
+Our primary motivation stems from field studies in PIM, where consolidating and working with data from multiple sources has remained a challenge for over a decade.
+Enabling PIM tools to support ad-hoc data integration has remained difficult because of the complex, theoretically difficult nature of data heterogeneity; specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
+As described by Alon Halvey:
 
@@
 
-A result of this problem is that users can experience data fragmentation~\cite{Jones05towardsa} across multiple web-based applications and services, and struggle to consolidate across them~\cite{bergman,boardmansasse}.
+Recently, users are experiencing data fragmentation~\cite{Jones05towardsa} across multiple web-based applications and services, and struggle to consolidate across them~\cite{bergman,boardmansasse}.
 In our own previously-conducted field study, (reference anonymised) we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage both online and offline data.
-Similar findings from other studies observed volunteers' ``homebrew databases'': manually-maintained information collated to handle the heterogeneous data requirements of groups needing to coordinate their activities~\cite{Voida:2011:HDC:1978942.1979078}.\looseness=-1
+Similar findings from other studies include Voida et al.'s observations of volunteer coordinators' ``homebrew databases'', manually-maintained information assemblages concocted to handle the widely varied and heterogeneous data collection requirements such groups needed to coordinate their activities \cite{Voida:2011:HDC:1978942.1979078}.
 
-A number of research PIM systems have demonstrated numerous ways that users cab benefit from better data integration support.
+A number of research PIM systems have demonstrated a great number of ways that users could benefit from better support for data integration.
 For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
@@
 
-%A different approach entirely has been taken in the web mash-up community; this community has largely focused on end-user interfaces that allow people to interactively reconcile data from multiple sources.
+% A different approach entirely has been taken in the web mash-up community; this community has largely focused on end-user interfaces that allow people to interactively reconcile data from multiple sources.
 Yahoo Pipes!
 is a visual programming language aimed at letting novice programmers easily create data-transformation workflows, but remained rather challenging to  use and effort-intensive for one-off information tasks.
-Mash-up makers such as Intel Mashmaker \cite{}, QDEWiki \cite{}, and TX2 \cite{}.
+Mash-up makers such as Intel Mashmaker \cite{intel_mashmaker}, QDEWiki \cite{}, and TX2 \cite{}.
 DataPalette's approach is most directly inspired by ``lightweight'' visual, user-interactive data collection, consolidation and alignment tools such as Potluck \cite{citeulike:3875264} and Cards \cite{Dontcheva:2007:RCS:1294211.1294224}.

### 2012-09-19 chi2012.tex
shortenend semantic web paragraph

--- 2012-09-19 chi2012.tex
@@
 
-A number of research PIM systems have demonstrated numerous ways that users cab benefit from better data integration support.
+A number of research PIM systems have demonstrated numerous ways that users can benefit from better data integration support.
 For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
@@
 The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} research communities have primarily focused efforts on purely-automatic approaches to the problem of data integration.
-Such approaches typically take the form of either \emph{ontology} or \emph{schema-level matching}, or \emph{direct} or \emph{instance-level} matching.
-We presently focus  on \emph{instance-level} matching, as schemas may not always be available, nor easily accessed by users.
-Approaches to automatic instance matching have included applied natural language processing (NLP) techniques for terminological matching (e.g., \cite{euzenat2004api}), machine-learning techniques that use both structural and terminological features for matching \cite{doan2003learning}, probabilistic approaches \cite{suchanek2011paris}, and instance schema comparisons using structural and logical definitions \cite{castano2006matching}.
-In real-world settings, however, manual and semi-automatic matching approaches are favoured over automatic because of the need for high degrees of accuracy.
+Such approaches typically take the form of either \emph{ontology/schema-level matching}~\cite{euzenat2004api,doan2003learning}, or \emph{direct/instance-level} matching~\cite{suchanek2011paris,castano2006matching}.
+We focus on \emph{instance-level} matching, as schemas may not always be available, nor easily accessible to users.
 

### 2012-09-19 chi2012.tex
design limits

@@
 
-\subsection{Integration Support}
-
-A key design decision we made at the outset was to target the most common, simpler forms of heterogeneity we observed in PS.2.
-Given the use of the interface mechanisms we observed, we wish to next broaden the scope of capabilities of DP so that it can become more useful in real settings.
-Among the integration capabilities we are considering are, first, field combination and splitting (by delimiters), in order to address the kinds of structural heterogeneity observed in several cases of PS.2, along with a framework for converting and reconciling units of measure, which would address several of the modelling issues.
-Extending the Do What I Mean capabilities of the histogram and bar-charting tools to be unit-aware would be a tremendous asset, we believe, for users who want to quickly compare things without paying close attention to the units used.
-A core problem to realising this, however, is that few data sources (besides the Weather category) explicitly annotated values with units used.
+A key design decision we made at the outset was to target only the most common forms of heterogeneity observed in PS.2.
+To expand DP's integration capabilities, we are considering a number of new interaction affordances for facilitating field combination and splitting, and unit of measure reconciliation.
+This would cover a majority of the structural and modeling-oriented heterogeneity cases observed in DP.2.
+Extending the Do What I Mean capabilities of the visualisation tools to be unit-aware would be a tremendous asset; a core problem, however, stems from the fact that few sources explicitly annotated values with units.
 

### 2012-09-19 chi2012.tex
Refined results first para

--- 2012-09-19 chi2012.tex
@@
 We selected the top restaurants for on Tripadvisor and Yelp, because they were the top two websites people quoted for comparing reviews when travelling.
-We selected the two university courses at random from a list taken form the complete university guide\footnote{The Complete University Guide - \url{www.thecompleteuniversityguide.co.uk}}, and selected the top six university for that subject because they provide a realistic shortlist and are comparable in terms of rankings and properties.
+We selected the two university courses at random from a list taken form the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}, and selected the top six university for that subject because they provide a realistic shortlist and are comparable in terms of rankings and properties.
 We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choice.
@@
 
-Twenty-six (26) participants responded, from which we selected the first 20, ten of which were female.
-All participants spoke English, but 12 of them were non-native English speakers.
-Seven were non-students, comprising University staff and alumni; out of the remaining 13, eight studied computer science and 5 studied other subjects.
+We received 26 responses to our call for participants, from which we selected the first 20 (ten of which were female).
+Seven were non-students, comprised of university staff and alumni; the remainder were students, 8 studying computer science, and 5 studying other subjects.
 On average, each study took 40 minutes to complete.
-The information in Table \ref{tbl:data} summarises the metrics that were captured of the trials for each participant.
-We analyse this data next.
+The information in Table \ref{tbl:data} summarises the captured metrics for each participant.
 

### 2012-09-19 chi2012.tex
Table caption slightly improved
shorten

--- 2012-09-19 chi2012.tex
@@
 The two pre-studies provided valuable insight towards this goal that allowed us to guide our design process.
-The first pre-study confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
-The second pre-study revealed that the most common inconsistencies that occurred in the APIs and data feeds available through from such data sources consisted predominantly of the simpler kinds of heterogeneity - terminological and simpler forms of structural heterogeneity, rather than the delicate semantic/modelling variety, which were comparatively rare.
+PS.1 confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
+PS.2, meanwhile, revealed that the most common kinds of heterogeneity across data feeds available today were the simpler kinds---terminological and simpler structural heterogeneity--- than the semantic/modelling variety.
 This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
@@
 This would cover a majority of the structural and modeling-oriented heterogeneity cases observed in DP.2.
-Extending the Do What I Mean capabilities of the visualisation tools to be unit-aware would be a tremendous asset; a core problem, however, stems from the fact that few sources explicitly annotated values with units.
+Extending the ``Do What I Mean'' capabilities of the visualisation features to be unit-of-measure-aware we think would be tremendously valuable; a core problem, however, stems from the fact that few sources explicitly annotated data values with such units.
 

### 2012-09-19 chi2012.tex

@@
 \end{center}
-\caption{Participants data.} \label{tbl:data}
+\caption{Data collected from each of our trials, for each participant.} \label{tbl:data}
 \normalsize

### 2012-09-19 chi2012.tex
shorter

@@
 %\section{Designing the DataPalette Interface}
-\section{Designing DataPalette : An Interface for Serendipitous Data Mixing}
+\section{DataPalette : An Interface for Data Mixing}
 
@@
 
-\subsection{Process}
+\subsection{Design Process}
 We followed a five-phase spiral model consisting of planning, designing, mocking-up, prototyping, and testing to derive the design of DataPalette.
@@
 
-\subsection{Interaction with DataPalette}
-In order to feel familiar to new users and focus our design efforts on integration-related aspects of the interface, we borrowed metaphors and interactions directly from MacOS Finder-style file managers.
-Specifically, in the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can arbitrarily drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
+In order to feel familiar to new users, we borrowed the overall  ``look and feel'' from MacOS Finder-style file managers.
+In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can arbitrarily drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
 (Although the design could be generalised to supporting nested groups, our prototype only supported a single level of grouping.)

### 2012-09-19 chi2012.tex
Tightening results, fixed broken cites

--- 2012-09-19 chi2012.tex
@@
 The two pre-studies provided valuable insight towards this goal that allowed us to guide our design process.
-PS.1 confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
-PS.2, meanwhile, revealed that the most common kinds of heterogeneity across data feeds available today were the simpler kinds---terminological and simpler structural heterogeneity--- than the semantic/modelling variety.
+The first pre-study confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
+The second pre-study revealed that the most common inconsistencies that occurred in the APIs and data feeds available through from such data sources consisted predominantly of the simpler kinds of heterogeneity - terminological and simpler forms of structural heterogeneity, rather than the delicate semantic/modelling variety, which were comparatively rare.
 This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
@@
 Our interface, condition A, took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).
-An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time (shown in Figure \ref{fig:time}).
+An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time.
 Although across trials, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor.
@@
 This would cover a majority of the structural and modeling-oriented heterogeneity cases observed in DP.2.
-Extending the ``Do What I Mean'' capabilities of the visualisation features to be unit-of-measure-aware we think would be tremendously valuable; a core problem, however, stems from the fact that few sources explicitly annotated data values with such units.
+Extending the Do What I Mean capabilities of the visualisation tools to be unit-aware would be a tremendous asset; a core problem, however, stems from the fact that few sources explicitly annotated values with units.
 
@@
 
-Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \cite{datajournalismhandbook}.
+Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \footnote{Data Journalism Hand Book: \url{datajournalismhandbook.org}}.
 In order for this to happen, however, tools that allow people to effectively combine data more effectively than current structured PIM tools and generic spreadsheet tools will become increasingly important.

### 2012-09-19 chi2012.tex
shorter

@@
 
-\section{Conclusion} %Key findings of the whole paper and its implications
+\section{Conclusions} %Key findings of the whole paper and its implications
 
@@
 
-Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \footnote{Data Journalism Hand Book: \url{datajournalismhandbook.org}}.
+% Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \cite{datajournalismhandbook}.
 In order for this to happen, however, tools that allow people to effectively combine data more effectively than current structured PIM tools and generic spreadsheet tools will become increasingly important.

### 2012-09-19 chi2012.tex

@@
 
+Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \footnote{Data Journalism Hand Book: \url{datajournalismhandbook.org}}.
+In order for this to happen, however, tools that allow people to effectively combine data more effectively than current structured PIM tools and generic spreadsheet tools will become increasingly important.
+
 \section{Acknowledgments}

### 2012-09-19 chi2012.tex
ps2 chop

@@
 \subsubsection{Results}
-Table \ref{tbl:prestudy2} shows the raw results of the process described.
-A number of features were quite striking about the data; First, while most of the data records were small and relatively simple, nearly one in each category produced substantially more complex records than the rest.
-In particular, Amazon, Soundcloud, Twitter and Weather Underground stood out in their respective categories.
-
-Looking at equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
+Table \ref{tbl:prestudy2} shows the results of the process described.
+A number of features were quite striking about the data; first, while most of the data records were small and relatively simple, nearly one in each category produced substantially larger records than the rest.
+Second, looking at equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
 While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.
@@
 
-The second finding is that equivalent properties were only rarely given the same name.
+The third finding is that equivalent properties were only rarely given the same name.
 Thus, terminological heterogeneity was prevalent across data records we observed.

### 2012-09-19 chi2012.tex
Filled in table!!!!
conclusions

@@
 
+Just as the wise Italian in Voltaire's poem \emph{La Begueule} predicted,``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have perhaps caused more scruffy, ``ad-hoc'' approaches to be overlooked.
+Based upon the results of our investigation, we believe that there are a number of advantages to user-driven data reconciliation over automatic to bespoke ones; specifically, this enables the user to combine data as part of his or her sensemaking process, using subjective judgement in the process
+
+%% the abililty to  In our investigation we found that complete approaches to data integration
+
 % - data integration literature demands `correct' but perfect is the enemy of the good.
 % ``perfect'' be the enemy of the good
-
 %% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
@@
 Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
-
 % Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \cite{datajournalismhandbook}.
@@
 
-\section{Acknowledgments}
+%% \section{Acknowledgments}
 

### 2012-09-19 chi2012.tex
conclusions

@@
 
-The 8 participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18-25 and 26-32.   
+The 8 participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
 % mention # that were computer scientists here?
@@
 In order to feel familiar to new users, we borrowed the overall  ``look and feel'' from MacOS Finder-style file managers.
-In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can arbitrarily drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
+In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can freely drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
 (Although the design could be generalised to supporting nested groups, our prototype only supported a single level of grouping.)
@@
 Unlike file managers, the basic unit of data is not a file, but an \emph{instance} - corresponding to an RDF instance or any small bit of structured data.
-In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data) into discrete instances with properties; for JSON or XML data feeds, we represented all subtrees with more than 1 child as its an instance.
+In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data) into discrete instances with properties (for JSON or XML data feeds, we represented all subtrees with more than 1 child as an instance).
 
@@
 chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, \newline 1.3, \newline 1.2& 4.2& 1.4, \newline 1.5, \newline 1.3& 3.2& 2.1, \newline 2.3, \newline 2.6& 3.5, \newline 3.3&2.6, \newline 2.3, \newline 2.5&1.6, \newline 1.5, \newline 1.3&4.1&3.3&2.4, \newline 2.5, \newline 2.6\\
-\# factors in choice &-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-\\
+\# factors in choice &6&5&4&4&2&4&8&8&6&3&4&8&5&4&7&7&4&5&5&5\\
 \hline\hline
@@
  chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4, \newline 1.2, \newline 1.6&2.1&2.3, \newline 2.6, \newline 2.5&3.5&2.2, \newline 2.3, \newline 2.6&3.5&1.1&4.1&1.6&4.2&2.6, \newline 2.3, \newline 2.5&2.3, \newline 2.5, \newline 2.6&1.1, \newline 1.3,  \newline 1.4&3.5\\
- \# factors in choice &-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-&-\\
+ \# factors in choice &4&3&2&3&1&6&5&7&7&5&4&7&6&3&6&4&3&6&4&5\\
  \hline

### 2012-09-19 chi2012.tex
conclusions

@@
 Just as the wise Italian in Voltaire's poem \emph{La Begueule} predicted,``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have perhaps caused more scruffy, ``ad-hoc'' approaches to be overlooked.
-Based upon the results of our investigation, we believe that there are a number of advantages to user-driven data reconciliation over automatic to bespoke ones; specifically, this enables the user to combine data as part of his or her sensemaking process, using subjective judgement in the process
+Based upon the results of our investigation, we see a number of benefits to user-driven data reconciliation approaches over automatic and bespoke ones.
+Specifically, letting the user freely merge and split data as part of his or her sensemaking process, gives users the power to see data combined or split up as best suited to the task
+
+ things as simple as coreference involved any subjective judgement,
 

### 2012-09-19 chi2012.tex
PS1 summary more to the point

--- 2012-09-19 chi2012.tex
@@
 \subsubsection{Summary of Findings}
-% TODO: how does this summary support the thesis of the paper?
-People said that they would like a website that could support the evolution of an event, that allowed them to post drafts of an event.
+
+All users experienced using multiple web sites in order to complete a task, and felt that the lack of existing integration between sites hindered their ability to make informed choices.
+Participants felt that they did not include particular aspects of a decision because the process to investigate them would be too difficult or take too much time.
+
+%People said that they would like a website that could support the evolution of an event, that allowed them to post drafts of an event.
 They felt that organising an event should be a process that changes over time.
@@
 We consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} list of most popular data feeds for this purpose.
-To increase the variety of sources, we selected from 5 categories: social network services, retailers, online event calendars, music sites, and weather, selecting 2-5 sources from each, for a total of 20.
+To increase the variety of sources, we selected from 5 categories: social network services, retailers, online event calendars, music sites, and weather, selecting 2--5 sources from each, for a total of 20.
 
-For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for retail sites this was product info, etc.
+For each feed, 3--5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for retail sites this was product info, etc.
 %Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.  
@@
 \subsubsection{Results}
-Table \ref{tbl:prestudy2} shows the results of the process described.
-A number of features were quite striking about the data; first, while most of the data records were small and relatively simple, nearly one in each category produced substantially larger records than the rest.
-Second, looking at equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
+Table \ref{tbl:prestudy2} shows the raw results of the process described.
+A number of features were quite striking about the data; First, while most of the data records were small and relatively simple, nearly one in each category produced substantially more complex records than the rest.
+In particular, Amazon, Soundcloud, Twitter and Weather Underground stood out in their respective categories.
+
+Looking at equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
 While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.
@@
 
-The third finding is that equivalent properties were only rarely given the same name.
+The second finding is that equivalent properties were only rarely given the same name.
 Thus, terminological heterogeneity was prevalent across data records we observed.
@@
 
-Just as the wise Italian in Voltaire's poem \emph{La Begueule} predicted,``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have perhaps caused more scruffy, ``ad-hoc'' approaches to be overlooked.
-Based upon the results of our investigation, we believe that there are a number of advantages to user-driven data reconciliation over automatic to bespoke ones; specifically, this enables the user to combine data as part of his or her sensemaking process, using subjective judgement in the process
-
-%% the abililty to  In our investigation we found that complete approaches to data integration
-
 % - data integration literature demands `correct' but perfect is the enemy of the good.
 % ``perfect'' be the enemy of the good
+
 %% In this paper, we take the position that there is a need for a difference in focus between the approaches taken to data integration between the approaches taken in database integration and that of PIM, because of vast differences in requirements.
@@
 Moreover, while a fully, uniform query capability might be nice, such capability is likely of less overall importance for a single quick task than simply being able to quickly and easily identify and cope with differences.
+
 % Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \cite{datajournalismhandbook}.
@@
 
-%% \section{Acknowledgments}
+\section{Acknowledgments}
 

### 2012-09-19 chi2012.tex
Fix Multi-path selection: link-sliding for heterogeneous sets

@@
 This achieves the ability to \emph{link slide} as introduced by Parallax~\cite{parallax}, but across sets of heterogeneous items by supporting multiple parallel dereferenced paths simultaneously.
+Multiple selection group-dereferencing results in two key benefits.
+First, that users can quickly compare all instances that have the same structure, and second, that instances with different property names and be quickly consolidated.
 
-Multiple selection group-dereferencing results in two effects:
-
-\begin{enumerate}
-\item Clicking on a property dereferences all corresponding values makes it easy to quickly compare all instances that have the same structure.
+%
+%\begin{enumerate}
+%\item Clicking on a property dereferences all corresponding values makes it easy to quickly compare all instances that have the same structure.
 Although we were concerned that users might be surprised that path selection applied to all objects within each group in parallel, users adapted very quickly and used this functionality to save them time.
-
-\item Supporting dereferencing along different paths \emph{by structure} allows instances with different property names, and even sub-structures to be quickly consolidated by separate dereferencing, thereby effectively coping with terminological and structural heterogeneity.
-\end{enumerate}
-
+%
+%\item Supporting dereferencing along different paths \emph{by structure} allows instances with different property names, and even sub-structures to be quickly consolidated by separate dereferencing, thereby effectively coping with terminological and structural heterogeneity.
+%\end{enumerate}
+%
 \subsubsection{Coreference consolidation: Drag-and-Drop Same-as}

### 2012-09-19 chi2012.tex
Task preparation cut down redundancy

@@
 We selected six because it enable us to scope our study, we found through two pilot studies that six data sources provided a sufficient challenge for users to explore the data and required them to use DP's main features.
-We manually collected the data from the nine selected websites into JSON files, and converted them to Microsoft Excel worksheets (which 40\% of the participants in PS.1 said that had previously used to compare data).
+We manually collected the key data points from the websites for use in DP and into Excel spreadsheets (which 40\% of the participants in PS.1 said that had previously used to compare data).
 
 % why was the B condition set up this way? What did we seek to achieve?
-The control (B) conditions were in a prepared Mac comprised of a web browser pre-loaded with a page with links to four live web sites, and an open instance of Excel with a workbook open containing spreadsheet versions of the same data as were loaded into DP for the A condition versions of the same tasks.
+In the control condition the participants were given the source web site URLs, as well as a spreadsheet containing the data taken from those web sites (the same data loaded into DP for condition A).
+Participants were free to use the web sites and spreadsheets however they wished.
+%The control (B) conditions were in a prepared Mac comprised of a web browser pre-loaded with a page with links to four live web sites, and an open instance of Excel with a workbook open containing spreadsheet versions of the same data as were loaded into DP for the A condition versions of the same tasks.
 Other Office applications were also available for use and participants were allowed to visit other sites than the pre-specified ones.

### 2012-09-19 chi2012.tex
foo

--- 2012-09-19 chi2012.tex
@@
 \subsubsection{Summary of Findings}
-
-All users experienced using multiple web sites in order to complete a task, and felt that the lack of existing integration between sites hindered their ability to make informed choices.
-Participants felt that they did not include particular aspects of a decision because the process to investigate them would be too difficult or take too much time.
-
-%People said that they would like a website that could support the evolution of an event, that allowed them to post drafts of an event.
+% TODO: how does this summary support the thesis of the paper?
+People said that they would like a website that could support the evolution of an event, that allowed them to post drafts of an event.
 They felt that organising an event should be a process that changes over time.
@@
 We consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} list of most popular data feeds for this purpose.
-To increase the variety of sources, we selected from 5 categories: social network services, retailers, online event calendars, music sites, and weather, selecting 2--5 sources from each, for a total of 20.
+To increase the variety of sources, we selected from 5 categories: social network services, retailers, online event calendars, music sites, and weather, selecting 2-5 sources from each, for a total of 20.
 
-For each feed, 3--5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for retail sites this was product info, etc.
+For each feed, 3-5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for retail sites this was product info, etc.
 %Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.  
@@
 
-Just as the wise Italian in Voltaire's poem \emph{La Begueule} predicted,``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have perhaps caused more scruffy, ``ad-hoc'' approaches to be overlooked.
-Based upon the results of our investigation, we believe that there are a number of advantages to user-driven data reconciliation over automatic to bespoke ones; specifically, this enables the user to combine data as part of his or her sensemaking process, using subjective judgement in the process
+Just as the wise Italian in Voltaire's poem \emph{La Begueule} said that ``the best is the enemy of the good'', we believe that complete approaches to data integration have perhaps caused less principled, ``ad-hoc'' approaches to be overlooked.
+Based upon the results of our investigation, we see a number of benefits to lightweight, user-driven approaches to coping with heterogeneity over automatic and bespoke ones.
+Just as the \emph{user-subjective approach} to PIM \cite{user-subjective} showed that subjective labels and attributes were more useful than physical ones for files and folders, a user-controlled approach to data integration, where data may be combined (and separated) at will may benefit the sensemaking or decision-making process.
+Towards this end, our results of our pre-studies and DataPalette show that, for a majority of the kinds of structured data feeds available on the Web today, the kinds of heterogeneity can be largely managed using simple interactive approaches that are intuitive and preceived as useful by end-users.
+
+%%  things as simple as coreference involved any subjective judgement,
 
@@
 
-Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \footnote{Data Journalism Hand Book: \url{datajournalismhandbook.org}}.
+%% Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \footnote{Data Journalism Hand Book: \url{datajournalismhandbook.org}}.
 In order for this to happen, however, tools that allow people to effectively combine data more effectively than current structured PIM tools and generic spreadsheet tools will become increasingly important.

### 2012-09-19 chi2012.tex
boo

@@
 
-Just as the wise Italian in Voltaire's poem \emph{La Begueule} said that ``the best is the enemy of the good'', we believe that complete approaches to data integration have perhaps caused less principled, ``ad-hoc'' approaches to be overlooked.
+Just as the Italian in Voltaire's poem \emph{La Begueule} said ``the best is the enemy of the good'', we believe that complete approaches to data integration have perhaps caused less principled, ``ad-hoc'' approaches to be overlooked.
 Based upon the results of our investigation, we see a number of benefits to lightweight, user-driven approaches to coping with heterogeneity over automatic and bespoke ones.

### 2012-09-19 chi2012.tex
conclusion

@@
 
-Just as the Italian in Voltaire's poem \emph{La Begueule} said ``the best is the enemy of the good'', we believe that complete approaches to data integration have perhaps caused less principled, ``ad-hoc'' approaches to be overlooked.
+As said by the Italian in Voltaire's poem \emph{La Begueule}, ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have, perhaps, caused less-principled, ``ad-hoc'' approaches to be overlooked.
+Based upon the results of our investigation, however, we see a number of benefits to lightweight, user-driven, interactive approaches over automatic and bespoke ones.
+Just as the \emph{user-subjective approach} to PIM \cite{user-subjective} showed that subjective attributes and labels were more useful than physical ones, we have shown that a user-driven approach to data integration, as achieved with DataPalette, affords users flexibility to combine and keep separate data at will, and may facilitate decision-making tasks.
+Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through simple, fast interactive approaches that are received as intuitive by end-users.
+
+%% Just as the Italian in Voltaire's poem \emph{La Begueule} said ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have perhaps caused less principled, ``ad-hoc'' approaches to be overlooked.
 Based upon the results of our investigation, we see a number of benefits to lightweight, user-driven approaches to coping with heterogeneity over automatic and bespoke ones.

### 2012-09-19 chi2012.tex
condition b cut down

--- 2012-09-19 chi2012.tex
@@
  paper used &n&n&n&n&n&n&y&n&y&y&n&y&n&y&y&n&n&y&n&n\\
- spreadsheet/website&&&&&&&&&&&&&&&&&&&&\\
+% spreadsheet/website&&&&&&&&&&&&&&&&&&&&\\
  number of spreadsheets used&5/5&4/4&2/5&1.5&2/5&5/5&4/4&2/4&3/4&&2/4&4/5&2/4&4/5&2/4&2/5&4/4&4/4&0&5/5\\
@@
 
-\subsection{Condition B - Results and Observations}
-In the B conditions, participants were given the option of using websites and/or a spreadsheet to make their choice.
+\subsection{Condition B: Results and Observations}
+
+Participants were given the option of using web sites and/or a spreadsheet to make their choices, using the same data as condition A.
+There was mixed use of websites and the spreadsheet, and we measured the fraction of the trial time spent in Excel versus looking at the sites in a 5-value range (0,25,50,75,100\%).
+Six individuals used Excel entirely (100\%) without consulting the web sites, while one avoided Excel entirely.
+The uses of Excel for external cognition ranged from collation behaviour to tallying annotations, to colour-coding cells to allow easy identification of  ``pros'' and ``cons''.
+The complete data collected is given in Table \ref{tbl:data}.
+In general, users struggled with both the websites and with the spreadsheet, although could generally find the information they required after a short period of frustration.
+
+%In the B conditions, participants were given the option of using websites and/or a spreadsheet to make their choice.
 For B trials, we measured the fraction of the trial time spent in Excel versus looking at the sites in a 5-value range (0,25,50,75,100\%).
@@
 
-Participants ran into a number of difficulties with the spreadsheet interface.
+%Participants ran into a number of difficulties with the spreadsheet interface.
 Participants struggled with looking up values across different worksheets (data sources).
@@
 
-While most participants only made minor formatting and column sizing changes to the worksheets to facilitate reading and interpretation,  four made extensive edits to the sheets to help them ``think through'' the task in various ways.
+%While most participants only made minor formatting and column sizing changes to the worksheets to facilitate reading and interpretation,  four made extensive edits to the sheets to help them ``think through'' the task in various ways.
 These uses of Excel for external cognition ranged from the collation behaviour just described to tallying annotations, to colour-coding cells to allow easy identification of  ``pros'' and ``cons''; ``I can't deal with so many textual values; I need colours to help me see'' (p6).
@@
 %% todo - look up to make sure it was really p7 here 
-Various difficulties were also encountered with use of the websites.
+%Various difficulties were also encountered with use of the websites.
 There was significant confusion over various attributes, especially in the University selection task.
@@
 
-Just as the Italian in Voltaire's poem \emph{La Begueule} said ``the best is the enemy of the good'', we believe that complete approaches to data integration have perhaps caused less principled, ``ad-hoc'' approaches to be overlooked.
-Based upon the results of our investigation, we see a number of benefits to lightweight, user-driven approaches to coping with heterogeneity over automatic and bespoke ones.
-Just as the \emph{user-subjective approach} to PIM \cite{user-subjective} showed that subjective labels and attributes were more useful than physical ones for files and folders, a user-controlled approach to data integration, where data may be combined (and separated) at will may benefit the sensemaking or decision-making process.
-Towards this end, our results of our pre-studies and DataPalette show that, for a majority of the kinds of structured data feeds available on the Web today, the kinds of heterogeneity can be largely managed using simple interactive approaches that are intuitive and preceived as useful by end-users.
-
-%%  things as simple as coreference involved any subjective judgement,
+Just as the wise Italian in Voltaire's poem \emph{La Begueule} predicted,``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have perhaps caused more scruffy, ``ad-hoc'' approaches to be overlooked.
+Based upon the results of our investigation, we believe that there are a number of advantages to user-driven data reconciliation over automatic to bespoke ones; specifically, this enables the user to combine data as part of his or her sensemaking process, using subjective judgement in the process
 
@@
 
-%% Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \footnote{Data Journalism Hand Book: \url{datajournalismhandbook.org}}.
+Just as journalism became peer-produced with the explosion of blogging, data journalism is on the brink of moving from skilled data analysts the likes of The Guardian Datablog staff\footnote{Guardian Datablog - \url{http://www.guardian.co.uk/news/datablog}} into the hands of the ordinary citizen \footnote{Data Journalism Hand Book: \url{datajournalismhandbook.org}}.
 In order for this to happen, however, tools that allow people to effectively combine data more effectively than current structured PIM tools and generic spreadsheet tools will become increasingly important.

### 2012-09-19 chi2012.tex

@@
 
-Just as the wise Italian in Voltaire's poem \emph{La Begueule} predicted,``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have perhaps caused more scruffy, ``ad-hoc'' approaches to be overlooked.
-Based upon the results of our investigation, we believe that there are a number of advantages to user-driven data reconciliation over automatic to bespoke ones; specifically, this enables the user to combine data as part of his or her sensemaking process, using subjective judgement in the process
+As said by the Italian in Voltaire's poem \emph{La Begueule}, ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have, perhaps, caused less-principled, ``ad-hoc'' approaches to be overlooked.
+Based upon the results of our investigation, however, we see a number of benefits to lightweight, user-driven, interactive approaches over automatic and bespoke ones.
+Just as the \emph{user-subjective approach} to PIM \cite{user-subjective} showed that subjective attributes and labels were more useful than physical ones, we have shown that a user-driven approach to data integration, as achieved with DataPalette, affords users flexibility to combine and keep separate data at will, and may facilitate decision-making tasks.
+Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through simple, fast interactive approaches that are received as intuitive by end-users.
+
+%% Just as the Italian in Voltaire's poem \emph{La Begueule} said ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have perhaps caused less principled, ``ad-hoc'' approaches to be overlooked.
+Based upon the results of our investigation, we see a number of benefits to lightweight, user-driven approaches to coping with heterogeneity over automatic and bespoke ones.
+Just as the \emph{user-subjective approach} to PIM \cite{user-subjective} showed that subjective labels and attributes were more useful than physical ones for files and folders, a user-controlled approach to data integration, where data may be combined (and separated) at will may benefit the sensemaking or decision-making process.
+Towards this end, our results of our pre-studies and DataPalette show that, for a majority of the kinds of structured data feeds available on the Web today, the kinds of heterogeneity can be largely managed using simple interactive approaches that are intuitive and preceived as useful by end-users.
+
+%%  things as simple as coreference involved any subjective judgement,
 

### 2012-09-19 chi2012.tex
voltaire in the house tonight

@@
 
-As said by the Italian in Voltaire's poem \emph{La Begueule}, ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have, perhaps, caused less-principled, ``ad-hoc'' approaches to be overlooked.
+As said in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have, perhaps, caused less-principled, ``ad-hoc'' approaches to be overlooked.
 Based upon the results of our investigation, however, we see a number of benefits to lightweight, user-driven, interactive approaches over automatic and bespoke ones.

### 2012-09-19 chi2012.tex
experimental setup and task prep combined

@@
 \subsubsection{Task preparation}
-% how data was prepared for each of the tasks; exactly how the A condition was done (full screen DP in a web browser, no outside access to apps),  B condition and why
-In order to prepare the data for our tasks, we identified websites that people commonly used for selecting universities and restaurants.
+
+In order to prepare the data for our tasks, we identified websites that people commonly used for selecting universities and restaurants (see Table \ref{tab:studyfactors}).
 Once we had compiled the list of websites, we select the top most used six for each task.
-We selected six because it enable us to scope our study, we found through two pilot studies that six data sources provided a sufficient challenge for users to explore the data and required them to use DP's main features.
+We selected six because it enables us to scope our study, we found through two pilot studies that six data sources provided a sufficient challenge for users to explore the data and required them to use DP's main features.
 We manually collected the key data points from the websites for use in DP and into Excel spreadsheets (which 40\% of the participants in PS.1 said that had previously used to compare data).
 
-% why was the B condition set up this way? What did we seek to achieve?
-In the control condition the participants were given the source web site URLs, as well as a spreadsheet containing the data taken from those web sites (the same data loaded into DP for condition A).
+The participants were given a standard desktop PC running MacOS X on a 22-inch monitor with a standard keyboard and mouse.
+Each participant was asked to complete two tasks, one A and one B condition.
+Each participant was given a maximum of 10 minutes to complete a task, but were told that if they finished earlier they could tell us and move on to the next condition.
+In order to select which tasks the participant was given, we used a Latin square to permutate the tasks and conditions.
+In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
+
+In condition B (control) the participants were given the source web site URLs, as well as a spreadsheet containing the data taken from those web sites (the same data loaded into DP for condition A).
 Participants were free to use the web sites and spreadsheets however they wished.
-%The control (B) conditions were in a prepared Mac comprised of a web browser pre-loaded with a page with links to four live web sites, and an open instance of Excel with a workbook open containing spreadsheet versions of the same data as were loaded into DP for the A condition versions of the same tasks.
-Other Office applications were also available for use and participants were allowed to visit other sites than the pre-specified ones.
-The purpose of this condition was to understand how people using standard desktop office software.
+
+\begin{table}[htbp]
+\begin{center}
+\small
+\begin{tabular}{|p{3cm}|p{4.7cm}|}
+\hline
+Task	 & Data Sources\\
+\hline
+Three universities they would apply to study 1)~Sport Science, and 2)~History & www.thecompleteuniversityguide.co.uk, www.ucas.ac.uk, unistats.direct.gov.uk, www.timeshighereducation.co.uk/world-university-rankings\\
+\hline
+Restaurant they would book for 12 friends in 3)~Cambridge, and 4)~Glasgow & www.yelp.co.uk, www.tripadvisor.co.uk, ratings.food.gov.uk, facebook and google plus.\\
+\hline
+\end{tabular}
+\end{center}
+\caption{For each type of task we used two variables, which affected which data we used from the relevant data sources.} \label{tab:studyfactors}
+\normalsize
+\end{table}
 
@@
 Participants were offered a small gift certificate from an online retailer for their time.
-
-\subsection{Experimental Setup}
-The participants were given a standard desktop PC running MacOS X on a 22-inch monitor with a standard keyboard and mouse.
-During the studies the Excel workbook and DP were initially set to full screen.
-Each participant was asked to complete two tasks, one A and one B condition.
-Each participant was given a maximum of 10 minutes to complete a task, but were told that if they finished earlier they could tell us and move on to the next condition.
-In order to select which tasks the participant was given, we used a Latin square to permutate the tasks and conditions.
-In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
-
-%Each task had an A and B condition.
-The A condition, the participant could only use DataPalette to make their decision; and in the B condition the participant was given the choice to using any websites they desired and or workbook in Microsoft Excel containing the data taken from the data sources to make their choice on individual spreadsheets.
-
-%Each study consisted of two tasks, one A and one B condition.
-Each participant was given 10 minutes to complete a task.
-In order to select which tasks the participant was given, we used a Latin square to permutate the tasks and conditions.
-In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
-
-\begin{table}[htbp]
-\begin{center}
-\small
-\begin{tabular}{|p{3cm}|p{4.7cm}|}
-\hline
-Task	 & Data Sources\\
-\hline
-Three universities they would apply to study 1)~Sport Science, and 2)~History & www.thecompleteuniversityguide.co.uk, www.ucas.ac.uk, unistats.direct.gov.uk, www.timeshighereducation.co.uk/world-university-rankings\\
-\hline
-Restaurant they would book for 12 friends in 3)~Cambridge, and 4)~Glasgow & www.yelp.co.uk, www.tripadvisor.co.uk, ratings.food.gov.uk, facebook and google plus.\\
-\hline
-\end{tabular}
-\end{center}
-\caption{For each type of task we used two variables, which affected which data we used from the relevant data sources.} \label{tab:studyfactors}
-\normalsize
-\end{table}
 

### 2012-09-19 chi2012.tex
Finish task preparation fixed

@@
 
-% why? why were these tasks selected? can we drop in the phrase ``multi-dimensional subjective choice task'' and refer back to PS.1 to justify the choice of tasks?
 We based our two tasks on choosing a university and picking a restaurant for a large social event, because in PS.1 the majority of people stated that they had used multiple websites for these tasks.
 We used different two different variables for both tasks, where the restaurant was located (Glasgow and Cambridge) and which subject people chose (sport science and history).
-The cities Glasgow and Cambridge were chosen because they are not geographically close to the participants, they are two heavily populated areas that have a choice of restaurants, and they contained detail data about the same restaurants on various review sites.
-When deciding on which cities to use, we found that different areas favoured different review websites and places such as Hull had sets of disjoint sets of restaurant reviews.
-We selected the top restaurants for on Tripadvisor and Yelp, because they were the top two websites people quoted for comparing reviews when travelling.
-We selected the two university courses at random from a list taken form the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}, and selected the top six university for that subject because they provide a realistic shortlist and are comparable in terms of rankings and properties.
-We selected a shortlist of six because it is typical for users to narrow down their options to a limited selection before analysing the details of their final choice.
+For the restaurant task, we chose the top restaurants in Glasgow and Cambridge from Tripadvisor and Yelp as the candidates.
+We chose these cities because they are not geographically close to the participants, to minimise bias).
+For the university condition we selected two university courses at random, and selected the top six universities for each subject.
+We selected a shortlist of six because it is typical for users to narrow down their options because analysing the final details of their choice.
 

### 2012-09-19 chi2012.tex
shortening

@@
 
+% why? why were these tasks selected? can we drop in the phrase ``multi-dimensional subjective choice task'' and refer back to PS.1 to justify the choice of tasks?
 We based our two tasks on choosing a university and picking a restaurant for a large social event, because in PS.1 the majority of people stated that they had used multiple websites for these tasks.
 We used different two different variables for both tasks, where the restaurant was located (Glasgow and Cambridge) and which subject people chose (sport science and history).
-For the restaurant task, we chose the top restaurants in Glasgow and Cambridge from Tripadvisor and Yelp as the candidates.
-We chose these cities because they are not geographically close to the participants, to minimise bias).
-For the university condition we selected two university courses at random, and selected the top six universities for each subject.
-We selected a shortlist of six because it is typical for users to narrow down their options because analysing the final details of their choice.
+For the restaurant task, the cities Glasgow and Cambridge were chosen because they were geographically distant to participants, and would therefore minimise likelihood of prior familiarity with the restaurants or location.
+For the University selection task, two university courses were selected at random from the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}, and the six, top-ranked universities for each subject were selected as candidates.
+In both tasks, we kept the number of candidate choices at six because we felt it was typical for users to narrow down their options to a limited selection before analysing the details of their final choice.
 

### 2012-09-19 chi2012.tex
mini mine moe

@@
 
-A number of research PIM systems have demonstrated numerous ways that users can benefit from better data integration support.
-For example, SEMEX's \cite{semex} rich uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
-Similarly, Haystack, a personal information platform, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
-Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to a pre-specified set of ontologies, or have the user provide a manual specify such a mapping translation themselves.
+A number of research PIM systems have demonstrated how users benefit from unified, integrated data models.
+SEMEX's \cite{semex} uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
+Similarly, Haystack, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
+% Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to pre-specified ontologies, or have the user specify such a mapping manually.
 
@@
 
-\section{Pre-studies: The Need for and Challenges in realising Ad-Hoc Data Integration in PIM Tools}
+\section{Pre-studies: Identifying Needs and Challenges}
 
@@
 
-In summary, while data records ranged in complexity, no data source subsumed the rest; all contained properties that other sources did not have.
-Moreover, the most common types of heterogeneity exhibited were terminological and value-related, although simpler kinds of structural heterogeneity (pertaining to representing substructures of values) were also common.
+In summary, while data records ranged in complexity, no source subsumed the rest; all contained unique properties other sources did not have.
+Moreover, the most common types of heterogeneity exhibited were terminological and value-related, which occurred vastly more than structural or modeling issues.
 

### 2012-09-19 chi2012.tex
mini mine moe
genrality

@@
 In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can freely drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
-(Although the design could be generalised to supporting nested groups, our prototype only supported a single level of grouping.)
-
-Unlike file managers, the basic unit of data is not a file, but an \emph{instance} - corresponding to an RDF instance or any small bit of structured data.
-In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data) into discrete instances with properties (for JSON or XML data feeds, we represented all subtrees with more than 1 child as an instance).
+Unlike file managers, however, the basic unit of data was not a file, but an \emph{instance} - corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
+In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data into discrete instances with properties.
+This was done at time of import, and could be done for every major Web data feed type (e.g., JSON, XML, RDF) without any loss of generality.
 

### 2012-09-19 chi2012.tex
boo

@@
 We asked the participants to follow a \emph{think-aloud} protocol as they worked, so that we could understand the reasoning behind their actions.
-At the end of the study, the participant completed an exit survey.
-
-We planned to analyse the qualitative and quantitative data using thematic and statistic analysis, respectively.
+At the end of the study, the participant completed a short exit survey.
+
+After the studies were complete, we generated transcripts of comments made by participants during the think-aloud protocols.
+At the same time, we populated a spreadsheet summarising quantitative metrics pertaining to how much various features of the tool were used (see Table \ref{tbl:data} for specific fields).
+After the transcriptions were complete, we categorised people's comments made during the study, suggests in their exit survey, and clustered them to common themes.
+
+%% We planned to analyse the qualitative and quantitative data using thematic and statistic analysis, respectively.
 In order to carry out our data analysis, we used the recorded transcripts to populate a spreadsheet summarising the quantitative metrics (see Table \ref{tbl:data} for specific fields).
@@
 We recruited participants through adverts posted near the University campus and e-mails to student mailing lists.
-We screened participants to be at least 18 years of age, but did not filter or preferentially select participants based on any other criteria.
+We screened participants to be at least 18 years of age, but did not filter on any other criteria.
 We took the first 20 that signed up and that successfully specified time slots that were available.
-As described below in results, we ended up getting a large number of international students, who had varying proficiency in English.
-Since all such participants had passed the requisite proficiency tests to enrol at University, we did not think this would substantially impact results.
 Participants were offered a small gift certificate from an online retailer for their time.
@@
 Seven were non-students, comprised of university staff and alumni; the remainder were students, 8 studying computer science, and 5 studying other subjects.
+Due to the large population of international students and staff at the University, we ended up getting a large number of non-native English speakers.
+However, since all students and staff had passed requisite proficiency tests, we did not think this would substantially impact results.
 On average, each study took 40 minutes to complete.

### 2012-09-19 chi2012.tex
boo

@@
 \subsubsection{Study and task design}
-To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'' which we describe below).
+To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'') as a control condition.
 We used upon the within-subjects A-B format in order to make both quantitative measurements about an individual's performance with DP (relative to baseline), but also to make qualitative observations of the processes used by each participant as they solved each task.
+Conditions and interface order were fully counterbalanced among participants to eliminate ordering and task bias.
+
+In the control conditions, participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from web sites pertaining to the candidates they were asked to choose from.
+Participants were told they were free to use the web sites and spreadsheets however they wished.
 Each task was paired with one of two datasets, depending on condition, so that we could evaluate whether the participant's performance of completing a task differed with different datasets.
-Conditions and interface order were fully counterbalanced among participants to eliminate ordering and task bias.
 
@@
 For the restaurant task, the cities Glasgow and Cambridge were chosen because they were geographically distant to participants, and would therefore minimise likelihood of prior familiarity with the restaurants or location.
-For the University selection task, two university courses were selected at random from the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}, and the six, top-ranked universities for each subject were selected as candidates.
-In both tasks, we kept the number of candidate choices at six because we felt it was typical for users to narrow down their options to a limited selection before analysing the details of their final choice.
-
-\subsubsection{Task preparation}
-
-In order to prepare the data for our tasks, we identified websites that people commonly used for selecting universities and restaurants (see Table \ref{tab:studyfactors}).
+For the University selection task, two university courses were selected at random from the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}, and the six top-ranked universities for each subject were selected as candidates.
+% \subsubsection{Task preparation} In order to prepare the data for our tasks, we identified websites that people commonly used for selecting universities and restaurants (see Table \ref{tab:studyfactors}).
 Once we had compiled the list of websites, we select the top most used six for each task.
@@
 In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
-
-In condition B (control) the participants were given the source web site URLs, as well as a spreadsheet containing the data taken from those web sites (the same data loaded into DP for condition A).
-Participants were free to use the web sites and spreadsheets however they wished.
 
@@
 
-As said in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have, perhaps, caused less-principled, ``ad-hoc'' approaches to be overlooked.
+As said in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have, perhaps, caused ``ad-hoc'' approaches to be overlooked.
 Based upon the results of our investigation, however, we see a number of benefits to lightweight, user-driven, interactive approaches over automatic and bespoke ones.

### 2012-09-19 chi2012.tex
shortened methodology

@@
 \subsubsection{Study and task design}
-To test these hypotheses, we designed a within-subjects lab-based A-B study in which participants were asked to perform two predefined tasks in separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'') as a control condition.
-We used upon the within-subjects A-B format in order to make both quantitative measurements about an individual's performance with DP (relative to baseline), but also to make qualitative observations of the processes used by each participant as they solved each task.
-Conditions and interface order were fully counterbalanced among participants to eliminate ordering and task bias.
+To test these hypotheses, we designed a within-subjects (repeated measures) lab-based A-B study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'').
+Each task was paired with one of two datasets, chosen depending on condition, which were fully counterbalanced in a Latin square design to eliminate potential ordering, task, and dataset biases.
+
+The participants were seated at a standard Macintosh desktop equipped with a 22-inch monitor with a standard keyboard and mouse.
+Each participant was given a maximum of 10 minutes to each of two tasks, but were told that if they finished earlier they could tell us and move on to the next condition.
+Prior to A conditions, participants were trained on how to use DP using a five-minute instructional video.
 
@@
 Participants were told they were free to use the web sites and spreadsheets however they wished.
-Each task was paired with one of two datasets, depending on condition, so that we could evaluate whether the participant's performance of completing a task differed with different datasets.
 
@@
 We manually collected the key data points from the websites for use in DP and into Excel spreadsheets (which 40\% of the participants in PS.1 said that had previously used to compare data).
-
-The participants were given a standard desktop PC running MacOS X on a 22-inch monitor with a standard keyboard and mouse.
-Each participant was asked to complete two tasks, one A and one B condition.
-Each participant was given a maximum of 10 minutes to complete a task, but were told that if they finished earlier they could tell us and move on to the next condition.
-In order to select which tasks the participant was given, we used a Latin square to permutate the tasks and conditions.
-In order to train the user on how to use our interface, they viewed a five-minute DataPalette tutorial video introducing its basic functionality, how to use histograms and maps, and linking instances.
 

### 2012-09-19 chi2012.tex
Cut down orphan

--- 2012-09-19 chi2012.tex
@@
 Assessing whether DP improved task performance (h3) was difficult given the small sample size of our study and large number of factors influencing each person's performance.
-While tasks were completed on average slightly faster than the control interface, these differences were not statistically significant.
+Tasks were completed on average slightly faster than the control interface, but not statistically significantly.
 However, more data sources seemed to be consulted (we use ``seemed'' carefully because these findings approached significance) during trials with DP over the control interface, which meant that people were looking at more diverse information than the control conditions.
@@
 Participants entertained more possibilities for longer in the DP trial, while participants were more prone to eliminate candidates early in the baseline interface.
-Finally, people took fewer notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP, which strongly indicates that DP made the task easier to complete.
+Finally, people took fewer notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP, which strongly indicates that DP made the task easier to complete.\looseness=-1
 

### 2012-09-19 chi2012.tex
Chunking

--- 2012-09-19 chi2012.tex
@@
 Unlike file managers, however, the basic unit of data was not a file, but an \emph{instance} - corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
-In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data into discrete instances with properties.
-This was done at time of import, and could be done for every major Web data feed type (e.g., JSON, XML, RDF) without any loss of generality.
+In order to make working with relational graphs manageable, we decided to break up RDF-like relational data into discrete instances with properties.
+Such chunking was implemented in a generic manner to work with all major Web data feed types (e.g., JSON, XML, RDF).
 

### 2012-09-19 chi2012.tex
more orphans killed

--- 2012-09-19 chi2012.tex
@@
 
-By making such consolidation user-driven, we gave the user freedom to combine things when it suited him or her.
-As we saw in the system evaluation, while in certain circumstances users preferred combining representations, in others people preferred to keep things separate.
-Thus, user-initiated consolidation afforded an additional flexibility that was ultimately deemed beneficial.
+By making consolidation user-driven, we gave users freedom to combine things when it suited them.
+As we saw in the system evaluation, while in certain circumstances users preferred combining representations, in others people preferred not to.
+Thus, user-initiated consolidation afforded an additional flexibility that was ultimately deemed beneficial.\looseness=-1
 
@@
 \subsubsection{Model-based brushing}
-Since it can get confusing having a single instance represented across several groups, we provided \emph{universal brushing} across the interface; hovering over any representation of any instance (an instance block, a map marker or a histogram bar), causes all other visible representations of that instance to lightly glow, so that the user can quickly identify all the places that a single instance is represented
+Since it can get confusing having a single instance represented across several groups, we provided \emph{universal brushing} across the interface; hovering over any representation of any instance (an instance block, a map marker or a histogram bar), causes all other visible representations of that instance to lightly glow, so that the user can quickly identify all the places that a single instance is represented.
 
@@
 \subsubsection{Study and task design}
-To test these hypotheses, we designed a within-subjects (repeated measures) lab-based A-B study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'').
-Each task was paired with one of two datasets, chosen depending on condition, which were fully counterbalanced in a Latin square design to eliminate potential ordering, task, and dataset biases.
+To test these hypotheses, we designed a within-subjects (repeated measures) lab-based A-B study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (condition A) and once with a baseline interface (condition B).
+Each task was paired with one of two datasets, depending on condition, which were fully counterbalanced in a Latin square design to eliminate potential ordering, task, and dataset biases.\looseness=-1
 
@@
 Participants spent an average of 8.9 minutes performing the tasks.
-Our interface, condition A, took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).
+Condition A (DP) took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).
 An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time.
@@
 
-\subsubsection{Candidate selection}
-
 \begin{figure}[htb]
@@
 
-Since the tasks were subjective-choice, we could not evaluate the `correctness' of people's answers.
+\subsubsection{Candidate selection}
+
+Since the tasks were subjective-choice, we could not evaluate the `correctness' of answers.
 However, to determine whether there was a difference in variability of answers between interface conditions, we plotted their choice on a histogram for each trial, per task and dataset for both conditions, visible in Figure \ref{fig:chosen_results}.
@@
 
-To understand whether the interface influenced choice of strategy, we measured the number of candidates that the participant considered \emph{at final result}.
-This we found was a strong marker for use of each strategy; people using the successive elimination strategy arrived with only 1-2 candidates at decision time, while those that tallied still maintained the entire starting set.
+To understand whether the interface influenced choice of strategy, we measured the number of candidates the participant considered \emph{at final result}.
+This is a strong marker for use of each strategy; people using the successive elimination strategy arrived with only 1--2 candidates at decision time, while those that tallied still maintained the entire starting set.\looseness=-1
 
@@
 Sixteen (80\%) used \emph{Same-As} as least once; among these participants, they used \emph{Same-As} an average of 4.6 times per trial.
-We counted a single ``use'' as a single drag-and-drop of an individual item, or a drag-and-drop operation of an entire group onto another.
-(We did not differentiate enumerated value consolidation from entity consolidation, as these operations were not discernibly different from the interface perspective.)
+We counted a single ``use'' as a single drag-and-drop of an individual item, or a drag-and-drop operation of an entire group onto another (we did not differentiate enumerated value consolidation from entity consolidation, as these are not discernibly different from the interface perspective).\looseness=-1
 
@@
 \subsection{Condition B: Results and Observations}
-
 Participants were given the option of using web sites and/or a spreadsheet to make their choices, using the same data as condition A.

### 2012-09-19 chi2012.tex
killing more orphans

--- 2012-09-19 chi2012.tex
@@
 
-The second finding is that equivalent properties were only rarely given the same name.
-Thus, terminological heterogeneity was prevalent across data records we observed.
-Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modelled as a subtree in the other.
-We also encountered only 2 examples where data was nested in textual description fields, both from records in the Weather category, which hints their intended use in RSS feed readers rather than in structured database applications.
-Examples of modelling/semantic heterogeneity instances were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and mismatches in units of measurement (commonly found in the Weather forecast measurements).
+The second finding is that equivalent properties were rarely given the same name.
+Thus, terminological heterogeneity was prevalent across the data records we observed.
+Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modelled as a subtree in another.
+We also encountered only 2 examples where data was nested in textual description fields, both from records in the Weather category, which hints at their intended use in RSS feed readers rather than in structured database applications.
+Examples of semantic heterogeneity were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and mismatches in units of measurement (commonly found in the Weather forecast measurements).
 
-Although not included in the table above, we also examined the source schemas of many of the data sources for property enumerated type values.
-There was substantial terminological inconsistency among such values, but also cases where some sources had values not represented in the other source's sets.
-For example, the property ``gender''  on Facebook allowed only supported ``male'' and ``female'' for values, while Google+ included a third option,``other''.
+Although not included in the table above, we also examined the source schemas of many of the data sources for enumerated type values.
+There was substantial terminological inconsistency among such values, but also cases where some sources had values not represented by other sources.
+For example,  ``gender''  on Facebook allowed only supported ``male'' and ``female'' for values, while Google+ included a third option,``other''.
 ``Relationship status'' was another which was similarly complex, as none of the string representations of enumerated value sets matched across data sources perfectly.
 
-In summary, while data records ranged in complexity, no source subsumed the rest; all contained unique properties other sources did not have.
+In summary, while data records ranged in complexity, no source subsumed the rest; all contained unique properties.
 Moreover, the most common types of heterogeneity exhibited were terminological and value-related, which occurred vastly more than structural or modeling issues.
@@
 In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can freely drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
-Unlike file managers, however, the basic unit of data was not a file, but an \emph{instance} - corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
+Unlike file managers, the basic unit of data was not a file, but an \emph{instance} - corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
 In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data into discrete instances with properties.

### 2012-09-19 chi2012.tex
shorter

--- 2012-09-19 chi2012.tex
@@
 
-The third metric measured the amount of overlap among the sources in each category.
-This was done by determining \emph{equivalent properties} between schemas, which we did manually by creating an alignment table for all properties across all data sources in each category.
-Each row of the alignment table comprised a property of a record from one data source, and all its closest matching properties from the others.
-We used both property names and attribute descriptions (in documentation, if available) to make matching decisions; two properties were aligned if they seemed to model the same piece of information.
+The third metric measured degree of overlap among sources in each category.
+This was done by first mapping \emph{equivalent properties} between schemas, which we did by creating alignment tables for all properties of data sources in each category.
+Each row comprised a property of a record of one data source, and all its closest matching properties from the others.
+We used both property names and attribute descriptions (in documentation, if available) to make matching decisions.
 
 Once established, equivalent properties were examined firstly for naming inconsistencies.
-Then, ignoring property name, we looked at whether the values were structurally comparable; if two fields were of the same data type and, if complex, had structurally equivalent subfields, they were considered comparable, otherwise they were considered \emph{structurally inconsistent}.
-The ``birthday'' fields in Facebook and LinkedIn demonstrate structural inconstancy; Facebook represents birthdays as a single string formatted ``MM/DD/YYYY'', while the LinkedIn representation is a sub-structure with separate ``Month'', ``Day'' and ``Year'' subfields.
-
+Then, ignoring property name, we examined whether the property's values were structurally compatible; if both values consisted of the data type and, if complex, had structurally equivalent subfields, they were considered comparable, otherwise marked \emph{structurally inconsistent}.
 Finally, we examined equivalent fields which were incomparable for reasons other than structural or terminological differences.
@@
 Thus, terminological heterogeneity was prevalent across data records we observed.
-Structural heterogeneity, however, was comparatively rare, with the most common case being similar to the birthday example above; a leaf property on one record modelled as a subtree in the other.
-We also encountered only 2 examples where data was nested in textual description fields, both from records in the Weather category, which hints their intended use in RSS feed readers rather than in structured database applications.
-Examples of modelling/semantic heterogeneity instances were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and mismatches in units of measurement (commonly found in the Weather forecast measurements).
+Structural heterogeneity, was slightly rarer.
+The most common case consisted of a leaf property from one data source modelled as a subtree in the other.
+We also encountered two examples where data was nested in textual description fields, which hints of their intended use in feed readers rather than in structured database applications.
+Examples of modelling/semantic heterogeneity instances were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and differences in units of measurement (as commonly found in the Weather data sources).
 
-Although not included in the table above, we also examined the source schemas of many of the data sources for property enumerated type values.
-There was substantial terminological inconsistency among such values, but also cases where some sources had values not represented in the other source's sets.
+We also examined the source schemas of data sources for enumerated value types.
+There was substantial terminological inconsistency among such value ranges of such types, but also cases where sources had values not represented in the other source's ranges.
 For example, the property ``gender''  on Facebook allowed only supported ``male'' and ``female'' for values, while Google+ included a third option,``other''.
-``Relationship status'' was another which was similarly complex, as none of the string representations of enumerated value sets matched across data sources perfectly.
+Few string representations of enumerated value sets matched across data sources perfectly.
 
@@
 The two pre-studies provided valuable insight towards this goal that allowed us to guide our design process.
-PS.1 confirmed the need for data integration by showing that people did, in fact, prefer to rely upon multiple sources of information for making important decisions, and identified the kinds of sources people most commonly used.
-PS.2, meanwhile, revealed that the most common kinds of heterogeneity across data feeds available today were the simpler kinds---terminological and simpler structural heterogeneity--- than the semantic/modelling variety.
-This gave us optimism that an interactional approach optimised around the these simpler kinds of heterogeneity would be effective.
-In the following sections we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.
+PS.1 confirmed the need for data integration by showing that people preferred to rely upon multiple sources of information for making important decisions, and identified the kinds of sources most commonly used.
+PS.2, meanwhile, revealed that the most common kinds of heterogeneity across data feeds were of the terminological structural varieties.
+As these were the ``simpler'' kinds of heterogeneity, this gave us optimism that an interactional approach would be feasible.
+In the following sections, we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.
 
@@
 We followed a five-phase spiral model consisting of planning, designing, mocking-up, prototyping, and testing to derive the design of DataPalette.
-This process was used because of the high degree of initial uncertainty associated with the mechanisms for facilitating the various aspects of effective data mixing that we describe in the next section.
-We approached this design challenge through alternative generation and elimination; alternatives were designed, drawn up and evaluated by expert colleagues (and in later iterations, non-expert potential users), who provided critiques of the alternatives.
+This process was used because of the high degree of initial uncertainty associated with how we'd achieve effective data mixing by end-users.
+We approached this design challenge through alternative generation and elimination; alternatives were designed, drawn up and evaluated by colleagues (and in later iterations, non-expert potential users), who critiqued the alternatives.
 The ease with which designs could be prototyped in HTML5 with modern web frameworks allowed us to perform this develop-test-iterate rapidly through seven iterations of this process.
@@
 
-By making consolidation user-driven, we gave users freedom to combine things when it suited them.
-As we saw in the system evaluation, while in certain circumstances users preferred combining representations, in others people preferred not to.
-Thus, user-initiated consolidation afforded an additional flexibility that was ultimately deemed beneficial.\looseness=-1
+By making such consolidation user-driven, we gave the user freedom to combine things when it suited him or her.
+As we saw in the system evaluation, while in certain circumstances users preferred combining representations, in others people preferred to keep things separate.
+Thus, user-initiated consolidation afforded an additional flexibility that was ultimately deemed beneficial.
 
@@
 \subsubsection{Model-based brushing}
-Since it can get confusing having a single instance represented across several groups, we provided \emph{universal brushing} across the interface; hovering over any representation of any instance (an instance block, a map marker or a histogram bar), causes all other visible representations of that instance to lightly glow, so that the user can quickly identify all the places that a single instance is represented.
+Since it can get confusing having a single instance represented across several groups, we provided \emph{universal brushing} across the interface; hovering over any representation of any instance (an instance block, a map marker or a histogram bar), causes all other visible representations of that instance to lightly glow, so that the user can quickly identify all the places that a single instance is represented
 
@@
 \subsubsection{Study and task design}
-To test these hypotheses, we designed a within-subjects (repeated measures) lab-based A-B study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (condition A) and once with a baseline interface (condition B).
-Each task was paired with one of two datasets, depending on condition, which were fully counterbalanced in a Latin square design to eliminate potential ordering, task, and dataset biases.\looseness=-1
+To test these hypotheses, we designed a within-subjects (repeated measures) lab-based A-B study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (``condition A'') and once with a baseline interface (``condition B'').
+Each task was paired with one of two datasets, chosen depending on condition, which were fully counterbalanced in a Latin square design to eliminate potential ordering, task, and dataset biases.
 
@@
 Participants spent an average of 8.9 minutes performing the tasks.
-Condition A (DP) took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).
+Our interface, condition A, took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).
 An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time.
@@
 
+\subsubsection{Candidate selection}
+
 \begin{figure}[htb]
@@
 
-\subsubsection{Candidate selection}
-
-Since the tasks were subjective-choice, we could not evaluate the `correctness' of answers.
+Since the tasks were subjective-choice, we could not evaluate the `correctness' of people's answers.
 However, to determine whether there was a difference in variability of answers between interface conditions, we plotted their choice on a histogram for each trial, per task and dataset for both conditions, visible in Figure \ref{fig:chosen_results}.
@@
 
-To understand whether the interface influenced choice of strategy, we measured the number of candidates the participant considered \emph{at final result}.
-This is a strong marker for use of each strategy; people using the successive elimination strategy arrived with only 1--2 candidates at decision time, while those that tallied still maintained the entire starting set.\looseness=-1
+To understand whether the interface influenced choice of strategy, we measured the number of candidates that the participant considered \emph{at final result}.
+This we found was a strong marker for use of each strategy; people using the successive elimination strategy arrived with only 1-2 candidates at decision time, while those that tallied still maintained the entire starting set.
 
@@
 Sixteen (80\%) used \emph{Same-As} as least once; among these participants, they used \emph{Same-As} an average of 4.6 times per trial.
-We counted a single ``use'' as a single drag-and-drop of an individual item, or a drag-and-drop operation of an entire group onto another (we did not differentiate enumerated value consolidation from entity consolidation, as these are not discernibly different from the interface perspective).\looseness=-1
+We counted a single ``use'' as a single drag-and-drop of an individual item, or a drag-and-drop operation of an entire group onto another.
+(We did not differentiate enumerated value consolidation from entity consolidation, as these operations were not discernibly different from the interface perspective.)
 
@@
 \subsection{Condition B: Results and Observations}
+
 Participants were given the option of using web sites and/or a spreadsheet to make their choices, using the same data as condition A.
@@
 Assessing whether DP improved task performance (h3) was difficult given the small sample size of our study and large number of factors influencing each person's performance.
-Tasks were completed on average slightly faster than the control interface, but not statistically significantly.
+While tasks were completed on average slightly faster than the control interface, these differences were not statistically significant.
 However, more data sources seemed to be consulted (we use ``seemed'' carefully because these findings approached significance) during trials with DP over the control interface, which meant that people were looking at more diverse information than the control conditions.
@@
 Participants entertained more possibilities for longer in the DP trial, while participants were more prone to eliminate candidates early in the baseline interface.
-Finally, people took fewer notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP, which strongly indicates that DP made the task easier to complete.\looseness=-1
+Finally, people took fewer notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP, which strongly indicates that DP made the task easier to complete.
 

### 2012-09-19 chi2012.tex

@@
 
-In summary, while data records ranged in complexity, no source subsumed the rest; all contained unique properties other sources did not have.
+In summary, while data records ranged in complexity, no source subsumed the rest; all contained unique properties.
 Moreover, the most common types of heterogeneity exhibited were terminological and value-related, which occurred vastly more than structural or modeling issues.
@@
 In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can freely drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
-Unlike file managers, however, the basic unit of data was not a file, but an \emph{instance} - corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
-In order to make working with relational graphs manageable, we decided to break up RDF-like relational data into discrete instances with properties.
-Such chunking was implemented in a generic manner to work with all major Web data feed types (e.g., JSON, XML, RDF).
+Unlike file managers, the basic unit of data was not a file, but an \emph{instance} - corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
+In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data into discrete instances with properties.
+This was done at time of import, and could be done for every major Web data feed type (e.g., JSON, XML, RDF) without any loss of generality.
 

### 2012-09-19 chi2012.tex
orphan massacre

--- 2012-09-19 chi2012.tex
@@
 In particular, Amazon, Soundcloud, Twitter and Weather Underground stood out in their respective categories.
-
 Looking at equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
 While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.
-Only Spotify, out of the 4 music providers, included information about the album(s) featuring a given song.
-Such omissions show us that these data APIs reflect each provider's own specific capabilities and needs, and suggests that integrating information from multiple sources may allow more complete and useful representations to be assembled.
+Only Spotify, out of the 4 music providers, included information about the album featuring a given song.
+Such omissions show us that these data APIs reflect each provider's own specific needs, and suggests that integrating information from multiple sources may allow more complete and useful representations to be assembled.
 

### 2012-09-19 chi2012.tex
fix table alignment
almost there

@@
 The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} research communities have primarily focused efforts on purely-automatic approaches to the problem of data integration.
-Such approaches typically take the form of either \emph{ontology/schema-level matching}~\cite{euzenat2004api,doan2003learning}, or \emph{direct/instance-level} matching~\cite{suchanek2011paris,castano2006matching}.
-We focus on \emph{instance-level} matching, as schemas may not always be available, nor easily accessible to users.
+Such approaches typically take the form of \emph{ontology/schema-level matching}~\cite{euzenat2004api,doan2003learning}, or \emph{direct/instance-level} matching~\cite{suchanek2011paris,castano2006matching}.
+We focus on \emph{instance-level} matching, as schemas may not always be available, or easily accessible to users.\looseness=-1
 
@@
 For our first pre-study, we sought an updated understanding on the use of multiple information sources in information gathering on the Web.
-There were two driving questions behind this inquiry; first, were people increasingly relying only on a handful of ``supersites'' (Facebook, Wikipedia), or searching across distributed sources of information?
-If the former were true, merely integrating PIM tools with the small number of supersites would provide greater benefit than tackling the substantially more challenging problem of integrating data from arbitrary sources on the Web.
+There were two driving questions behind this inquiry: first, were people increasingly relying only on a handful of ``supersites'' (Facebook, Wikipedia), or searching across distributed sources of information?
+If the former were true, merely integrating PIM tools with the small number of supersites would provide greater benefit than tackling the more challenging problem of integrating data from arbitrary sources on the Web.
 Second, why did people choose a single or multiple sources for information gathering?
@@
 %\subsubsection{Question 1: Heterogenous Data Tasks}
-All participants reported regular use of multiple websites to accomplish tasks, examples of which including: shopping (including food, consumer electronics, hotels and flights), choosing restaurants, job-seeking, selecting a university, seeking recipes, and finding answers to technical problems.
+All participants reported regular use of multiple websites to accomplish tasks, examples of which including: shopping (food, consumer electronics, hotels and flights), choosing restaurants, job-seeking, selecting a university, seeking recipes, and finding answers to technical problems.
 When beginning a task, it was common for participants to use Google to discover several related websites, which would then be used in different ways depending on the task at hand.
@@
 Also, friends' preferences in price and food choices were not seen as a priority; they would try to be inclusive but only if it meant small changes to the plan.
-Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
+Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.\looseness=-1
 
@@
 
-\begin{table*}
+\begin{table*}[htb]
 \small
@@
 Finally, we examined equivalent fields which were incomparable for reasons other than structural or terminological differences.
-These \emph{modelling differences} arose from a variety of causes, including, but not limited to: measurement unit inconsistencies, measurement method (e.g., dew point vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').\looseness=-1
+These \emph{modelling differences} arose from a variety of causes, including: unit inconsistencies, measurement method (e.g., dew point vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').\looseness=-1
 
@@
 There was substantial terminological inconsistency among such value ranges of such types, but also cases where sources had values not represented in the other source's ranges.
-For example, the property ``gender''  on Facebook allowed only supported ``male'' and ``female'' for values, while Google+ included a third option,``other''.
+For example, ``gender''  on Facebook allowed only ``male'' and ``female'' for values, while Google+ included a third option,``other''.
 Few string representations of enumerated value sets matched across data sources perfectly.

### 2012-09-19 chi2012.tex
aha
10 pages

--- 2012-09-19 chi2012.tex
@@
 When organising an event, the weather was not that important because they did not trust the accuracy of predictions.
-Also, friends' preferences in price and food choices were not seen as a priority; they would try to be inclusive but only if it meant small changes to the plan.
+% Also, friends' preferences in price and food choices were not seen as a priority; they would try to be inclusive but only if it meant small changes to the plan.
 Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
@@
 The third metric measured degree of overlap among sources in each category.
-This was done by first mapping \emph{equivalent properties} between schemas, which we did by creating alignment tables for all properties of data sources in each category.
+This was done by first mapping \emph{equivalent properties} between schemas, which was done manually by creating alignment tables for all properties of data sources in each category.
 Each row comprised a property of a record of one data source, and all its closest matching properties from the others.
@@
 Once established, equivalent properties were examined firstly for naming inconsistencies.
-Then, ignoring property name, we examined whether the property's values were structurally compatible; if both values consisted of the data type and, if complex, had structurally equivalent subfields, they were considered comparable, otherwise marked \emph{structurally inconsistent}.
+Then, ignoring property names, we examined whether the property's values were structurally compatible.
+If both values consisted of the data type and same equivalent subfields, they were considered comparable; otherwise they were considered \emph{structurally inconsistent}.
 Finally, we examined equivalent fields which were incomparable for reasons other than structural or terminological differences.
@@
 The second finding is that equivalent properties were only rarely given the same name.
-Thus, terminological heterogeneity was prevalent across data records we observed.
-Structural heterogeneity, was slightly rarer.
-The most common case consisted of a leaf property from one data source modelled as a subtree in the other.
-We also encountered two examples where data was nested in textual description fields, which hints of their intended use in feed readers rather than in structured database applications.
-Examples of modelling/semantic heterogeneity instances were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and differences in units of measurement (as commonly found in the Weather data sources).
+Thus, terminological heterogeneity was prevalent across properties and records we observed.
+Structural heterogeneity, was slightly rarer; the most common case consisted of a leaf property from one data source modelled as a subtree in the other.
+We also encountered two examples where data was nested in textual description fields, which hints of their intended use in feed readers rather than in structured data applications.
+Examples of modelling differences were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and differences in units of measurement.
 
@@
 When this is done, DataPalette attempts to identify matching pairs among items in the dragged and target groups by comparing each pairs' labels.
-Only exact matches (modulo white space/underscores) are considered matches and automatically combined.
-The rest are added to the target group as distinct elements, and may be explicitly combined by hand.
-We are considering more sophisticated / less fragile approaches to matching in the future, which would make the system assistive.
+Only exact matches (modulo white space/underscores) were considered matches and automatically combined.
+The rest are added to the target group as distinct elements, and may be explicitly combined by hand.\looseness=-1
 
+%We are considering more sophisticated / less fragile approaches to matching in the future, which would make the system assistive.
 By making consolidation user-driven, we gave users freedom to combine things when it suited them.
 As we saw in the system evaluation, while in certain circumstances users preferred combining representations, in others people preferred not to.
-Thus, user-initiated consolidation afforded an additional flexibility that was ultimately deemed beneficial.\looseness=-1
+Thus, user-initiated consolidation afforded an additional flexibility that was ultimately deemed beneficial.
 
@@
 Participants entertained more possibilities for longer in the DP trial, while participants were more prone to eliminate candidates early in the baseline interface.
-Finally, people took fewer notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP, which strongly indicates that DP made the task easier to complete.\looseness=-1
+Finally, people took fewer paper notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP.\looseness=-1
 

### 2012-09-19 chi2012.tex
10 pages

@@
 %\subsubsection{Question 2: Process Used to Organise a Large Social Event}
+
 Participants said they would organise a large social event by first conferring with friends about their ideas, preferences and recommendations.
@@
 They considered how to get there and looked up train or bus times on the Web, or organised carpools.
+%When organising an event, the weather was not that important because they did not trust the accuracy of predictions.
+Also, friends' preferences in price and food choices were not seen as a priority; they would try to be inclusive but only if it meant small changes to the plan.
+Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
+
+% old:
+% Participants said they would organise a large social event by first conferring with friends about their ideas, preferences and recommendations.
+They selected venues, locations, restaurants and activities through a number of sources: their own prior experience, friends' recommendations, searching Google maps, dedicated reviews sites (such as Yelp for restaurants), and through web search.
+People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.
+They considered how to get there and looked up train or bus times on the Web, or organised carpools.
 When organising an event, the weather was not that important because they did not trust the accuracy of predictions.
-% Also, friends' preferences in price and food choices were not seen as a priority; they would try to be inclusive but only if it meant small changes to the plan.
-Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.
+Also, friends' preferences in price and food choices were not seen as a priority; they would try to be inclusive but only if it meant small changes to the plan.
+Choosing food and timings were often made on-the-fly, by either with the use of mobile applications or by walking past a location.\looseness=-1
 
@@
 Second, giving participants a time limit of 10 minutes on trials may have affected the strategies they used to make their choices.
-For this reason, in our follow-up study, we will try to give some participants unlimited time on tasks of similar complexity.
+For this reason, in our follow-up study, we will give some participants unlimited time on tasks of similar complexity.
 A second departure from a realistic task was the way in which we pre-determined the data sources for participants to use.

### 2012-09-19 chi2012.tex
results header
post hoc

@@
 To test these hypotheses, we designed a within-subjects (repeated measures) lab-based A-B study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (condition A) and once with a baseline interface (condition B).
-Each task was paired with one of two datasets, depending on condition, which were fully counterbalanced in a Latin square design to eliminate potential ordering, task, and dataset biases.\looseness=-1
+Each task was paired with one of two datasets, depending on condition, which were counterbalanced in a full-factorial design to eliminate potential ordering, task, and dataset biases.\looseness=-1
 
@@
 Participants spent an average of 8.9 minutes performing the tasks.
-Condition A (DP) took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).
-An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time.
-Although across trials, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor.
+Condition A (DP) took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$); however, a 2-way repeated measures ANOVA test of time taken by interface condition, blocked by participant ID, revealed that interface condition did not have a significant effect on task completion time.
+This analysis also showed no significant effect of participant ID on task completion.
+Comparing tasks, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), but this difference was not significant.
 
@@
 The second metric was the number of factors the participant considered in making their final choice.
-To determine this, we asked participants to explain precisely why they made their choice, and matched their justifications to properties in the data.
-This analysis revealed the interface used made a significant impact on the number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
+To determine this, we asked participants to explain why they made their choice(s), and recorded the number of distinct data dimensions mentioned.
+An ANOVA revealed a significant effect of interface on the number of such factors mentioned; post-hoc analysis using Tukey-Schiffe adjustments indicated that participants in the DP condition mentioned significantly more dimensions in their final decisions ($M=5.5; SD=1.7$) than those in the control condition ($M=4; SD=1.73$), $F(1,19)=3.95; p<0.10$.
+A strong significant effect was also observed between participant ID and number of dimensions mentioned, meaning people differed significantly on the number of reasons they gave ($F(19,19)=4.53 p < 0.001$).
 

### 2012-09-19 chi2012.tex
post hoc

@@
 To determine this, we asked participants to explain why they made their choice(s), and recorded the number of distinct data dimensions mentioned.
-An ANOVA revealed a significant effect of interface on the number of such factors mentioned; post-hoc analysis using Tukey-Schiffe adjustments indicated that participants in the DP condition mentioned significantly more dimensions in their final decisions ($M=5.5; SD=1.7$) than those in the control condition ($M=4; SD=1.73$), $F(1,19)=3.95; p<0.10$.
-A strong significant effect was also observed between participant ID and number of dimensions mentioned, meaning people differed significantly on the number of reasons they gave ($F(19,19)=4.53 p < 0.001$).
+A 2-way ANOVA test of the effect of interface and participant ID on the number of factors mentioned revealed a significant effect; post-hoc analysis with Tukey-Schiffe adjustments indicated that participants in the DP condition mentioned significantly more factors in their final decisions ($M=5.5; SD=1.7$) than those in the control condition ($M=4; SD=1.73$), $F(1,19)=3.95; p<0.10$.
+A strong significant effect was also observed between participant ID and number of dimensions mentioned, meaning some participants mentioned significantly more factors than others ($F(19,19)=4.53; p < 0.001$).
 
 \subsubsection{Diversity: Data sources consulted}
-The third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
-In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04$ $p\approx0.10$).
+The third metric we used to gauge the task performance is the number of different data sources consulted during the trial; our rationale for this metric was that more informed decisions derived from more thorough consideration of available data.
+In this metric, participants used an average of 88\% ($SD=0.13$) of data sources provided in the DP case, compared to 80\% ($SD=0.18$)in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance ($F(1,19)=0.04$ $p\approx0.10$).
 
 \subsubsection{Effort: External cognition}
-Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much participants had to keep track of during the study.
-We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22; p\approx0.10$).
-
-\begin{figure}[htb]
+Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much information participants had to keep track of during the study.
+We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in a 2-way ANOVA blocked by participant ID ($F(1,19)= 8.22; p\approx0.10$).
+
+\begin{figure}[tbp]
 \begin{center}
@@
 \subsubsection{Candidate selection}
-
 Since the tasks were subjective-choice, we could not evaluate the `correctness' of answers.

### 2012-09-19 chi2012.tex
minor

--- 2012-09-19 chi2012.tex
@@
 
-We hypothesise that end-users will benefit from tools that allow them effectively browse, use and consume heterogeneous structured data from diverse sources.
+We hypothesise that end-users will benefit from tools that allow them to effectively browse, use and consume heterogeneous structured data from diverse sources.
 In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data.
-We used a three-stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using a plurality of information sources, and the processes that they relied upon to perform such tasks.
+We used a three-stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using multiple information sources, and the processes that they relied upon to perform such tasks.
 Our results suggested that people rely on multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and for alternative perspectives.
@@
 In this section, we motivate our work using three fields: PIM, data integration, and end-user mashups and toolkits.
-Our primary motivation stems from field studies in PIM, which foregrounded work to consolidate data from multiple sources and this still remains a challenge.
+Our primary motivation stems from field studies in PIM, which foregrounded work to consolidate data from multiple sources and which still remains a challenge.
 Specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
@@
 A number of research PIM systems have demonstrated how users benefit from unified, integrated data models.
-SEMEX's \cite{semex} uniform query interface allowed users to quickly trace references of people, places and things across files, folders and repositories.
-Similarly, Haystack, allowed users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
+SEMEX's~\cite{semex} uniform query interface allows users to quickly trace references of people, places and things across files, folders and repositories.
+Similarly, Haystack  allows users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
 % Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to pre-specified ontologies, or have the user specify such a mapping manually.
@@
 
+\newpage
 \section{Results}

### 2012-09-19 chi2012.tex
goo

--- 2012-09-19 chi2012.tex
@@
 To determine this, we asked participants to explain why they made their choice(s), and recorded the number of distinct data dimensions mentioned.
-A 2-way ANOVA test of the effect of interface and participant ID on the number of factors mentioned revealed a significant effect; post-hoc analysis with Tukey-Schiffe adjustments indicated that participants in the DP condition mentioned significantly more factors in their final decisions ($M=5.5; SD=1.7$) than those in the control condition ($M=4; SD=1.73$), $F(1,19)=3.95; p<0.10$.
+A 2-way ANOVA test of the effect of interface and participant ID on the number of factors mentioned revealed a significant effect; post-hoc analysis with Tukey-Scheff\'{e} adjustments indicated that participants in the DP condition mentioned significantly more factors in their final decisions ($M=5.5; SD=1.7$) than those in the control condition ($M=4; SD=1.73$), $F(1,19)=3.95; p<0.10$.
 A strong significant effect was also observed between participant ID and number of dimensions mentioned, meaning some participants mentioned significantly more factors than others ($F(19,19)=4.53; p < 0.001$).
@@
 We found that the use of DP strongly influenced people to keep all candidates around, while participants in the B condition eliminated choices early.
-An ANOVA test demonstrated a significant effect between condition and number of candidates maintained ($F(1,19)=5.323; p<0.05$); a post-hoc Tukey HSD test confirmed that the number of candidates used in the final choice for participants in the A condition ($M=5.85;SD=0.45$) differed significantly from the B condition ($M=4.95; SD=2.26$).
+An ANOVA test demonstrated a significant effect between condition and number of candidates maintained; a post-hoc analysis with Tukey-Scheff\'{e} adjustments confirmed that the number of candidates used in the final choice for participants in the DP condition was greater ($M=5.85;SD=0.45$) than the control condition ($M=4.95; SD=2.26$) $F(1,19)=5.323; p<0.05$.
 

### 2012-09-19 chi2012.tex
minor proof fixes

--- 2012-09-19 chi2012.tex
@@
 For the first pre-study, we conducted a semi-structured interview to understand the reasons and ways that people drew information from diverse information sources.
-We interviewed 8 participants, asking them to identify tasks that performed recently that required use of multiple information sources, the kinds of sources used, and the tools used to manage the resulting information.
+We interviewed 8 participants, asking them to identify tasks they performed recently that required use of multiple information sources, the kinds of sources used, and the tools used to manage the resulting information.
 We then inquired as to how each participant would go about planning various aspects of a large party, such as date, location and entertainment.
@@
 
-The 8 participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
+The 8 participants, recruited informally via word of mouth, consisted of 7 males and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
 % mention # that were computer scientists here?
@@
 One participant said that they shared their outlook and gmail calendars, but only for work events and did not record social activities on them.
-To provide a full coverage of information, websites of different types were used, such as manufacturers website for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
+To provide a full coverage of information, websites of different types were used, such as manufacturers websites for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
 
@@
 				& etsy			& 31						& 1.13					& a:11			& a:10					& a:1						& a:2 \\
-				& amazon			& 108					&						&				& 						&						& \\
+				& amazon			& 108					& 1.96					&				& 						&						& \\
 \hline
@@
 \end{tabular}
-\caption{\emph{PS.2 Analysis of Data Feeds} - Structural statistics about data sampled from each API organised by category.
+\caption{\emph{PS.2 Analysis of Data Feeds}: Structural statistics about data sampled from each API organised by category.
 The first column lists the category and type of data record that was examined.
-Cols 3 and 4 list the average number of properties per data record, and number of levels from root to leaf.
+Columns 3 and 4 list the average number of properties per data record, and number of levels from root to leaf.
 The remaining fields show results from a pairwise analysis of data records from each of the data sources with those of the others in each category; values are prefixed with the first letter(s) of the target schema the row schema was compared against.
@@
 Our analysis of structured data feeds started with identifying a set of candidate sources to examine.
-We consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} list of most popular data feeds for this purpose.
+We consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb: \url{http://www.programmableweb.com/apis/directory}} list of most popular data feeds for this purpose.
 To increase the variety of sources, we selected from 5 categories: social network services, retailers, online event calendars, music sites, and weather, selecting 2--5 sources from each, for a total of 20.
@@
 As these were the ``simpler'' kinds of heterogeneity, this gave us optimism that an interactional approach would be feasible.
-In the following sections, we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.
+In the following sections, we describe the design process we used to create the prototype of DataPalette, followed by a detailed description of the design of the prototype and justification.
 

### 2012-09-19 chi2012.tex
minor changes

@@
 
-In order to feel familiar to new users, we borrowed the overall  ``look and feel'' from MacOS Finder-style file managers.
+In order to feel familiar to new users, we borrowed the overall  ``look and feel'' from OS X Finder-style file managers.
 In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can freely drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
@@
 
-The participants were seated at a standard Macintosh desktop equipped with a 22-inch monitor with a standard keyboard and mouse.
+The participants were seated at a standard OS X desktop equipped with a 22-inch monitor with a standard keyboard and mouse.
 Each participant was given a maximum of 10 minutes to each of two tasks, but were told that if they finished earlier they could tell us and move on to the next condition.

### 2012-09-19 chi2012.tex
minor

@@
 In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can freely drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
-Unlike file managers, the basic unit of data was not a file, but an \emph{instance} - corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
+Unlike file managers, the basic unit of data was not a file, but an \emph{instance} --- corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
 In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data into discrete instances with properties.
-This was done at time of import, and could be done for every major Web data feed type (e.g., JSON, XML, RDF) without any loss of generality.
+This was done at time of import, and could be done for every major Web data feed type (JSON, XML, RDF) without any loss of generality.
 

### 2012-09-19 chi2012.tex
de abstract

--- 2012-09-19 chi2012.tex
@@
 
-While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
-In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data on the Web.
-We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies examining the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
-These observations led to our design of DataPalette, an interface that introduced simple drag-and-drop and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
-Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions, drawing upon more data sources with DataPalette than a control condition during multidimensional, subjective-choice decisions.
+While people naturally draw from diverse information sources in the course of routine decision-making tasks, complementary support for managing heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
+In this paper, we present an initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data on the Web.
+We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources.
+These observations led to our design of DataPalette, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
+Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions during subjective-choice tasks when using DataPalette than when using a control interface, drawing information from more data sources in the process.
 
@@
 For the first pre-study, we conducted a semi-structured interview to understand the reasons and ways that people drew information from diverse information sources.
-We interviewed 8 participants, asking them to identify tasks they performed recently that required use of multiple information sources, the kinds of sources used, and the tools used to manage the resulting information.
+We interviewed 8 participants, asking them to identify tasks that performed recently that required use of multiple information sources, the kinds of sources used, and the tools used to manage the resulting information.
 We then inquired as to how each participant would go about planning various aspects of a large party, such as date, location and entertainment.
@@
 
-The 8 participants, recruited informally via word of mouth, consisted of 7 males and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
+The 8 participants, recruited informally via word of mouth, consisted of 7 male and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
 % mention # that were computer scientists here?
@@
 One participant said that they shared their outlook and gmail calendars, but only for work events and did not record social activities on them.
-To provide a full coverage of information, websites of different types were used, such as manufacturers websites for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
+To provide a full coverage of information, websites of different types were used, such as manufacturers website for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
 
@@
 				& etsy			& 31						& 1.13					& a:11			& a:10					& a:1						& a:2 \\
-				& amazon			& 108					& 1.96					&				& 						&						& \\
+				& amazon			& 108					&						&				& 						&						& \\
 \hline
@@
 \end{tabular}
-\caption{\emph{PS.2 Analysis of Data Feeds}: Structural statistics about data sampled from each API organised by category.
+\caption{\emph{PS.2 Analysis of Data Feeds} - Structural statistics about data sampled from each API organised by category.
 The first column lists the category and type of data record that was examined.
-Columns 3 and 4 list the average number of properties per data record, and number of levels from root to leaf.
+Cols 3 and 4 list the average number of properties per data record, and number of levels from root to leaf.
 The remaining fields show results from a pairwise analysis of data records from each of the data sources with those of the others in each category; values are prefixed with the first letter(s) of the target schema the row schema was compared against.
@@
 Our analysis of structured data feeds started with identifying a set of candidate sources to examine.
-We consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb: \url{http://www.programmableweb.com/apis/directory}} list of most popular data feeds for this purpose.
+We consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb - \url{http://www.programmableweb.com/apis/directory}} list of most popular data feeds for this purpose.
 To increase the variety of sources, we selected from 5 categories: social network services, retailers, online event calendars, music sites, and weather, selecting 2--5 sources from each, for a total of 20.
@@
 As these were the ``simpler'' kinds of heterogeneity, this gave us optimism that an interactional approach would be feasible.
-In the following sections, we describe the design process we used to create the prototype of DataPalette, followed by a detailed description of the design of the prototype and justification.
+In the following sections, we describe the design process we used to create prototype DataPalette, followed by a detailed description of the design of the prototype and justification.
 
@@
 The second metric was the number of factors the participant considered in making their final choice.
-To determine this, we asked participants to explain precisely why they made their choice, and matched their justifications to properties in the data.
-This analysis revealed the interface used made a significant impact on the number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
+To determine this, we asked participants to explain why they made their choice(s), and recorded the number of distinct data dimensions mentioned.
+A 2-way ANOVA test of the effect of interface and participant ID on the number of factors mentioned revealed a significant effect; post-hoc analysis with Tukey-Scheff\'{e} adjustments indicated that participants in the DP condition mentioned significantly more factors ($M=5.5; SD=1.7$) than those in the control condition ($M=4; SD=1.73$), $F(1,19)=3.95; p<0.10$.
+A strong significant effect was also observed between participant ID and number of dimensions mentioned, meaning some participants mentioned significantly more factors than others ($F(19,19)=4.53; p < 0.001$).
 
 \subsubsection{Diversity: Data sources consulted}
-The third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
-In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04$ $p\approx0.10$).
+The third metric we used to gauge the task performance is the number of different data sources consulted during the trial; our rationale for this metric was that more informed decisions derived from more thorough consideration of available data.
+In this metric, participants used an average of 88\% ($SD=0.13$) of data sources provided in the DP case, compared to 80\% ($SD=0.18$)in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance ($F(1,19)=0.04$ $p\approx0.10$).
 
 \subsubsection{Effort: External cognition}
-Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much participants had to keep track of during the study.
-We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22; p\approx0.10$).
-
-\begin{figure}[htb]
+Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much information participants had to keep track of during the study.
+We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in a 2-way ANOVA blocked by participant ID ($F(1,19)= 8.22; p\approx0.10$).
+
+\begin{figure}[tbp]
 \begin{center}
@@
 Just as the \emph{user-subjective approach} to PIM \cite{user-subjective} showed that subjective attributes and labels were more useful than physical ones, we have shown that a user-driven approach to data integration, as achieved with DataPalette, affords users flexibility to combine and keep separate data at will, and may facilitate decision-making tasks.
-Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through simple, fast interactive approaches that are received as intuitive by end-users.
+Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through easy to use, and intuitive interactive approaches for end-users.
 

### 2012-09-19 chi2012.tex

@@
 
-In order to feel familiar to new users, we borrowed the overall  ``look and feel'' from MacOS Finder-style file managers.
+In order to feel familiar to new users, we borrowed the overall  ``look and feel'' from OS X Finder-style file managers.
 In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can freely drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
-Unlike file managers, the basic unit of data was not a file, but an \emph{instance} - corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
+Unlike file managers, the basic unit of data was not a file, but an \emph{instance} --- corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
 In order to make working with relational graphs manageable, an early decision we made was to break up RDF-like relational data into discrete instances with properties.
-This was done at time of import, and could be done for every major Web data feed type (e.g., JSON, XML, RDF) without any loss of generality.
+This was done at time of import, and could be done for every major Web data feed type (JSON, XML, RDF) without any loss of generality.
 

### 2012-09-19 chi2012.tex
minor

@@
 Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, identifiers used by each network were system-specific and inconsistent.
-Thus, we sought to make instead make it user-driven, and easy to do manually.
+Thus, we sought to instead make it user-driven, and easy to perform manually.
 
@@
 The third metric we used to gauge the task performance is the number of different data sources consulted during the trial; our rationale for this metric was that more informed decisions derived from more thorough consideration of available data.
-In this metric, participants used an average of 88\% ($SD=0.13$) of data sources provided in the DP case, compared to 80\% ($SD=0.18$)in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance ($F(1,19)=0.04$ $p\approx0.10$).
+In this metric, participants used an average of 88\% ($SD=0.13$) of data sources provided in the DP case, compared to 80\% ($SD=0.18$) in the control condition, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance ($F(1,19)=0.04$ $p < 0.15$).
 

### 2012-09-19 chi2012.tex
minor

@@
 
-While people naturally draw from diverse information sources in the course of routine decision-making tasks, complementary support for managing heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
-In this paper, we present an initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data on the Web.
-We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources.
-These observations led to our design of DataPalette, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
-Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions during subjective-choice tasks when using DataPalette than when using a control interface, drawing information from more data sources in the process.
+While people naturally draw from diverse information sources in the course of routine decision-making, complementary support for managing diverse, heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
+In this paper, we present an exploratory initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data on the Web.
+We conducted an exploratory, sequential, mixed-method study, starting with two pre-studies examining the data integration needs and challenges, respectively, of effectively combining data from diverse Web-based structured data sources.
+These observations led to our design of DataPalette, an interface that introduced simple drag-and-drop and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
+Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions, drawing upon more data sources with DataPalette than a control condition during multidimensional, subjective-choice decisions.
 
@@
 To test these hypotheses, we designed a within-subjects (repeated measures) lab-based A-B study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (condition A) and once with a baseline interface (condition B).
-Each task was paired with one of two datasets, depending on condition, which were counterbalanced in a full-factorial design to eliminate potential ordering, task, and dataset biases.\looseness=-1
+Each task was paired with one of two datasets, depending on condition, which were fully counterbalanced in a Latin square design to eliminate potential ordering, task, and dataset biases.\looseness=-1
 
@@
 
-In the control conditions, participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from web sites pertaining to the candidates they were asked to choose from.
+In the control condition (B), participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from web sites pertaining to the candidates they were asked to choose from.
 Participants were told they were free to use the web sites and spreadsheets however they wished.
@@
 Participants spent an average of 8.9 minutes performing the tasks.
-Condition A (DP) took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$); however, a 2-way repeated measures ANOVA test of time taken by interface condition, blocked by participant ID, revealed that interface condition did not have a significant effect on task completion time.
-This analysis also showed no significant effect of participant ID on task completion.
-Comparing tasks, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), but this difference was not significant.
+Condition A (DP) took slightly less time ($M=8.7$min, $SD=1.79$) per trial than condition B ($M=9.0$min; $SD=1.44$).
+An ANOVA test of time taken by interface condition blocking by participant did not find interface use a significant predictor of task completion time.
+Although across trials, the restaurant selection task took on average slightly less time ($M=8.8$min,$SD=1.58$) than the university trials ($M=8.9$min,$SD=1.66$), the aforementioned ANOVA test found task type to not be a significant predictor.
 
@@
 The second metric was the number of factors the participant considered in making their final choice.
-To determine this, we asked participants to explain why they made their choice(s), and recorded the number of distinct data dimensions mentioned.
-A 2-way ANOVA test of the effect of interface and participant ID on the number of factors mentioned revealed a significant effect; post-hoc analysis with Tukey-Scheff\'{e} adjustments indicated that participants in the DP condition mentioned significantly more factors ($M=5.5; SD=1.7$) than those in the control condition ($M=4; SD=1.73$), $F(1,19)=3.95; p<0.10$.
-A strong significant effect was also observed between participant ID and number of dimensions mentioned, meaning some participants mentioned significantly more factors than others ($F(19,19)=4.53; p < 0.001$).
+To determine this, we asked participants to explain precisely why they made their choice, and matched their justifications to properties in the data.
+This analysis revealed the interface used made a significant impact on the number of factors weighed ($F(1,19)=3.95; p<0.10$); participants in the DP condition listed an average of 5.5 choices in their decision ($SD=1.7$) versus 4 ($SD=1.73$) in the standard condition.
 
 \subsubsection{Diversity: Data sources consulted}
-The third metric we used to gauge the task performance is the number of different data sources consulted during the trial; our rationale for this metric was that more informed decisions derived from more thorough consideration of available data.
-In this metric, participants used an average of 88\% ($SD=0.13$) of data sources provided in the DP case, compared to 80\% ($SD=0.18$) in the control condition, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance ($F(1,19)=0.04$ $p < 0.15$).
+The third metric we used to gauge how well the participant performed their tasks is the number of different data sources the participant consulted during the trial.
+In this metric, participants used an average of 88\% of data sources provided in the DP case, compared to 80\% in the standard interface, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance in predicting the number of sources consulted ($F(1,19)=0.04$ $p\approx0.10$).
 
 \subsubsection{Effort: External cognition}
-Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much information participants had to keep track of during the study.
-We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in a 2-way ANOVA blocked by participant ID ($F(1,19)= 8.22; p\approx0.10$).
-
-\begin{figure}[tbp]
+Although we did not have an explicit measure of effort for this experiment (such as NASA's TLX \cite{tlx}) the use of paper notes provided an indicator of how much participants had to keep track of during the study.
+We found that participants took notes less on average (25\%) with the DP interface than the standard interface (35\%), a difference which approached significance in an ANOVA blocked by participant ($F(1,19)= 8.22; p\approx0.10$).
+
+\begin{figure}[htb]
 \begin{center}
@@
 \subsubsection{Candidate selection}
+
 Since the tasks were subjective-choice, we could not evaluate the `correctness' of answers.
@@
 We found that the use of DP strongly influenced people to keep all candidates around, while participants in the B condition eliminated choices early.
-An ANOVA test demonstrated a significant effect between condition and number of candidates maintained; a post-hoc analysis with Tukey-Scheff\'{e} adjustments confirmed that the number of candidates used in the final choice for participants in the DP condition was greater ($M=5.85;SD=0.45$) than the control condition ($M=4.95; SD=2.26$) $F(1,19)=5.323; p<0.05$.
+An ANOVA test demonstrated a significant effect between condition and number of candidates maintained ($F(1,19)=5.323; p<0.05$); a post-hoc Tukey HSD test confirmed that the number of candidates used in the final choice for participants in the A condition ($M=5.85;SD=0.45$) differed significantly from the B condition ($M=4.95; SD=2.26$).
 
@@
 Just as the \emph{user-subjective approach} to PIM \cite{user-subjective} showed that subjective attributes and labels were more useful than physical ones, we have shown that a user-driven approach to data integration, as achieved with DataPalette, affords users flexibility to combine and keep separate data at will, and may facilitate decision-making tasks.
-Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through easy to use, and intuitive interactive approaches for end-users.
+Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through simple, fast interactive approaches that are received as intuitive by end-users.
 

### 2012-09-19 chi2012.tex
minor

--- 2012-09-19 chi2012.tex
@@
 We based our two tasks on choosing a university and picking a restaurant for a large social event, because in PS.1 the majority of people stated that they had used multiple websites for these tasks.
-We used different two different variables for both tasks, where the restaurant was located (Glasgow and Cambridge) and which subject people chose (sport science and history).
+We used two different variables for both tasks, where the restaurant was located (Glasgow and Cambridge) and which subject people chose (sport science and history).
 For the restaurant task, the cities Glasgow and Cambridge were chosen because they were geographically distant to participants, and would therefore minimise likelihood of prior familiarity with the restaurants or location.

### 2012-11-19 rebuttal.txt
rebuttal

We would like to thank the reviewers & AC for their extremely detailed suggestions and comments. We would like to address the suggestions and concerns in order of importance suggested by the AC as follows: evaluation and discussion, similarity to "mash-up" makers, lack of automation, formatting and figures, and future work.

Evaluation - We think R2's comments suggest that our description of the tasks and experimental set-up were perhaps too parsimonious, and could be expanded to be made clearer.  The extra space that will be gained from condensing Table 1 (and possibly eliminating Table 3 altogether, as suggested by R2) will allow us to describe the tasks and methodology adequately, as well as to go into our task design justification with greater detail.  We believe that R2's comments also indicate a need to expand the description of how measures were taken, particularly pertaining to task performance (e.g. that the number of dimensions each participant considered was derived from an analysis of  speak-aloud dictation). Addressing the relevance of evaluation (R2), plan to connect with better descriptive text, the hypotheses (h1-3) with the goals initially set out in the introduction, and similarly, the measures taken with the hypotheses.

Distinction with "mashup-makers" - While we originally cited the Mashmaker, Vegemite and Marmite work, in the process of condensing related/background work, decided to cut these due to difference in emphasis of capability and interface metaphors/data flow.  Seeking to enable reuse of data "off of any web page", a major capability of mashup makers is unstructured data acquisition, e.g., page scraping, tabular aggregation and shaping.  DataPalette, meanwhile, starts with the observation that already structured data can be acquired directly from most, if not all major web sites and platform providers today.  Second,

Automation - We completely agree with R1 and R3 that careful introduction of automation could improve the user experience and reduce tedium, potentially helping people arrive at the view(s) and conclusions they want faster.  In fact, we tried a few approaches at automatic same-as reconciliation and visualisation view-inference (DWIM approaches) during our prototype design phase, and ran into a hurdles which essentially caused us to go back to the largely manual design for the final evaluation.

The first issue pertained to our observation that participants only sometimes wanted consolidated views of data -- depending on task (and sometimes even on a per-instance basis).  Un-aggregated views were preferred for reasons ranging from "it's simply easier to think about 'facebook data' vs 'what's in my address book'", to reasons of provenance ("Yelp is more trustworthy so I want to refer to it separately"). Thus, even if reconciliation was made partially automatic, it would need to be user-initiated, so that support for separated (un-aggregated) views was maintained.  
The second issue pertained to helping users cope with errors -- that debugging incorrectly-combined (false positives) or left-separated (false negatives) causes immediate confusion for non-programmers.  We observed this during testing even without automatic aggregation, when participants inadvertantly combined records when accidentally dragging and dropping items - they simply had no idea what they did, and later could not figure out why dissimilar data items were combined.

In summary we think that thus, intorduction of automation bears further design investigation, and will introduce these thoughts and possible workarounds in a "Future Work" section following "Study Limitations"

### 2012-12-26 rebuttal-2.txt
getting ready to do final submission

--- 2012-11-19 rebuttal.txt
@@
 that the number of dimensions each participant considered was derived from an analysis of  speak-aloud dictation).
-Addressing the relevance of evaluation (R2), plan to connect with better descriptive text, the hypotheses (h1-3) with the goals initially set out in the introduction, and similarly, the measures taken with the hypotheses.
+Addressing the relevance of evaluation (R2), plan to connect with better descriptive text, the hypotheses (h1-3) the goals initially set out in the introduction, and similarly, the connection between the measures taken and the hypotheses.
 
-Distinction with "mashup-makers" - While we originally cited the Mashmaker, Vegemite and Marmite work, in the process of condensing related/background work, decided to cut these due to difference in emphasis of capability and interface metaphors/data flow.
-Seeking to enable reuse of data "off of any web page", a major capability of mashup makers is unstructured data acquisition, e.g., page scraping, tabular aggregation and shaping.
-DataPalette, meanwhile, starts with the observation that already structured data can be acquired directly from most, if not all major web sites and platform providers today.
-Second,
+Distinction with "mashup-makers" - We originally cited the Mashmaker, Vegemite and Marmite work, but in the process of condensing related/background work, decided to cut these due to difference in emphasis of capability.
+Seeking to enable reuse of data "off of any web page", a heavy  emphases of this work is on data acquisition, e.g., page scraping, unstructured data manipulation and shaping.
+DataPalette, meanwhile, while similar in the sense of encouraging data re-use, starts with the observation that already structured data can be acquired directly from most, if not all major web sites and platform providers today.
 

### 2012-12-26 rebuttal.txt
getting ready to do final submission

@@
 We would like to thank the reviewers & AC for their extremely detailed suggestions and comments.
-We would like to address the suggestions and concerns in order of importance suggested by the AC as follows: evaluation and discussion, similarity to "mash-up" makers, lack of automation, formatting and figures, and future work.
+In this rebuttal, we will briefly address the suggestions and concerns in order of importance (as described by the AC): evaluation, prestudy 1 sample bias, similarity to "mash-up" makers, potential for automation, and, paper formatting & figures.
 
-Evaluation - We think R2's comments suggest that our description of the tasks and experimental set-up were perhaps too parsimonious, and could be expanded to be made clearer.
-The extra space that will be gained from condensing Table 1 (and possibly eliminating Table 3 altogether, as suggested by R2) will allow us to describe the tasks and methodology adequately, as well as to go into our task design justification with greater detail.
-We believe that R2's comments also indicate a need to expand the description of how measures were taken, particularly pertaining to task performance (e.g.
-that the number of dimensions each participant considered was derived from an analysis of  speak-aloud dictation).
-Addressing the relevance of evaluation (R2), plan to connect with better descriptive text, the hypotheses (h1-3) with the goals initially set out in the introduction, and similarly, the measures taken with the hypotheses.
+Evaluation - R2's comments suggest that the description of the study design was perhaps too parsimonious, and should be made clearer.
+The extra space gained from condensing Table 1 (and possibly eliminating Table 3 altogether, as suggested by R2) should give us space to adequately describe the tasks and methodology, as well as to provide some design justification to the choice of tasks and format, including choice of measures.
+We believe R2's comments also indicate a need to clarify the ways each of the measures were taken; for example, we will clarify that the number of properties considered were measured by counting distinct properties mentioned during the think-aloud process.
+We can also provide a discussion of the potential limitations of such approaches (e.g., of think-aloud for task analysis, and how we think this technique may under-estimate results but may be still suitable for comparative analysis used consistently across conditions).
+Addressing the relevance of evaluation (R2), we would like to improve the descriptive text that connects the hypotheses (h1-3) with the goals set out in the intro, and similarly, the measures taken with the hypotheses.
+We will also call out the potential bias resulting from the subject selection process of PS1 (as suggested by R1) which involved 5 computer scientists - a substantially different source population than that for our main study.
 
-Distinction with "mashup-makers" - While we originally cited the Mashmaker, Vegemite and Marmite work, in the process of condensing related/background work, decided to cut these due to difference in emphasis of capability and interface metaphors/data flow.
-Seeking to enable reuse of data "off of any web page", a major capability of mashup makers is unstructured data acquisition, e.g., page scraping, tabular aggregation and shaping.
-DataPalette, meanwhile, starts with the observation that already structured data can be acquired directly from most, if not all major web sites and platform providers today.
-Second,
+Distinction with "mashup-makers" - While we intended to cite Mashmaker, Vegemite & Marmite, we omitted them due to perceived differences in emphasis & approach.
+First, these systems focused on unstructured data acquisition, e.g., scraping and collecting data from web pages.
+With DataPalette, we started with pre-structured data from APIs and feeds allowing for a focus on structured reconciliation problems instead.
+Second, DataPalette gets users to work directly with instances, while Vegemite, Marmite & Yahoo!
+Pipes required users to construct data-flows (using visual programming paradigms) or scripts (through programming by demonstration).
+We believe that directly working with instances is simpler and more accessible, although perhaps more limited, than these "simplified programming" approaches, because users avoid having to think about generalizing their actions using abstract operators.
+A direct-manipulation approach also better supports small-scale exploratory sensemaking by providing direct, immediate feedback.
+We will to mention these differences, & re-introduce references as suggested.
 
-Automation - We completely agree with R1 and R3 that careful introduction of automation could improve the user experience and reduce tedium, potentially helping people arrive at the view(s) and conclusions they want faster.
-In fact, we tried a few approaches at automatic same-as reconciliation and visualisation view-inference (DWIM approaches) during our prototype design phase, and ran into a hurdles which essentially caused us to go back to the largely manual design for the final evaluation.
+Automation - We agree w/ R1 & R3 that automation could improve DataPalette and potentially help people arrive at conclusions faster.
+We considered a number of approaches to automating same-as & view selection during the design process, but ran into challenges.
+First, testers only sometimes wanted the consolidate data from multiple places, on a seemingly task & individual basis.
+Un-consolidated instances were preferred sometimes for reasons of simplicity "it's easier to think about it separately", or for reasons of reputation & trust (e.g., "Yelp reviews are better so I separate them").
+Thus, even if reconciliation were automatic, selectively (manually) initiating it might be important.
+The second, greater challenge pertained to error handling: errors in consolidation were difficult to detect.
+Even with our manual-only approach, discovering and correcting incorrectly-combined (false positives) or incompletely-combined instances (e.g., resulting from dragging & dropping items in the wrong place(s)) caused people to be confused.
+Thus we must consider ways to make this combination process more visible and understandable.
+In summary, re: automation, we think that there is much potential for automation, but it also  presents substantial interaction challenges we need to understand better.
+We will mention these challenges in "Future work" at the end of the paper.
 
-The first issue pertained to our observation that participants only sometimes wanted consolidated views of data -- depending on task (and sometimes even on a per-instance basis).
-Un-aggregated views were preferred for reasons ranging from "it's simply easier to think about 'facebook data' vs 'what's in my address book'", to reasons of provenance ("Yelp is more trustworthy so I want to refer to it separately").
-Thus, even if reconciliation was made partially automatic, it would need to be user-initiated, so that support for separated (un-aggregated) views was maintained.
-The second issue pertained to helping users cope with errors -- that debugging incorrectly-combined (false positives) or left-separated (false negatives) causes immediate confusion for non-programmers.
-We observed this during testing even without automatic aggregation, when participants inadvertantly combined records when accidentally dragging and dropping items - they simply had no idea what they did, and later could not figure out why dissimilar data items were combined.
+Formatting and figures - In order to make figures more readable (AC), we will expand to full-width the screen shot in Fig.1, and make the fonts of Fig.2 and Fig.5 larger as well.
+As described, we will consolidate T1 and eliminate T3 to make room for the above mentioned improvements.
+In its place, we will (slightly) expand the textual discussion of PS2 (originally explanatory text for T1) to focus on observations that most informed the design of the system.
+We will perform additional proofreading and corrections, including the minor edits suggested by R2 and R3.
 
-In summary we think that thus, intorduction of automation bears further design investigation, and will introduce these thoughts and possible workarounds in a "Future Work" section following "Study Limitations"
+In summary, we hope to provide a clearer, more readable and better justified final copy, using space gained from Tables 1 & 3, and improving the final discussion of results.

### 2013-01-04 chi2013-webbox.tex
new abstract

--- 2012-12-26 chi2013-webbox-submitted.tex
@@
 
-%% \numberofauthors{0}
-%% \author{
-%%   \alignauthor 1st Author Name\\
-%%     \affaddr{Affiliation}\\
-%%     \affaddr{Address}\\
-%%     \email{e-mail address}\\
-%%     \affaddr{Optional phone number}
-%%   \alignauthor 2nd Author Name\\
-%%     \affaddr{Affiliation}\\
-%%     \affaddr{Address}\\
-%%     \email{e-mail address}\\
-%%     \affaddr{Optional phone number}    
-%%   \alignauthor 3rd Author Name\\
-%%     \affaddr{Affiliation}\\
-%%     \affaddr{Address}\\
-%%     \email{e-mail address}\\
-%%     \affaddr{Optional phone number}
-%% }
+\numberofauthors{1}
+ \author{
+   \alignauthor Max Van Kleek, Daniel A. Smith, Heather R. Packer, James Skinner, Nigel R. Shadbolt \\
+     \affaddr{School of Electronics and Computer Science}\\
+     \affaddr{University of Southampton, Southampton UK}\\
+     \email{\{emax, ds, hp3, js40g09, nrs\}@ecs.soton.ac.uk}}
 
@@
 \begin{abstract}
-
-While people naturally draw from diverse information sources in the course of routine decision-making tasks, complementary support for managing heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
+The information processing capabilities of humans enable us to opportunistically draw evidence from available information sources to carry out tasks effectively.
+However, in the digital realm, serendipitously combining structured data from different sources without prior careful alignment or integration remains difficult due to problems of heterogeneity that arise when data modeled and represented differently are brought together.
+In this paper, we investigate the feasibility of extending digital Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, towards enabling users to more effectively and routinely take advantage of the vast emerging ecosystems of structured data on the Web.
+We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources.
+These observations led to our design of \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
+Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions during subjective-choice tasks when using DataPalette than when using a control interface, drawing information from more data sources in the process.
+
+% While people naturally draw from diverse information sources in the course of routine decision-making tasks, complementary support for managing heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
 In this paper, we present an initial investigation of user-driven, ad-hoc data integration approaches for PIM, towards enabling greater use of the emerging ecosystems of structured data on the Web.
@@
 A result of this problem is that users can experience data fragmentation~\cite{Jones05towardsa} across multiple web-based applications and services, and struggle to consolidate across them~\cite{bergman,boardmansasse}.
-In our own previously-conducted field study, (reference anonymised) we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage both online and offline data.
+In our own previously-conducted field study, \cite{infoscraps} we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage both online and offline data.
 Similar findings from other studies observed volunteers' ``homebrew databases'': manually-maintained information collated to handle the heterogeneous data requirements of groups needing to coordinate their activities~\cite{Voida:2011:HDC:1978942.1979078}.\looseness=-1

### 2013-01-04 chi2013-webbox.tex
new abst

@@
  \author{
-   \alignauthor Max Van Kleek, Daniel A. Smith, Heather R. Packer, James Skinner, Nigel R. Shadbolt \\
+   \alignauthor Max Van Kleek, Daniel A. Smith, Heather R. Packer, Jim Skinner, Nigel R. Shadbolt \\
      \affaddr{School of Electronics and Computer Science}\\
-     \affaddr{University of Southampton, Southampton UK}\\
+     \affaddr{University of Southampton, Southampton, UK}\\
      \email{\{emax, ds, hp3, js40g09, nrs\}@ecs.soton.ac.uk}}
@@
 \begin{abstract}
-The information processing capabilities of humans enable us to opportunistically draw evidence from available information sources to carry out tasks effectively.
-However, in the digital realm, serendipitously combining structured data from different sources without prior careful alignment or integration remains difficult due to problems of heterogeneity that arise when data modeled and represented differently are brought together.
-In this paper, we investigate the feasibility of extending digital Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, towards enabling users to more effectively and routinely take advantage of the vast emerging ecosystems of structured data on the Web.
+The information processing capabilities of humans enable people to opportunistically draw and integrate knowledge from nearly any information sources that become available to them.
+However, supporting similar, opportunistic integration of digital, structured data from diverse sources remains difficult without tedious, manual integration, due to problems of heterogeneity that arise when data modeled and represented differently are brought together.
+In this paper, we investigate the feasibility of minimally extending digital Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, towards enabling users to take advantage of the vast emerging ecosystems of structured data being made available.
 We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources.

### 2013-01-05 chi2013-webbox.tex
new abst
intro work

@@
 %}
+
+\toappear{
+Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page.
+To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee.\\
+\emph{CHI 2013}, April 27--May 2, 2013, Paris, France.\\
+Copyright 2013 ACM 978-1-4503-1899-0/13/04...\$15.00.}
 
@@
  \author{
-   \alignauthor Max Van Kleek, Daniel A. Smith, Heather R. Packer, Jim Skinner, Nigel R. Shadbolt \\
+   \alignauthor Max Van Kleek, Daniel A. Smith, Heather S. Packer, Jim Skinner, Nigel R. Shadbolt \\
      \affaddr{School of Electronics and Computer Science}\\
@@
 \begin{abstract}
-The information processing capabilities of humans enable people to opportunistically draw and integrate knowledge from nearly any information sources that become available to them.
+The information processing capabilities of humans enable us to opportunistically draw and integrate knowledge from nearly any information sources that become available.
+However, enabling similar, opportunistic integration of digital, structured data from diverse sources has remained difficult without tedious, manual integration, due to problems of heterogeneity that arise when data modeled separately and differently are brought together.
+In this paper, we present an investigation of the feasibility of extending digital Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, towards enabling users to take advantage of the vast emerging ecosystems of structured data being made available.
+We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources.
+These observations led to our design of \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
+Our lab study showed, first, that participants understood the new interaction mechanisms introduced, and, second, that they made more carefully-justified decisions weighing a greater number of factors, while spending less effort and taking the same amount of time, during subjective-choice tasks when using DataPalette than a control (web page and spreadsheet) interface.
+%The information processing capabilities of humans enable people to opportunistically draw and integrate knowledge from nearly any information sources that become available to them.
 However, supporting similar, opportunistic integration of digital, structured data from diverse sources remains difficult without tedious, manual integration, due to problems of heterogeneity that arise when data modeled and represented differently are brought together.
@@
 Our lab study showed that participants understood the interaction mechanisms introduced, and made more carefully-justified decisions during subjective-choice tasks when using DataPalette than when using a control interface, drawing information from more data sources in the process.
-
 % While people naturally draw from diverse information sources in the course of routine decision-making tasks, complementary support for managing heterogeneous data in digital Personal Information Management (PIM) tools remains poor.
@@
 \section{Introduction}
-
 In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web.
-This information has emerged from many channels, from the APIs now offered by most of the major retailers, apps, and social media sites, to the massive publishing of datasets by governmental and other institutions, to even the emergence of new kinds of data sources, such as wearable and smartphone-based bio-sensors.
-A primary goal of making much of this data available has been to give end-user citizens the ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
-This data is predominately used by specialised app developers, journalists and other ``data specialists'' but remains inaccessible to citizens because of the available tools.
-For example, the most commonly used structured personal information management (PIM) tools either manage a small, fixed set of data types (such as digital calendaring tools and  to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{infoscraps}.
+This explosion of structured personal data APIs and downloadable data sets has arrived from many different sectors, including retail, banking, social networking, opinion and recommendation sites, mobile and desktop apps, as well as a number of entirely new kinds of sources such as wearable activity and bio-sensors.
+Perhaps even more significantly, governmental departments and authorities, from the smallest local councils to entire federal departments, have started to release large quantities of data as a way to provide transparency and accountability to citizens.
+A common goal for the release of this data, however, has been to provide end-user citizens an ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
+Thus far, however, this data has been predominately used by app developers, journalists and other ``data specialists'' to perform ``one-off'' data analyses and investigations, but has remained inaccessible to citizens because of a lack of suitable tools with which these data sets can be easily manipulated and explored.
+For example, the most familiar kinds of personal information management (PIM) tools only provide facilities for managing small, fixed sets of data types (such as digital calendaring tools and to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{infoscraps}.
 
@@
 In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data.
-We used a three-stage sequential exploratory mixed-method approach: first, we performed a qualitative analysis of a semi-structured interview pre-study to understand the types of tasks people performed using multiple information sources, and the processes that they relied upon to perform such tasks.
-Our results suggested that people rely on multiple and diverse sources, then singular, integrated ones for a number of reasons, including coverage, reliability, and for alternative perspectives.
+We used a three-stage sequential exploratory mixed-method approach: first, we performed a semi-structured interview pre-study to understand the types of tasks people performed using multiple information sources, and the processes that they relied upon to perform such tasks.
+Our results suggested that people routinely drew upon multiple and diverse sources for reasons including coverage, reliability, and to gain multiple perspectives or subjective assessments.
 
@@
 
-Second, we conducted a structural analysis of various popular personal data sources available today to identify potential barriers to unifying data.
-Drawing upon definitions from the database integration and automatic schema matching literature, we characterised the types of heterogeneity exhibited across data sources in six domains: contacts, events, music, shopping, social networking and weather, finding that terminological heterogeneity dominates the simpler data feeds (including social networking and restaurant recommendations), while structural issues pervade the more complex schemas of online retailers' product catalogues.
+Second, we conducted a structural analysis of data records from personal data sources to identify the most common kinds of problems likely to arise from integrating data from these types of sources.
+We examined data sources in six common domains: contacts, events, music, shopping, social networking and weather.
+Our analysis found that while terminological heterogeneity was the most common issue for unifying simpler data records (such as address book contacts and social network profiles), structural differences were exhibit among more complex schemas, such as online retailers' product catalogues.
 
@@
 \section{Background}
-
 In this section, we motivate our work using three fields: PIM, data integration, and end-user mashups and toolkits.

### 2013-01-05 chi2013-webbox.tex
intro work

@@
 In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web.
-This explosion of structured personal data APIs and downloadable data sets has arrived from many different sectors, including retail, banking, social networking, opinion and recommendation sites, mobile and desktop apps, as well as a number of entirely new kinds of sources such as wearable activity and bio-sensors.
-Perhaps even more significantly, governmental departments and authorities, from the smallest local councils to entire federal departments, have started to release large quantities of data as a way to provide transparency and accountability to citizens.
-A common goal for the release of this data, however, has been to provide end-user citizens an ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
+The explosion of structured data APIs and downloadable data sets has arrived from many different sectors, including retail, banking, social networking, opinion and recommendation sites, mobile and desktop apps, as well as from new kinds sensors and devices such as wearable activity and bio-sensors.
+In addition to commercial apps and services, governments have embraced open data as a way of providing transparency and accountability to their citizens, causing a huge push to release data at scales ranging from the smallest local councils to entire national departments.
+A common goal for the release of such data has been to provide end-user citizens an ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
 Thus far, however, this data has been predominately used by app developers, journalists and other ``data specialists'' to perform ``one-off'' data analyses and investigations, but has remained inaccessible to citizens because of a lack of suitable tools with which these data sets can be easily manipulated and explored.

### 2013-01-05 chi2013-webbox.tex
mor etext

@@
 The explosion of structured data APIs and downloadable data sets has arrived from many different sectors, including retail, banking, social networking, opinion and recommendation sites, mobile and desktop apps, as well as from new kinds sensors and devices such as wearable activity and bio-sensors.
-In addition to commercial apps and services, governments have embraced open data as a way of providing transparency and accountability to their citizens, causing a huge push to release data at scales ranging from the smallest local councils to entire national departments.
+Beyond commercial apps and services, governments have embraced releasing data as a way of providing transparency and accountability to their citizens, causing a huge push to release ``open data'' to the public, at scales ranging from the smallest local councils to entire national departments.
+
 A common goal for the release of such data has been to provide end-user citizens an ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
-Thus far, however, this data has been predominately used by app developers, journalists and other ``data specialists'' to perform ``one-off'' data analyses and investigations, but has remained inaccessible to citizens because of a lack of suitable tools with which these data sets can be easily manipulated and explored.
-For example, the most familiar kinds of personal information management (PIM) tools only provide facilities for managing small, fixed sets of data types (such as digital calendaring tools and to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{infoscraps}.
+Thus far, however, this data has been predominately been consumed by app developers, journalists and other ``data specialists'' in one-off data analyses, apps and investigations, remaining still far from influencing the activities of ordinary end-user citizens.
+A core reason for this, we posit, is a lack of suitable tools for accessing and interrogating these heterogeneous data sets, and a means by which these disparate information sources can be effectively brought together and cross-referenced.
+For example, most personal information management (PIM) tools only provide facilities for managing small, pre-fixed sets of data types (such as digital calendaring tools and to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{infoscraps}.
 

### 2013-01-05 chi2013-webbox.tex
mor etext

@@
 
-We hypothesise that end-users will benefit from tools that allow them to effectively browse, use and consume heterogeneous structured data from diverse sources.
-In this paper, we present an investigation into extending PIM tools to effectively use emerging ecosystems of personal data.
-We used a three-stage sequential exploratory mixed-method approach: first, we performed a semi-structured interview pre-study to understand the types of tasks people performed using multiple information sources, and the processes that they relied upon to perform such tasks.
+% We hypothesise that end-users will benefit from tools that allow them to effectively browse, use and consume heterogeneous structured data from diverse sources. 
+In this paper, we present an investigation into ways that PIM tools might be extended to allow data to be brought in in an ad-hoc fashion from any structured data source, combined with other data, and interrogated to effectively allow end-user citizens to harness knowledge from the vast, emerging and future ecosystems of personal data.
+We used a three-stage sequential exploratory mixed-method approach in this investigation: first, we performed a semi-structured interview pre-study to understand the types of tasks people performed using multiple information sources, and the processes that they relied upon to perform such tasks.
 Our results suggested that people routinely drew upon multiple and diverse sources for reasons including coverage, reliability, and to gain multiple perspectives or subjective assessments.

### 2013-01-06 chi2013-webbox.tex
related work

@@
 % We hypothesise that end-users will benefit from tools that allow them to effectively browse, use and consume heterogeneous structured data from diverse sources. 
-In this paper, we present an investigation into ways that PIM tools might be extended to allow data to be brought in in an ad-hoc fashion from any structured data source, combined with other data, and interrogated to effectively allow end-user citizens to harness knowledge from the vast, emerging and future ecosystems of personal data.
+In this paper, we present an investigation into ways that PIM tools might be extended to allow data to be brought together in an ad-hoc fashion from arbitrary structured data sources, and interrogated, so as to enable end-user citizens to effectively harness knowledge from the vast, emerging ecosystems of data.
 We used a three-stage sequential exploratory mixed-method approach in this investigation: first, we performed a semi-structured interview pre-study to understand the types of tasks people performed using multiple information sources, and the processes that they relied upon to perform such tasks.
@@
 
-A result of this problem is that users can experience data fragmentation~\cite{Jones05towardsa} across multiple web-based applications and services, and struggle to consolidate across them~\cite{bergman,boardmansasse}.
+The user experience benefits of conquering such data integration challenges have been demonstrated in PIM prototype systems with integrative data models.
+For example, the uniform query capabilities provided by SEMEX~\cite{semex} and \emph{Stuff I've Seen}\cite{stuffiveseen}, showed the speed, efficiency and re-findability gained when a user is able to quickly trace and cross-reference information about the people, places and things mentioned throughout his or her files, folders, e-mail, and other information repositories.
+Similarly, Haystack \cite{haystack} demonstrated the reduced effort and increased efficiency of being able to view and re-use information easily in multiple task contexts without fragmentation or loss of consistency when a single, consolidated data model was used across applications.
+
+Therefore, despite the multifaceted challenges of data integration, various fields have sought methods to resolve them.
+The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} fields, for example, have pursued automatic, recently machine-learning approaches to data integration, at either the level of \emph{ontology/schema-level matching}~\cite{euzenat2004api,doan2003learning}, or \emph{instance matching}~\cite{suchanek2011paris,castano2006matching}.
+In contrast, the end-user programming community has pursued user-driven approaches, in which the user orchestrates the process of reconciliation at various levels of specificity.
+Systems that use this approach include ``mash-up makers'' (such as Mashmaker, \cite{mashmaker}, Marmite \cite{marmite}, and Vegemite \cite{vegemite}) and visual programming languages such as Yahoo Pipes \cite{pipes}.
+
+Our approach is most directly inspired by systems that, like the end-user programming approaches, allow users to easily perform data reconciliation using concrete instances as examples, but which, by generalising directly from examples, avoids the need for any sort of ``programming'' by the user.
+Prior systems in this space include ``Cards, Relations and Templates''\cite{Dontcheva:2007:RCS:1294211.1294224} and Potluck \cite{citeulike:3875264} which both allow users to reconcile instances from different schemas into a uniform representation using drag-and-drop gestures to specify relations among fields.
+Our system, \emph{DataPalette} extends this approach to a data workbench environment which aims to allow people to easily consolidate and compare instances by their properties.
+
+% algorithms that train on a corpus of alignments and produce new alignments on unseen data schemas or instances, respectively
+
+%A result of this problem is that users can experience data fragmentation~\cite{Jones05towardsa} across multiple web-based applications and services, and struggle to consolidate across them~\cite{bergman,boardmansasse}.
 In our own previously-conducted field study, \cite{infoscraps} we observed variations of manual ``coping strategies'' (or, if programmers, elaborate, custom-coded one-off solutions) to allow people to manage both online and offline data.
@@
 
-A number of research PIM systems have demonstrated how users benefit from unified, integrated data models.
-SEMEX's~\cite{semex} uniform query interface allows users to quickly trace references of people, places and things across files, folders and repositories.
-Similarly, Haystack  allows users to visualise and customise how they worked with their data in a uniform, consolidated manner \cite{haystack}.
+%We focus on \emph{instance-level} matching, as schemas may not always be available, or easily accessible to users.\looseness=-1
+
 % Pertaining to each system's ability to integrate information from arbitrary new data sources, however, both systems generally required new sources to either already conform to pre-specified ontologies, or have the user specify such a mapping manually.
@@
 % i think we do not need to differentiate between semantic web and database integration here, we're saying they're both bad, let's lump them together to keep the argument flowing and not kill our readers
-The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} research communities have primarily focused efforts on purely-automatic approaches to the problem of data integration.
-Such approaches typically take the form of \emph{ontology/schema-level matching}~\cite{euzenat2004api,doan2003learning}, or \emph{direct/instance-level} matching~\cite{suchanek2011paris,castano2006matching}.
-We focus on \emph{instance-level} matching, as schemas may not always be available, or easily accessible to users.\looseness=-1
 

### 2013-01-06 chi2013-webbox.tex
better rel work

@@
 The user experience benefits of conquering such data integration challenges have been demonstrated in PIM prototype systems with integrative data models.
-For example, the uniform query capabilities provided by SEMEX~\cite{semex} and \emph{Stuff I've Seen}\cite{stuffiveseen}, showed the speed, efficiency and re-findability gained when a user is able to quickly trace and cross-reference information about the people, places and things mentioned throughout his or her files, folders, e-mail, and other information repositories.
+For example, the uniform query capabilities provided by SEMEX~\cite{semex} and \emph{Stuff I've Seen}\cite{Dumais:2003:SIS:860435.860451}, showed the speed, efficiency and re-findability gained when a user is able to quickly trace and cross-reference information about the people, places and things mentioned throughout his or her files, folders, e-mail, and other information repositories.
 Similarly, Haystack \cite{haystack} demonstrated the reduced effort and increased efficiency of being able to view and re-use information easily in multiple task contexts without fragmentation or loss of consistency when a single, consolidated data model was used across applications.
@@
 In contrast, the end-user programming community has pursued user-driven approaches, in which the user orchestrates the process of reconciliation at various levels of specificity.
-Systems that use this approach include ``mash-up makers'' (such as Mashmaker, \cite{mashmaker}, Marmite \cite{marmite}, and Vegemite \cite{vegemite}) and visual programming languages such as Yahoo Pipes \cite{pipes}.
+Systems that use this approach include ``mash-up makers'' (such as Mashmaker, \cite{intel_mashmaker}, Marmite \cite{Lin:2009:EPM:1502650.1502667}, and Vegemite \cite{Wong:2007:MMM:1240624.1240842}) and visual programming languages such as Yahoo Pipes \cite{citeulike:8887891}.
 
@@
 \section{Pre-studies: Identifying Needs and Challenges}
-
 %% not sure where this goes >> but i want to keep it

### 2013-01-06 chi2013-webbox.tex
mum edits

@@
 \begin{abstract}
-The information processing capabilities of humans enable us to opportunistically draw and integrate knowledge from nearly any information sources that become available.
-However, enabling similar, opportunistic integration of digital, structured data from diverse sources has remained difficult without tedious, manual integration, due to problems of heterogeneity that arise when data modeled separately and differently are brought together.
-In this paper, we present an investigation of the feasibility of extending digital Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, towards enabling users to take advantage of the vast emerging ecosystems of structured data being made available.
-We conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources.
+The information processing capabilities of humans enable them to opportunistically draw and integrate knowledge from nearly any information source that becomes available.
+However, the integration of digital, structured data from diverse sources remains difficult, due to problems of heterogeneity that arise when data modeled separately are brought together.
+In this paper, we present an investigation of the feasibility of extending the scope of Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, with the objective of allowing  users to take advantage of the vast, emerging ecosystems of structured data currently becoming available.
+In this study, we conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources.
 These observations led to our design of \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
-Our lab study showed, first, that participants understood the new interaction mechanisms introduced, and, second, that they made more carefully-justified decisions weighing a greater number of factors, while spending less effort and taking the same amount of time, during subjective-choice tasks when using DataPalette than a control (web page and spreadsheet) interface.
+Our lab study showed, first, that participants readily understood the new interaction mechanisms which were introduced.
+Secondly, participants made more carefully justified decisions, even while weighing a greater number of factors, while expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up (web page and spreadsheet interface).
 %The information processing capabilities of humans enable people to opportunistically draw and integrate knowledge from nearly any information sources that become available to them.
@@
 
-Our approach is most directly inspired by systems that, like the end-user programming approaches, allow users to easily perform data reconciliation using concrete instances as examples, but which, by generalising directly from examples, avoids the need for any sort of ``programming'' by the user.
+Our approach is most similar to systems that, like the end-user programming approaches, allow users to easily perform data reconciliation using concrete instances as examples, but that avoid the need for any sort of ``programming'' by the user.
 Prior systems in this space include ``Cards, Relations and Templates''\cite{Dontcheva:2007:RCS:1294211.1294224} and Potluck \cite{citeulike:3875264} which both allow users to reconcile instances from different schemas into a uniform representation using drag-and-drop gestures to specify relations among fields.

### 2013-01-06 chi2013-webbox.tex
mum edits

@@
 \begin{abstract}
-The information processing capabilities of humans enable them to opportunistically draw and integrate knowledge from nearly any information source that becomes available.
+The information processing capabilities of humans enable them to opportunistically draw and integrate knowledge from nearly any information source.
 However, the integration of digital, structured data from diverse sources remains difficult, due to problems of heterogeneity that arise when data modeled separately are brought together.
@@
 Our lab study showed, first, that participants readily understood the new interaction mechanisms which were introduced.
-Secondly, participants made more carefully justified decisions, even while weighing a greater number of factors, while expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up (web page and spreadsheet interface).
+Secondly, participants made more carefully justified decisions, even while weighing a greater number of factors, moreover expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up (web page and spreadsheet interface).
 %The information processing capabilities of humans enable people to opportunistically draw and integrate knowledge from nearly any information sources that become available to them.
@@
 
-A common goal for the release of such data has been to provide end-user citizens an ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
-Thus far, however, this data has been predominately been consumed by app developers, journalists and other ``data specialists'' in one-off data analyses, apps and investigations, remaining still far from influencing the activities of ordinary end-user citizens.
-A core reason for this, we posit, is a lack of suitable tools for accessing and interrogating these heterogeneous data sets, and a means by which these disparate information sources can be effectively brought together and cross-referenced.
-For example, most personal information management (PIM) tools only provide facilities for managing small, pre-fixed sets of data types (such as digital calendaring tools and to-do list managers) or provide little or no support for structured data at all, such as text editors and sketching/drawing tools \cite{infoscraps}.
+A common goal for the release of such data has been to provide end-users the ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
+Thus far, however, this data has been predominately used by app developers, journalists and other data specialists in one-off data analyses, apps and investigations.
+Hence, as yet, these data still remain far from making a positive impact upon the activities of ordinary people.
+A core reason for this, we posit, is a lack of suitable tools for accessing and interrogating these heterogeneous data sources, as well as a means by which these disparate information sources can be effectively brought together and cross-referenced.
+For example, most personal information management (PIM) tools only provide the capacity to manage small, pre-fixed sets of data types (such as calendar appointments, address book entires,  tools and to-do list items).
+Alternatively, they may provide little or no support for structured data, as exhibited in word processors, text editors, and sketching/drawing tools \cite{infoscraps}.
 
 % We hypothesise that end-users will benefit from tools that allow them to effectively browse, use and consume heterogeneous structured data from diverse sources. 
-In this paper, we present an investigation into ways that PIM tools might be extended to allow data to be brought together in an ad-hoc fashion from arbitrary structured data sources, and interrogated, so as to enable end-user citizens to effectively harness knowledge from the vast, emerging ecosystems of data.
-We used a three-stage sequential exploratory mixed-method approach in this investigation: first, we performed a semi-structured interview pre-study to understand the types of tasks people performed using multiple information sources, and the processes that they relied upon to perform such tasks.
-Our results suggested that people routinely drew upon multiple and diverse sources for reasons including coverage, reliability, and to gain multiple perspectives or subjective assessments.
+In this paper, we present an investigation of ways that PIM tools might be extended to allow data to be brought together in an ad-hoc fashion from arbitrary structured data sources, so that end-users can effectively harness knowledge emerging from the vast ecosystems of data.
+For this purpose, we employed an exploratory, mixed-method approach with three sequential stages.
+First, in a pre-study, we conducted a series of semi-structured interviews in order to understand the various types of tasks people performed using multiple information sources, and the processes that they relied upon to perform such tasks.
+Our results suggested that people typically drew upon multiple, diverse sources, for a number of reasons, including increasing the breadth of information, assessing the reliability of information, and gaining multiple perspectives.
 

### 2013-01-07 chi2013-webbox.tex
mum edits

@@
 
-Second, we conducted a structural analysis of data records from personal data sources to identify the most common kinds of problems likely to arise from integrating data from these types of sources.
-We examined data sources in six common domains: contacts, events, music, shopping, social networking and weather.
-Our analysis found that while terminological heterogeneity was the most common issue for unifying simpler data records (such as address book contacts and social network profiles), structural differences were exhibit among more complex schemas, such as online retailers' product catalogues.
-
-Third, we developed DataPalette, an interface that enables serendipitous ``data mixing,'' negating the need for people to write bespoke code to effectively combine and compare heterogenous data.
-It was principally designed to support people's data integration needs for data heterogeneity, using observations made from the previous studies.
-In particular, we designed a user-centric interface to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
-We then performed a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we devised, and that this interface effectively improved people's ability to perform multifaceted decision making tasks using multiple sources.
+Second, we carried out a structural analysis of data records from personal data sources to identify the most common kinds of problems likely to arise from integrating data from these multiple sources.
+For this, we examined data sources from six common domains: contacts, events, music, shopping, social networking and weather.
+We discovered in our analysis that while terminological heterogeneity was the most common issue for unifying simpler data records (such as address book contacts and social network profiles), structural differences were exhibited among more complex schemas, such as online retailers' product catalogues.
+
+Third, we developed DataPalette, an interface that enables serendipitous ``data mixing,'' eliminating the need for people to write bespoke code to effectively combine and compare heterogeneous data.
+This interface was principally designed to support people's data integration needs for data heterogeneity, using observations made during the pre-studies.
+In particular, DataPalette was designed to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
+We then conducted a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we had devised, and that it effectively improved people's ability to perform multifaceted decision-making tasks using multiple sources.
 
@@
 The user experience benefits of conquering such data integration challenges have been demonstrated in PIM prototype systems with integrative data models.
-For example, the uniform query capabilities provided by SEMEX~\cite{semex} and \emph{Stuff I've Seen}\cite{Dumais:2003:SIS:860435.860451}, showed the speed, efficiency and re-findability gained when a user is able to quickly trace and cross-reference information about the people, places and things mentioned throughout his or her files, folders, e-mail, and other information repositories.
-Similarly, Haystack \cite{haystack} demonstrated the reduced effort and increased efficiency of being able to view and re-use information easily in multiple task contexts without fragmentation or loss of consistency when a single, consolidated data model was used across applications.
+For example, the uniform query capabilities provided by SEMEX~\cite{semex} and SIS \cite{Dumais:2003:SIS:860435.860451}, showed the speed, efficiency and re-findability gained when a user is able to quickly trace and cross-reference information about the people, places and things mentioned throughout his or her files, folders, e-mail, and other information repositories.
+Similarly, Haystack \cite{haystack} demonstrated the benefit of viewing and re-using information easily in multiple task contexts without fragmentation or loss of consistency when a single, consolidated data model was used across applications.
 
@@
 Prior systems in this space include ``Cards, Relations and Templates''\cite{Dontcheva:2007:RCS:1294211.1294224} and Potluck \cite{citeulike:3875264} which both allow users to reconcile instances from different schemas into a uniform representation using drag-and-drop gestures to specify relations among fields.
-Our system, \emph{DataPalette} extends this approach to a data workbench environment which aims to allow people to easily consolidate and compare instances by their properties.
+Our system, \emph{DataPalette} extends this approach to a data workbench environment which allows users to easily consolidate and compare instances by their properties.
 
@@
 
-\begin{table*}[htb]
-\small
-\begin{tabular}{p{2.0cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}  p{1.8cm}}
-				&				& {\bf average}				& {\bf average }				& 				& {\bf prop name}			& {\bf structural} 			& {\bf model } \\
-				& {\bf source}		& {\bf props/record}			& {\bf depth/record}			& {\bf equivalent to}	& {\bf inconsistent}			& {\bf inconsistency} 			& {\bf inconsistency} \\
-\hline
-social networking	& google+			& 12						& 1.52					& tw:5, li:4, fb:4		& tw:3, li:2, fb:2				& tw:0, li:1, fb:1				& tw:0, li:0, fb:0 \\
-(profiles)			& twitter			& 39						& 1.37					& li:4, fb:4 			& li:2, fb:1					& li:0, fb:0					& li:0, fb:0 \\
-				& linkedin			& 18						& 2.42					& fb:7			& fb:5					& fb:3					& fb:0 \\
-				& facebook		& 22						& 1.5						&				&						&						& \\
-\hline
-calendars			& google calendar	& 15						& 1.46					& eb:9			& eb:6					& eb:1					& eb:1 \\
-(events)			& eventbrite		& 32						& 1.43					&				&						& 						& \\
-\hline
-retail				& milo			& 5						& 1.2						& z:2, s:3, e:1, a:3	& z:2, s:2, e:1, a:3			& z:0, s:0, e:0, a:2			& z:0, s:0, e:0, a:0 \\
-(products)			& zappos			& 6						& 1						& s:2, e:2, a:3		& s:2, e:2, a:2				& s:0, e:0, a:0				& s:0, e:0, a:0 \\
-				& shopping.com	& 11						& 3.74					& e:2, a:8			& e:2, a:8					& e:0, a:0					& e:0, a:0 \\
-				& etsy			& 31						& 1.13					& a:11			& a:10					& a:1						& a:2 \\
-				& amazon			& 108					& 1.96					&				& 						&						& \\
-\hline
-contacts			& google contacts	& 7						& 1.46					& yp:2			& yp:2					& yp: 0					& yp: 0 \\
-				& yellowpages		& 25						& 1						&				&						&						& \\
-\hline
-music			& echonest		& 4						& 1.83					& so:3, r:3, sp:3		& so:3, r:3, sp:3				& so:0, r:0, sp:0				& so:0, r:0, sp:0 \\
-(songs/			& soundcloud		& 42						& 2.14					& r:12, sp:7		& r:11, sp:6 				& r:1, sp:1					& r:1, sp:0 \\
-tracks)			& rdio			& 27						& 2						& sp: 7			& sp: 6					& sp: 1					& sp: 1 \\
-				& spotify			& 10						& 2.44					&				&						&						& \\
-\hline
-weather			& weatherbug		& 8						& 2						& wu: 5, y: 5		& wu: 5, y: 5				& wu: 5, y: 4				& wu: 0, y: 0 \\
-(forecasts)		& wundergound	& 52						& 1.38					& y:12			& y:12					& y:7						& y:3 \\
-				& yahoo weather	& 22						& 1.58					&				&						&						& \\
-\end{tabular}
-\caption{\emph{PS.2 Analysis of Data Feeds}: Structural statistics about data sampled from each API organised by category.
-The first column lists the category and type of data record that was examined.
-Columns 3 and 4 list the average number of properties per data record, and number of levels from root to leaf.
-The remaining fields show results from a pairwise analysis of data records from each of the data sources with those of the others in each category; values are prefixed with the first letter(s) of the target schema the row schema was compared against.
-For example ``Tw:5'' in the 5th column of Google+ row indicates that Facebook and Twitter records had 5 equivalent fields.
-}\label{tbl:prestudy2}
-\end{table*}
-
 \subsection{PS.2 - Technical Challenges of Data Integration}
@@
 \subsubsection{Results}
-Table \ref{tbl:prestudy2} shows the raw results of the process described.
-A number of features were quite striking about the data; First, while most of the data records were small and relatively simple, nearly one in each category produced substantially more complex records than the rest.
-In particular, Amazon, Soundcloud, Twitter and Weather Underground stood out in their respective categories.
-Looking at equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
+While most of the data records were small and relatively simple, nearly one in each category provided access to substantially larger and more complex records than the rest.
+In particular, Amazon, Soundcloud, Twitter and Weather Underground were the most complex in each of their respective categories.
+Examining equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
 While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.

### 2013-01-07 chi2013-webbox.tex
pre background

@@
 In particular, Amazon, Soundcloud, Twitter and Weather Underground were the most complex in each of their respective categories.
-Examining equivalent properties, we found overall little overlap among records; that is, except for the smallest records (which had few properties to begin with), most of the data records had more unique properties than ones in common with others.
-While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent.
-Only Spotify, out of the 4 music providers, included information about the album featuring a given song.
-Such omissions show us that these data APIs reflect each provider's own specific needs, and suggests that integrating information from multiple sources may allow more complete and useful representations to be assembled.
+Mapping equivalent properties among the instance records for each category separately, revealed overall little overlap among records, except among the smallest records (which had few properties to begin with).
+The fact that most of the data records had more unique properties than properties that were redundant with records from other sources, suggests that consolidating such records would effectively, substantially increase the amount of information held by each.
+While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent in other sources' records.
+For example, among the 4 music data providers examined, only Spotify listed the source album that featured a particular song.
+Such omissions reveal that data APIs often reflect each provider's own specific needs, and that integrating information across multiple sources may often be necessary to create more complete and useful representations.
 
-The second finding is that equivalent properties were only rarely given the same name.
-Thus, terminological heterogeneity was prevalent across properties and records we observed.
-Structural heterogeneity, was slightly rarer; the most common case consisted of a leaf property from one data source modelled as a subtree in the other.
+The second finding is that equivalent properties were only rarely given the same name; thus, terminological (or naming) heterogeneity was prevalent across nearly all properties of records observed.
+Structural inconsistencies, were slightly rarer, the most common case of which occurred in approximately 1 in 4 records observed, consisted of a simple literal property from one data record expanded out as a sub-structure in the other.
 We also encountered two examples where data was nested in textual description fields, which hints of their intended use in feed readers rather than in structured data applications.

### 2013-01-07 chi2013-webbox.tex
past tense intro

@@
 In this study, we conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources.
-These observations led to our design of \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
+Observations from these pre-studies led to our designing \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
 Our lab study showed, first, that participants readily understood the new interaction mechanisms which were introduced.
@@
 
-A common goal for the release of such data has been to provide end-users the ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
+A common goal for the release of such data has been to provide end-users with the ability to make more informed decisions pertaining to their health, wealth, and well-being \cite{Shadbolt:2006:SWR:1155313.1155373}.
 Thus far, however, this data has been predominately used by app developers, journalists and other data specialists in one-off data analyses, apps and investigations.
@@
 \section{Background}
-In this section, we motivate our work using three fields: PIM, data integration, and end-user mashups and toolkits.
-Our primary motivation stems from field studies in PIM, which foregrounded work to consolidate data from multiple sources and which still remains a challenge.
-Specifically, reconciling the differences that arise among representations when data is created by different people, at different times, for different purposes.
+Our work was initiated by examining three fields of inquiry, comprising PIM, database integration, and end-user mashups and toolkits.
+Our primary motivation stemmed from field studies in PIM, where the need for consolidating data was observed.
+This was an area in which many challenges remained, however, specifically, in reconciling the differences that arose among representations when data was created by different people, at different times, and for different purposes.
 As described by Alon Halevy:
@@
 
-The user experience benefits of conquering such data integration challenges have been demonstrated in PIM prototype systems with integrative data models.
-For example, the uniform query capabilities provided by SEMEX~\cite{semex} and SIS \cite{Dumais:2003:SIS:860435.860451}, showed the speed, efficiency and re-findability gained when a user is able to quickly trace and cross-reference information about the people, places and things mentioned throughout his or her files, folders, e-mail, and other information repositories.
-Similarly, Haystack \cite{haystack} demonstrated the benefit of viewing and re-using information easily in multiple task contexts without fragmentation or loss of consistency when a single, consolidated data model was used across applications.
+The benefits to the end-user in conquering data integration were demonstrated in PIM prototype systems that employed integrative data models.
+For example, the uniform query capabilities provided by SEMEX~\cite{semex} and SIS \cite{Dumais:2003:SIS:860435.860451}, showed the speed, efficiency and re-findability gained when a user was able to quickly trace and cross-reference information about people, places and things mentioned throughout files, folders, e-mail, and other information repositories.
+Similarly, Haystack \cite{haystack} demonstrated that, when a single, consolidated data model was used across applications, the results included increased efficiency, the elimination of information fragmentation, and reduced effort, with the user being able to view and re-use information easily in multiple task contexts.
 
-Therefore, despite the multifaceted challenges of data integration, various fields have sought methods to resolve them.
-The database integration and Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373} fields, for example, have pursued automatic, recently machine-learning approaches to data integration, at either the level of \emph{ontology/schema-level matching}~\cite{euzenat2004api,doan2003learning}, or \emph{instance matching}~\cite{suchanek2011paris,castano2006matching}.
-In contrast, the end-user programming community has pursued user-driven approaches, in which the user orchestrates the process of reconciliation at various levels of specificity.
-Systems that use this approach include ``mash-up makers'' (such as Mashmaker, \cite{intel_mashmaker}, Marmite \cite{Lin:2009:EPM:1502650.1502667}, and Vegemite \cite{Wong:2007:MMM:1240624.1240842}) and visual programming languages such as Yahoo Pipes \cite{citeulike:8887891}.
+Because of the multifaceted challenges of data integration, contrasting approaches were investigated in different research communities.
+In the fields of database integration and the Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373}, for example, automatic, machine-learning approaches, at either the level of \emph{ontology/schema-level matching}~\cite{euzenat2004api,doan2003learning}, or \emph{instance matching}~\cite{suchanek2011paris,castano2006matching} were pursued.
+In contrast, the end-user programming community preferred user-driven approaches, in which the user orchestrated the process of reconciliation at various levels of specificity.
+Systems that used this approach included ``mash-up makers'' (such as Mashmaker, \cite{intel_mashmaker}, Marmite \cite{Lin:2009:EPM:1502650.1502667}, and Vegemite \cite{Wong:2007:MMM:1240624.1240842}) and visual programming language environments such as Yahoo Pipes \cite{citeulike:8887891}.
 
-Our approach is most similar to systems that, like the end-user programming approaches, allow users to easily perform data reconciliation using concrete instances as examples, but that avoid the need for any sort of ``programming'' by the user.
-Prior systems in this space include ``Cards, Relations and Templates''\cite{Dontcheva:2007:RCS:1294211.1294224} and Potluck \cite{citeulike:3875264} which both allow users to reconcile instances from different schemas into a uniform representation using drag-and-drop gestures to specify relations among fields.
-Our system, \emph{DataPalette} extends this approach to a data workbench environment which allows users to easily consolidate and compare instances by their properties.
+With DataPalette, our approach bore similarities with the end-user programming approach in that the user could easily perform data reconciliation using concrete instances as examples, but also avoided the need for any sort of programming by the user.
+Previous systems utilising this approach included ``Cards, Relations and Templates'' \cite{Dontcheva:2007:RCS:1294211.1294224} and Potluck \cite{citeulike:3875264} both of which permitted the user to reconcile instances from different schemas into a uniform representation using drag-and-drop gestures to specify relations among fields.
+\emph{DataPalette} extended this approach to a data workbench environment which allowed people to easily consolidate and compare instances by their properties.
 

### 2013-01-07 chi2013-webbox.tex
imperfect

@@
 In this paper, we present an investigation of ways that PIM tools might be extended to allow data to be brought together in an ad-hoc fashion from arbitrary structured data sources, so that end-users can effectively harness knowledge emerging from the vast ecosystems of data.
-For this purpose, we employed an exploratory, mixed-method approach with three sequential stages.
-First, in a pre-study, we conducted a series of semi-structured interviews in order to understand the various types of tasks people performed using multiple information sources, and the processes that they relied upon to perform such tasks.
-Our results suggested that people typically drew upon multiple, diverse sources, for a number of reasons, including increasing the breadth of information, assessing the reliability of information, and gaining multiple perspectives.
+For this purpose, we have employed an exploratory, mixed-method approach with three sequential stages.
+First, in a pre-study, we have conducted a series of semi-structured interviews in order to understand the various types of tasks people perform using multiple information sources, and the processes that they rely upon to perform such tasks.
+Our results suggest that people typically draw upon multiple, diverse sources for a number of reasons, including increasing the breadth of information, assessing the reliability of information, and gaining multiple perspectives.
 
@@
 
-Second, we carried out a structural analysis of data records from personal data sources to identify the most common kinds of problems likely to arise from integrating data from these multiple sources.
-For this, we examined data sources from six common domains: contacts, events, music, shopping, social networking and weather.
-We discovered in our analysis that while terminological heterogeneity was the most common issue for unifying simpler data records (such as address book contacts and social network profiles), structural differences were exhibited among more complex schemas, such as online retailers' product catalogues.
+Second, we have carried out a structural analysis of data records from personal data sources to identify the most common kinds of problems likely to arise from integrating data from these multiple sources.
+For this, we have examined data sources from six common domains: contacts, events, music, shopping, social networking and weather.
+We have discovered, in our analysis, that while terminological heterogeneity is the most common issue for unifying simpler data records (such as address book contacts and social network profiles), structural differences are exhibited among more complex schemas, such as online retailers' product catalogues.
 
-Third, we developed DataPalette, an interface that enables serendipitous ``data mixing,'' eliminating the need for people to write bespoke code to effectively combine and compare heterogeneous data.
-This interface was principally designed to support people's data integration needs for data heterogeneity, using observations made during the pre-studies.
-In particular, DataPalette was designed to facilitate simple, ``light-touch'' integration tasks of diverse, heterogeneous data sources.
-We then conducted a usability evaluation of DataPalette, which revealed that most users were comfortable with the interactive integration mechanisms we had devised, and that it effectively improved people's ability to perform multifaceted decision-making tasks using multiple sources.
+Third, we have developed DataPalette, an interface that enables serendipitous ``data mixing,'' eliminating the need for people to write bespoke code to effectively combine and compare heterogeneous data.
+DataPalette facilitates basic integration of diverse, heterogeneous data sources using simple interface gestures.
+A usability evaluation of DataPalette has revealed that most users are comfortable with its interactive integration mechanisms, and that it effectively improves people's ability to perform multifaceted decision-making tasks using multiple sources.
 
 \section{Background}
-Our work was initiated by examining three fields of inquiry, comprising PIM, database integration, and end-user mashups and toolkits.
-Our primary motivation stemmed from field studies in PIM, where the need for consolidating data was observed.
-This was an area in which many challenges remained, however, specifically, in reconciling the differences that arose among representations when data was created by different people, at different times, and for different purposes.
+Our work has been informed by three fields of inquiry, comprising PIM, database integration, and end-user mashups and toolkits.
+Our primary motivation stems from field studies in PIM, where the need for consolidating data has been observed.
+This is an area in which many challenges remain, however, specifically, in reconciling the differences that arise among representations when data is created by different people, at different times, and for different purposes.
 As described by Alon Halevy:
@@
 
-The benefits to the end-user in conquering data integration were demonstrated in PIM prototype systems that employed integrative data models.
-For example, the uniform query capabilities provided by SEMEX~\cite{semex} and SIS \cite{Dumais:2003:SIS:860435.860451}, showed the speed, efficiency and re-findability gained when a user was able to quickly trace and cross-reference information about people, places and things mentioned throughout files, folders, e-mail, and other information repositories.
-Similarly, Haystack \cite{haystack} demonstrated that, when a single, consolidated data model was used across applications, the results included increased efficiency, the elimination of information fragmentation, and reduced effort, with the user being able to view and re-use information easily in multiple task contexts.
+The benefits to the end-user in conquering data integration have been demonstrated in PIM prototype systems that employ integrative data models.
+For example, the uniform query capabilities provided by SEMEX~\cite{semex} and SIS \cite{Dumais:2003:SIS:860435.860451}, show the speed, efficiency and re-findability gained when a user is able to quickly trace and cross-reference information about people, places and things mentioned throughout files, folders, e-mail, and other information repositories.
+Similarly, Haystack \cite{haystack} demonstrates that, when a single, consolidated data model is used across applications, the results include increased efficiency, the elimination of information fragmentation, and reduced effort, with the user being able to view and re-use information easily in multiple task contexts.
 
-Because of the multifaceted challenges of data integration, contrasting approaches were investigated in different research communities.
-In the fields of database integration and the Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373}, for example, automatic, machine-learning approaches, at either the level of \emph{ontology/schema-level matching}~\cite{euzenat2004api,doan2003learning}, or \emph{instance matching}~\cite{suchanek2011paris,castano2006matching} were pursued.
-In contrast, the end-user programming community preferred user-driven approaches, in which the user orchestrated the process of reconciliation at various levels of specificity.
-Systems that used this approach included ``mash-up makers'' (such as Mashmaker, \cite{intel_mashmaker}, Marmite \cite{Lin:2009:EPM:1502650.1502667}, and Vegemite \cite{Wong:2007:MMM:1240624.1240842}) and visual programming language environments such as Yahoo Pipes \cite{citeulike:8887891}.
+Because of the multifaceted challenges of data integration, contrasting approaches have been investigated in different research communities.
+In the fields of database integration and the Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373}, for example, automatic, machine-learning approaches, at either the level of \emph{ontology/schema matching}~\cite{euzenat2004api,doan2003learning}, or \emph{instance matching}~\cite{suchanek2011paris,castano2006matching} have been pursued.
+In contrast, the end-user programming community has preferred user-driven approaches, in which the user orchestrates the process of reconciliation at various levels of specificity.
+Systems that use this approach include ``mash-up makers'' (such as Mashmaker, \cite{intel_mashmaker}, Marmite \cite{Lin:2009:EPM:1502650.1502667}, and Vegemite \cite{Wong:2007:MMM:1240624.1240842}) and visual programming language environments such as Yahoo Pipes \cite{citeulike:8887891}.
 
-With DataPalette, our approach bore similarities with the end-user programming approach in that the user could easily perform data reconciliation using concrete instances as examples, but also avoided the need for any sort of programming by the user.
-Previous systems utilising this approach included ``Cards, Relations and Templates'' \cite{Dontcheva:2007:RCS:1294211.1294224} and Potluck \cite{citeulike:3875264} both of which permitted the user to reconcile instances from different schemas into a uniform representation using drag-and-drop gestures to specify relations among fields.
-\emph{DataPalette} extended this approach to a data workbench environment which allowed people to easily consolidate and compare instances by their properties.
+DataPalette's approach exhibits similarities with the user-driven methods just mentioned, in that the user performs data reconciliation using concrete instances as examples, but also avoids the need for any sort of programming.
+Previous systems utilising this approach have included ``Cards, Relations and Templates'' \cite{Dontcheva:2007:RCS:1294211.1294224} and Potluck \cite{citeulike:3875264} both of which permit the user to reconcile instances from different schemas into a uniform representation using drag-and-drop gestures to specify relations among fields.
+\emph{DataPalette} extends this approach to a data workbench environment which allows people to easily consolidate and compare instances by their properties.
 

### 2013-01-07 chi2013-webbox.tex
through ps1

@@
 
-We performed an exploratory sequential mixed-method study to help us design our contribution.
-For our first pre-study, we sought an updated understanding on the use of multiple information sources in information gathering on the Web.
-There were two driving questions behind this inquiry: first, were people increasingly relying only on a handful of ``supersites'' (Facebook, Wikipedia), or searching across distributed sources of information?
-If the former were true, merely integrating PIM tools with the small number of supersites would provide greater benefit than tackling the more challenging problem of integrating data from arbitrary sources on the Web.
-Second, why did people choose a single or multiple sources for information gathering?
-If a number of websites were used, why was this preferred over a single source?
+In the first of two pre-studies, we aimed to get an updated understanding of the use of multiple information sources in information gathering on the Web.
+Several basic questions drove our inquiry.
+First, were people increasingly relying only on a handful of ``supersites'' (Facebook, Wikipedia), or searching across distributed sources of information?
+If the former were true, merely integrating PIM tools with the small number of supersites would provide greater benefit than tackling the more challenging problem of integrating data from an array of arbitrary sources on the Web.
+In a related vein, why did some people choose a single source for information gathering, while others selected multiple sources?
+What advantages were gained in consulting multiple sources?
 
@@
 
-The second pre-study focused on the characteristics of the data available from the kinds of sites used by the first pre-study's participants for the tasks described.
-The purpose of this follow-on was not to identify specific characteristics of particular sites in question, but to identify, in general, the degree of complexity of the data feeds currently available across a variety of domains, and typical integration problems that might arise when mixing this data.
-In the next two sections, we present our methodologies and results for our pre-studies.
+The second pre-study focused on the characteristics of the data available from the selected sites.
+The purpose of this exercise was not to identify specific characteristics of particular sites, but to establish general characteristics across a variety of domains, so that typical integration problems that might arise when mixing these data could be identified.
+In the next two sections, we present our methodologies and the results of our pre-studies.
 
@@
 
-For the first pre-study, we conducted a semi-structured interview to understand the reasons and ways that people drew information from diverse information sources.
-We interviewed 8 participants, asking them to identify tasks they performed recently that required use of multiple information sources, the kinds of sources used, and the tools used to manage the resulting information.
-We then inquired as to how each participant would go about planning various aspects of a large party, such as date, location and entertainment.
+In the first pre-study, we held semi-structured interviews to better understand the reasons that people drew information from single or diverse information sources.
+We interviewed 8 participants, asking them to identify tasks they had performed recently, which had required the use of multiple information sources.
+This was followed by determining the kinds of sources which were used, and the tools that were needed to manage the resulting information.
+We then inquired how each participant would go about planning a hypothetical social event entailing details such as scheduling a  date, finding a suitable location and selecting the appropriate entertainment.
 

### 2013-01-09 chi2013-webbox.tex
more improvements

@@
 
-For each feed, 3--5 ``typical'' records of a single type were obtained; for social networking sites this was user profiles, for retail sites this was product info, etc.
+For each feed, 3--5 typical records of a particular type were obtained from each service's API or feed.
+For social networking sites, user profiles were the type of record examined, while for retail sites, event tracking sites, music sites and weather sites, product information, event time/date info, song listing information, and weather forecasts were examined, respectively.
 %Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.  
-For each such data record, two simple complexity metrics were computed: the \emph{average width} and \emph{average depth}.
-Determining the width required first finding the appropriate level for comparison; for example, Amazon's retail product ``Item'' root record had few attributes about the actual product compared to its ``ItemAttributes'' sub-property, which was used instead.
-The average depth was simply the mean depth to each leaf node from the chosen comparison level.
-% This step was done manually, with an effort to select most similar components for comparison.  
-% Depth was computed as the average number of levels from record root to leaves across all properties, and was measured, with width, to convey the overall size/complexity of each record.
-
+For each such data record, two complexity metrics were computed: the \emph{average width}, corresponding to the number of properties, and \emph{average depth}, corresponding to the number of nested structures, of each record.
 The third metric measured degree of overlap among sources in each category.
@@
 
+%Determining the width required first finding the appropriate level for comparison; for example, for Amazon's retail product ``Item'' records, the sub-structure under the ``ItemAttributes'' property was used instead of the root-level record.
+The average depth was simply the mean depth to each leaf node from the chosen comparison level.
+% This step was done manually, with an effort to select most similar components for comparison.  
+% Depth was computed as the average number of levels from record root to leaves across all properties, and was measured, with width, to convey the overall size/complexity of each record.
+
 Once established, equivalent properties were examined firstly for naming inconsistencies.
-Then, ignoring property names, we examined whether the property's values were structurally compatible.
-If both values consisted of the data type and same equivalent subfields, they were considered comparable; otherwise they were considered \emph{structurally inconsistent}.
-Finally, we examined equivalent fields which were incomparable for reasons other than structural or terminological differences.
+Then, ignoring names, we examined whether the property's values were structurally compatible, defined as follows: if both values consisted of the same literal data type or had one or more equivalent subfields, they were considered compatible.
+Finally, we examined incompatibilities not covered by naming or structural differences alone.
 These \emph{modelling differences} arose from a variety of causes, including, but not limited to: measurement unit inconsistencies, measurement method (e.g., dew point vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').\looseness=-1
@@
 \subsubsection{Results}
-While most of the data records were small and relatively simple, nearly one in each category provided access to substantially larger and more complex records than the rest.
+While most of the data records were small and relatively simple, nearly one in each category was substantially larger and more complex.
 In particular, Amazon, Soundcloud, Twitter and Weather Underground were the most complex in each of their respective categories.
-Mapping equivalent properties among the instance records for each category separately, revealed overall little overlap among records, except among the smallest records (which had few properties to begin with).
-The fact that most of the data records had more unique properties than properties that were redundant with records from other sources, suggests that consolidating such records would effectively, substantially increase the amount of information held by each.
+Mapping equivalent properties among the instance records for each category revealed little overall overlap among records, except among the smallest records (which had few properties to begin with).
+This finding suggests that consolidating such records would effectively substantially increase the amount of information held by each.
 While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent in other sources' records.
@@
 
-The second finding is that equivalent properties were only rarely given the same name; thus, terminological (or naming) heterogeneity was prevalent across nearly all properties of records observed.
-Structural inconsistencies, were slightly rarer, the most common case of which occurred in approximately 1 in 4 records observed, consisted of a simple literal property from one data record expanded out as a sub-structure in the other.
-We also encountered two examples where data was nested in textual description fields, which hints of their intended use in feed readers rather than in structured data applications.
-Examples of modelling differences were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and differences in units of measurement.
+The second finding is that equivalent properties were only rarely given the same name; thus, terminological (or naming) heterogeneity was prevalent across nearly all records observed.
+Structural inconsistencies were rarer, with the most common case occuring where a record that represented a value as a simple literal was expanded out in another as an entire sub-structure.
+This most commonly occurred for fields with common string serialisations, such as dates, prices, and intervals.
+Meanwhile, examples of modelling differences were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and differences in units of measurement.
+% We also encountered two examples of records in which multiple data fields were nested as a single literal, textual ``description'' field.
+This behaviour stems from the intended use of such records only by human readers, rather than in structured data applications.
 
-We also examined the source schemas of data sources for enumerated value types.
-There was substantial terminological inconsistency among such value ranges of such types, but also cases where sources had values not represented in the other source's ranges.
-For example, ``gender''  on Facebook allowed only ``male'' and ``female'' for values, while Google+ included a third option,``other''.
-Few string representations of enumerated value sets matched across data sources perfectly.
+To examine value consistency, we compared source schemas provided in the API and feed documentation of the data providers.
+There was substantial terminological inconsistency among the values names for enumerated types, with exact matches occurring less than 10\% of examined value pairs.
+There were also range omissions, corresponding to cases where values represented in one record were not represented as valid values for the equivalent property in the target schema.
+A simple example of this was the ``gender'' property  on Facebook, which allowed only ``male'' and ``female'' values, while Google+ included a third option,``other''.
 
@@
 At a high level, the overall goal of DataPalette was to empower end-users to mix arbitrary structured data feeds on-the-fly while solving their particular information task(s) --- be them sense-making or decision-making related --- in order to more effectively accomplish these tasks.
-Unlike Mashup makers, including Yahoo Pipes!, the goal was to facilitate such integration without a need for an explicit separate mapping or integration step.
-Instead, DataPalette's approach is most directly inspired by ``lightweight'' visual, user-interactive data collection, consolidation and alignment tools such as Potluck \cite{citeulike:3875264} and Cards \cite{Dontcheva:2007:RCS:1294211.1294224}.
+Unlike Mashup makers, including Yahoo Pipes!, the goal was to facilitate such integration without a need for an explicit separate scraping, shaping,, mapping or integration step.
+
+% Instead, DataPalette's approach is most directly inspired by ``lightweight'' visual, user-interactive data collection, consolidation and alignment tools such as Potluck \cite{citeulike:3875264} and Cards \cite{Dontcheva:2007:RCS:1294211.1294224}.
 
@@
 
-In order to feel familiar to new users, we borrowed the overall  ``look and feel'' from OS X Finder-style file managers.
+In order to feel familiar to new users, we started with the look and feel from OS X Finder-style file managers.
 In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can freely drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.

### 2013-01-09 chi2013-webbox.tex
more improvements

@@
 
-In order to determine whether the interaction affordances introduced in DataPalette (henceforth DP) allow users to perform serendipitous data mixing on real data feeds by end-users, we devised a within-subjects A-B trial lab study which we describe next.
+In order to determine whether the interaction affordances introduced in DataPalette (henceforth DP) allow users to perform serendipitous data mixing on real data feeds by end-users, we devised a within-subjects controlled lab study which we describe next.
 
@@
 \subsubsection{Study and task design}
-To test these hypotheses, we designed a within-subjects (repeated measures) lab-based A-B study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (condition A) and once with a baseline interface (condition B).
+To test these hypotheses, we designed a within-subjects (repeated measures) controlled lab study in which participants were asked to perform predefined subjective-choice tasks in two separate timed trials, once with the DP interface (condition A) and once with a baseline interface (condition B).
 Each task was paired with one of two datasets, depending on condition, which were counterbalanced in a full-factorial design to eliminate potential ordering, task, and dataset biases.\looseness=-1

### 2013-01-09 chi2013-webbox.tex
latest with mum

@@
 DataPalette facilitates basic integration of diverse, heterogeneous data sources using simple interface gestures.
-A usability evaluation of DataPalette has revealed that most users are comfortable with its interactive integration mechanisms, and that it effectively improves people's ability to perform multifaceted decision-making tasks using multiple sources.
+A usability evaluation of DataPalette has revealed that most users are comfortable with its interactive integration mechanisms, and that it effectively improves people's abilities to perform multi-faceted decision-making tasks using multiple sources.
 
@@
 
-Because of the multifaceted challenges of data integration, contrasting approaches have been investigated in different research communities.
+Because of the multi-faceted challenges of data integration, contrasting approaches have been investigated in different research communities.
 In the fields of database integration and the Semantic Web \cite{Shadbolt:2006:SWR:1155313.1155373}, for example, automatic, machine-learning approaches, at either the level of \emph{ontology/schema matching}~\cite{euzenat2004api,doan2003learning}, or \emph{instance matching}~\cite{suchanek2011paris,castano2006matching} have been pursued.
@@
 %\subsubsection{Question 1: Heterogenous Data Tasks}
-All participants reported regular use of multiple websites to accomplish tasks, examples of which including: shopping (food, consumer electronics, hotels and flights), choosing restaurants, job-seeking, selecting a university, seeking recipes, and finding answers to technical problems.
-When beginning a task, it was common for participants to use Google to discover several related websites, which would then be used in different ways depending on the task at hand.
-For shopping-oriented tasks, balancing price, quality and speed of delivery was the main aim.
-For product reviews, single sources were not trusted due to possible bias and incomplete feature coverage.
+All participants reported regular usage of multiple websites to accomplish tasks, examples of which included: shopping (food, consumer electronics, hotels and flights), choosing a restaurant, job-seeking, selecting a university, seeking a recipe, or finding answers to technical problems.
+When initiating a task, it was common for participants to use Google to discover several related websites, which would then be explored in different ways, depending on the task at hand.
+For shopping-oriented tasks, a balance of price, quality of merchandise, and speed of delivery was sought.
+For product reviews, single sources alone were not trusted due to possible biases, or incomplete coverage of products' desired features.
 % Jim - Not sure if this is relevant. If it is, put it back in.
@@
 One participant said that they shared their outlook and gmail calendars, but only for work events and did not record social activities on them.
-To provide a full coverage of information, websites of different types were used, such as manufacturers websites for technical details, review aggregators for a range of opinions, and Google maps for location-based decisions.
+To benefit from the fullest coverage of information, many different websites of different types were consulted, such as manufacturers' websites for technical details, and review aggregates for a range of opinions.
 
@@
 
-Participants said they would organise a large social event by first conferring with friends about their ideas, preferences and recommendations.
-They selected venues, locations, restaurants and activities through a number of sources: their own prior experience, friends' recommendations, searching Google maps, dedicated reviews sites (such as Yelp for restaurants), and through web search.
-People stated that the cost, price and easy to get to locations were the most important factors when choosing a venue.
-They considered how to get there and looked up train or bus times on the Web, or organised carpools.
+In response to our question of how participants would plan a large social event, most stated they would first confer with friends to get their ideas, preferences and recommendations.
+Then, they would select the venue, locations, restaurants and activities through a number of sources: their own prior experience, friends' recommendations, searching Google maps, and dedicated reviews sites (such as Yelp for restaurants).
+The participants cited the cost and ease of accessing a location as the most important factors when choosing the venue.
+% They considered how to get there and looked up train or bus times on the Web, or organised carpools.  
 %When organising an event, the weather was not that important because they did not trust the accuracy of predictions.
@@
 
-All users experienced using multiple web sites in order to complete a task, and felt that the lack of existing integration between sites hindered their ability to make informed choices.
-Participants felt that they did not include particular aspects of a decision because the process to investigate them would be too difficult or take too much time.
+All of the participants used multiple web sites in order to complete their tasks, and all felt that the lack of existing integration between sites hindered their ability to make informed choices.
+Participants prioritised the sources they used by both the ease and immediacy of accessing them.
+Conversely, sources that took much effort or time to investigate were often excluded.
+
+% details of their decision making process were excluded because participants felt it was too much effort to investigate them.
+% Participants felt that they did not include particular aspects of a decision because the process to investigate them would be too difficult or take too much time.
 
@@
 
-We received 26 responses to our call for participants, from which we selected the first 20 (ten of which were female).
+We received 26 responses to our call, from which we selected the first 20 (ten of whom were female).
 Seven were non-students, comprised of university staff and alumni; the remainder were students, 8 studying computer science, and 5 studying other subjects.
-Due to the large population of international students and staff at the University, we ended up getting a large number of non-native English speakers.
-However, since all students and staff had passed requisite proficiency tests, we did not think this would substantially impact results.
+Due to the large population of international students and staff at our university, we ended up with a large sample (12) of non-native English speakers; however, since all had passed requisite proficiency tests, we did not think this would substantially impact results.
 On average, each study took 40 minutes to complete.

### 2013-01-09 chi2013-webbox.tex
mum edits ps

@@
 Participants prioritised the sources they used by both the ease and immediacy of accessing them.
-Conversely, sources that took much effort or time to investigate were often excluded.
+As a result, sources that took much effort or time to investigate were often excluded.
 
@@
 We consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb: \url{http://www.programmableweb.com/apis/directory}} list of most popular data feeds for this purpose.
-To increase the variety of sources, we selected from 5 categories: social network services, retailers, online event calendars, music sites, and weather, selecting 2--5 sources from each, for a total of 20.
+To examine the many varieties of personal data sources, we selected from 5 categories: social network services, retailers, online event calendars, music sites, and weather, selecting 2--5 sources from each, for a total of 20.
 
-For each feed, 3--5 typical records of a particular type were obtained from each service's API or feed.
-For social networking sites, user profiles were the type of record examined, while for retail sites, event tracking sites, music sites and weather sites, product information, event time/date info, song listing information, and weather forecasts were examined, respectively.
+For each source, 3--5 typical records of a particular type were obtained from each service's API or feed.
+For social networking sites, user profiles were the type of record examined, while for retail sites, it was product information; for event tracking sites, we focused on event time/date information; for music sites, on song listing information; and for weather sites, forecast records were selected for our study.
 %Records were located either by first querying for popular key terms using each site's keyword query API (if available), or a ``top products'' query if available.  
-For each such data record, two complexity metrics were computed: the \emph{average width}, corresponding to the number of properties, and \emph{average depth}, corresponding to the number of nested structures, of each record.
-The third metric measured degree of overlap among sources in each category.
-This was done by first mapping \emph{equivalent properties} between schemas, which was done manually by creating alignment tables for all properties of data sources in each category.
+For each such data record, two complexity metrics were computed: the \emph{average width}, corresponding to the number of properties, and the \emph{average depth}, corresponding to the number of nested structures.
+In addition, a third metric measured degree of overlap among the sources in each category.
+This was accomplished by first mapping \emph{equivalent properties} between schemas, performed manually by creating alignment tables for all properties of data sources in each category.
 Each row comprised a property of a record of one data source, and all its closest matching properties from the others.
-We used both property names and attribute descriptions (in documentation, if available) to make matching decisions.
+We used both property names and attribute descriptions (in documentation, if available) to determine which properties were semantically equivalent.
 
@@
 
-Once established, equivalent properties were examined firstly for naming inconsistencies.
-Then, ignoring names, we examined whether the property's values were structurally compatible, defined as follows: if both values consisted of the same literal data type or had one or more equivalent subfields, they were considered compatible.
-Finally, we examined incompatibilities not covered by naming or structural differences alone.
-These \emph{modelling differences} arose from a variety of causes, including, but not limited to: measurement unit inconsistencies, measurement method (e.g., dew point vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').\looseness=-1
+Once established, equivalent properties were examined, first, for naming inconsistencies.
+Then, disregarding names, we examined whether the property's values were structurally compatible.
+That is, if both values consisted of the same literal data type, or had one or more equivalent subfields, they were considered compatible.
+Finally, we examined incompatibilities not covered by either naming or structural differences alone.
+These \emph{modelling differences} arose from a variety of reasons, including, but not limited to, measurement unit inconsistencies, measurement method disagreements (e.g., dew point vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').\looseness=-1
 

### 2013-01-10 chi2013-webbox.tex
mum edits ps2

@@
 \subsection{PS.2 - Technical Challenges of Data Integration}
-Our analysis of structured data feeds started with identifying a set of candidate sources to examine.
+For Pre-Study 2, our analysis of structured data feeds started with identifying a set of candidate sources to examine.
 We consulted the ProgrammableWeb API directory\footnote{ProgrammableWeb: \url{http://www.programmableweb.com/apis/directory}} list of most popular data feeds for this purpose.
@@
 \subsubsection{Results}
-While most of the data records were small and relatively simple, nearly one in each category was substantially larger and more complex.
+While most of the data records examined in PS.2 were small and relatively simple, nearly one in each category was substantially larger and more complex.
 In particular, Amazon, Soundcloud, Twitter and Weather Underground were the most complex in each of their respective categories.
-Mapping equivalent properties among the instance records for each category revealed little overall overlap among records, except among the smallest records (which had few properties to begin with).
-This finding suggests that consolidating such records would effectively substantially increase the amount of information held by each.
-While some of these `unique records' were primarily for internal, service-specific use (such as service-specific IDs) others were useful properties that were simply absent in other sources' records.
+Mapping equivalent properties among the instance records for each category revealed little overall overlap, except among the smallest records (which had few properties to begin with).
+This finding suggests that consolidating such records would substantially increase the amount of information held by each.
+Some of the unique properties were primarily for internal use (such as service-specific IDs).
+Some, however, contained useful properties which were simply absent in alternate sources.
 For example, among the 4 music data providers examined, only Spotify listed the source album that featured a particular song.
-Such omissions reveal that data APIs often reflect each provider's own specific needs, and that integrating information across multiple sources may often be necessary to create more complete and useful representations.
+Such omissions revealed that data APIs often were reflective of each provider's own specific needs, necessitating the integration of information across multiple sources to create more complete and useful representations.
 
-The second finding is that equivalent properties were only rarely given the same name; thus, terminological (or naming) heterogeneity was prevalent across nearly all records observed.
+A second observation is that equivalent properties were only rarely given the same name; thus, terminological (or naming) heterogeneity was prevalent across nearly all records observed.
 Structural inconsistencies were rarer, with the most common case occuring where a record that represented a value as a simple literal was expanded out in another as an entire sub-structure.

### 2013-01-10 chi2013-webbox.tex
mum edits ps2

@@
 
-A second observation is that equivalent properties were only rarely given the same name; thus, terminological (or naming) heterogeneity was prevalent across nearly all records observed.
+A second observation in PS.2 was that equivalent properties were only rarely given the same name.
+Thus, terminological (or naming) heterogeneity was prevalent across nearly all records observed.
 Structural inconsistencies were rarer, with the most common case occuring where a record that represented a value as a simple literal was expanded out in another as an entire sub-structure.
 This most commonly occurred for fields with common string serialisations, such as dates, prices, and intervals.
-Meanwhile, examples of modelling differences were rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts on one site not mapping perfectly to those of another site, and differences in units of measurement.
+Meanwhile, examples of modelling differences were even rarer than those of structural heterogeneity, with the few we found mostly being attributed to concepts from one source not mapping perfectly to those of another source, for example, due to differences in methods of measurement, or measurement units.
 % We also encountered two examples of records in which multiple data fields were nested as a single literal, textual ``description'' field.
@@
 To examine value consistency, we compared source schemas provided in the API and feed documentation of the data providers.
-There was substantial terminological inconsistency among the values names for enumerated types, with exact matches occurring less than 10\% of examined value pairs.
-There were also range omissions, corresponding to cases where values represented in one record were not represented as valid values for the equivalent property in the target schema.
+There was substantial terminological inconsistency among the value names for enumerated types, with exact matches occurring less than 10\% of examined value pairs.
+There were also range inconsistencies, corresponding to cases where values represented in one record were not represented as valid values in the range of the equivalent property in the target schema.
 A simple example of this was the ``gender'' property  on Facebook, which allowed only ``male'' and ``female'' values, while Google+ included a third option,``other''.
 
-In summary, while data records ranged in complexity, no source subsumed the rest; all contained unique properties.
+In summary, we found in our PS.2, that while data records ranged in complexity, no single source subsumed any others; all contained unique properties.
 Moreover, the most common types of heterogeneity exhibited were terminological and value-related, which occurred vastly more than structural or modeling issues.
@@
 
-At a high level, the overall goal of DataPalette was to empower end-users to mix arbitrary structured data feeds on-the-fly while solving their particular information task(s) --- be them sense-making or decision-making related --- in order to more effectively accomplish these tasks.
-Unlike Mashup makers, including Yahoo Pipes!, the goal was to facilitate such integration without a need for an explicit separate scraping, shaping,, mapping or integration step.
+The primary goal of DataPalette was to empower end-users so that they are able to mix arbitrary, structured data feeds on-the-fly while solving particular information task(s) more effectively.
+Unlike mashup makers, such as Yahoo Pipes!, the aim was to facilitate such integration without the need for a separate step involving programming, scraping, shaping, mapping or integration.
 
@@
 
-The two pre-studies provided valuable insight towards this goal that allowed us to guide our design process.
-PS.1 confirmed the need for data integration by showing that people preferred to rely upon multiple sources of information for making important decisions, and identified the kinds of sources most commonly used.
-PS.2, meanwhile, revealed that the most common kinds of heterogeneity across data feeds were of the terminological structural varieties.
-As these were the ``simpler'' kinds of heterogeneity, this gave us optimism that an interactional approach would be feasible.
-In the following sections, we describe the design process we used to create the prototype of DataPalette, followed by a detailed description of the design of the prototype and justification.
+The two pre-studies above provided valuable insights towards this goal, which, in turn, informed our design process.
+PS.1 confirmed the need for data integration, by showing that people preferred to rely upon multiple sources of information for making important decisions, and identified the kinds of sources most commonly used.
+PS.2, meanwhile, revealed that the most common kinds of heterogeneity across data sources were of the terminological or structural varieties.
+As these were the ``simpler'' kinds of heterogeneity, this gave us optimism that an interaction-based approach would be feasible.
+
+In the following sections, we describe the design process of creating the prototype of DataPalette.
+This is followed detailed descriptions of prototype with justifications grounded in the pre-studies.
 

### 2013-01-10 chi2013-webbox.tex
Figure 1 full width
Remove Table 3 and references to it.

@@
 
-\begin{figure*}[tbp]
+\begin{figure*}[thb]
 \begin{center}
@@
 After the studies were complete, we generated transcripts of comments made by participants during the think-aloud protocols.
-At the same time, we populated a spreadsheet summarising quantitative metrics pertaining to how much various features of the tool were used (see Table \ref{tbl:data} for specific fields).
+At the same time, we populated a spreadsheet summarising quantitative metrics pertaining to how much various features of the tool were used.
 After the transcriptions were complete, we categorised people's comments made during the study, suggests in their exit survey, and clustered them to common themes.
@@
 
-\newpage
+%\newpage
 \section{Results}
@@
 On average, each study took 40 minutes to complete.
-The information in Table \ref{tbl:data} summarises the captured metrics for each participant.
+%The information in Table \ref{tbl:data} summarises the captured metrics for each participant.
 
-\begin{table*}[htbp]
-\small
-\begin{center}
-\begin{tabular}{|r|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|}
-\hline
-Participants & 1 &2&3&4&5&6&7&8&9&10&11&12&13&14&15&16&17&18&19&20\\
-\hline\hline
-gender&f&m&f&f&f&f&f&m&f&f&m&f&m&f&m&m&m&m&m&m\\
-A condition first task &y&y&y&n&n&y&n&n&y&y&n&n&y&y&n&n&n&n&n&n\\
-\hline\hline
-A Condition: task &4&2&3&4&3&&2&1&4&1&4&1&3&2&3&2&1&4&3&2\\
-paper used&n&n&n&n&n&n&y&n&y&y&n&y&n&n&y&n&n&n&n&n\\
-charts used &1&2&2&0&2&4&5&0&0&0&2&3&2&0&0&0&3&3&4&1\\
-maps used &1&1&2&2&2&1&1&0&1&0&1&1&2&1&0&1&0&0&2&1\\ 
-same as used &4&3&3&6&0&2&0&2&1&0&6&5&6&2&7&0&6&7&7&7\\
-\# data sources used &4&3&4&4&4&3&4&3&4&4&3&3&5&4&5&4&4&5&5&4\\
-chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, \newline 1.3, \newline 1.2& 4.2& 1.4, \newline 1.5, \newline 1.3& 3.2& 2.1, \newline 2.3, \newline 2.6& 3.5, \newline 3.3&2.6, \newline 2.3, \newline 2.5&1.6, \newline 1.5, \newline 1.3&4.1&3.3&2.4, \newline 2.5, \newline 2.6\\
-\# factors in choice &6&5&4&4&2&4&8&8&6&3&4&8&5&4&7&7&4&5&5&5\\
-\hline\hline
- B Condition:task &3&&4&3&4&2&1&2&2&3&2&3&1&4&1&4&2&2&1&3\\
- paper used &n&n&n&n&n&n&y&n&y&y&n&y&n&y&y&n&n&y&n&n\\
-% spreadsheet/website&&&&&&&&&&&&&&&&&&&&\\
- number of spreadsheets used&5/5&4/4&2/5&1.5&2/5&5/5&4/4&2/4&3/4&&2/4&4/5&2/4&4/5&2/4&2/5&4/4&4/4&0&5/5\\
- number of websites used&1&1&4&3&3&0&2&3&2&&1&2&3&1&2&3&2&0&4&0\\
- chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4, \newline 1.2, \newline 1.6&2.1&2.3, \newline 2.6, \newline 2.5&3.5&2.2, \newline 2.3, \newline 2.6&3.5&1.1&4.1&1.6&4.2&2.6, \newline 2.3, \newline 2.5&2.3, \newline 2.5, \newline 2.6&1.1, \newline 1.3,  \newline 1.4&3.5\\
- \# factors in choice &4&3&2&3&1&6&5&7&7&5&4&7&6&3&6&4&3&6&4&5\\
- \hline
-\end{tabular}
-\end{center}
-\caption{Data collected from each of our trials, for each participant.} \label{tbl:data}
-\normalsize
-\end{table*}
+%\begin{table*}[htbp]
+%\small
+%\begin{center}
+%\begin{tabular}{|r|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|p{0.3cm}|}
+%\hline
+%Participants & 1 &2&3&4&5&6&7&8&9&10&11&12&13&14&15&16&17&18&19&20\\
+%\hline\hline
+%gender&f&m&f&f&f&f&f&m&f&f&m&f&m&f&m&m&m&m&m&m\\
+%A condition first task &y&y&y&n&n&y&n&n&y&y&n&n&y&y&n&n&n&n&n&n\\
+%\hline\hline
+%A Condition: task &4&2&3&4&3&&2&1&4&1&4&1&3&2&3&2&1&4&3&2\\
+%paper used&n&n&n&n&n&n&y&n&y&y&n&y&n&n&y&n&n&n&n&n\\
+%charts used &1&2&2&0&2&4&5&0&0&0&2&3&2&0&0&0&3&3&4&1\\
+%maps used &1&1&2&2&2&1&1&0&1&0&1&1&2&1&0&1&0&0&2&1\\ 
+%same as used &4&3&3&6&0&2&0&2&1&0&6&5&6&2&7&0&6&7&7&7\\
+%\# data sources used &4&3&4&4&4&3&4&3&4&4&3&3&5&4&5&4&4&5&5&4\\
+%chosen restaurant/uni &4.1&2.4&3.5&4.3&3.3&&2.4&1.4&4.2&1.6, \newline 1.3, \newline 1.2& 4.2& 1.4, \newline 1.5, \newline 1.3& 3.2& 2.1, \newline 2.3, \newline 2.6& 3.5, \newline 3.3&2.6, \newline 2.3, \newline 2.5&1.6, \newline 1.5, \newline 1.3&4.1&3.3&2.4, \newline 2.5, \newline 2.6\\
+%\# factors in choice &6&5&4&4&2&4&8&8&6&3&4&8&5&4&7&7&4&5&5&5\\
+%\hline\hline
+% B Condition:task &3&&4&3&4&2&1&2&2&3&2&3&1&4&1&4&2&2&1&3\\
+% paper used &n&n&n&n&n&n&y&n&y&y&n&y&n&y&y&n&n&y&n&n\\
+%% spreadsheet/website&&&&&&&&&&&&&&&&&&&&\\
+% number of spreadsheets used&5/5&4/4&2/5&1.5&2/5&5/5&4/4&2/4&3/4&&2/4&4/5&2/4&4/5&2/4&2/5&4/4&4/4&0&5/5\\
+% number of websites used&1&1&4&3&3&0&2&3&2&&1&2&3&1&2&3&2&0&4&0\\
+% chosen restaurant/uni&3.5&&4.2&3.3&4.3&2.1&1.4, \newline 1.2, \newline 1.6&2.1&2.3, \newline 2.6, \newline 2.5&3.5&2.2, \newline 2.3, \newline 2.6&3.5&1.1&4.1&1.6&4.2&2.6, \newline 2.3, \newline 2.5&2.3, \newline 2.5, \newline 2.6&1.1, \newline 1.3,  \newline 1.4&3.5\\
+% \# factors in choice &4&3&2&3&1&6&5&7&7&5&4&7&6&3&6&4&3&6&4&5\\
+% \hline
+%\end{tabular}
+%\end{center}
+%\caption{Data collected from each of our trials, for each participant.} \label{tbl:data}
+%\normalsize
+%\end{table*}
 
@@
 The uses of Excel for external cognition ranged from collation behaviour to tallying annotations, to colour-coding cells to allow easy identification of  ``pros'' and ``cons''.
-The complete data collected is given in Table \ref{tbl:data}.
 In general, users struggled with both the websites and with the spreadsheet, although could generally find the information they required after a short period of frustration.

### 2013-01-13 chi2013-webbox.tex
better conclusion

@@
 \subsection{Does DataPalette enable data integration?}
-
 We set out to test three hypotheses with the DataPalette study.
@@
 
-As said in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for complete approaches to data integration have, perhaps, caused ``ad-hoc'' approaches to be overlooked.
-Based upon the results of our investigation, however, we see a number of benefits to lightweight, user-driven, interactive approaches over automatic and bespoke ones.
-Just as the \emph{user-subjective approach} to PIM \cite{user-subjective} showed that subjective attributes and labels were more useful than physical ones, we have shown that a user-driven approach to data integration, as achieved with DataPalette, affords users flexibility to combine and keep separate data at will, and may facilitate decision-making tasks.
-Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through easy to use, and intuitive interactive approaches for end-users.
+As said in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for fully-automatic, complete approaches to data integration have casued simpler, manual approaches to be overlooked.
+Based upon the results of our investigation, we see a number of benefits to interactive, lightweight approaches over automatic and bespoke methods, at least for several classes of common, information-gathering for decision-making tasks.
+
+We believe that the key affordances for the suitability of such tools are flexibility, simplicity, and ease of use.
+Just as work in the ``user-subjective approach'' to PIM (e.g., \cite{user-subjective}) showed that personal, user-specified attributes were ultimately the most useful, granting users the flexibility to combine or separate information items as needed allowed the tool to support the varied strategies participants employed to compare and examine their options.
+Second, the low cost of use  (in terms of time and effort required) meant that integration could be done (and undone) repeatedly within the ``inner loop'' of their exploratory decision-making process.
+The importance of this was that it was the process itself that determined, incrementally, what information was ultimately needed, including the sources that were consulted.
+Therefore, the use of any integration approach that required sources to be determined \emph{a priori}, or that required a time or effort-intensive integration step would not have been feasible.
+
+Based on these findings, we plan to continue investigating lightweight, \emph{in-situ} approaches to supporting ad-hoc data integration during the course of sensemaking and decision-making activities.
+In particular, we are examining gestural input methods for the purposes of supporting more fluid exploration and mixing.
+Based upon feedback from reviewers, we have also begun considering semi-automatic and crowd-powered methods to help users quickly resolve challenging reconciliation problems they might encounter.
+
+% instead of being a separate step requiring pre-planning one that and that information sources could be brought in or rmeoved as needed.
+
+%% this study demonstrated that user-driven and-defined approach to information integration provided the flexibility to allow people to organize and support information gathering tasks in different ways.
+The result, that  people weighed more factors from a greater number of information sources, while expending equal or less time and effort, demonstrated that such tools could broaden the information gathering capacities and allow people to make better use of available information sources during decision-making activities.
+
+%% this was having a sufficiently low barrier of effort to use, provided both the flexibillity and facility to support information gathering tasks associated with decision-making activities.
+The result was people weighed more factors from a greater number of information sources, broading the scope of their information gathering for such decisions, while expending comparable or less effort and time.
+
+%% Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through easy to use, and intuitive interactive approaches for end-users.
+
+\section{Acknowledgements}
+
+The work presented in this paper was sponsored by the EPSRC project SOCIAM (Ref.
+EP/J017728/1).
+We would like to thank our collaborators Wendy Hall, mc schraefel, Tim Berners-Lee, Kieron O'Hara, Susan Davies, Maraim Elbadri, Tom Smith, Yang Yang, Igor Popov, and Shauna VanKleek whose help and advice have shaped this work throughout.
 

### 2013-01-14 chi2013-webbox.tex
edits thru methodology

@@
 In summary, we found in our PS.2, that while data records ranged in complexity, no single source subsumed any others; all contained unique properties.
-Moreover, the most common types of heterogeneity exhibited were terminological and value-related, which occurred vastly more than structural or modeling issues.
+Moreover, the most common types of heterogeneity exhibited were terminological and value-related, which occurred vastly more often than structural or modeling issues.
 
@@
 
-The primary goal of DataPalette was to empower end-users so that they are able to mix arbitrary, structured data feeds on-the-fly while solving particular information task(s) more effectively.
+The primary goal of DataPalette (henceforth DP) was to empower end-users so that they were able to mix arbitrary, structured data feeds on-the-fly, while solving particular information task(s) more effectively.
 Unlike mashup makers, such as Yahoo Pipes!, the aim was to facilitate such integration without the need for a separate step involving programming, scraping, shaping, mapping or integration.
@@
 The two pre-studies above provided valuable insights towards this goal, which, in turn, informed our design process.
-PS.1 confirmed the need for data integration, by showing that people preferred to rely upon multiple sources of information for making important decisions, and identified the kinds of sources most commonly used.
+PS.1 confirmed the need for data integration, by showing that people preferred to rely upon multiple sources of information for making important decisions; PS.1 also identified the kinds of sources most commonly used.
 PS.2, meanwhile, revealed that the most common kinds of heterogeneity across data sources were of the terminological or structural varieties.
@@
 
-In the following sections, we describe the design process of creating the prototype of DataPalette.
-This is followed detailed descriptions of prototype with justifications grounded in the pre-studies.
+In the following sections, we describe the design process of creating the prototype of DP.  This is followed by detailed descriptions of the prototype, with justifications grounded in the pre-studies.
 
 \subsection{Design Process}
-We followed a five-phase spiral model consisting of planning, designing, mocking-up, prototyping, and testing to derive the design of DataPalette.
-This process was used because of the high degree of initial uncertainty associated with how we'd achieve effective data mixing by end-users.
+We followed a five-phase spiral model, consisting of planning, designing, mocking-up, prototyping, and testing, to derive the final design of DP.
+This multi-phase process was used, due to the high degree of initial uncertainty associated with how users would achieve effective data mixing.
 We approached this design challenge through alternative generation and elimination; alternatives were designed, drawn up and evaluated by colleagues (and in later iterations, non-expert potential users), who critiqued the alternatives.
-The ease with which designs could be prototyped in HTML5 with modern web frameworks allowed us to perform this develop-test-iterate rapidly through seven iterations of this process.
-
-In order to feel familiar to new users, we started with the look and feel from OS X Finder-style file managers.
-In the DataPalette workspace, shown in Figure \ref{fig:workspace}, the user can freely drag and drop entities from data sources (on the left) into windows that they create, resize, arrange and label.
+The ease with which designs could be prototyped in HTML5 with modern web frameworks allowed us to perform this process of develop-test-iterate rapidly through seven iterations.
+
+In order to allow new users to be comfortable with our interface, we adopted the look and feel from OS X Finder-style file managers.
+In the DP workspace, shown in Figure \ref{fig:workspace}, the user would be able to freely drag and drop entities from data sources (on the left) into windows that they have created, resized, arranged and labeled.
 Unlike file managers, the basic unit of data was not a file, but an \emph{instance} --- corresponding to a single data record, JSON object, RDF instance, or any small bit of structured data.
@@
 \subsubsection{Multi-path selection: link-sliding for heterogeneous sets}
-
 Typically, users are interested in viewing and comparing the \emph{properties} of instances; for example, a user might be keen to examine a product's price, rating, manufacturer and so on.
-An instance's properties are displayed beneath it; selecting a property creates a \emph{path selection}, which dereferences all of the instances in a group that contain the selected property and displays their corresponding values.
-For example, when a user selects the \emph{manufacturer} property for a product, all products in the same group with a manufacturer property will be dereferenced, causing the manufactures to be displayed alongside their corresponding products.
-When such a path selection is applied, the properties displayed are updated to be the set of properties of the \emph{terminating value of the path}, so that this value can be dereferenced further.
-For example, once selecting a manufacturer, a user may want to know the product's manufacturer's reputation and selects their ``reputation score'' property, causing the previous path selection to be extended.
+To support such comparisons, an instance's properties would be displayed beneath it; selecting a property would create a \emph{path selection}, which would dereference all of the instances in a group that contained the selected property and displayed their corresponding values.
+For example, when a user selected the \emph{manufacturer} property for a product, all products in the same group with a manufacturer property would be dereferenced, causing the manufacturers to be displayed alongside their corresponding products.
+When such a path selection was applied, the properties displayed were updated to be the set of properties of the \emph{terminating value of the path}, so that this value could be dereferenced further.
+Having selected a manufacturer, a user might want to be informed of its reputation.
+By selecting the ``reputation score'' property, the previous path selection would then be extended.
 This process is illustrated in Figure \ref{fig:multipathing}.
@@
 
-When instances cannot be dereferenced by the chosen path selection, an additional path selection can be created.
-As with the first path selection, it causes all matching instances to be dereferenced.
-Each group can carry an arbitrary number of path selections, and each path selection can be extended continually by selecting successive properties of values.
-This achieves the ability to \emph{link slide} as introduced by Parallax~\cite{parallax}, but across sets of heterogeneous items by supporting multiple parallel dereferenced paths simultaneously.
-Multiple selection group-dereferencing results in two key benefits.
-First, that users can quickly compare all instances that have the same structure, and second, that instances with different property names and be quickly consolidated.
+When instances could not be dereferenced by the chosen path selection, an additional path selection could be created.
+As with the first path selection, this would cause all matching instances to be dereferenced.
+Each group would be able to carry an arbitrary number of path selections, and each path selection could be extended continually by selecting successive properties of values.
+This would achieve the ability to \emph{link slide} multiple instances simultaneously, as introduced by Parallax~\cite{parallax}, extended to support sets of heterogeneous items through multiple parallel simultaneous paths.
+Multiple-path group-dereferencing would result in two key benefits.
+First, users would be able to quickly compare all instances that have the same structure, and second, instances with different property names would be easily consolidated.
 
@@
 \subsubsection{Coreference consolidation: Drag-and-Drop Same-as}
-
-A major challenge with integration from multiple sources is coreference consolidation, or combining representations that co-refer to the same entity into one.
-For example, consolidating one's friends from Facebook, Google+, and LinkedIn, one might quickly want to de-duplicate and combine friends that use multiple services so that they each have one representation.
-Although we considered a way to do this automatically, PS.2 revealed no way to reliably identify co-referring instances; due to a lack of standard URIs across social networks, identifiers used by each network were system-specific and inconsistent.
+A major challenge with integration from multiple sources is coreference consolidation, or combining into one, representations that co-refer to the same entity.
+For example, when consolidating one's friends from Facebook, Google+, and LinkedIn, one might wish to eliminate duplication and combine friends' profiles across multiple services.
+Although we considered ways to do this automatically, PS.2 revealed no viable method to reliably identify the co-referring instances; due to a lack of standard URIs across social networks, identifiers used by each network were system-specific and inconsistent.
 Thus, we sought to instead make it user-driven, and easy to perform manually.
@@
 
-To remain consistent with the drag-and-drop interactions used throughout the interface, we introduced the ability to simply drag one instance on top of another to specify that two things should be combined.
-This action also caused the \emph{same-as display} in the lower left of the screen to display this new combination relationship.
+To remain consistent with the drag-and-drop interactions used throughout the interface, we introduced the ability to simply drag one instance directly on top of another, to specify that these two items should be combined.
+This action also caused the \emph{same-as display} in the lower left of the screen to indicate this new combined relationship.
 The display also acted as a mechanism to delete/undo combinations, which reverted representations of instances back to their original, separate forms.
 An illustration of this interaction is provided in Figure \ref{fig:sameas}.
-Although general, this can be tedious for groups of items.
-Thus, to support more efficient combination, we extended this to allow entire groups to be dropped on other groups.
-When this is done, DataPalette attempts to identify matching pairs among items in the dragged and target groups by comparing each pairs' labels.
-Only exact matches (modulo white space/underscores) were considered matches and automatically combined.
-The rest are added to the target group as distinct elements, and may be explicitly combined by hand.\looseness=-1
+Since individually dragging and dropping items could be tedious for large groups of items, we extended this to allow entire groups to be dropped on to other groups.
+DP would then identify matching pairs among items in the dragged and target groups by comparing the labels of items in each group, repsectively.
+Only exact matches (excepting white space and underscores) were considered to correspond to label matches, and these were automatically combined.
+The rest were added to the target group as distinct elements, which could be manually combined by hand.\looseness=-1
 
@@
 \subsubsection{Enumerated-type value consolidation}
-PS.2 revealed that enumerated values of properties were often inconsistent, e.g.
-``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property.
+PS.2 revealed that enumerated values of properties were often inconsistent, e.g., ``Casual dining'' vs ``Relaxed'' for a restaurant's ``atmosphere'' property.
 To reconcile and consolidate corresponding values from different  data sources, we extended the drag and drop support of co-reference combination to also allow users to drag and drop to combine enumerated literal values as well.
@@
 
-To make it easier for people to quickly visually compare aspects of instances, we created a charting tool and integrated a mapping feature.
-To make these tools suitable for fast use, e.g.
-data exploration, we designed these tools to automatically configure display parameters instances or groups of were dropped on them, rather than making users manually specify them.
-In the current prototype, the DWIM behaviour of the chart automatically sets whether the plot is a histogram representation (counts of values) or numeric value display based upon the types of the dereferenced values being plotted; values that are non-numeric are counted, while numeric values are displayed.
+To make it easier for people to visually compare aspects of instances, we created a charting tool and mapping feature.
+To ensure that these tools were suitable for rapid utilisation, (e.g., data exploration), they were designed to automatically configure display parameters based on the instances or groups that were dropped onto them.
+In the current prototype, this ``Do What I Mean'' behaviour automatically would set whether the plot was a histogram representation (counts of values) or a numeric value display, based upon the types of the dereferenced values being plotted.
+Hence, non-numeric values would be histogrammed, while numeric values were displayed as a bar or line chart.
 Similarly, the map was made to detect address strings and latitude/longitude pairs, geocoding them where appropriate, and determining the optimal bounding box to display all items at maximum zoom.
@@
 \subsubsection{Model-based brushing}
-Since it can get confusing having a single instance represented across several groups, we provided \emph{universal brushing} across the interface; hovering over any representation of any instance (an instance block, a map marker or a histogram bar), causes all other visible representations of that instance to lightly glow, so that the user can quickly identify all the places that a single instance is represented.
+To avoid the confusion of having a single instance represented across several groups, we provided \emph{universal brushing} across the interface.
+Hovering over any representation of any instance (an instance block, a map marker or a histogram bar), would cause all other visible representations of that instance to lightly glow, enabling the user to instantly identify all the places that a single instance was represented.
 
@@
 
-In order to determine whether the interaction affordances introduced in DataPalette (henceforth DP) allow users to perform serendipitous data mixing on real data feeds by end-users, we devised a within-subjects controlled lab study which we describe next.
+In order to determine whether or not the interaction affordances introduced in DP allowed users to perform serendipitous data mixing on real data sources, we devised a within-subjects controlled lab study, which we describe next.
 
@@
 
-We first describe our methodology, starting with the three hypotheses we set out to test:
+We began with three hypotheses pertaining to people's ability to use DP, and its impact on task performance:
 %These hypothesis pertain to the usability of DP by end-users (h1), whether DP achieves its stated goal of enabling end-users to integrate data (h2), and finally if it has an effect on task performance (h3) - either in how well people perform a task, how quickly they perform it, or if it reduces task effort, as follows:
@@
 The participants were seated at a standard OS X desktop equipped with a 22-inch monitor with a standard keyboard and mouse.
-Each participant was given a maximum of 10 minutes to each of two tasks, but were told that if they finished earlier they could tell us and move on to the next condition.
-Prior to A conditions, participants were trained on how to use DP using a five-minute instructional video.
-
-In the control condition (B), participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from web sites pertaining to the candidates they were asked to choose from.
+Each participant was given a maximum of 10 minutes to complete each of two tasks, but were told that if they finished early, they could inform us and move on to the next condition.
+Prior to the A condition trials, participants were trained on how to use DP using a five-minute instructional video.
+
+In the control (B) condition trials, participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from web sites pertaining to the candidates they were asked to choose from.
 Participants were told they were free to use the web sites and spreadsheets however they wished.
@@
 % why? why were these tasks selected? can we drop in the phrase ``multi-dimensional subjective choice task'' and refer back to PS.1 to justify the choice of tasks?
-We based our two tasks on choosing a university and picking a restaurant for a large social event, because in PS.1 the majority of people stated that they had used multiple websites for these tasks.
-We used two different variables for both tasks, where the restaurant was located (Glasgow and Cambridge) and which subject people chose (sport science and history).
+
+We established that our two tasks would entail choosing a university and selecting a restaurant for a large social event.
+These tasks were selected because in PS.1, the majority of people stated that they had used multiple websites for such tasks.
+We introduced a variable for each task.
+For the University selection task, we varied the course for which to select a University, between history and sports science, which we had selected randomly from the course list of the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}.
+For the restaurant selection task, the variable selected where the restaurants were located; for this experiment, the cities Glasgow and Cambridge were chosen because they were geographically distant from participants, and would therefore minimise prior familiarity.
+
+For each of the two tasks, the final candidates from which participants made their selection were pre-selected by us.
+For the university task, the six top-ranked universities for each subject according to The Complete University Guide were selected as candidates.
+We selected six in order to limit the difficulty of the task given the fixd time allocated for each trial.
+We manually collected the key statistics about each university from each of the data sources, and collated them into Excel spreadsheets.
+% In addition, for each subject, we provided a hypothetical student's A/AS-level scores and personal preferences pertaining to distance from home, %%
+
+For the restaurant task, we selected six restaurants for each location randomly from Yelp's list of popular eateries for each of the two cities.
+In addition, for each of the two cities, we generated 12 social network profiles for hypothetical friends who lived in either city, to accompany the participant to the restaraunt.
+Each friend's profile expressed their personal preferences for favourite cuisine types, which could be taken into consideration by participants during the selection process.
+
+% We established that our two tasks would entail choosing a university and selecting a restaurant for a large social event.
+These tasks were selected because in PS.1, the majority of people stated that they had used multiple websites for such tasks.
+We used two different variables for each task.
+where the restaurant was located (Glasgow and Cambridge) and which subject people chose (sport science and history).
 For the restaurant task, the cities Glasgow and Cambridge were chosen because they were geographically distant to participants, and would therefore minimise likelihood of prior familiarity with the restaurants or location.
 For the University selection task, two university courses were selected at random from the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}, and the six top-ranked universities for each subject were selected as candidates.
-% \subsubsection{Task preparation} In order to prepare the data for our tasks, we identified websites that people commonly used for selecting universities and restaurants (see Table \ref{tab:studyfactors}).
-Once we had compiled the list of websites, we select the top most used six for each task.
 We selected six because it enables us to scope our study, we found through two pilot studies that six data sources provided a sufficient challenge for users to explore the data and required them to use DP's main features.
@@
 \subsection{Data Collection and Analysis}
-Each study was overseen by a facilitator and an observer:  the role of the facilitator was to explain the study's protocol to the participant and to answer any questions; and the role of the observer was to observe without interacting with the user, and to take notes.
-Both the facilitator and observer were trained on the purpose of the study, their role and how they could and couldn't influence the study.
-Explicitly both the facilitator and observer where trained in identifying processes that people used during a task, so that they could thematically categorise them.
+Each study was overseen by a facilitator and an observer: the role of the facilitator was to explain the study's protocols to the participant and to answer any questions; the role of the observer was to monitor the study and record observations.
+Both the facilitator and observer were trained on the purpose of the study, their respective roles, and how they could or couldn't influence the study.
+Explicitly, the facilitator and the observer were trained to identify processes that participants might use during a task, so that they could thematically categorise them.
 Initially, we identified the following themes: organisation of data, eliminating candidates, co-referencing, and visualisations.
-The studies were run over 4 days.
-At the end of each day the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.
+The studies were conducted over a period of 4 days.
+At the end of each day, the facilitators and observers met to discuss the studies' data and to revisit protocols, if necessary.
 During the lab study, we recorded the audio and the participant's actions on screen.
@@
 After the studies were complete, we generated transcripts of comments made by participants during the think-aloud protocols.
-At the same time, we populated a spreadsheet summarising quantitative metrics pertaining to how much various features of the tool were used.
-After the transcriptions were complete, we categorised people's comments made during the study, suggests in their exit survey, and clustered them to common themes.
+At the same time, we created a spreadsheet summarising quantitative metrics pertaining to how much of the various features of DP were used.
+After the transcriptions were completed, we categorised comments made by the participants during the study, together with suggestions from their exit survey, and clustered them to common themes.
 
@@
 \subsection{Participant Recruitment}
-We recruited participants through adverts posted near the University campus and e-mails to student mailing lists.
-We screened participants to be at least 18 years of age, but did not filter on any other criteria.
-We took the first 20 that signed up and that successfully specified time slots that were available.
+
+We recruited participants through adverts posted near the Southampton University campus and via e-mails to student mailing lists.
+We screened participants to ensure they were at least 18 years of age, but did not filter candidates based on any other criteria.
+We received 26 responses to our call, from which we accepted the first 20 applicants (10 of whom were female).
+Seven were non-students, comprised of university staff and alumni; the remainder were students, 8 studying computer science, and 5 studying other subjects.
+Due to the large population of international students and staff at our university, we ended up with a large sample (12) of non-native English speakers; however, since all had passed requisite proficiency tests, we did not think this would substantially impact results.
+On average, each study took 40 minutes to complete.
 Participants were offered a small gift certificate from an online retailer for their time.
@@
 
-We received 26 responses to our call, from which we selected the first 20 (ten of whom were female).
+% We received 26 responses to our call, from which we selected the first 20 (ten of whom were female).
 Seven were non-students, comprised of university staff and alumni; the remainder were students, 8 studying computer science, and 5 studying other subjects.
@@
 A second departure from a realistic task was the way in which we pre-determined the data sources for participants to use.
-We did this in order to focus participants' efforts on looking at and working with their choices, rather than retrieving it.
+We did this in order to focus participants' efforts on looking at and working with their choices, rather than retrieving them.
 However, as supporting effective retrieval and serendipitous discovery will be important for DP to be useful in practice, we plan to look at supporting these stages in the future.
+
 % \subsection{Integration Limitations}
-
 % In the A condition, where participants used DataPalette to complete their tasks, they used it to collate the data on-screen, and compare multiple values directly.
@@
 %% \subsection{Better Supporting the Task}
-
 %% - are people more thorough
@@
 
-As said in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for fully-automatic, complete approaches to data integration have casued simpler, manual approaches to be overlooked.
-Based upon the results of our investigation, we see a number of benefits to interactive, lightweight approaches over automatic and bespoke methods, at least for several classes of common, information-gathering for decision-making tasks.
-
-We believe that the key affordances for the suitability of such tools are flexibility, simplicity, and ease of use.
+As expressed in Voltaire's poem \emph{La B\'{e}gueule}, ``the best is the enemy of the good'', we believe that the quest for fully-automatic, complete approaches to data integration have caused simpler, manual approaches to be overlooked.
+Based upon the results of our investigation, we have seen a number of benefits to interactive, lightweight approaches over automatic and bespoke methods, at least for a frequently-occurring class of information gathering for decision-making tasks.
+
+In particular, the results suggest that the key affordances for the suitability of such tools are flexibility, simplicity, and ease of use.
 Just as work in the ``user-subjective approach'' to PIM (e.g., \cite{user-subjective}) showed that personal, user-specified attributes were ultimately the most useful, granting users the flexibility to combine or separate information items as needed allowed the tool to support the varied strategies participants employed to compare and examine their options.
-Second, the low cost of use  (in terms of time and effort required) meant that integration could be done (and undone) repeatedly within the ``inner loop'' of their exploratory decision-making process.
-The importance of this was that it was the process itself that determined, incrementally, what information was ultimately needed, including the sources that were consulted.
-Therefore, the use of any integration approach that required sources to be determined \emph{a priori}, or that required a time or effort-intensive integration step would not have been feasible.
-
-Based on these findings, we plan to continue investigating lightweight, \emph{in-situ} approaches to supporting ad-hoc data integration during the course of sensemaking and decision-making activities.
+Second, the low cost of use  (in terms of time and effort required) meant that integration could be done (and undone) as appropriate throughout the ``inner loop'' of users' exploratory processes.
+The reason this was significant was that this process itself determined, incrementally, the information and sources that were ultimately considered for the final decision.
+Therefore, any integration approach that required sources to be specified \emph{a priori}, or that was time or effort-intensive at each integration step, was simply not appropriate for this context.
+
+Due to the promising findings of this study, we plan to continue investigating lightweight, \emph{in-situ} approaches to supporting ad-hoc data integration during the course of sensemaking and decision-making activities.
 In particular, we are examining gestural input methods for the purposes of supporting more fluid exploration and mixing.
-Based upon feedback from reviewers, we have also begun considering semi-automatic and crowd-powered methods to help users quickly resolve challenging reconciliation problems they might encounter.
+On the basis of feedback from reviewers, we have also begun considering semi-automatic and crowd-powered methods to help users quickly resolve challenging reconciliation problems they might encounter.
 

### 2013-01-14 chi2013-webbox.tex
webbox

@@
 These tasks were selected because in PS.1, the majority of people stated that they had used multiple websites for such tasks.
-We introduced a variable for each task.
-For the University selection task, we varied the course for which to select a University, between history and sports science, which we had selected randomly from the course list of the complete university guide\footnote{The Complete University Guide:\\\url{www.thecompleteuniversityguide.co.uk}}.
-For the restaurant selection task, the variable selected where the restaurants were located; for this experiment, the cities Glasgow and Cambridge were chosen because they were geographically distant from participants, and would therefore minimise prior familiarity.
+We introduced a dependent \emph{dataset} variable for each task.
+For the University selection task, this variable corresponded to the course for which to select a University, between history and sports science, and corresponding university candidates for each course.
+For the restaurant selection task, the variable corresponded to the city and respective restaurants located therein; for this experiment, Glasgow and Cambridge were chosen because they were geographically distant from participants, and would therefore minimise prior familiarity.
 
@@
 We selected six in order to limit the difficulty of the task given the fixd time allocated for each trial.
-We manually collected the key statistics about each university from each of the data sources, and collated them into Excel spreadsheets.
-% In addition, for each subject, we provided a hypothetical student's A/AS-level scores and personal preferences pertaining to distance from home, %%
-
-For the restaurant task, we selected six restaurants for each location randomly from Yelp's list of popular eateries for each of the two cities.
+In addition, for each course, we provided participants with the profile of a hypothetical student applying to the course, for whom they should base their choice.
+This profile included the student's entrance admissions tests scores (corresponding to A/AS-levels), location of the student's hometown, and personal preferences pertaining to how far they wanted to be from home, tuition they could afford, gender balance, and social-athletic-academic balance.
+
+For the restaurant task, we selected six restaurants at random from Yelp's list of popular eateries for each of the two cities.
 In addition, for each of the two cities, we generated 12 social network profiles for hypothetical friends who lived in either city, to accompany the participant to the restaraunt.
-Each friend's profile expressed their personal preferences for favourite cuisine types, which could be taken into consideration by participants during the selection process.
+Each friend's profile expressed their street address and preference for favourite cuisine type which could be taken into consideration by participants during the venue selection process.
+
+For the B conditions we prepared, for each dataset condition in each task, a spreadsheet listing key statistics for each of the relevant candidates (e.g., universities or restaurants) assembled from each of a pre-selected set of 6 popular data sources (university and restaurant guides), and collated them into Excel spreadsheets.
 

### 2013-01-14 chi2013-webbox.tex
Updated epsrc acknowledgement to standard form and added heather's EPSRC acknowledgement

@@
 %}
+
+\hyphenation{EPSRC SOCIAM}
 
@@
 
-The work presented in this paper was sponsored by the EPSRC project SOCIAM (Ref.
-EP/J017728/1).
+The work presented in this paper was supported by the SOCIAM project, funded by the Engineering and Physical Sciences Research Council under contract EP/J017728/1.
+Dr Packer is grateful for funding from the EPSRC in the form of a Post Doctoral Award.
 We would like to thank our collaborators Wendy Hall, mc schraefel, Tim Berners-Lee, Kieron O'Hara, Susan Davies, Maraim Elbadri, Tom Smith, Yang Yang, Igor Popov, and Shauna VanKleek whose help and advice have shaped this work throughout.

### 2013-01-14 chi2013-webbox.tex
Replace "School of ECS" with current name
edits

@@
    \alignauthor Max Van Kleek, Daniel A. Smith, Heather S. Packer, Jim Skinner, Nigel R. Shadbolt \\
-     \affaddr{School of Electronics and Computer Science}\\
+     \affaddr{Web and Internet Science Research Group, Electronics and Computer Science}\\
      \affaddr{University of Southampton, Southampton, UK}\\
@@
 
-All of the participants used multiple web sites in order to complete their tasks, and all felt that the lack of existing integration between sites hindered their ability to make informed choices.
+All of the participants used multiple websites in order to complete their tasks, and all felt that the lack of existing integration between sites hindered their ability to make informed choices.
 Participants prioritised the sources they used by both the ease and immediacy of accessing them.
@@
 
-In the control (B) condition trials, participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from web sites pertaining to the candidates they were asked to choose from.
-Participants were told they were free to use the web sites and spreadsheets however they wished.
+In the control (B) condition trials, participants were given a web browser with tabs open to all of the data sources they were provided for use, as well as a spreadsheet containing the data taken from websites pertaining to the candidates they were asked to choose from.
+Participants were told they were free to use the websites and spreadsheets however they wished.
 
@@
 Seven were non-students, comprised of university staff and alumni; the remainder were students, 8 studying computer science, and 5 studying other subjects.
-Due to the large population of international students and staff at our university, we ended up with a large sample (12) of non-native English speakers; however, since all had passed requisite proficiency tests, we did not think this would substantially impact results.
-On average, each study took 40 minutes to complete.
-Participants were offered a small gift certificate from an online retailer for their time.
+Due to the large population of international students and staff at our university, we ended up with a large sample (12) of non-native English speakers; however, since all had passed requisite proficiency tests, we did not think this would substantially impact the results.
+On average, each study took 40 minutes to complete, after which participants were offered a small gift certificate from an online retailer for their time.
 
@@
 \subsubsection{Thoroughness: Factors weighed in final choice}
-The second metric was the number of factors the participant considered in making their final choice.
+The second metric related to the number of factors each participant considered in making his or her final choice.
 To determine this, we asked participants to explain why they made their choice(s), and recorded the number of distinct data dimensions mentioned.
@@
 \subsubsection{Diversity: Data sources consulted}
-The third metric we used to gauge the task performance is the number of different data sources consulted during the trial; our rationale for this metric was that more informed decisions derived from more thorough consideration of available data.
-In this metric, participants used an average of 88\% ($SD=0.13$) of data sources provided in the DP case, compared to 80\% ($SD=0.18$) in the control condition, although an ANOVA test blocking on participant demonstrated that interface choice only approached significance ($F(1,19)=0.04$ $p < 0.15$).
+The third metric we employed to gauge the task performance was the number of different data sources consulted during the trial; our rationale for this metric was that more informed decisions derived from more thorough consideration of available data.
+In this metric, participants used an average of 88\% ($SD=0.13$) of data sources provided in the DP case, compared to 80\% ($SD=0.18$) in the control condition, although an ANOVA test blocking on participant ID demonstrated that interface choice only approached significance ($F(1,19)=0.04$ $p < 0.15$).
 
@@
 Since the tasks were subjective-choice, we could not evaluate the `correctness' of answers.
-However, to determine whether there was a difference in variability of answers between interface conditions, we plotted their choice on a histogram for each trial, per task and dataset for both conditions, visible in Figure \ref{fig:chosen_results}.
-As can be seen there is considerable variability and no discernible difference in variability between conditions.
+However, to determine whether there was a difference in variability of answers between interface conditions, we plotted the choices on a histogram for each trial, per task and dataset for both conditions, visible in Figure \ref{fig:chosen_results}.
+As can be seen, there was less agreement among universities than restaurants, although little agreement overall in both conditions.
+This is likely due to the many subjective factors involved in such choice tasks.
 
 \subsection{Strategies: Successive Elimination vs Tallying}
-As described in later this section, we noticed several different strategies people used to evaluate each candidate selection.
-One common strategy was \emph{successive elimination} - to look at a single dimension at a time, starting with the most important, ruling out candidates that did not meet the required minimum for that value.
-However, when there is no clear ``most important'' aspect, this  ``greedy-style'' can result in suboptimal decisions.
-Another strategy, which we called \emph{tallying the pros and cons}, kept all candidates around, and counted up the advantages and disadvantages of each, potentially weighing each factor by its perceived importance.
-This strategy is less vulnerable to getting stuck in local maxima than the former.
-
-To understand whether the interface influenced choice of strategy, we measured the number of candidates the participant considered \emph{at final result}.
-This is a strong marker for use of each strategy; people using the successive elimination strategy arrived with only 1--2 candidates at decision time, while those that tallied still maintained the entire starting set.\looseness=-1
+As described later in this section, we noticed several different strategies the participants used to evaluate each candidate selection.
+One common strategy was \emph{successive elimination}, that is, to regard a single dimension or factor at a time, starting with the most important, and ruling out candidates not meeting the minimum requirements for that value.
+However, when there was no clear aspect that was ``most important'', this  ``greedy'' strategy can result in suboptimal decisions.
+Another tactic, which we called \emph{tallying the pros and cons}, was to consider all candidate choices together, totaling up the advantages and disadvantages of each, potentially weighed by its perceived importance.
+This strategy was less vulnerable to getting stuck in local maxima than the former.
+
+To understand whether the interface influenced the choice of strategy, we measured the number of candidates the participant considered at their final choice.
+This was a strong marker for use of each strategy; people who used the successive elimination strategy arrived with only 1--2 candidates at decision time, while those that tallied still maintained the entire starting set.\looseness=-1
 
@@
 
-\subsection{Use of Data Integration Features}
+\subsection{Condition A: Use of Data Integration Features}
 
@@
 Five participants (25\%) deliberately used multi-path selection, defined as selecting more than one active path per group at the same time to display common values across heterogeneous items.
-(We did not count accidental use of multi-path, such as happened a few times when a participant had an already active path and wanted to switch to a separate path accidentally).
+(We did not take into account accidental uses of multi-path, such as when a participant had an already active path and wanted to switch to a separate path accidentally).
 
@@
 \subsection{Use of Visualisation Features}
-To determine to what extent visualisation tools were used in condition A, 13 (65\%) used charting tools at least once, while a majority (15, 75\%) used the map.
-Determining the impact these had on performance, we performed a multiple regression on time taken with the number of uses of charting and mapping as variables.
+In our determination of the extent to which visualisation tools were used in condition A, 13 (65\%) used charting tools at least once, while a majority (15, 75\%) used the map.
+To assess the impact use of these features had on performance, we carried out a multiple regression on the time taken, with the number of uses of charting and mapping as variables.
 We found significant effects of both charting and mapping variables on time taken, ($R^2_{adj} = 0.04; F(2,17)=3.79; p<0.05$), with coefficients ($Y_0=8.91;\beta_{charts}=0.49;\beta_{maps}=-1.05$) demonstrating that charting positively influenced time taken, while the use of maps led to shorter times.
@@
 \subsection{Condition B: Results and Observations}
-Participants were given the option of using web sites and/or a spreadsheet to make their choices, using the same data as condition A.
-There was mixed use of websites and the spreadsheet, and we measured the fraction of the trial time spent in Excel versus looking at the sites in a 5-value range (0,25,50,75,100\%).
-Six individuals used Excel entirely (100\%) without consulting the web sites, while one avoided Excel entirely.
-The uses of Excel for external cognition ranged from collation behaviour to tallying annotations, to colour-coding cells to allow easy identification of  ``pros'' and ``cons''.
-In general, users struggled with both the websites and with the spreadsheet, although could generally find the information they required after a short period of frustration.
+In condition B, participants were given the option of using websites and/or a spreadsheet containing the same data as condition A, to make their choices.
+A mixed usage of websites and the spreadsheet was observed, and we measured the fraction of the trial time spent in Excel versus looking at the sites in a 5-value range (0,25,50,75,100\%).
+Six individuals used Excel entirely (100\%) without consulting the websites, while one avoided Excel entirely.
+The uses of Excel for external cognition ranged from collation behaviour, to tallying annotations, to colour-coding cells for making easier the identification of  ``pros'' and ``cons''.
+In general, users struggled with both the websites and with the spreadsheet, although they could usually find the information they required after a short period of frustration.
 
@@
 
-When asked to rate the DP interface, all participants responded that they felt they understood the system, with sixteen (80\%) responding that they agreed \emph{very strongly}.
+When asked to rate the DP interface in the exit survey, all participants responded that they felt they understood the system, with sixteen out of the twenty participants (80\%) responding that they agreed \emph{very strongly} that they felt they understood it.
 Eighteen (90\%) reported that they thought  DP was \emph{easy}, as well as \emph{useful}, with one person being neutral on each.
-On the statement ``DP is confusing'',  11 strongly disagreed, 6 disagreed, and 3 agreed.
-Revisiting the time trials, we discovered that the 3 participants who rated DP as confusing completed the trials statistically significantly slower than the rest of the group.
-Survey results were visualised in Figure \ref{fig:survey}.
+As to the statement ``DP is confusing'',  11 (55\%) strongly disagreed, 6 (30\%) disagreed, and 3 (15\%) agreed.
+On reviewing the time trials, we discovered that the 3 participants who rated DP as confusing completed the trials significantly more slowly than the rest of the group.
+The survey results were visualised in Figure \ref{fig:survey}.
+
+% (ANOVA?)
 
@@
 
-In the exit survey, the users noted that they would use an interface like DataPalette in the future, for finding a property to rent/buy, choosing a job and purchasing electronic devices.
-However, one participant said that they would not use it for things like deciding on a restaurant because ``it would cheapen the experience'' by removing the element of spontaneity of choice, but would use it for making more important decisions.
-In order to improve our solutions the participants suggested making the origin of the information clearer, particularly when combining multiple data sources.
+The users also noted that they would use an interface like DataPalette in the future, for example, in finding a property to rent/buy, in choosing a job and for purchasing electronic devices.
+Interestingly, one participant indicated that he/she would not use it for decisions such as selecting a restaurant, because ``it would cheapen the experience'' by removing the element of spontaneity of choice; they would use it, however, for making more ``important'' decisions.
+As for ways to improve DP, the participants suggested making the origin of the information clearer, particularly when combining multiple data sources.
 The participants also requested the ability to sort instances, as well as explanations of ratings and scoring systems (however some of these are not even transparent on websites).
@@
 \section{Discussion and Limitations}
-
-In this section, we first re-visit our study hypotheses introduced in Methodology, using observations from results to support views on each.
-We follow this with a discussion of limitations regarding the design of the interface, the state of the prototype, and the design and execution of the study.
+In this section, we first revisit the three hypotheses introduced in Methodology, incorporating observations from the study's results to support views on each.
+We follow this with a discussion of the limitations regarding the design of the interface, the state of the prototype, and the design and execution of the study.
 Finally, we discuss our current and follow-on plans for continuing this line of research.
@@
 We set out to test three hypotheses with the DataPalette study.
-The first was that end-user citizens would understand the interface mechanisms of DP (h1).
-The second, h2, was that DP enables people to mix and integrate heterogeneous data in the process of performing their tasks.
-The third, h3, was that using DP facilitated task completion (h3).
-Using real data obtained from typical data sources identified during our pre-study (PS.1), we tested the DP prototype and its integration interaction mechanisms with ``real'' users, who were students, staff and alumni of our University who had varying backgrounds and levels of expertise with computers.
-
-We find substantial support for h1.
+The first was that users would understand the interface mechanisms of DP (h1).
+The second, h2, was that DP would enable people to mix and integrate heterogeneous data in the process of performing their tasks.
+The third, h3, was that using DP would facilitate task completion.
+% Using real data obtained from typical data sources identified during our pre-study (PS.1), we tested the DP prototype and its integration interaction mechanisms with ``real'' users, who were students, staff and alumni of our University who had varying backgrounds and levels of expertise with computers.
+
+The study produced substantial support for h1.
 All participants reported that they understood the tool, and that it was easy to use.
@@
 
-Additional evidence that participants had a solid understanding of the system was their feedback, specifically feature requests and desired capabilities.
-Several participants requested search functionality, sort and filter functionality, the ability to display multiple properties for a single instance simultaneously, and visualise multiple attributes in a 2 or 3-dimensional visualisation.
-Perhaps the most interesting suggestion was also the most common -- the ability to selectively view the provenance of information after instances have been combined.
+Additional evidence that participants had a solid understanding of the system was reflected in their feedback, particularly in their requests for features and desired capabilities.
+Several participants requested search functionality, sort and filter functionality, the ability to display multiple properties for a single instance simultaneously, and the ability to display multiple attributes together in a 2-or-3 dimensional visualisation.
+Perhaps the most interesting suggestion was also the most common -- the ability to selectively view the provenance of information after instances were combined.
 For example, after his trial, P15 said:
@@
 With respect to integration specifically, a majority (80\%) of the participants successfully and deliberately integrated data using drag-and-drop SameAs capabilities.
-While single-paths were used, multi-path features were not as widely used, only by 4 participants.
+While single-paths were readily used, multi-path features were not as widely used (only by 4 participants).
 We believe that participants may not have realised that creating multiple path selections per group was possible, instead assuming it to be similar to the single ``Current path'' state per window of most file managers.
 
-Assessing whether DP improved task performance (h3) was difficult given the small sample size of our study and large number of factors influencing each person's performance.
-Tasks were completed on average slightly faster than the control interface, but not statistically significantly.
-However, more data sources seemed to be consulted (we use ``seemed'' carefully because these findings approached significance) during trials with DP over the control interface, which meant that people were looking at more diverse information than the control conditions.
+Assessing whether or not DP improved task performance (h3) was difficult, given the small sample size of our study, and the large number of factors influencing each person's choice.
+Tasks were completed on average slightly faster than the control interface, but not significantly.
+However, more data sources seemed to be consulted (we use ``seemed'' carefully because these findings approached significance) during trials with DP over the control interface, which meant that participants were looking at more diverse information than the control conditions.
 We also found that participants justified their choices with a greater number of factors in DP than the control interface, suggesting that decisions may have been made considering a greater number of factors.

### 2013-01-14 chi2013-webbox.tex
edits

@@
 However, more data sources seemed to be consulted (we use ``seemed'' carefully because these findings approached significance) during trials with DP over the control interface, which meant that participants were looking at more diverse information than the control conditions.
-We also found that participants justified their choices with a greater number of factors in DP than the control interface, suggesting that decisions may have been made considering a greater number of factors.
-Participants entertained more possibilities for longer in the DP trial, while participants were more prone to eliminate candidates early in the baseline interface.
-Finally, people took fewer paper notes in the DP condition than the baseline, suggesting that there was less need for external cognitive support in DP.\looseness=-1
+We also found that participants justified their choices with a greater number of factors in DP than with the control interface, suggesting that their decisions may have been made considering a greater number of factors.
+Participants entertained more possibilities for longer in the DP trial, while they were more prone to eliminate candidates early in the baseline interface.
+Finally, people took fewer paper notes in the DP condition than with the baseline, suggesting that there was less need for external cognitive support in DP.\looseness=-1
 
@@
 \subsection{Design Limitations}
-
-In addition to the aforementioned feature requests, we observed a number of ``pain points'' in the design of the DP interface.
-One such pain point was window layout and screen real estate management; participants were prone to opening up a large number of windows, including maps and charts.
-When their screen started to fill up, we found that participants spent a lot of time arranging windows, and second, that people tended to forget what each window represented -- ``what was this again?''.
-We plan to address these issue in the next version of DataPalette.
+In addition to the aforementioned feature requests, we observed a number of potential areas for improvement in the design of the DP interface.
+One such area pertained to window layout and screen real-estate management: participants were prone to opening up a large number of windows, including maps and charts.
+When their workspaces started to fill up, we found that participants spent a lot of time arranging windows, and that they tended to forget what each window represented.
+We plan to address these issues in the next version of DataPalette.
 

### 2013-01-14 chi2013-webbox.tex
edits

@@
 This would cover a majority of the structural and modeling-oriented heterogeneity cases observed in DP.2.
-Extending the ``Do What I Mean'' capabilities of the visualisation features to be unit-of-measure-aware we think would be tremendously valuable; a core problem, however, stems from the fact that few sources explicitly annotated data values with such units.
+Extending the ``Do What I Mean'' capabilities of the visualisation features to be unit-of-measure-aware, we believe, would be valuable to users.
+A core problem with realizing such an capability, however, is that few sources explicitly annotated data values with measurement units.
 
@@
 
-Our study had several limitations.
-First, the small sample size of our study (20 participants) meant that we could not confirm some of the effects observed with statistical significance.
-Second, giving participants a time limit of 10 minutes on trials may have affected the strategies they used to make their choices.
-For this reason, in our follow-up study, we will give some participants unlimited time on tasks of similar complexity.
-A second departure from a realistic task was the way in which we pre-determined the data sources for participants to use.
+Concerning limitations of the study, the small sample size of our study (20 participants) meant that we could not confirm some of the effects observed with statistical significance.
+Second, providing participants a time limit of 10 minutes on trials may have affected the strategies they used to make their choices.
+For this reason, in our follow-up study, we will give participants a greater amount of time for tasks of similar complexity.
+A departure from realistic tasks was the way in which we pre-determined the data sources for participants to use.
 We did this in order to focus participants' efforts on looking at and working with their choices, rather than retrieving them.

### 2013-01-14 chi2013-webbox.tex
Minor typo fix

--- 2013-01-14 chi2013-webbox.tex
@@
 A core reason for this, we posit, is a lack of suitable tools for accessing and interrogating these heterogeneous data sources, as well as a means by which these disparate information sources can be effectively brought together and cross-referenced.
-For example, most personal information management (PIM) tools only provide the capacity to manage small, pre-fixed sets of data types (such as calendar appointments, address book entires,  tools and to-do list items).
+For example, most personal information management (PIM) tools only provide the capacity to manage small, pre-fixed sets of data types (such as calendar appointments, address book entries,  tools and to-do list items).
 Alternatively, they may provide little or no support for structured data, as exhibited in word processors, text editors, and sketching/drawing tools \cite{infoscraps}.

### 2013-01-14 chi2013-webbox.tex
edits

--- 2013-01-14 chi2013-webbox.tex
@@
 
-\begin{table}[htbp]
-\begin{center}
-\small
-\begin{tabular}{|p{3cm}|p{4.7cm}|}
-\hline
-Task	 & Data Sources\\
-\hline
-Three universities they would apply to study 1)~Sport Science, and 2)~History & www.thecompleteuniversityguide.co.uk, www.ucas.ac.uk, unistats.direct.gov.uk, www.timeshighereducation.co.uk/world-university-rankings\\
-\hline
-Restaurant they would book for 12 friends in 3)~Cambridge, and 4)~Glasgow & www.yelp.co.uk, www.tripadvisor.co.uk, ratings.food.gov.uk, facebook and google plus.\\
-\hline
-\end{tabular}
-\end{center}
-\caption{For each type of task we used two variables, which affected which data we used from the relevant data sources.} \label{tab:studyfactors}
-\normalsize
-\end{table}
+%% \begin{table}[htbp]
+%% \begin{center}
+%% \small
+%% \begin{tabular}{|p{3cm}|p{4.7cm}|}
+%% \hline
+%% Task	 & Data Sources\\
+%% \hline
+%% Three universities they would apply to study 1)~Sport Science, and 2)~History & www.thecompleteuniversityguide.co.uk, www.ucas.ac.uk, unistats.direct.gov.uk, www.timeshighereducation.co.uk/world-university-rankings\\
+%% \hline
+%% Restaurant they would book for 12 friends in 3)~Cambridge, and 4)~Glasgow & www.yelp.co.uk, www.tripadvisor.co.uk, ratings.food.gov.uk, facebook and google plus.\\
+%% \hline
+%% \end{tabular}
+%% \end{center}
+%% \caption{For each type of task we used two variables, which affected which data we used from the relevant data sources.} \label{tab:studyfactors}
+%% \normalsize
+%% \end{table}
 

### 2013-01-14 chi2013-webbox.tex
Spellcheck corrections.

@@
 The information processing capabilities of humans enable them to opportunistically draw and integrate knowledge from nearly any information source.
-However, the integration of digital, structured data from diverse sources remains difficult, due to problems of heterogeneity that arise when data modeled separately are brought together.
+However, the integration of digital, structured data from diverse sources remains difficult, due to problems of heterogeneity that arise when data modelled separately are brought together.
 In this paper, we present an investigation of the feasibility of extending the scope of Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, with the objective of allowing  users to take advantage of the vast, emerging ecosystems of structured data currently becoming available.
@@
 %The information processing capabilities of humans enable people to opportunistically draw and integrate knowledge from nearly any information sources that become available to them.
-However, supporting similar, opportunistic integration of digital, structured data from diverse sources remains difficult without tedious, manual integration, due to problems of heterogeneity that arise when data modeled and represented differently are brought together.
+However, supporting similar, opportunistic integration of digital, structured data from diverse sources remains difficult without tedious, manual integration, due to problems of heterogeneity that arise when data modelled and represented differently are brought together.
 In this paper, we investigate the feasibility of minimally extending digital Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, towards enabling users to take advantage of the vast emerging ecosystems of structured data being made available.
@@
 Several basic questions drove our inquiry.
-First, were people increasingly relying only on a handful of ``supersites'' (Facebook, Wikipedia), or searching across distributed sources of information?
-If the former were true, merely integrating PIM tools with the small number of supersites would provide greater benefit than tackling the more challenging problem of integrating data from an array of arbitrary sources on the Web.
+First, were people increasingly relying only on a handful of ``super-sites'' (Facebook, Wikipedia), or searching across distributed sources of information?
+If the former were true, merely integrating PIM tools with the small number of super-sites would provide greater benefit than tackling the more challenging problem of integrating data from an array of arbitrary sources on the Web.
 In a related vein, why did some people choose a single source for information gathering, while others selected multiple sources?
@@
 Thus, terminological (or naming) heterogeneity was prevalent across nearly all records observed.
-Structural inconsistencies were rarer, with the most common case occuring where a record that represented a value as a simple literal was expanded out in another as an entire sub-structure.
+Structural inconsistencies were rarer, with the most common case occurring where a record that represented a value as a simple literal was expanded out in another as an entire sub-structure.
 This most commonly occurred for fields with common string serialisations, such as dates, prices, and intervals.
@@
 In summary, we found in our PS.2, that while data records ranged in complexity, no single source subsumed any others; all contained unique properties.
-Moreover, the most common types of heterogeneity exhibited were terminological and value-related, which occurred vastly more often than structural or modeling issues.
+Moreover, the most common types of heterogeneity exhibited were terminological and value-related, which occurred vastly more often than structural or modelling issues.
 
@@
 Since individually dragging and dropping items could be tedious for large groups of items, we extended this to allow entire groups to be dropped on to other groups.
-DP would then identify matching pairs among items in the dragged and target groups by comparing the labels of items in each group, repsectively.
+DP would then identify matching pairs among items in the dragged and target groups by comparing the labels of items in each group, respectively.
 Only exact matches (excepting white space and underscores) were considered to correspond to label matches, and these were automatically combined.
@@
 In the current prototype, this ``Do What I Mean'' behaviour automatically would set whether the plot was a histogram representation (counts of values) or a numeric value display, based upon the types of the dereferenced values being plotted.
-Hence, non-numeric values would be histogrammed, while numeric values were displayed as a bar or line chart.
+Hence, non-numeric values would be shown on a histogram, while numeric values were displayed as a bar or line chart.
 Similarly, the map was made to detect address strings and latitude/longitude pairs, geocoding them where appropriate, and determining the optimal bounding box to display all items at maximum zoom.
@@
 For the university task, the six top-ranked universities for each subject according to The Complete University Guide were selected as candidates.
-We selected six in order to limit the difficulty of the task given the fixd time allocated for each trial.
+We selected six in order to limit the difficulty of the task given the fixed time allocated for each trial.
 In addition, for each course, we provided participants with the profile of a hypothetical student applying to the course, for whom they should base their choice.
@@
 For the restaurant task, we selected six restaurants at random from Yelp's list of popular eateries for each of the two cities.
-In addition, for each of the two cities, we generated 12 social network profiles for hypothetical friends who lived in either city, to accompany the participant to the restaraunt.
+In addition, for each of the two cities, we generated 12 social network profiles for hypothetical friends who lived in either city, to accompany the participant to the restaurant.
 Each friend's profile expressed their street address and preference for favourite cuisine type which could be taken into consideration by participants during the venue selection process.
@@
 However, when there was no clear aspect that was ``most important'', this  ``greedy'' strategy can result in suboptimal decisions.
-Another tactic, which we called \emph{tallying the pros and cons}, was to consider all candidate choices together, totaling up the advantages and disadvantages of each, potentially weighed by its perceived importance.
+Another tactic, which we called \emph{tallying the pros and cons}, was to consider all candidate choices together, totalling up the advantages and disadvantages of each, potentially weighed by its perceived importance.
 This strategy was less vulnerable to getting stuck in local maxima than the former.
@@
 To expand DP's integration capabilities, we are considering a number of new interaction affordances for facilitating field combination and splitting, and unit of measure reconciliation.
-This would cover a majority of the structural and modeling-oriented heterogeneity cases observed in DP.2.
+This would cover a majority of the structural and modelling-oriented heterogeneity cases observed in DP.2.
 Extending the ``Do What I Mean'' capabilities of the visualisation features to be unit-of-measure-aware, we believe, would be valuable to users.
-A core problem with realizing such an capability, however, is that few sources explicitly annotated data values with measurement units.
+A core problem with realising such an capability, however, is that few sources explicitly annotated data values with measurement units.
 
@@
 
-Due to the promising findings of this study, we plan to continue investigating lightweight, \emph{in-situ} approaches to supporting ad-hoc data integration during the course of sensemaking and decision-making activities.
+Due to the promising findings of this study, we plan to continue investigating lightweight, \emph{in-situ} approaches to supporting ad-hoc data integration during the course of sense-making and decision-making activities.
 In particular, we are examining gestural input methods for the purposes of supporting more fluid exploration and mixing.
@@
 
-% instead of being a separate step requiring pre-planning one that and that information sources could be brought in or rmeoved as needed.
-
-%% this study demonstrated that user-driven and-defined approach to information integration provided the flexibility to allow people to organize and support information gathering tasks in different ways.
+% instead of being a separate step requiring pre-planning one that and that information sources could be brought in or removed as needed.
+
+%% this study demonstrated that user-driven and-defined approach to information integration provided the flexibility to allow people to organise and support information gathering tasks in different ways.
 The result, that  people weighed more factors from a greater number of information sources, while expending equal or less time and effort, demonstrated that such tools could broaden the information gathering capacities and allow people to make better use of available information sources during decision-making activities.
 
-%% this was having a sufficiently low barrier of effort to use, provided both the flexibillity and facility to support information gathering tasks associated with decision-making activities.
-The result was people weighed more factors from a greater number of information sources, broading the scope of their information gathering for such decisions, while expending comparable or less effort and time.
+%% this was having a sufficiently low barrier of effort to use, provided both the flexibility and facility to support information gathering tasks associated with decision-making activities.
+The result was people weighed more factors from a greater number of information sources, broadening the scope of their information gathering for such decisions, while expending comparable or less effort and time.
 
@@
 
-%% the abililty to  In our investigation we found that complete approaches to data integration
+%% the ability to  In our investigation we found that complete approaches to data integration
 

### 2013-01-14 chi2013-webbox.tex
Versus consistency fix

@@
 Finally, we examined incompatibilities not covered by either naming or structural differences alone.
-These \emph{modelling differences} arose from a variety of reasons, including, but not limited to, measurement unit inconsistencies, measurement method disagreements (e.g., dew point vs relative humidity), differences in scale and granularity, and mismatches in what was being modelled (``offer'' versus ``listing'').\looseness=-1
+These \emph{modelling differences} arose from a variety of reasons, including, but not limited to, measurement unit inconsistencies, measurement method disagreements (e.g., ``dew point'' versus ``relative humidity''), differences in scale and granularity, and mismatches in what was being modelled (e.g., ``offer'' versus ``listing'').\looseness=-1
 

### 2013-01-14 chi2013-webbox.tex
candb
abstract short

@@
 However, the integration of digital, structured data from diverse sources remains difficult, due to problems of heterogeneity that arise when data modelled separately are brought together.
-In this paper, we present an investigation of the feasibility of extending the scope of Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, with the objective of allowing  users to take advantage of the vast, emerging ecosystems of structured data currently becoming available.
+In this paper, we present an investigation of the feasibility of extending Personal Information Management (PIM) tools to support lightweight, user-driven mixing of previously un-integrated data, with the objective of allowing  users to take advantage of the emerging ecosystems of structured data currently becoming available.
 In this study, we conducted an exploratory, sequential, mixed-method investigation, starting with two pre-studies of the data integration needs and challenges, respectively, of Web-based data sources.
-Observations from these pre-studies led to our designing \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
+Observations from these pre-studies led to \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
 Our lab study showed, first, that participants readily understood the new interaction mechanisms which were introduced.
-Secondly, participants made more carefully justified decisions, even while weighing a greater number of factors, moreover expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up (web page and spreadsheet interface).
+Secondly, participants made more carefully justified decisions, even while weighing a greater number of factors, moreover expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up.
+
 %The information processing capabilities of humans enable people to opportunistically draw and integrate knowledge from nearly any information sources that become available to them.
@@
 
-\keywords{Personal Information Management, Personal Data Management, User-driven data integration}
-
-\category{H.5.m.}{Information Interfaces and Presentation (e.g. HCI)}{Miscellaneous}
-
-\terms{Human Factors; Design; Measurement.}
+\keywords{Personal information management; end-user data integration; mash-ups; sensemaking with data}
+
+\category{H.5.2.}{Interaction styles}{Direct manipulation}
+\category{H.2.5.}{Heterogeneous Databases}{Data translation}
+\category{H.5.m.}{Information Interfaces and Presentation}{Miscellaneous}
+
+% \terms{Human Factors; Design; Measurement.}
 

### 2013-01-14 chi2013-webbox.tex
abstract short

@@
 
-For the B conditions we prepared, for each dataset condition in each task, a spreadsheet listing key statistics for each of the relevant candidates (e.g., universities or restaurants) assembled from each of a pre-selected set of 6 popular data sources (university and restaurant guides), and collated them into Excel spreadsheets.
+For the B conditions, we prepared a spreadsheet listing key statistics for each of the relevant candidates (e.g., universities or restaurants) using data from of a pre-selected set of 6 popular data sources (university and restaurant guides), and collated them into Excel spreadsheets.
 

### 2013-01-14 chi2013-webbox.tex
abstract short
Refs \small
foo

@@
 In recent years, an unprecedented quantity and variety of information has been made available as structured data on the Web.
-The explosion of structured data APIs and downloadable data sets has arrived from many different sectors, including retail, banking, social networking, opinion and recommendation sites, mobile and desktop apps, as well as from new kinds sensors and devices such as wearable activity and bio-sensors.
+The explosion of structured data APIs and downloadable data sets has arrived from many different sectors, including retail, banking, social networking, opinion and recommendation sites, mobile and desktop apps, as well as from new kinds of sensors and devices such as wearable activity and bio-sensors.
 Beyond commercial apps and services, governments have embraced releasing data as a way of providing transparency and accountability to their citizens, causing a huge push to release ``open data'' to the public, at scales ranging from the smallest local councils to entire national departments.
@@
 Our primary motivation stems from field studies in PIM, where the need for consolidating data has been observed.
-This is an area in which many challenges remain, however, specifically, in reconciling the differences that arise among representations when data is created by different people, at different times, and for different purposes.
+This is an area in which many challenges remain, specifically, in reconciling the differences that arise among representations when data is created by different people, at different times, and for different purposes.
 As described by Alon Halevy:
@@
 Previous systems utilising this approach have included ``Cards, Relations and Templates'' \cite{Dontcheva:2007:RCS:1294211.1294224} and Potluck \cite{citeulike:3875264} both of which permit the user to reconcile instances from different schemas into a uniform representation using drag-and-drop gestures to specify relations among fields.
-\emph{DataPalette} extends this approach to a data workbench environment which allows people to easily consolidate and compare instances by their properties.
+DataPalette extends this approach to a data workbench environment which allows people to easily consolidate and compare instances by their properties.
 
@@
 
-In the first of two pre-studies, we aimed to get an updated understanding of the use of multiple information sources in information gathering on the Web.
+In the first pre-study, we aimed to get an updated understanding of the use of multiple information sources in information gathering on the Web.
 Several basic questions drove our inquiry.
@@
 
-The 8 participants, recruited informally via word of mouth, consisted of 7 males and 1 female. They were split evenly into two age groups of 18--25 and 26--32.   
+The 8 participants, recruited informally via word of mouth, consisted of 7 males and 1 female, ranging in ages of 18--32.   
 % mention # that were computer scientists here?
@@
 %\subsubsection{Question 1: Heterogenous Data Tasks}
-All participants reported regular usage of multiple websites to accomplish tasks, examples of which included: shopping (food, consumer electronics, hotels and flights), choosing a restaurant, job-seeking, selecting a university, seeking a recipe, or finding answers to technical problems.
+All participants reported regular usage of multiple websites to accomplish tasks, examples of which included: shopping, choosing a restaurant, job-seeking, selecting a university, seeking a recipe, or finding answers to technical problems.
 When initiating a task, it was common for participants to use Google to discover several related websites, which would then be explored in different ways, depending on the task at hand.
@@
 To remain consistent with the drag-and-drop interactions used throughout the interface, we introduced the ability to simply drag one instance directly on top of another, to specify that these two items should be combined.
-This action also caused the \emph{same-as display} in the lower left of the screen to indicate this new combined relationship.
+This action also caused the \emph{Same-As display} in the lower left of the screen to indicate this new combined relationship.
 The display also acted as a mechanism to delete/undo combinations, which reverted representations of instances back to their original, separate forms.
@@
 % uncomment the following line:
-% \small
+\small
 \bibliographystyle{acm-sigchi}

### 2013-01-14 chi2013-webbox.tex
foo
latest

@@
 Observations from these pre-studies led to \emph{DataPalette}, an interface that introduced simple co-reference and group multi-path-selection mechanisms for working with terminologically and structurally heterogeneous data.
-Our lab study showed, first, that participants readily understood the new interaction mechanisms which were introduced.
-Secondly, participants made more carefully justified decisions, even while weighing a greater number of factors, moreover expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up.
+Our lab study showed that participants readily understood the new interaction mechanisms which were introduced.
+Participants made more carefully justified decisions, even while weighing a greater number of factors, moreover expending less effort, during subjective-choice tasks when using DataPalette, than with a control set-up.
 
@@
 Mapping equivalent properties among the instance records for each category revealed little overall overlap, except among the smallest records (which had few properties to begin with).
-This finding suggests that consolidating such records would substantially increase the amount of information held by each.
+This finding suggested that consolidating such records would substantially increase the amount of information held by each.
 Some of the unique properties were primarily for internal use (such as service-specific IDs).
@@
 
-When asked to rate the DP interface in the exit survey, all participants responded that they felt they understood the system, with sixteen out of the twenty participants (80\%) responding that they agreed \emph{very strongly} that they felt they understood it.
+When asked to rate the DP interface in the exit survey, all participants responded that they felt they understood the system, with sixteen out of the twenty participants (80\%) responding that they agreed \emph{very strongly}.
 Eighteen (90\%) reported that they thought  DP was \emph{easy}, as well as \emph{useful}, with one person being neutral on each.
@@
 On reviewing the time trials, we discovered that the 3 participants who rated DP as confusing completed the trials significantly more slowly than the rest of the group.
-The survey results were visualised in Figure \ref{fig:survey}.
+(See  Figure \ref{fig:survey}.)
 
@@
 When their workspaces started to fill up, we found that participants spent a lot of time arranging windows, and that they tended to forget what each window represented.
-We plan to address these issues in the next version of DataPalette.
+We plan to address these issues in the next version of DP.
 
@@
 To expand DP's integration capabilities, we are considering a number of new interaction affordances for facilitating field combination and splitting, and unit of measure reconciliation.
-This would cover a majority of the structural and modelling-oriented heterogeneity cases observed in DP.2.
+This would cover a majority of the structural and modelling-oriented heterogeneity cases observed in PS.2.
 Extending the ``Do What I Mean'' capabilities of the visualisation features to be unit-of-measure-aware, we believe, would be valuable to users.
@@
 %% Moreover, through our pre-studies and DataPalette evaluation, we have shown that much of heterogeneity exhibited by structured data feeds available on the Web today can be managed reasonably through easy to use, and intuitive interactive approaches for end-users.
+
+\newpage
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
Skeleton version of PCC workshop paper

\documentclass{chi-ext}
% Please be sure that you have the dependencies (i.e., additional LaTeX packages) to compile this example.
% See http://personales.upv.es/luileito/chiext/

\copyrightinfo{
  Copyright is held by the author/owner(s).\\
  \emph{CHI'13}, April 27 -- May 2, 2013, Paris, France.\\
  ACM 978-1-XXXX-XXXX-X/XX/XX.\\
}

\title{Personal Social Machines}

\numberofauthors{3}
% Notice how author names are alternately typesetted to appear ordered in 2-column format;
% i.e., the first 4 autors on the first column and the other 4 auhors on the second column.
% Actually, it's up to you to strictly adhere to this author notation.
\author{
  \vspace{-1.5em} % lisatolles: The abstract heading should start at the time height on the page as the authors names
  \alignauthor{
  	\textbf{Daniel A. Smith}\\
  	\affaddr{Web and Internet Science Research Group}\\
%  	\affaddr{Electronics and Computer Science}\\
  	\affaddr{University of Southampton}\\
  	\affaddr{Southampton, UK}\\
  	\email{ds@ecs.soton.ac.uk}
  }\alignauthor{
  	\textbf{Max Van Kleek}\\
  	\affaddr{Web and Internet Science Research Group}\\
%  	\affaddr{Electronics and Computer Science}\\
  	\affaddr{University of Southampton}\\
  	\affaddr{Southampton, UK}\\
  	\email{emax@ecs.soton.ac.uk}
  }
  \vfil
  \alignauthor{
  	\textbf{Nigel R. Shadbolt}\\
  	\affaddr{Web and Internet Science Research Group}\\
%  	\affaddr{Electronics and Computer Science}\\
  	\affaddr{University of Southampton}\\
  	\affaddr{Southampton, UK}\\
  	\email{nrs@ecs.soton.ac.uk}
  }
  \vfil
  \vfil
  \vfil
  \vfil
}

% Paper metadata (use plain text, for PDF inclusion and later re-using, if desired)
\def\plaintitle{Personal Social Machines}
\def\plainauthor{Daniel A. Smith, Max Van Kleek, Nigel R. Shadbolt}
\def\plainkeywords{social machines, personal data, linked data}
\def\plaingeneralterms{Social Machines, Personal Data}

\hypersetup{
  % Your metadata go here
  pdftitle={\plaintitle},
  pdfauthor={\plainauthor},  
  pdfkeywords={\plainkeywords},
  pdfsubject={\plaingeneralterms},
  % Quick access to color overriding:
  %citecolor=black,
  %linkcolor=black,
  %menucolor=black,
  %urlcolor=black,
}

\usepackage{graphicx}   % for EPS use the graphics package instead
\usepackage{balance}    % useful for balancing the last columns
\usepackage{bibspacing} % save vertical space in references

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\keywords{\plainkeywords}
%\textcolor{red}{Mandatory section to be included in your final version.}

\category{H.5.m}{Information interfaces and presentation (e.g., HCI)}{Miscellaneous}. 
%See \cite{ACMCCS} 
%See: \url{http://www.acm.org/about/class/1998/} 
%or help using the ACM Classification system.
%\textcolor{red}{Mandatory section to be included in your final version.}

%\terms{\plaingeneralterms}
%\textcolor{red}{Optional section to be included in your final version.}

\section{Introduction}

%% =============================================================================
%\section{Introduction}
%% =============================================================================
%This format is to be used for submissions that are published in the conference extended abstracts.  
%We wish to give this volume a consistent, high-quality appearance. 
%We therefore ask that authors follow some simple guidelines. 
%In essence, you should format your paper exactly like this document. 
%The easiest way to do this is simply to download a template from the conference website and replace the content with your own material.
%
%% =============================================================================
%\section{Copyright}
%% =============================================================================
%For publications in the CHI Extended Abstracts, copyright remains with the author.  
%The publication is not considered an archival publication; however, it does go into the ACM Digital Library. 
%Because you retain copyright, as the author you are free to use this material as you like, including submitting a paper based on this work to other conferences or journals.  
%Authors grant unrestricted permission for ACM to publish the accepted submission in the CHI Extended Abstracts without additional consideration or remuneration.
%
%% =============================================================================
%\section{Text formatting}
%% =============================================================================
%Please use an 8.5-point Verdana font, or other sans serifs font as close as possible in appearance to Verdana in which these guidelines have been set. 
%Arial 9-point font is a reasonable substitute for Verdana as it has a similar x-height. 
%Please use serif or non-proportional fonts only for special purposes, such as distinguishing source code text.
%Additionally, here is an example of footnoted text.\footnote{Use footnotes sparingly, if at all.}
%As stated in the footnote, footnotes should rarely be used.
%
%\subsection{Language, style, and content}
%% -----------------------------------------------------------------------------
%The written and spoken language of SIGCHI is English. 
%Spelling and punctuation may use any dialect of English (e.g., British, Canadian, US, etc.) provided this is done consistently. 
%Hyphenation is optional. 
%To ensure suitability for an international audience, please pay attention to the following:
%
%\begin{itemize}\compresslist
%\item 	
%Write in a straightforward style. 
%Use simple sentence structure. 
%Try to avoid long sentences and complex sentence structures. 
%Use semicolons carefully.
%\item 	
%Use common and basic vocabulary (e.g., use the word ``unusual" rather than the word ``arcane").
%\item 	
%Briefly define or explain all technical terms. 
%The terminology common to your practice/discipline may be different in other design practices/disciplines.
%\item 	
%Spell out all acronyms the first time they are used in your text. 
%For example, ``World Wide Web (WWW)".
%\item 	
%Explain local references (e.g., not everyone knows all city names in a particular country).
%\item 	
%Explain ``insider" comments. 
%Ensure that your whole audience understands any reference whose meaning you do not describe (e.g., do not assume that everyone has used a Macintosh or a particular application).
%\item 	
%Explain colloquial language and puns. 
%Understanding phrases like ``red herring" requires a cultural knowledge of English. 
%Humor and irony are difficult to translate.
%\item 	
%Use unambiguous forms for culturally localized concepts, such as times, dates, currencies and numbers (e.g., ``1-5-97" or ``5/1/97" may mean 5 January or 1 May, and ``seven o'clock" may mean 7:00 am or 19:00).
%\item 	
%Be careful with the use of gender-specific pronouns (he, she) and other gender-specific words (chairman, manpower, man-months). 
%Use inclusive language (e.g., she or he, they, chair, staff, staff-hours, person-years) that is gender-neutral. 
%If necessary, you may be able to use ``he" and ``she" in alternating sentences, so that the two genders occur equally often~\cite{Schwartz95}. 
%\end{itemize}
%
%
%% =============================================================================
%\section{Figures}
%% =============================================================================
%The examples on this and following pages should help you get a feel for how screen-shots and other figures should be placed in the template. 
%Be sure to make images large enough so the important details are legible and clear.
%
%%\begin{figure}
%%  \centering
%%  \includegraphics[width=\linewidth]{sample.jpg}
%%  \caption{Insert a caption below each figure.}
%%  \label{fig:sample}
%%\end{figure}
%
%Your document may use color figures, which are included in the page limit; the figures must be usable when printed in black and white.
%You can use the \LaTeX's \texttt{marginpar} command to insert figures in the (right) margin side of the document (see \autoref{fig:marginparsample}).
%
%
%% =============================================================================
%\section{References and Citations}
%% =============================================================================
%Use a numbered list of references at the end of the article, ordered alphabetically by first author, and referenced by numbers in brackets \cite{Anderson92,Klemmer02,Mather00,Zellweger01}
%For papers from conference proceedings, include the title of the paper and an abbreviated name of the conference (e.g., for Interact 2003 proceedings, use Proc. Interact 2003). 
%Do not include the location of the conference or the exact date; do include the page numbers if available. 
%See the examples of citations at the end of this document. 
%
%Your references should be published materials accessible to the public.  
%Internal technical reports may be cited only if they are easily accessible (i.e., you provide the address for obtaining the report within your citation) and may be obtained by any reader for a nominal fee.  
%Proprietary information may not be cited. 
%Private communications should be acknowledged in the main text, not referenced  (e.g., [Robertson, personal communication]).
%
%
%% =============================================================================
%\section{Producing and testing PDF files}
%% =============================================================================
%We recommend that you produce a PDF version of your submission well before the final deadline. 
%Besides making sure that you are able to produce a PDF, you will need to check that (a) the length of the file remains within the submission category's page limit, (b) the PDF file size is 4 megabytes or less, and (c) the file can be read and printed using Adobe Acrobat Reader. 
%Test your PDF file by viewing or printing it with the same software we will use when we receive it, Adobe Acrobat Reader Version 7. 
%This is widely available at no cost from~\cite{Acrobat7}.  
%Note that most reviewers will use a North American/European version of Acrobat reader, which cannot handle documents containing non-North American or non-European fonts (e.g. Asian fonts).  
%Please therefore do not use Asian fonts, and verify this by testing with a North American/European Acrobat reader (obtainable as above). Something as minor as including a space or punctuation character in a two-byte font can render a file unreadable.
%
%
%% =============================================================================
%\section{Dummy text}
%% =============================================================================
%Lorem ipsum dolor sit amet, consectetur adipiscing elit. Duis ut eros semper lectus vehicula elementum. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices posuere cubilia Curae; Aliquam quis mi sapien. Suspendisse potenti. Mauris ultrices euismod velit sed dictum. Nullam auctor, nulla tincidunt dapibus suscipit, velit leo convallis metus, vel commodo libero erat in dolor. In laoreet porttitor ligula, porta blandit lectus consequat quis. 
%
%Nam ut eros dui. Mauris volutpat elit metus, euismod pellentesque purus. In hac habitasse platea dictumst. Nullam consectetur lacinia interdum. Sed nec blandit nisi. Proin in nulla purus, sit amet iaculis tortor. Ut dapibus pellentesque nulla in interdum. Nunc at velit felis. Curabitur euismod neque eu orci varius in pharetra sem interdum. Morbi in mauris ac risus iaculis dapibus id in magna. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
%
%%\marginpar{
%%\begin{figure}
%%  \begin{center}
%%  \includegraphics[width=\marginparwidth]{sample.jpg}
%%  \caption{A marginal figure.}
%%  \label{fig:marginparsample}
%%  \end{center}  
%%\end{figure}
%%}
%
%Aliquam consectetur quam sed odio varius vitae rhoncus urna fermentum. Phasellus viverra diam non justo porttitor varius. Integer ultrices accumsan lectus eget mollis. Nulla et leo sit amet nunc ornare rutrum sit amet ac dui. Cras vehicula accumsan purus nec fermentum. Mauris viverra condimentum metus, ut posuere quam laoreet nec. Phasellus massa tellus, ullamcorper nec porta sed, aliquet vitae nulla. Phasellus non tortor mauris. Cras ullamcorper egestas erat, vel rutrum elit viverra a. Donec in nisl ut est facilisis blandit. Quisque congue accumsan risus, ut venenatis magna vulputate vel. Nam commodo sapien vel mauris adipiscing nec dictum quam congue. Phasellus tempor vestibulum nisl quis blandit. Nullam condimentum auctor nibh, quis elementum libero tristique.

%\section{Acknowledgements}
%We thank all DUX 2003 publications support and staff who wrote this document originally and allowed us to modify it for this conference.
%This template was based on Manas Tungare's \texttt{chi.cls}, and rewritten by Luis A. Leiva.

\balance
\bibliographystyle{acm-sigchi}
\bibliography{sample}

\end{document}

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
Correct title/idea for this workshop

@@
 
-\title{Personal Social Machines}
+\title{Using Distributed Information Stores for Patient-Clinician Health Data Sharing}
 
@@
 % Paper metadata (use plain text, for PDF inclusion and later re-using, if desired)
-\def\plaintitle{Personal Social Machines}
+\def\plaintitle{Using Distributed Information Stores for Patient-Clinician Health Data Sharing}
 \def\plainauthor{Daniel A. Smith, Max Van Kleek, Nigel R. Shadbolt}
-\def\plainkeywords{social machines, personal data, linked data}
-\def\plaingeneralterms{Social Machines, Personal Data}
+\def\plainkeywords{personal data, patient care, healthcare, information stores}
+\def\plaingeneralterms{Personal Data, Patient Care, Healthcare, Information Stores}
 
@@
 \begin{abstract}
+
+Electronic integrated health records have...
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
Abstract start

@@
 
-Electronic integrated health records have...
+Existing approaches to electronic integrated health records have largely focused on centralised information architectures, where patient data is held in a single repository, and accessed over the internet.
+However, there are a number of benefits to both patients and clinicians in a decentralised information approach to health records.
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
some intro blah

@@
 \section{Introduction}
+
+A key goal in improving patient-clinician communication is
+establishing a common context for communication with which a shared
+understanding can be established.  This understanding is
+bi-directional, in the sense that, both the patient needs to be able
+to properly interpret what the clincian is trying to convey to him or
+her, while the clinician needs to be able to arrive at an accurate
+characterisation of the patient's situation, as updated since the last
+time he or she met with the patient.
+
+One of the most promising technologies to help provide a common
+context of communication is the patient-centric electronic health
+record\cite{ehr} which constitute a complete medical history of
+a patient as recorded through his or her entire life. (...)
+
+% challenges to 
+% background on authors
+%
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
Adding quantified self intro and product references.

@@
 
+\usepackage{url}
+
 \begin{document}
@@
 %
+
+\section{Personal Electronic Sensors}
+
+There has been a surge in popularity of personal electronic sensors as
+part of the ``quantified-self'' movement to measure and improve one's health.
+Wireless sensors such as the Fit-bit and Nike FuelBand and small and discreet,
+yet provide the wearer with a wealth of information about their own activity level.
+In particular, wearers have a desire to improve their health and fitness, based on
+the feedback they receive from their device.
+
+The measurements from such devices are of particular interest because of the
+frequency and granularity of measurement that they perform - showing minute-by-minute updates
+throughout the day, with no clinical involvement.
+
+In addition to fitness monitors, the market for quantified self sensors has
+expanded into areas of personal health. For example, there are sensors for
+measuring blood oxygenation~\footnote{SpO2 Blood Oxygenation sensor: \url{http://ispo2.com}},
+Heart Rate Variability and Respiratory Rate~\footnote{Zensorium Tinke, heart and respiratory rate sensor: \url{http://www.zensorium.com/tinke/}} and
+Temperature and Galvanic Skin Response~\footnote{BodyMedia Core 2, temperature and GSR sensor: \url{http://www.bodymedia.com}}.
+There is even a wireless fork that informs you how long you spent eating~\footnote{HapiLabs HapiFork: \url{http://hapilabs.com/products-hapifork.asp}}.
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
Minor aesthetic fixes

@@
 expanded into areas of personal health. For example, there are sensors for
-measuring blood oxygenation~\footnote{SpO2 Blood Oxygenation sensor: \url{http://ispo2.com}},
-Heart Rate Variability and Respiratory Rate~\footnote{Zensorium Tinke, heart and respiratory rate sensor: \url{http://www.zensorium.com/tinke/}} and
-Temperature and Galvanic Skin Response~\footnote{BodyMedia Core 2, temperature and GSR sensor: \url{http://www.bodymedia.com}}.
-There is even a wireless fork that informs you how long you spent eating~\footnote{HapiLabs HapiFork: \url{http://hapilabs.com/products-hapifork.asp}}.
+measuring blood oxygenation\footnote{SpO2 Blood Oxygenation sensor: \url{http://ispo2.com}},
+Heart Rate Variability and Respiratory Rate\footnote{Zensorium Tinke, heart and respiratory rate sensor: \newline \indent \indent \indent \url{http://www.zensorium.com/tinke/}} and
+Temperature and Galvanic Skin Response\footnote{BodyMedia Core 2, temperature and GSR sensor: \newline \indent \indent \indent  \url{http://www.bodymedia.com}}.
+The diet aspect of weight loss is also covered, by products such as a wireless fork that informs you how long you spent eating\footnote{HapiLabs HapiFork: \newline \indent \indent \indent  \url{http://hapilabs.com/products-hapifork.asp}}.
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
More about electronic sensors

@@
 The diet aspect of weight loss is also covered, by products such as a wireless fork that informs you how long you spent eating\footnote{HapiLabs HapiFork: \newline \indent \indent \indent  \url{http://hapilabs.com/products-hapifork.asp}}.
+
+Currently these products tend to work in isolation, where the readings from a single product
+are presented either on the device themselves, or via a smartphone or website. However,
+their use could be integrated with clinical care, in order to monitor patients, and to give an
+insight into their lifestyle.
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
foo

--- 2013-01-15 patient-clinician-workshop/pccworkshop.tex
@@
 
-\usepackage{url}
-
 \begin{document}
@@
 understanding can be established.  This understanding is
-bi-directional, in the sense that, both the patient needs to be able
-to properly interpret what the clincian is trying to convey to him or
-her, while the clinician needs to be able to arrive at an accurate
-characterisation of the patient's situation, as updated since the last
-time he or she met with the patient.
+bi-directional, in the sense that the patient needs to be able to
+effectively interpret what the clincian is trying to convey to him or
+her, but also the clinician needs to be able to arrive at an accurate
+characterisation of the patient's situation (often under time
+constraints) as updated since the last time the clinician met with the
+patient.
 
-One of the most promising technologies to help provide a common
-context of communication is the patient-centric electronic health
-record\cite{ehr} which constitute a complete medical history of
-a patient as recorded through his or her entire life. (...)
+One of the most promising technologies that has the potential to help
+provide such a common context is the patient-centric electronic health
+record (henceforth EHR)\cite{ehr}, a digital representation of the
+complete medical history of an individual as recorded by clinicians
+throughout a patient's life. The EHR promises to enable both
+patients and their clinicians unfettered access to the patient's
+medical data, including diagnostics, reports, and history of
+interventions.  This means that clinicians can more easily review a
+patient's long-term history in the course of basing interpretations
+and interventions. Moreover, since patients have access to their own
+medical histories, there is (theoretically) a basis by which patients
+can understand the justifications for clinicial prognoses.
+
+% proactive activity recording to provide sketches of a person's
+% activity profiles outside of the hospital, streaming data
+
+Yet so far, the ``medical history'' that EHRs have been used to store
+have constituted only the official diagnoses and diagnostic reports
+prepared by clinicians, technicians and nurses
 
@@
 %
-
-\section{Personal Electronic Sensors}
-
-There has been a surge in popularity of personal electronic sensors as
-part of the ``quantified-self'' movement to measure and improve one's health.
-Wireless sensors such as the Fit-bit and Nike FuelBand and small and discreet,
-yet provide the wearer with a wealth of information about their own activity level.
-In particular, wearers have a desire to improve their health and fitness, based on
-the feedback they receive from their device.
-
-The measurements from such devices are of particular interest because of the
-frequency and granularity of measurement that they perform - showing minute-by-minute updates
-throughout the day, with no clinical involvement.
-
-In addition to fitness monitors, the market for quantified self sensors has
-expanded into areas of personal health. For example, there are sensors for
-measuring blood oxygenation~\footnote{SpO2 Blood Oxygenation sensor: \url{http://ispo2.com}},
-Heart Rate Variability and Respiratory Rate~\footnote{Zensorium Tinke, heart and respiratory rate sensor: \url{http://www.zensorium.com/tinke/}} and
-Temperature and Galvanic Skin Response~\footnote{BodyMedia Core 2, temperature and GSR sensor: \url{http://www.bodymedia.com}}.
-There is even a wireless fork that informs you how long you spent eating~\footnote{HapiLabs HapiFork: \url{http://hapilabs.com/products-hapifork.asp}}.
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex

@@
 %
+
+\section{Personal Electronic Sensors}
+
+There has been a surge in popularity of personal electronic sensors as
+part of the ``quantified-self'' movement to measure and improve one's health.
+Wireless sensors such as the Fit-bit and Nike FuelBand and small and discreet,
+yet provide the wearer with a wealth of information about their own activity level.
+In particular, wearers have a desire to improve their health and fitness, based on
+the feedback they receive from their device.
+
+The measurements from such devices are of particular interest because of the
+frequency and granularity of measurement that they perform - showing minute-by-minute updates
+throughout the day, with no clinical involvement.
+
+In addition to fitness monitors, the market for quantified self sensors has
+expanded into areas of personal health. For example, there are sensors for
+measuring blood oxygenation\footnote{SpO2 Blood Oxygenation sensor: \url{http://ispo2.com}},
+Heart Rate Variability and Respiratory Rate\footnote{Zensorium Tinke, heart and respiratory rate sensor: \newline \indent \indent \indent \url{http://www.zensorium.com/tinke/}} and
+Temperature and Galvanic Skin Response\footnote{BodyMedia Core 2, temperature and GSR sensor: \newline \indent \indent \indent  \url{http://www.bodymedia.com}}.
+The diet aspect of weight loss is also covered, by products such as a wireless fork that informs you how long you spent eating\footnote{HapiLabs HapiFork: \newline \indent \indent \indent  \url{http://hapilabs.com/products-hapifork.asp}}.
+
+Currently these products tend to work in isolation, where the readings from a single product
+are presented either on the device themselves, or via a smartphone or website. However,
+their use could be integrated with clinical care, in order to monitor patients, and to give an
+insight into their lifestyle.
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
phrs

--- 2013-01-15 patient-clinician-workshop/pccworkshop.tex
@@
 \section{Introduction}
-
 A key goal in improving patient-clinician communication is
@@
 understanding can be established.  This understanding is
-bi-directional, in the sense that, both the patient needs to be able
-to properly interpret what the clincian is trying to convey to him or
-her, while the clinician needs to be able to arrive at an accurate
-characterisation of the patient's situation, as updated since the last
-time he or she met with the patient.
+bi-directional, in the sense that the patient needs to be able to
+effectively interpret what the clincian is trying to convey to him or
+her, but also that the clinician needs to be able to arrive at an
+accurate characterisation of the patient's situation and condition, as
+updated since the patient was last seen, often under time pressure and
+other constraints.
 
-One of the most promising technologies to help provide a common
-context of communication is the patient-centric electronic health
-record\cite{ehr} which constitute a complete medical history of
-a patient as recorded through his or her entire life. (...)
+One of the most promising technologies that has the potential to help
+provide such a common context is the patient-centric electronic health
+record (henceforth EHR)\cite{ehr}, a digital representation of the
+complete medical history of an individual as recorded by clinicians
+throughout a patient's life. The EHR has enabled both patients and
+their clinicians unfettered access to the patient's medical data,
+including diagnostics, reports, and history of interventions.  This
+means that clinicians can more easily review a patient's long-term
+history even under time pressure and constriants. Moreover, since
+patients have access to their own medical histories, they can access
+the raw information that was the basis for a prognosis or decision.
 
-% challenges to 
-% background on authors
-%
+% proactive activity recording to provide sketches of a person's
+% activity profiles outside of the hospital, streaming data
+
+Yet, so far, EHRs have primarily been limited to the storage of
+official diagnoses and diagnostic reports prepared by clinicians,
+technicians and nurses.  These communications are both highly
+technical, making them difficult to interpret by end-user patients,
+but also highly limited in temporal resolution, reflecting only cases
+where a person was attended to and diagnostic-tested by a medical
+professional.  Although this may be frequent for patients in inpatient
+care, those that come to see their clinician, say, once a month end up
+with extremely low-resolution views of a patient's condition.
+
+In this position paper, we look at the potential and problems of
+augmenting the next generation of EHRs with \emph{personal activity
+  diaries}, portraits of a person's dynamic condition captured at high
+temporal fidelity.  Such activity diaries, we posit, could be derived
+from a combination of automatically-sensed activity and vital
+statistics through a wide array of available consumer-grade
+network-enabled activity senors and diagnostic tools, as well as
+through manual annotations provided by patients about their day.
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
phrs

@@
 accurate characterisation of the patient's situation and condition, as
-updated since the patient was last seen, often under time pressure and
-other constraints.
+updated since the patient was last seen. This is often under time
+pressure and with limited access to information about how the patient
+has been in the interim.
 
@@
 Yet, so far, EHRs have primarily been limited to the storage of
-official diagnoses and diagnostic reports prepared by clinicians,
-technicians and nurses.  These communications are both highly
-technical, making them difficult to interpret by end-user patients,
-but also highly limited in temporal resolution, reflecting only cases
-where a person was attended to and diagnostic-tested by a medical
-professional.  Although this may be frequent for patients in inpatient
-care, those that come to see their clinician, say, once a month end up
-with extremely low-resolution views of a patient's condition.
+official diagnostic reports prepared by clinicians, technicians and
+nurses.  These communications are both highly technical, making them
+difficult to interpret by end-user patients, but also highly limited
+in temporal resolution, reflecting only cases where a person was
+attended to and diagnostic-tested by a medical professional.  Although
+this may be frequent for patients in inpatient care, those that come
+to see their clinician, say, once a month end up with extremely
+low-resolution views of a patient's condition.
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
phrs

@@
 to see their clinician, say, once a month end up with extremely
-low-resolution views of a patient's condition.
+low-resolution samples of a patient's condition.
 

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
phrs

@@
   diaries}, portraits of a person's dynamic condition captured at high
-temporal fidelity.  Such activity diaries, we posit, could be derived
-from a combination of automatically-sensed activity and vital
-statistics through a wide array of available consumer-grade
-network-enabled activity senors and diagnostic tools, as well as
-through manual annotations provided by patients about their day.
+temporal fidelity.  Such activity diaries are intended to supplement,
+rather than supplant, the existing diagnostic reports performed by
+professionals, first and foremost, to help clinicians derive a more
+complete picture of the person throughout a period of care, such as an
+intervention.  We believe that such activity diaries could potentially
+also be beneficial to individuals as well, making people more aware of
+the evolution of their health.
 
-\section{Personal Electronic Sensors}
+Nonetheless, challenges abound. These primarily pertain to capturing
+these activity diaries, effectively storing these diaries, and privacy
+controls for giving patients the ability to disclose desired quantities
+of their diary to their clinicians.  We discuss these, next.
 
+\section{Capturing activity diaries}
 There has been a surge in popularity of personal electronic sensors as

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
phrs

@@
 \section{Capturing activity diaries}
+
+\subsection{Personal Electronic Sensors}
 There has been a surge in popularity of personal electronic sensors as

### 2013-01-15 patient-clinician-workshop/pccworkshop.tex
phrs

@@
 Nonetheless, challenges abound. These primarily pertain to capturing
-these activity diaries, effectively storing these diaries, and privacy
-controls for giving patients the ability to disclose desired quantities
-of their diary to their clinicians.  We discuss these, next.
+these activity diaries, effectively storing these diaries, giving
+users easy-to-use access to these diaries, and privacy controls for
+giving patients the ability to disclose desired quantities of their
+diary to their clinicians.  We discuss these, next.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
more edits

@@
 
-\subsection{Personal Electronic Sensors}
-There has been a surge in popularity of personal electronic sensors as
-part of the ``quantified-self'' movement to measure and improve one's health.
-Wireless sensors such as the Fit-bit and Nike FuelBand and small and discreet,
-yet provide the wearer with a wealth of information about their own activity level.
-In particular, wearers have a desire to improve their health and fitness, based on
-the feedback they receive from their device.
-
-The measurements from such devices are of particular interest because of the
-frequency and granularity of measurement that they perform - showing minute-by-minute updates
-throughout the day, with no clinical involvement.
+\subsection{Automatic activity sensing and digital well-being}
+The recent proliferation in personal wearable activity sensors and
+network and storage- enabled digital health measurement devices has
+been driven by both the falling costs of embeddable electronics, and
+the demand for self-measurement from the ``quantified-self''\cite{}
+movement.  Wireless sensors such as the Fit-bit\cite{} and Nike
+FuelBand\cite{} and small, discreet, and low-cost, providing the
+wearer with high-resolution temporal data about their activity levels,
+with little or no effort.
 
@@
 insight into their lifestyle.
+
+\section{Storing activity diaries}
+
+\section{Representing activity diaries}
+
+\section{Privacy Controls}
+
+\section{Conclusion}
+
+\section{Biographies}
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
PCC Workshop update

@@
 
+The data collected by patients' devices has the potential to create a large volume of data, and therefore particular
+consideration should be paid about who and where the data will be stored. In particular, it is likely
+that existing providers of storage for electronic health records will not want to be burdened with the
+administration costs of storing additional data. The longevity of third-party storage providers is
+also a concern, because there has been a history of short-notice closures of EHR providers (such as Google Health).
+
 \section{Representing activity diaries}
 
+Particular care should be taken when deciding on a format to store the activity diaries. Ideally
+a format will be machine-readable, using open standards (for example XML, RDF or JSON) and
+following existing schematic/ontological knowledge modelling conventions. There are a 
+number of medical ontologies that already exist, as well as vocabularies and systems (such as
+SNOMED\cite{} and IDC-10) that focus on the medical domain. The hope is that by using
+such systems, that the data collected by patients can be analysed more rapidly by clinical systems
+and staff.
+
 \section{Privacy Controls}
+
+One of the major concerns with storage of patient data is privacy. Indeed, there is also potentially
+legal requirements on storing personal health data, depending on the nature of the data being
+stored.
 
@@
 \section{Biographies}
+?
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
More about representation in PCC Workshop

@@
 
+One option is to embrace the Semantic Web, because it offers several key benefits:
+
+\begin{itemize}
+
+\item Distributed system with universal addresses and standard mature protocols
+\item Open extensible ontology system and existing domain ontologies
+\item Existing library and tool support and expertise
+\item Interoperability with existing systems
+
+\end{itemize}
+
+Even if personal health records or systems do not use the Semantic Web, there are existing
+techniques to integrate with their data.
+
 \section{Privacy Controls}
@@
 stored.
+
+However, it is important that control of data be sufficiently granular so that patients can choose
+to sure it with clinicians, and that clinicians can refer to it in their notes in the patient's health
+record.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
capturing

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 
-\subsection{Automatic activity sensing and digital well-being}
 The recent proliferation in personal wearable activity sensors and
 network and storage- enabled digital health measurement devices has
-been driven by both the falling costs of embeddable electronics, and
-the demand for self-measurement from the ``quantified-self''\cite{}
-movement.  Wireless sensors such as the Fit-bit\cite{} and Nike
-FuelBand\cite{} and small, discreet, and low-cost, providing the
-wearer with high-resolution temporal data about their activity levels,
-with little or no effort.
+been driven by both the falling costs of embeddable electronics, and a
+surging public interest in ``quantified-self''-style self-monitoring
+for self-improvement.  Many of these sensors are lightweight,
+can be worn discreetly, measure several physiological statistics
+simultaneously, and most require little maintenance. Table
+\ref{sensors} lists a variety of such sensors, currently available
+readily from online retailers.
 
-In addition to fitness monitors, the market for quantified self sensors has
-expanded into areas of personal health. For example, there are sensors for
-measuring blood oxygenation\footnote{SpO2 Blood Oxygenation sensor: \url{http://ispo2.com}},
-Heart Rate Variability and Respiratory Rate\footnote{Zensorium Tinke, heart and respiratory rate sensor: \newline \indent \indent \indent \url{http://www.zensorium.com/tinke/}} and
-Temperature and Galvanic Skin Response\footnote{BodyMedia Core 2, temperature and GSR sensor: \newline \indent \indent \indent  \url{http://www.bodymedia.com}}.
-The diet aspect of weight loss is also covered, by products such as a wireless fork that informs you how long you spent eating\footnote{HapiLabs HapiFork: \newline \indent \indent \indent  \url{http://hapilabs.com/products-hapifork.asp}}.
+\begin{table*}[htb]
+\small
+\begin{tabular}{p{5.0cm}  p{4cm} l}
+
+ measurement & products & price range \\
+\hline
+general activity (walking/jogging) & FitBit, Nike FuelBand  & 49-199 \\ 
+weight & Writhings WiFi Scale & 200 \\
+blood pressure & Writhings BPM & 200 \\ 
+sleep duration & Zeo Pro, WakeMate, & 100-300  \\
+heart rate variability & Zensorium Tinke & 200 \\
+blood oxygenation & Zensorium Tinke & 200 \\ 
+respiratory rate & Zensorium Tinke & 200 \\ 
+galvanic skin repsonse & BodyMedia CORE 2 &  50-100 \\
+body temperature & BodyMedia CORE 2 &  50-100 \\ 
+Eating & Hapifork & (not yet released) \\ 
+\end{tabular}
+\caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurememnt sensors with automatic data uplink to cloud services }
+\end{table*}
+
+While few of these devices had been independently evaluated for
+accuracy for clinical use, as they have matured, several of these
+manufacturers begun rigorous evaluaton of their products for achieving
+certification. For example, BodyMedia advertise themselves as
+providing the ``FDA-approved, clincally validated, most accurate
+wireless activity measurement''.  A separate study of activity-based
+sleep monitoring tools revealed that ``low cost actigraphy-based
+approaches'' correlated well with baseline (in-lab EEG-based
+measurements) for sleep duration, although less with sleep quality
+metrics.
+
+%% In addition to fitness monitors, the market for quantified self sensors has
+%% expanded into areas of personal health. For example, there are sensors for
+%% measuring blood oxygenation\footnote{SpO2 Blood Oxygenation sensor: \url{http://ispo2.com}},
+%% Heart Rate Variability and Respiratory Rate\footnote{Zensorium Tinke, heart and respiratory rate sensor: \newline \indent \indent \indent \url{http://www.zensorium.com/tinke/}} and
+%% Temperature and Galvanic Skin Response\footnote{BodyMedia Core 2, temperature and GSR sensor: \newline \indent \indent \indent  \url{http://www.bodymedia.com}}.
+%% The diet aspect of weight loss is also covered, by products such as a wireless fork that informs you how long you spent eating\footnote{HapiLabs HapiFork: \newline \indent \indent \indent  \url{http://hapilabs.com/products-hapifork.asp}}.
+
+\section{Storing activity diaries}
 
@@
 insight into their lifestyle.
-
-\section{Storing activity diaries}
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
table in margin
added label
More collection words:
pivot

@@
 
-\begin{table*}[htb]
+\newpage
+
+\marginpar{
+\begin{table}
 \small
-\begin{tabular}{p{5.0cm}  p{4cm} l}
+\begin{tabular}{p{3cm}  p{6cm} p{1.5cm}}
 
- measurement & products & price range \\
+product & measurement indices & price range (approx. USD) \\
 \hline
-general activity (walking/jogging) & FitBit, Nike FuelBand  & 49-199 \\ 
-weight & Writhings WiFi Scale & 200 \\
-blood pressure & Writhings BPM & 200 \\ 
-sleep duration & Zeo Pro, WakeMate, & 100-300  \\
-heart rate variability & Zensorium Tinke & 200 \\
-blood oxygenation & Zensorium Tinke & 200 \\ 
-respiratory rate & Zensorium Tinke & 200 \\ 
-galvanic skin repsonse & BodyMedia CORE 2 &  50-100 \\
-body temperature & BodyMedia CORE 2 &  50-100 \\ 
-Eating & Hapifork & (not yet released) \\ 
+FitBit & general activity (walking/jogging) & \$50 \\
+Nike FuelBand & general activity (walking/jogging) & \$199 \\ 
+Writhings WiFi Scale & weight & \$200 \\
+Writhings BPM & blood pressure & \$200 \\ 
+Zeo Pro & sleep duration \& quality & \$200 \\
+WakeMate & sleep duration & \$50  \\
+Zensorium Tinkne & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
+BodyMedia CORE 2 & galvanic skin repsonse, body temperature, sleep, respiratory rate, heart rate & \$50-100 \\
+body temperature & BodyMedia CORE 2 &  (TBD) \\ 
+Hapifork & eating time, rate, frequency & (TBD) \\ 
 \end{tabular}
 \caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurememnt sensors with automatic data uplink to cloud services }
-\end{table*}
+\label{sensors}
+\end{table}
+}
+\vspace{6cm}
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
snomed ref

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 
-\begin{table*}[htb]
+\newpage
+
+\marginpar{
+\begin{table}
 \small
@@
 \caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurememnt sensors with automatic data uplink to cloud services }
-\end{table*}
+\label{sensors}
+\end{table}
+}
+
+\vspace{5cm}
 
@@
 
-The data collected by patients' devices has the potential to create a large volume of data, and therefore particular
+In order to do this, the data from each device should be collected and archived, so it can
+be viewed together in an appropriate interface. The data collected by patients' devices has
+the potential to create a large volume of data, and therefore particular
 consideration should be paid about who and where the data will be stored. In particular, it is likely
@@
 number of medical ontologies that already exist, as well as vocabularies and systems (such as
-SNOMED\cite{} and IDC-10) that focus on the medical domain. The hope is that by using
+SNOMED CT \cite{stearns2001snomed} and IDC-10) that focus on the medical domain. The hope is that by using
 such systems, that the data collected by patients can be analysed more rapidly by clinical systems
@@
 \bibliographystyle{acm-sigchi}
-\bibliography{sample}
+\bibliography{pccworkshop}
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Added mockup to margin
rewrite storing

@@
 
-Currently these products tend to work in isolation, where the readings from a single product
-are presented either on the device themselves, or via a smartphone or website. However,
-their use could be integrated with clinical care, in order to monitor patients, and to give an
-insight into their lifestyle.
+Currently these products are designed to work in isolation, that is, to upload sensor measurements to dedicated web sites set up by each device's manufacturer.
+Such sites are poorly suited for the long-term storage for the highly personal activity data for several reasons; first, none of the sites provide any long-term retention or access guarantees of user data (nor HIPAA security guarantees).
+Second, it is difficult for a user to gain a unified understanding of their activity across their devices through the disparate interfaces provided by these services, due of the widely different visual and interface representations provided.
 
-In order to do this, the data from each device should be collected and archived, so it can
-be viewed together in an appropriate interface. The data collected by patients' devices has
-the potential to create a large volume of data, and therefore particular
-consideration should be paid about who and where the data will be stored. In particular, it is likely
-that existing providers of storage for electronic health records will not want to be burdened with the
-administration costs of storing additional data. The longevity of third-party storage providers is
-also a concern, because there has been a history of short-notice closures of EHR providers (such as Google Health).
+Fortunately, many of these sites have APIs, which makes it feasible to access captured data from the devices by way of these web services.
+We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single reprsentation kept with their EHR.
+In particular, grounding this activity data in a common taxonomy with associated description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{snowmed} and ICD-10 \cite{icd}, could enable these diaries to interpreted directly by clinical medical systems, such as primary care information management systems.
+
+%% Such a representation might enable GPs to more quickly understand a   such as for providing
+%% like the SNOWMEDif Building such a representation, however, makes several assumptions, the 
+%%   We propose that   to be harvested and stored into unified representations, such as to be stored with a person's 
+%%   single product are peruploaded by the devices to 
+%% are presented either on the device themselves, or via a smartphone or website. However,
+%% their use could be integrated with clinical care, in order to monitor patients, and to give an
+%% insight into their lifestyle.
+
+%% In order to do this, the data from each device should be collected and archived, so it can
+%% be viewed together in an appropriate interface. The data collected by patients' devices has
+%% the potential to create a large volume of data, and therefore particular
+%% consideration should be paid about who and where the data will be stored. In particular, it is likely
+%% that existing providers of storage for electronic health records will not want to be burdened with the
+%% administration costs of storing additional data. The longevity of third-party storage providers is
+%% also a concern, because there has been a history of short-notice closures of EHR providers (such as Google Health).
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
bwah

@@
 
+\marginpar{
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=\marginparwidth]{img/energy-scenario}
+  \caption{Mock-up of an activity diary dashboard, showing data from multiple devices.}
+  \label{fig:mockup}
+  \end{center}  
+\end{figure}
+}
+
 Yet, so far, EHRs have primarily been limited to the storage of
@@
 Second, it is difficult for a user to gain a unified understanding of their activity across their devices through the disparate interfaces provided by these services, due of the widely different visual and interface representations provided.
+Finally, many of these visualisations represent user activity in arbitrary, 'user-friendly' units, such as 'Sleep score', 'Fuel', ``Zen level'', which, being not rigorously defined are difficult or impossible to compare across services.
 
-Fortunately, many of these sites have APIs, which makes it feasible to access captured data from the devices by way of these web services.
+Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data by way of these services.
 We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single reprsentation kept with their EHR.
-In particular, grounding this activity data in a common taxonomy with associated description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{snowmed} and ICD-10 \cite{icd}, could enable these diaries to interpreted directly by clinical medical systems, such as primary care information management systems.
+In particular, grounding this activity data in a common taxonomy with associated description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{icd}, could enable these diaries to interpreted directly by clinical medical systems, such as primary care information management systems.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
bwah

@@
 
-Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data by way of these services.
+Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data.
 We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single reprsentation kept with their EHR.

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
bwah

@@
 We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single reprsentation kept with their EHR.
-In particular, grounding this activity data in a common taxonomy with associated description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{icd}, could enable these diaries to interpreted directly by clinical medical systems, such as primary care information management systems.
+In particular, grounding this activity data in a uniform, common taxonomy and description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{icd}, would enable these diaries to interpretable by future clinical medical systems, such as primary care information management systems.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Added refs

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 bi-directional, in the sense that the patient needs to be able to
-effectively interpret what the clincian is trying to convey to him or
+effectively interpret what the clinician is trying to convey to him or
 her, but also that the clinician needs to be able to arrive at an
@@
 provide such a common context is the patient-centric electronic health
-record (henceforth EHR)\cite{ehr}, a digital representation of the
+record (henceforth EHR)\cite{hayrinen2008definition}, a digital representation of the
 complete medical history of an individual as recorded by clinicians
@@
 means that clinicians can more easily review a patient's long-term
-history even under time pressure and constriants. Moreover, since
+history even under time pressure and constraints. Moreover, since
 patients have access to their own medical histories, they can access
@@
 We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single reprsentation kept with their EHR.
-In particular, grounding this activity data in a common taxonomy with associated description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{icd}, could enable these diaries to interpreted directly by clinical medical systems, such as primary care information management systems.
+In particular, grounding this activity data in a common taxonomy with associated description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, could enable these diaries to interpreted directly by clinical medical systems, such as primary care information management systems.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex

@@
 
-Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data by way of these services.
+Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data.
 We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single reprsentation kept with their EHR.
-In particular, grounding this activity data in a common taxonomy with associated description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, could enable these diaries to interpreted directly by clinical medical systems, such as primary care information management systems.
+In particular, grounding this activity data in a uniform, common taxonomy and description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, would enable these diaries to interpretable by future clinical medical systems, such as primary care information management systems.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Enbiggen the image
boo

@@
 \begin{figure}
-  \begin{center}
-  \includegraphics[width=\marginparwidth]{img/energy-scenario}
+  \hspace{-11mm}
+  \includegraphics[width=5.5cm]{img/energy-scenario}
   \caption{Mock-up of an activity diary dashboard, showing data from multiple devices.}
   \label{fig:mockup}
-  \end{center}  
 \end{figure}
@@
 Zensorium Tinkne & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
-BodyMedia CORE 2 & galvanic skin repsonse, body temperature, sleep, respiratory rate, heart rate & \$50-100 \\
-body temperature & BodyMedia CORE 2 &  (TBD) \\ 
+BodyMedia CORE 2 & galvanic skin repsonse, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
 Hapifork & eating time, rate, frequency & (TBD) \\ 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
using

@@
 
+\section{Using activity diaries}
+
+We imagine integrated activity diaries to be used by both clinicans
+and patients in several ways.  First, for nearly all types of
+conditions, monitoring a person's overall daily activity and sleep
+duration can easily indicate how much rest, exercise, exposure to
+others, and the outside world he or she might be getting.  Daily
+records of the time(s) that a person gets up, eats, works, exercises
+and relaxes can be analysed by clincians for irregular behaviours or
+habits that might be affecting a person's well-being. For more
+specific conditions, daily measurements of a person's blood pressure,
+blood oxidation, heart rate, breathing, weight, and galvanic skin
+response could be useful for tracking progress along an clinician
+prescribed intervention, recovery from illness, or onset of new
+conditions, be used to pinpoint sources of prolonged stress.
+
 \section{Capturing activity diaries}
@@
 surging public interest in ``quantified-self''-style self-monitoring
-for self-improvement.  Many of these sensors are lightweight,
-can be worn discreetly, measure several physiological statistics
+for self-improvement.  Many of these sensors are lightweight, can be
+worn discreetly, measure several physiological statistics
 simultaneously, and most require little maintenance. Table
@@
 Currently these products are designed to work in isolation, that is, to upload sensor measurements to dedicated web sites set up by each device's manufacturer.
-Such sites are poorly suited for the long-term storage for the highly personal activity data for several reasons; first, none of the sites provide any long-term retention or access guarantees of user data (nor HIPAA security guarantees).
+Such sites are poorly suited for the long-term storage for the highly personal activity data for several reasons; first, none of the sites provide any long-term retention or access guarantees of user data (nor HIPAA security guarantees!).
 Second, it is difficult for a user to gain a unified understanding of their activity across their devices through the disparate interfaces provided by these services, due of the widely different visual and interface representations provided.

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Storing and representation together

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 
-\section{Storing activity diaries}
+\section{Storing and Representing Activity Diaries}
 
@@
 In particular, grounding this activity data in a uniform, common taxonomy and description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, would enable these diaries to interpretable by future clinical medical systems, such as primary care information management systems.
+
+The data collected by patients' devices has the potential to create a large volume of highly structured data. This is in contrast to
+EHRs which tend to contain individual records for each clinical session.
+
+To this challenge we bring our contribution of a common ontology for recording and encompassing patient activities, which we have begun to
+connect to SNOMED-CT, in order to enable semantic interoperability with existing systems.
 
@@
 %% administration costs of storing additional data. The longevity of third-party storage providers is
-%% also a concern, because there has been a history of short-notice closures of EHR providers (such as Google Health).
+%% also a concern, because there has b een a history of short-notice closures of EHR providers (such as Google Health).
 
-\section{Representing activity diaries}
-
-Particular care should be taken when deciding on a format to store the activity diaries. Ideally
-a format will be machine-readable, using open standards (for example XML, RDF or JSON) and
-following existing schematic/ontological knowledge modelling conventions. There are a 
-number of medical ontologies that already exist, as well as vocabularies and systems (such as
-SNOMED CT \cite{stearns2001snomed} and IDC-10) that focus on the medical domain. The hope is that by using
-such systems, that the data collected by patients can be analysed more rapidly by clinical systems
-and staff.
-
-One option is to embrace the Semantic Web, because it offers several key benefits:
-
-\begin{itemize}
-
-\item Distributed system with universal addresses and standard mature protocols
-\item Open extensible ontology system and existing domain ontologies
-\item Existing library and tool support and expertise
-\item Interoperability with existing systems
-
-\end{itemize}
-
-Even if personal health records or systems do not use the Semantic Web, there are existing
-techniques to integrate with their data.
+%Particular care should be taken when deciding on a format to store the activity diaries. Ideally
+%a format will be machine-readable, using open standards (for example XML, RDF or JSON) and
+%following existing schematic/ontological knowledge modelling conventions. There are a 
+%number of medical ontologies that already exist, as well as vocabularies and systems (such as
+%SNOMED CT \cite{stearns2001snomed} and IDC-10) that focus on the medical domain. The hope is that by using
+%such systems, that the data collected by patients can be analysed more rapidly by clinical systems
+%and staff.
+%
+%One option is to embrace the Semantic Web, because it offers several key benefits:
+%
+%\begin{itemize}
+%
+%\item Distributed system with universal addresses and standard mature protocols
+%\item Open extensible ontology system and existing domain ontologies
+%\item Existing library and tool support and expertise
+%\item Interoperability with existing systems
+%
+%\end{itemize}
+% 
+%Even if personal health records or systems do not use the Semantic Web, there are existing
+%techniques to integrate with their data.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
using

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 conditions, be used to pinpoint sources of prolonged stress.
+
+Merely providing access to such activity records back to individuals
+could have direct effects on people's behavior, motivation and
+perceived well-being.  For example, just as asking individuals to
+introspect and justify their opinions and feelings has been found to
+be both disruptive and reinforcing of their opinions
+\cite{wilson1989introspection}, allowing people to reflect upon their
+historical records of their activities might help people realize or
+change their minds about how or why they feel particular ways. We feel
+that there is significant room for investigation here, and potential
+for research pertaining to self motivation and improvement.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Storing/representing update

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 Second, it is difficult for a user to gain a unified understanding of their activity across their devices through the disparate interfaces provided by these services, due of the widely different visual and interface representations provided.
-Finally, many of these visualisations represent user activity in arbitrary, 'user-friendly' units, such as 'Sleep score', 'Fuel', ``Zen level'', which, being not rigorously defined are difficult or impossible to compare across services.
+Finally, many of these visualisations represent user activity in arbitrary, ``user-friendly'' units, such as ``Sleep score'', ``Fuel'', ``Zen level'', which, being not rigorously defined are difficult or impossible to compare across services.
 
 Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data.
-We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single reprsentation kept with their EHR.
-In particular, grounding this activity data in a uniform, common taxonomy and description logic, like those provided for medical diagnoses and diseases through SNOWMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, would enable these diaries to interpretable by future clinical medical systems, such as primary care information management systems.
-
-The data collected by patients' devices has the potential to create a large volume of highly structured data. This is in contrast to
-EHRs which tend to contain individual records for each clinical session.
+We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single representation kept with their EHR.
+In particular, grounding this activity data in a uniform, common taxonomy and description logic, like those provided for medical diagnoses and diseases through SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, would enable these diaries to interpretable by future clinical medical systems, such as primary care information management systems.
 
 To this challenge we bring our contribution of a common ontology for recording and encompassing patient activities, which we have begun to
-connect to SNOMED-CT, in order to enable semantic interoperability with existing systems.
+connect to SNOMED-CT, in order to enable semantic interoperability with existing systems, and provide a common vocabulary across devices.
+
+The data collected by patients' devices has the potential to create a large volume of highly structured data. Some sensors
+are collecting minute-by-minute data values form multiple sensors 24 hours a day, whereas 
+EHRs tend to contain individual records for each clinical session. Therefore, existing methods of storing EHRs may
+prove to be unsuitable for the large volumes of data from activity diaries. However, integrating with existing EHRs
+would enable clinicians and patients to view the data in context.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex

@@
 
+\section{Using activity diaries}
+
+We imagine integrated activity diaries to be used by both clinicans
+and patients in several ways.  First, for nearly all types of
+conditions, monitoring a person's overall daily activity and sleep
+duration can easily indicate how much rest, exercise, exposure to
+others, and the outside world he or she might be getting.  Daily
+records of the time(s) that a person gets up, eats, works, exercises
+and relaxes can be analysed by clincians for irregular behaviours or
+habits that might be affecting a person's well-being. For more
+specific conditions, daily measurements of a person's blood pressure,
+blood oxidation, heart rate, breathing, weight, and galvanic skin
+response could be useful for tracking progress along an clinician
+prescribed intervention, recovery from illness, or onset of new
+conditions, be used to pinpoint sources of prolonged stress.
+
+Merely providing access to such activity records back to individuals
+could have direct effects on people's behavior, motivation and
+perceived well-being.  For example, just as asking individuals to
+introspect and justify their opinions and feelings has been found to
+be both disruptive and reinforcing of their opinions
+\cite{wilson1989introspection}, allowing people to reflect upon their
+historical records of their activities might help people realize or
+change their minds about how or why they feel particular ways. We feel
+that there is significant room for investigation here, and potential
+for research pertaining to self motivation and improvement.
+
 \section{Capturing activity diaries}

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
table fix
proofs

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 conditions, monitoring a person's overall daily activity and sleep
-duration can easily indicate how much rest, exercise, exposure to
+duration could easily indicate how much rest, exercise, exposure to
 others, and the outside world he or she might be getting.  Daily
@@
 blood oxidation, heart rate, breathing, weight, and galvanic skin
-response could be useful for tracking progress along an clinician
-prescribed intervention, recovery from illness, or onset of new
-conditions, be used to pinpoint sources of prolonged stress.
+response could be useful for tracking progress on a prescribed
+intervention, recovery from an illness, or onset of a new conditions,
+for example, or be used to pinpoint sources of prolonged stress.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 
-\section{Capturing activity diaries}
-
-The recent proliferation in personal wearable activity sensors and
-network and storage- enabled digital health measurement devices has
-been driven by both the falling costs of embeddable electronics, and a
-surging public interest in ``quantified-self''-style self-monitoring
-for self-improvement.  Many of these sensors are lightweight, can be
-worn discreetly, measure several physiological statistics
-simultaneously, and most require little maintenance. Table
-\ref{sensors} lists a variety of such sensors, currently available
-readily from online retailers.
-
-\newpage
-
 \marginpar{
@@
 \small
+\vspace{0.5cm}
 \begin{tabular}{p{3cm}  p{6cm} p{1.5cm}}
@@
 \vspace{6cm}
+
+\section{Capturing activity diaries}
+
+The recent proliferation in personal wearable activity sensors and
+network and storage- enabled digital health measurement devices has
+been driven by both the falling costs of embeddable electronics, and a
+surging public interest in ``quantified-self''-style self-monitoring
+for self-improvement.  Many of these sensors are lightweight, can be
+worn discreetly, measure several physiological statistics
+simultaneously, and most require little maintenance. Table
+\ref{sensors} lists a variety of such sensors, currently available
+readily from online retailers.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
webbox!

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 would enable clinicians and patients to view the data in context.
+
+To this challenge we present our own personal data platform ``WebBox.'' We have used WebBox with EHRs
+to enable patient collected data to be augmented with hospital-provided EHR data. WebBox is able to combine
+data from multiple sources, and enables patients to make their own annotations against data, and
+to optionally share these annotations, along with their activity diaries, with their clinician. This
+then enables clinicians and patients to enter into a dialog about how well the patient is doing,
+and any lifestyle changes they can make to better their situation.
 
@@
 
-\section{Biographies}
-?
+By augmenting data from personal health sensors with EHRs, both patients and clinicians
+can gain insight into the lifestyle and typical activities of a patient. Challenges in implementation
+of such a system do exist, but through the use of domain ontologies and peer-to-peer
+sharing...
+
+%\section{Biographies}
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Webbox ref

@@
 
-To this challenge we present our own personal data platform ``WebBox.'' We have used WebBox with EHRs
+To this challenge we present our own personal data platform ``WebBox \cite{webbox}.'' We have used WebBox with EHRs
 to enable patient collected data to be augmented with hospital-provided EHR data. WebBox is able to combine

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 
+To this challenge we present our own personal data platform ``WebBox \cite{webbox}.'' We have used WebBox with EHRs
+to enable patient collected data to be augmented with hospital-provided EHR data. WebBox is able to combine
+data from multiple sources, and enables patients to make their own annotations against data, and
+to optionally share these annotations, along with their activity diaries, with their clinician. This
+then enables clinicians and patients to enter into a dialog about how well the patient is doing,
+and any lifestyle changes they can make to better their situation.
+
 %% Such a representation might enable GPs to more quickly understand a   such as for providing

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
oops 5 pages

@@
 
-Currently these products are designed to work in isolation, that is, to upload sensor measurements to dedicated web sites set up by each device's manufacturer.
+Currently, these products are designed to work separately by themselves, uploading sensor measurements to dedicated web sites set up by each device's manufacturer.
 Such sites are poorly suited for the long-term storage for the highly personal activity data for several reasons; first, none of the sites provide any long-term retention or access guarantees of user data (nor HIPAA security guarantees).
@@
 
-Fortunately, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data.
+However, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data.
 We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single representation kept with their EHR.
@@
 
-To this challenge we bring our contribution of a common ontology for recording and encompassing patient activities, which we have begun to
-connect to SNOMED-CT, in order to enable semantic interoperability with existing systems, and provide a common vocabulary across devices.
+Since SNOWMED-CT nor ICD cover concepts pertaining to routine measured activity, we are devising a supplemental ontology for these concepts we call \emph{Activity-SENSE}.
+This ontology encompasses patient activities, standard units of measure with which these activities are measured, and the ability to describe how and what kinds of sensors were used to sense the activity measured.
+The intention is for Activity-SENSE records to, like those of SNOMED-CT, eventually be translatable to different languages and mapped to other similiar ontologies of behaviour.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
oops 5 pages

@@
 
-Since SNOWMED-CT nor ICD cover concepts pertaining to routine measured activity, we are devising a supplemental ontology for these concepts we call \emph{Activity-SENSE}.
-This ontology encompasses patient activities, standard units of measure with which these activities are measured, and the ability to describe how and what kinds of sensors were used to sense the activity measured.
+Since SNOWMED-CT nor ICD cover routine measured activity, we are devising a supplemental ontology for these concepts called \emph{Activity-SENSE}.
+This ontology encompasses patient activities, and standard units of measure with which activities are measured, as well as the ability to describe the sensors used.
 The intention is for Activity-SENSE records to, like those of SNOMED-CT, eventually be translatable to different languages and mapped to other similiar ontologies of behaviour.
 
-The data collected by patients' devices has the potential to create a large volume of highly structured data. Some sensors
+The data collected by patients' devices has the potential to create a large volumes of highly structured data. Some sensors
 are collecting minute-by-minute data values form multiple sensors 24 hours a day, whereas 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
boo

@@
 
-The data collected by patients' devices has the potential to create a large volumes of highly structured data. Some sensors
-are collecting minute-by-minute data values form multiple sensors 24 hours a day, whereas 
-EHRs tend to contain individual records for each clinical session. Therefore, existing methods of storing EHRs may
-prove to be unsuitable for the large volumes of data from activity diaries. However, integrating with existing EHRs
-would enable clinicians and patients to view the data in context.
+\section{The WebBox EHR Platform}
 
+The high-resolution data collected by activity-sensing devices, accumulated over a long period such as a patient's lifespan, can result in a substantially greater volume of data than current EHR platforms are designed to handle.
 To this challenge we present our own personal data platform ``WebBox \cite{webbox}.'' We have used WebBox with EHRs
@@
 and any lifestyle changes they can make to better their situation.
+
+%% Furthermore, Therefore, existing platforms for storing EHRs may
+%% prove to be unsuitable for the large volumes of data from activity diaries. However, integrating with existing EHRs
+%% would enable clinicians and patients to view the data in context.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Nigel bio

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 
-%\section{Biographies}
+\section{Biographies}
+
+\subsection{Daniel A. Smith and Max Van Kleek}
+
+\subsection{Nigel R. Shadbolt}
+
+Nigel Shadbolt (FBCS FREng) is Professor of Artificial Intelligence and Head of the Web and Internet Science Group, Electronics and Computer Science, University of Southampton.
+He is the Chairman and Co-Founder of the Open Data Institute, Director of the Web Science Trust, and of the Web Foundation.
+In 2009 the Prime Minister appointed him and Sir Tim Berners-Lee as Information Advisors to transform access to Public Sector Information, leading to the highly acclaimed data.gov.uk.
+He is a member of the Public Sector Transparency Board, overseeing Open Data releases across the public sector.
+He is Chair of the UK Government’s midata programme, which seeks to empower consumers by releasing their data back to them.
+
+\section{Key Technology: WebBox}
+
+WebBox is...
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
boo

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 The high-resolution data collected by activity-sensing devices, accumulated over a long period such as a patient's lifespan, can result in a substantially greater volume of data than current EHR platforms are designed to handle.
-To this challenge we present our own personal data platform ``WebBox \cite{webbox}.'' We have used WebBox with EHRs
-to enable patient collected data to be augmented with hospital-provided EHR data. WebBox is able to combine
-data from multiple sources, and enables patients to make their own annotations against data, and
-to optionally share these annotations, along with their activity diaries, with their clinician. This
-then enables clinicians and patients to enter into a dialog about how well the patient is doing,
-and any lifestyle changes they can make to better their situation.
+To address this deficiency, we have devised a personal data platform ``WebBox \cite{webbox}'', optimised to handle the secure, long-term storage of structured data at large volume.
+Thus far, WebBox has been used to archive clinician (hospital and GP-provided) data, including both raw structured data (such as clinical diagnostic results) and accompanying report documents.
+Being optimised for structured data capture and integration, our current work extends this platform to the aforementioned activity profiles, created from retrieving raw sensed data from digital health device APIs and translating such observations to the Activity-SENSE representation.
+
+%Following this, to optionally share these annotations, along with their activity diaries, with their clinician.
+This then enables clinicians and patients to enter into a dialog about how well the patient is doing, and any lifestyle changes they can make to better their situation.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
foo

@@
 To address this deficiency, we have devised a personal data platform ``WebBox \cite{webbox}'', optimised to handle the secure, long-term storage of structured data at large volume.
-Thus far, WebBox has been used to archive clinician (hospital and GP-provided) data, including both raw structured data (such as clinical diagnostic results) and accompanying report documents.
+Thus far, WebBox has been used to archive clinician (hospital and GP-provided) data, including both raw structured data (such as clinical diagnostic results) and traditional reports prepared by clinicians.
 Being optimised for structured data capture and integration, our current work extends this platform to the aforementioned activity profiles, created from retrieving raw sensed data from digital health device APIs and translating such observations to the Activity-SENSE representation.

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
bios and ack

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 
-\section{The WebBox EHR Platform}
+The data collected by patients' devices has the potential to create a large volumes of highly structured data. Some sensors
+are collecting minute-by-minute data values form multiple sensors 24 hours a day, whereas 
+EHRs tend to contain individual records for each clinical session. Therefore, existing methods of storing EHRs may
+prove to be unsuitable for the large volumes of data from activity diaries. However, integrating with existing EHRs
+would enable clinicians and patients to view the data in context.
 
-The high-resolution data collected by activity-sensing devices, accumulated over a long period such as a patient's lifespan, can result in a substantially greater volume of data than current EHR platforms are designed to handle.
-To address this deficiency, we have devised a personal data platform ``WebBox \cite{webbox}'', optimised to handle the secure, long-term storage of structured data at large volume.
-Thus far, WebBox has been used to archive clinician (hospital and GP-provided) data, including both raw structured data (such as clinical diagnostic results) and accompanying report documents.
-Being optimised for structured data capture and integration, our current work extends this platform to the aforementioned activity profiles, created from retrieving raw sensed data from digital health device APIs and translating such observations to the Activity-SENSE representation.
-
-%Following this, to optionally share these annotations, along with their activity diaries, with their clinician.
-This then enables clinicians and patients to enter into a dialog about how well the patient is doing, and any lifestyle changes they can make to better their situation.
-
-%% Furthermore, Therefore, existing platforms for storing EHRs may
-%% prove to be unsuitable for the large volumes of data from activity diaries. However, integrating with existing EHRs
-%% would enable clinicians and patients to view the data in context.
+To this challenge we present our own personal data platform ``WebBox \cite{webbox}.'' We have used WebBox with EHRs
+to enable patient collected data to be augmented with hospital-provided EHR data. WebBox is able to combine
+data from multiple sources, and enables patients to make their own annotations against data, and
+to optionally share these annotations, along with their activity diaries, with their clinician. This
+then enables clinicians and patients to enter into a dialog about how well the patient is doing,
+and any lifestyle changes they can make to better their situation.
 
@@
 
-%\section{Biographies}
+\section{Biographies}
+
+\subsection{Daniel A. Smith and Max Van Kleek}
+
+Daniel and Max are postdoctoral researchers on the SOCIAM project at the University of Southampton, and the principal architects of the WebBox
+personal information store. They have research interests in HCI, the Semantic Web and Personal Information Management.
+
+\subsection{Nigel R. Shadbolt}
+
+Nigel Shadbolt (FBCS FREng) is Professor of Artificial Intelligence and Head of the Web and Internet Science Group, University of Southampton.
+He is the Chairman and Co-Founder of the Open Data Institute, Director of the Web Science Trust, and of the Web Foundation.
+He and Sir Tim Berners-Lee were appointmented as Information Advisors to the UK Government, leading to \url{data.gov.uk}.
+He is a member of the Public Sector Transparency Board, and Chair of the UK Government’s midata programme.
+
+\section{Key Technology: WebBox}
+
+WebBox is...
+
+\section{Acknowledgement}
+
+This work is supported under SOCIAM: The Theory and Practice of Social Machines.
+The SOCIAM Project is funded by the UK Engineering and Physical Research Council (ESPRC) under grant number EP/J017728/1 and comprises the Universities of Southampton, Oxford and Edinburgh.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
typos

@@
 This ontology encompasses patient activities, and standard units of measure with which activities are measured, as well as the ability to describe the sensors used.
-The intention is for Activity-SENSE records to, like those of SNOMED-CT, eventually be translatable to different languages and mapped to other similiar ontologies of behaviour.
+The intention is for Activity-SENSE records to, like those of SNOMED-CT, eventually be translatable to different languages and mapped to other similar ontologies of behaviour.
 
@@
 He is the Chairman and Co-Founder of the Open Data Institute, Director of the Web Science Trust, and of the Web Foundation.
-He and Sir Tim Berners-Lee were appointmented as Information Advisors to the UK Government, leading to \url{data.gov.uk}.
-He is a member of the Public Sector Transparency Board, and Chair of the UK Government’s midata programme.
+He and Sir Tim Berners-Lee were appointented as Information Advisors to the UK Government, leading to \url{data.gov.uk}.
+He is a member of the Public Sector Transparency Board, and Chair of the UK Government's midata programme.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
so close

@@
 
-\section{Privacy Controls}
+%% \section{Privacy Controls}
 
-One of the major concerns with storage of patient data is privacy. Indeed, there is also potentially
-legal requirements on storing personal health data, depending on the nature of the data being
-stored.
+%% One of the major concerns with storage of patient data is privacy. Indeed, there is also potentially
+%% legal requirements on storing personal health data, depending on the nature of the data being
+%% stored.
 
-However, it is important that control of data be sufficiently granular so that patients can choose
-to sure it with clinicians, and that clinicians can refer to it in their notes in the patient's health
-record.
+%% However, it is important that control of data be sufficiently granular so that patients can choose
+%% to sure it with clinicians, and that clinicians can refer to it in their notes in the patient's health
+%% record.
 
-\section{Conclusion}
+%% \section{Conclusion}
 
-By augmenting data from personal health sensors with EHRs, both patients and clinicians
-can gain insight into the lifestyle and typical activities of a patient. Challenges in implementation
-of such a system do exist, but through the use of domain ontologies and peer-to-peer
-sharing...
+%% By augmenting data from personal health sensors with EHRs, both patients and clinicians
+%% can gain insight into the lifestyle and typical activities of a patient. Challenges in implementation
+%% of such a system do exist, but through the use of domain ontologies and peer-to-peer
+%% sharing...
 
@@
 Nigel Shadbolt (FBCS FREng) is Professor of Artificial Intelligence and Head of the Web and Internet Science Group, University of Southampton.
-He is the Chairman and Co-Founder of the Open Data Institute, Director of the Web Science Trust, and of the Web Foundation.
-He and Sir Tim Berners-Lee were appointented as Information Advisors to the UK Government, leading to \url{data.gov.uk}.
+He is the Chairman and Co-Founder of the Open Data Institute, Director of the Web Science Trust, and the Web Foundation.
+He and Sir Tim Berners-Lee were Information Advisors to the UK Government, which led to \url{data.gov.uk}.
 He is a member of the Public Sector Transparency Board, and Chair of the UK Government's midata programme.
 
-\section{Key Technology: WebBox}
+%% \section{Key Technology: WebBox}
 
-WebBox is...
+%% WebBox is...
 
@@
 
-This work is supported under SOCIAM: The Theory and Practice of Social Machines.
-The SOCIAM Project is funded by the UK Engineering and Physical Research Council (ESPRC) under grant number EP/J017728/1 and comprises the Universities of Southampton, Oxford and Edinburgh.
+This work is supported by the SOCIAM Project (EPSRC grant EP/J017728/1).
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
nigel bio fix
trims

@@
 
-The high-resolution data collected by activity-sensing devices, accumulated over a long period such as a patient's lifespan, can result in a substantially greater volume of data than current EHR platforms are designed to handle.
-To address this deficiency, we have devised a personal data platform ``WebBox \cite{webbox}'', optimised to handle the secure, long-term storage of structured data at large volume.
+The high-resolution data collected by activity-sensing devices can result in a greater volume of data than current EHR platforms are designed to handle when  accumulated over a patient's lifetime.
+To address this need, we have devised  ``WebBox \cite{webbox}'', a personal data platform optimised to handle the secure, long-term storage of large volumes of structured data.
 Thus far, WebBox has been used to archive clinician (hospital and GP-provided) data, including both raw structured data (such as clinical diagnostic results) and traditional reports prepared by clinicians.
@@
 
-Nigel Shadbolt (FBCS FREng) is Professor of Artificial Intelligence and Head of the Web and Internet Science Group, University of Southampton.
+Nigel Shadbolt (FBCS FREng) is Professor of Artificial Intelligence and Head of the Web and Internet Science Group, at the University of Southampton.
 He is the Chairman and Co-Founder of the Open Data Institute, Director of the Web Science Trust, and the Web Foundation.

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex

@@
 
-\subsection{Daniel A. Smith and Max Van Kleek}
+\emph{Daniel A.
+Smith} and \emph{Max Van Kleek} are postdoctoral researchers on the SOCIAM project at the University of Southampton, and the principal architects of the WebBox personal information store.
+They have research interests in HCI, the Semantic Web and Personal Information Management.
 
-Daniel and Max are postdoctoral researchers on the SOCIAM project at the University of Southampton, and the principal architects of the WebBox
-personal information store. They have research interests in HCI, the Semantic Web and Personal Information Management.
-
-\subsection{Nigel R. Shadbolt}
-
-Nigel Shadbolt (FBCS FREng) is Professor of Artificial Intelligence and Head of the Web and Internet Science Group, at the University of Southampton.
-He is the Chairman and Co-Founder of the Open Data Institute, Director of the Web Science Trust, and the Web Foundation.
-He and Sir Tim Berners-Lee were Information Advisors to the UK Government, which led to \url{data.gov.uk}.
-He is a member of the Public Sector Transparency Board, and Chair of the UK Government's midata programme.
+\emph{Nigel Shadbolt} (FBCS FREng) is Professor of Artificial Intelligence and Head of the Web and Internet Science Group, University of Southampton.
+As a member of the Public Sector Transparency Board, Chair of the UK Government's midata programme, and Chairman and Co-Founder of the Open Data Institute, Nigel has led the effort towards assisting the NHS to begin the continual release of health-care related datasets.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
sleep

@@
 In this position paper, we look at the potential and problems of
-augmenting the next generation of EHRs with \emph{personal activity
+augmenting the next generation of EHRs with personal \emph{activity
   diaries}, portraits of a person's dynamic condition captured at high
@@
 \end{tabular}
-\caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurememnt sensors with automatic data uplink to cloud services }
+\caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurement sensors with automatic data uplink to cloud services }
 \label{sensors}
@@
 simultaneously, and most require little maintenance. Table
-\ref{sensors} lists a variety of such sensors, currently available
-readily from online retailers.
+\ref{sensors} lists a sampling of the sensors currently available
+directly from retailers.
 
-While few of these devices had been independently evaluated for
-accuracy for clinical use, as they have matured, several of these
-manufacturers begun rigorous evaluaton of their products for achieving
-certification. For example, BodyMedia advertise themselves as
-providing the ``FDA-approved, clincally validated, most accurate
-wireless activity measurement''.  A separate study of activity-based
-sleep monitoring tools revealed that ``low cost actigraphy-based
-approaches'' correlated well with baseline (in-lab EEG-based
-measurements) for sleep duration, although less with sleep quality
-metrics.
+While few of these devices had initially been evaluated for accuracy
+or certified for clinical use, increasingly device manufacturers have
+begun to get independent certification. For example, BodyMedia
+advertise themselves as providing the ``FDA-approved, clincally
+validated, most accurate wireless activity measurement''.  A separate
+study of activity-based sleep monitoring tools revealed that ``low
+cost actigraphy-based approaches'' correlated well with baseline
+(in-lab EEG-based measurements) for sleep duration, although less with
+sleep quality metrics \cite{reid1999correlation}.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
almost

@@
 begun to get independent certification. For example, BodyMedia
-advertise themselves as providing the ``FDA-approved, clincally
-validated, most accurate wireless activity measurement''.  A separate
-study of activity-based sleep monitoring tools revealed that ``low
-cost actigraphy-based approaches'' correlated well with baseline
-(in-lab EEG-based measurements) for sleep duration, although less with
-sleep quality metrics \cite{reid1999correlation}.
+advertises its device as ``FDA-approved, clincally validated, most
+accurate wireless activity measurement''.  A separate study of
+activity-based sleep monitoring tools revealed that ``low cost
+actigraphy-based approaches'' correlated well with baseline (in-lab
+EEG-based measurements) for sleep duration, although less with sleep
+quality metrics \cite{reid1999correlation}.
 
@@
 \section{Storing and Representing Activity Diaries}
+Currently, these devices are designed to upload measurements only to separate, dedicated web sites set up by their manufacturer.
+Such sites are poorly suited for the long-term storage of people's activity data for several reasons.
+First, none of the sites providelong-term retention or access guarantees of user data (nor HIPAA security guarantees!).
+Second, it is difficult for a user to gain a unified understanding of their activity through the disparate interfaces provided, in part due of the widely different visual and interface representations.
+Finally, many of interfaces represent activity in arbitrary, ``user-friendly'' units, such as ``Sleep score'', ``Fuel'', ``Zen level'' etc., which are difficult or impossible for users to compare.
 
-Currently, these products are designed to work separately by themselves, uploading sensor measurements to dedicated web sites set up by each device's manufacturer.
-Such sites are poorly suited for the long-term storage for the highly personal activity data for several reasons; first, none of the sites provide any long-term retention or access guarantees of user data (nor HIPAA security guarantees).
-Second, it is difficult for a user to gain a unified understanding of their activity across their devices through the disparate interfaces provided by these services, due of the widely different visual and interface representations provided.
-Finally, many of these visualisations represent user activity in arbitrary, ``user-friendly'' units, such as ``Sleep score'', ``Fuel'', ``Zen level'', which, being not rigorously defined are difficult or impossible to compare across services.
+However, many of these sites have APIs that provide programmatic access to raw sensor data.
+Our approach has been touse such APIs to retrieve a patient's information from all services corresponding to their devices.
+In particular, we propose that consolidating such information into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd} for diseases, would enable such diaries to be interpretable and used directly by future clinical medical systems, such as primary care information management systems.
 
-However, many of these sites have APIs that provide access to the raw sensor data, making it feasible to access and interpret captured data.
-We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single representation kept with their EHR.
-In particular, grounding this activity data in a uniform, common taxonomy and description logic, like those provided for medical diagnoses and diseases through SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, would enable these diaries to interpretable by future clinical medical systems, such as primary care information management systems.
-
-Since SNOWMED-CT nor ICD cover routine measured activity, we are devising a supplemental ontology for these concepts called \emph{Activity-SENSE}.
-This ontology encompasses patient activities, and standard units of measure with which activities are measured, as well as the ability to describe the sensors used.
-The intention is for Activity-SENSE records to, like those of SNOMED-CT, eventually be translatable to different languages and mapped to other similar ontologies of behaviour.
+Since SNOWMED-CT nor ICD cover activity-related concepts, we devised a supplemental ontology called \emph{Activity-SENSE}.
+This ontology encompasses patient activities, and standard units of measure with which activities measurements were taken, and concepts for describing sensors used.
+The intention is for records in Activity-SENSE, like those of SNOMED-CT, will be eventually translatable to dozens of different languages and mapped to similar activity ontologies, allowing international use and greater operability.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
typos etc

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 provide such a common context is the patient-centric electronic health
-record (henceforth EHR)\cite{hayrinen2008definition}, a digital representation of the
+record (henceforth EHR)~\cite{hayrinen2008definition}, a digital representation of the
 complete medical history of an individual as recorded by clinicians
@@
 
-\section{Using activity diaries}
+\section{Using Activity Diaries}
 
-We imagine integrated activity diaries to be used by both clinicans
+We imagine integrated activity diaries to be used by both clinicians
 and patients in several ways.  First, for nearly all types of
@@
 duration could easily indicate how much rest, exercise, exposure to
-others, and the outside world he or she might be getting.  Daily
+others, and the outside world they might be getting.  Daily
 records of the time(s) that a person gets up, eats, works, exercises
-and relaxes can be analysed by clincians for irregular behaviours or
+and relaxes can be analysed by clinicians for irregular behaviours or
 habits that might be affecting a person's well-being. For more
@@
 Merely providing access to such activity records back to individuals
-could have direct effects on people's behavior, motivation and
+could have direct effects on people's behaviour, motivation and
 perceived well-being.  For example, just as asking individuals to
 introspect and justify their opinions and feelings has been found to
-be both disruptive and reinforcing of their opinions
-\cite{wilson1989introspection}, allowing people to reflect upon their
-historical records of their activities might help people realize or
+be both disruptive and reinforcing of their opinions~\cite{wilson1989introspection},
+allowing people to reflect upon their
+historical records of their activities might help people realise or
 change their minds about how or why they feel particular ways. We feel
@@
 Nike FuelBand & general activity (walking/jogging) & \$199 \\ 
-Writhings WiFi Scale & weight & \$200 \\
-Writhings BPM & blood pressure & \$200 \\ 
+WiThings WiFi Scale & weight & \$200 \\
+WiThings BPM & blood pressure & \$200 \\ 
 Zeo Pro & sleep duration \& quality & \$200 \\
 WakeMate & sleep duration & \$50  \\
-Zensorium Tinkne & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
-BodyMedia CORE 2 & galvanic skin repsonse, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
+Zensorium Tinnke & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
+BodyMedia CORE 2 & galvanic skin response, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
 Hapifork & eating time, rate, frequency & (TBD) \\ 
 \end{tabular}
-\caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurememnt sensors with automatic data uplink to cloud services }
+\caption{\emph{Consumer-grade activity sensors}: Activity and physiological measurement sensors with automatic data uplink to cloud services }
 \label{sensors}
@@
 The recent proliferation in personal wearable activity sensors and
-network and storage- enabled digital health measurement devices has
+network- and storage- enabled digital health measurement devices has
 been driven by both the falling costs of embeddable electronics, and a
@@
 accuracy for clinical use, as they have matured, several of these
-manufacturers begun rigorous evaluaton of their products for achieving
+manufacturers begun rigorous evaluation of their products for achieving
 certification. For example, BodyMedia advertise themselves as
-providing the ``FDA-approved, clincally validated, most accurate
+providing the ``FDA-approved, clinically validated, most accurate
 wireless activity measurement''.  A separate study of activity-based
@@
 We propose that these APIs be used to consolidate information from among all of the services corresponding to a patient's personal health devices, into a single representation kept with their EHR.
-In particular, grounding this activity data in a uniform, common taxonomy and description logic, like those provided for medical diagnoses and diseases through SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}, would enable these diaries to interpretable by future clinical medical systems, such as primary care information management systems.
+In particular, grounding this activity data in a uniform, common taxonomy and description logic, like those provided for medical diagnoses and diseases through SNOMED-CT~\cite{stearns2001snomed} and ICD-10~\cite{world1993icd}, would enable these diaries to interpretable by future clinical medical systems, such as primary care information management systems.
 
-Since SNOWMED-CT nor ICD cover routine measured activity, we are devising a supplemental ontology for these concepts called \emph{Activity-SENSE}.
+Since SNOMED-CT nor ICD cover routine measured activity, we are devising a supplemental ontology for these concepts called \emph{Activity-SENSE}.
 This ontology encompasses patient activities, and standard units of measure with which activities are measured, as well as the ability to describe the sensors used.
@@
 The high-resolution data collected by activity-sensing devices can result in a greater volume of data than current EHR platforms are designed to handle when  accumulated over a patient's lifetime.
-To address this need, we have devised  ``WebBox \cite{webbox}'', a personal data platform optimised to handle the secure, long-term storage of large volumes of structured data.
+To address this need, we have devised  ``WebBox''~\cite{webbox}, a personal data platform optimised to handle the secure, long-term storage of large volumes of structured data.
 Thus far, WebBox has been used to archive clinician (hospital and GP-provided) data, including both raw structured data (such as clinical diagnostic results) and traditional reports prepared by clinicians.

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
almost

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 
-However, many of these sites have APIs that provide programmatic access to raw sensor data.
-Our approach has been touse such APIs to retrieve a patient's information from all services corresponding to their devices.
-In particular, we propose that consolidating such information into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd} for diseases, would enable such diaries to be interpretable and used directly by future clinical medical systems, such as primary care information management systems.
+However, most of these sites have APIs providing programmatic access to raw sensor data.
+Our approach has been to leverage such APIs to retrieve a patient's information from all their devices, consolidating it into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}.
+A uniform representation enables such data to be interpretabled and used directly by future clinical medical systems, such as primary care information management systems.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 Nike FuelBand & general activity (walking/jogging) & \$199 \\ 
-Writhings WiFi Scale & weight & \$200 \\
-Writhings BPM & blood pressure & \$200 \\ 
+WiThings WiFi Scale & weight & \$200 \\
+WiThings BPM & blood pressure & \$200 \\ 
 Zeo Pro & sleep duration \& quality & \$200 \\
 WakeMate & sleep duration & \$50  \\
-Zensorium Tinkne & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
-BodyMedia CORE 2 & galvanic skin repsonse, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
+Zensorium Tinnke & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
+BodyMedia CORE 2 & galvanic skin response, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
 Hapifork & eating time, rate, frequency & (TBD) \\ 
@@
 Our approach has been touse such APIs to retrieve a patient's information from all services corresponding to their devices.
-In particular, we propose that consolidating such information into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd} for diseases, would enable such diaries to be interpretable and used directly by future clinical medical systems, such as primary care information management systems.
+In particular, we propose that consolidating such information into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT~\cite{stearns2001snomed} and ICD-10~\cite{world1993icd} for diseases, would enable such diaries to be interpretable and used directly by future clinical medical systems, such as primary care information management systems.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex

@@
 
-However, many of these sites have APIs that provide programmatic access to raw sensor data.
-Our approach has been touse such APIs to retrieve a patient's information from all services corresponding to their devices.
-In particular, we propose that consolidating such information into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT~\cite{stearns2001snomed} and ICD-10~\cite{world1993icd} for diseases, would enable such diaries to be interpretable and used directly by future clinical medical systems, such as primary care information management systems.
+However, most of these sites have APIs providing programmatic access to raw sensor data.
+Our approach has been to leverage such APIs to retrieve a patient's information from all their devices, consolidating it into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT~\cite{stearns2001snomed} and ICD-10~\cite{world1993icd}.
+A uniform representation enables such data to be interpretabled and used directly by future clinical medical systems, such as primary care information management systems.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
edits

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 \section{Storing and Representing Activity Diaries}
-Currently, these devices are designed to upload measurements only to separate, dedicated web sites set up by their manufacturer.
+Currently, these devices are designed to upload measurements only to separate, dedicated web sites run by their manufacturers.
 Such sites are poorly suited for the long-term storage of people's activity data for several reasons.
@@
 Our approach has been to leverage such APIs to retrieve a patient's information from all their devices, consolidating it into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}.
-A uniform representation enables such data to be interpretabled and used directly by future clinical medical systems, such as primary care information management systems.
+The goal of such a formalisation is to provide a standard descriptive mechanism by which this data could be interpreted, such as by future clinical medical systems.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 However, most of these sites have APIs providing programmatic access to raw sensor data.
-Our approach has been to leverage such APIs to retrieve a patient's information from all their devices, consolidating it into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT~\cite{stearns2001snomed} and ICD-10~\cite{world1993icd}.
-A uniform representation enables such data to be interpretabled and used directly by future clinical medical systems, such as primary care information management systems.
+Our approach has been to leverage such APIs to retrieve a patient's information from all their devices, consolidating it into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}.
+The goal of such a formalisation is to provide a standard descriptive mechanism by which this data could be interpreted, such as by future clinical medical systems.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
edits

@@
 conditions, monitoring a person's overall daily activity and sleep
-duration could easily indicate how much rest, exercise, exposure to
-others, and the outside world they might be getting.  Daily
-records of the time(s) that a person gets up, eats, works, exercises
-and relaxes can be analysed by clinicians for irregular behaviours or
-habits that might be affecting a person's well-being. For more
-specific conditions, daily measurements of a person's blood pressure,
-blood oxidation, heart rate, breathing, weight, and galvanic skin
-response could be useful for tracking progress on a prescribed
-intervention, recovery from an illness, or onset of a new conditions,
-for example, or be used to pinpoint sources of prolonged stress.
+duration could easily indicate how much rest, exercise, and social
+exposure they might be getting.  Moreover, longitudinal records of the
+time(s) that a person gets up, eats, works, exercises and relaxes can
+be analysed for irregular behaviours or habits that might be affecting
+their well-being. For more specific conditions, daily
+measurements of a person's blood pressure, blood oxidation, heart
+rate, breathing, weight, and galvanic skin response could be useful
+for tracking progress on a prescribed intervention, recovery from an
+illness, or onset of a new conditions, for example, or be used to
+pinpoint sources of prolonged stress.
 
@@
 
-Since SNOWMED-CT nor ICD cover activity-related concepts, we devised a supplemental ontology called \emph{Activity-SENSE}.
+Since SNOWMED-CT nor ICD-10 cover activity-related concepts, we devised a supplemental ontology called \emph{Activity-SENSE}.
 This ontology encompasses patient activities, and standard units of measure with which activities measurements were taken, and concepts for describing sensors used.

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Spelling

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 WakeMate & sleep duration & \$50  \\
-Zensorium Tinnke & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
+Zensorium Tink\'{e} & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
 BodyMedia CORE 2 & galvanic skin response, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
@@
 begun to get independent certification. For example, BodyMedia
-advertises its device as ``FDA-approved, clincally validated, most
+advertises its device as ``FDA-approved, clinically validated, most
 accurate wireless activity measurement''.  A separate study of
@@
 \section{Storing and Representing Activity Diaries}
-Currently, these devices are designed to upload measurements only to separate, dedicated web sites run by their manufacturers.
+Currently, these devices are designed to upload measurements only to separate, dedicated web sites set up by their manufacturer.
 Such sites are poorly suited for the long-term storage of people's activity data for several reasons.
-First, none of the sites providelong-term retention or access guarantees of user data (nor HIPAA security guarantees!).
+First, none of the sites provide long-term retention or access guarantees of user data (nor HIPAA security guarantees!).
 Second, it is difficult for a user to gain a unified understanding of their activity through the disparate interfaces provided, in part due of the widely different visual and interface representations.
@@
 However, most of these sites have APIs providing programmatic access to raw sensor data.
-Our approach has been to leverage such APIs to retrieve a patient's information from all their devices, consolidating it into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT \cite{stearns2001snomed} and ICD-10 \cite{world1993icd}.
-The goal of such a formalisation is to provide a standard descriptive mechanism by which this data could be interpreted, such as by future clinical medical systems.
+Our approach has been to leverage such APIs to retrieve a patient's information from all their devices, consolidating it into a uniform, common taxonomy and description logic, like those provided by SNOMED-CT~\cite{stearns2001snomed} and ICD-10~\cite{world1993icd}.
+A uniform representation enables such data to be interpretable and used directly by future clinical medical systems, such as primary care information management systems.
 
-Since SNOWMED-CT nor ICD cover activity-related concepts, we devised a supplemental ontology called \emph{Activity-SENSE}.
+Since SNOMED-CT nor ICD cover activity-related concepts, we devised a supplemental ontology called \emph{Activity-SENSE}.
 This ontology encompasses patient activities, and standard units of measure with which activities measurements were taken, and concepts for describing sensors used.

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
Spelling

@@
 Such sites are poorly suited for the long-term storage of people's activity data for several reasons.
-First, none of the sites providelong-term retention or access guarantees of user data (nor HIPAA security guarantees!).
+First, none of the sites provide long-term retention or access guarantees of user data (nor HIPAA security guarantees!).
 Second, it is difficult for a user to gain a unified understanding of their activity through the disparate interfaces provided, in part due of the widely different visual and interface representations.
@@
 
-Since SNOWMED-CT nor ICD-10 cover activity-related concepts, we devised a supplemental ontology called \emph{Activity-SENSE}.
+Since SNOMED-CT nor ICD-10 cover activity-related concepts, we devised a supplemental ontology called \emph{Activity-SENSE}.
 This ontology encompasses patient activities, and standard units of measure with which activities measurements were taken, and concepts for describing sensors used.

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
boo

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 be analysed for irregular behaviours or habits that might be affecting
-their well-being. For more specific conditions, daily
+his or her well-being. For more specific conditions, daily
 measurements of a person's blood pressure, blood oxidation, heart
@@
 
-Merely providing access to such activity records back to individuals
-could have direct effects on people's behaviour, motivation and
-perceived well-being.  For example, just as asking individuals to
+Simply letting individuals see their own activity diaries could have
+direct effects on people's behaviour, motivation and perceived
+well-being, as well.  For example, just as asking individuals to
 introspect and justify their opinions and feelings has been found to
-be both disruptive and reinforcing of their opinions~\cite{wilson1989introspection},
-allowing people to reflect upon their
-historical records of their activities might help people realise or
+be both disruptive and reinforcing of their
+opinions~\cite{wilson1989introspection}, allowing people to reflect
+upon their historical activity records might help them realise or
 change their minds about how or why they feel particular ways. We feel
 that there is significant room for investigation here, and potential
-for research pertaining to self motivation and improvement.
+for research contributions pertaining to motivation and self-improvement.
 
@@
 WakeMate & sleep duration & \$50  \\
-Zensorium Tink\'{e} & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
+Zensorium Tinnke & heart rate variability, blood oxygenation, respiratory rate & \$200 \\
 BodyMedia CORE 2 & galvanic skin response, body temperature, sleep, respiratory rate, heart rate & (TBD) \\
@@
 begun to get independent certification. For example, BodyMedia
-advertises its device as ``FDA-approved, clinically validated, most
+advertises its device as ``FDA-approved, clincally validated, most
 accurate wireless activity measurement''.  A separate study of

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
boo

@@
   diaries}, portraits of a person's dynamic condition captured at high
-temporal fidelity.  Such activity diaries are intended to supplement,
-rather than supplant, the existing diagnostic reports performed by
-professionals, first and foremost, to help clinicians derive a more
-complete picture of the person throughout a period of care, such as an
-intervention.  We believe that such activity diaries could potentially
-also be beneficial to individuals as well, making people more aware of
-the evolution of their health.
+temporal fidelity.  Such activity diaries are intended, first and
+foremost, to supplement, rather than supplant the existing diagnostic
+reports performed by professionals, in order provide clinicians with a
+more complete picture of a patient's life throughout a period of care.
+We believe that such activity diaries could potentially also be
+beneficial to individuals as well, making people more aware of the
+evolution of their health.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
title

@@
 
-\title{Using Distributed Information Stores for Patient-Clinician Health Data Sharing}
+\title{Using Personal Activity Diaries to Enhance Electronic Health Records}
 
@@
 % Paper metadata (use plain text, for PDF inclusion and later re-using, if desired)
-\def\plaintitle{Using Distributed Information Stores for Patient-Clinician Health Data Sharing}
+\def\plaintitle{Using Personal Activity Diaries to Enhance Electronic Health Records}
 \def\plainauthor{Daniel A. Smith, Max Van Kleek, Nigel R. Shadbolt}

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
abstract

@@
 Existing approaches to electronic integrated health records have largely focused on centralised information architectures, where patient data is held in a single repository, and accessed over the internet.
-However, there are a number of benefits to both patients and clinicians in a decentralised information approach to health records.
+Many patients are wearing personal sensors that report high-frequency and high-resolution data about their lifestyle, which could benefit patients and clinicians when taken in clinical context.
+We present an architecture that enables electronic health records to be augmented with activity diaries from personal sensors, and shared between patients and clinicians, to enhance communication and understanding about a patient's condition and how lifestyle changes can improve their situation.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
abstract shoterend
boo

--- 2013-01-16 patient-clinician-workshop/pccworkshop.tex
@@
 
-Existing approaches to electronic integrated health records have largely focused on centralised information architectures, where patient data is held in a single repository, and accessed over the internet.
-However, there are a number of benefits to both patients and clinicians in a decentralised information approach to health records.
+% Existing approaches to electronic integrated health records have largely focused on centralised information architectures, where patient data is held in a single repository, and accessed over the internet.
+
+%Many patients are wearing personal sensors that report high-frequency and high-resolution data about their lifestyle, which could benefit patients and clinicians when taken in clinical context.
+
+We posit that significant gains in patient-clinician communications could be made by helping clinicians better understand the patient's life between visits to the doctor's.
+In this paper, we present an architecture for enabling electronic health records to be augmented with \emph{personal activity diaries}, logs of a patient's day-to-day life that are derived automatically from consolidated readings from personal activity sensors.
+These activity diaries establish a shared context by which a patient and his or her clinician can discuss his or her lifestyle and condition.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
new

@@
 
-We posit that significant gains in patient-clinician communications could be made by helping clinicians better understand the patient's life between visits to the doctor's.
-In this paper, we present an architecture for enabling electronic health records to be augmented with \emph{personal activity diaries}, logs of a patient's day-to-day life that are derived automatically from consolidated readings from personal activity sensors.
-These activity diaries establish a shared context by which a patient and his or her clinician can discuss his or her lifestyle and condition.
+We posit that significant gains in patient-clinician communications can be made by helping clinicians to better understand a patient's life between their visits to the doctor.
+In this paper, we present an architecture for enabling electronic health records to be augmented with \emph{personal activity diaries}, high resolution records of a patient's daily life, derived automatically from personal activity sensors.
+These activity diaries establish a shared context between the patient and clinician for discussing the patient's lifestyle and condition.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
boo

@@
 % Existing approaches to electronic integrated health records have largely focused on centralised information architectures, where patient data is held in a single repository, and accessed over the internet.
-
 %Many patients are wearing personal sensors that report high-frequency and high-resolution data about their lifestyle, which could benefit patients and clinicians when taken in clinical context.
@@
 We posit that significant gains in patient-clinician communications can be made by helping clinicians to better understand a patient's life between their visits to the doctor.
-In this paper, we present an architecture for enabling electronic health records to be augmented with \emph{personal activity diaries}, high resolution records of a patient's daily life, derived automatically from personal activity sensors.
-These activity diaries establish a shared context between the patient and clinician for discussing the patient's lifestyle and condition.
+In this paper, we propose a mechanism for augmenting electronic health records with \emph{personal activity diaries}, high resolution records of a patient's daily life, derived automatically from personal activity sensors.
+These activity diaries establish a shared context for discussing the patient's lifestyle and condition.
 

### 2013-01-16 patient-clinician-workshop/pccworkshop.tex
highres

@@
 We posit that significant gains in patient-clinician communications can be made by helping clinicians to better understand a patient's life between their visits to the doctor.
-In this paper, we propose a mechanism for augmenting electronic health records with \emph{personal activity diaries}, high resolution records of a patient's daily life, derived automatically from personal activity sensors.
+In this paper, we propose a mechanism for augmenting electronic health records with \emph{personal activity diaries}, high-resolution records of a patient's daily life, derived automatically from personal activity sensors.
 These activity diaries establish a shared context for discussing the patient's lifestyle and condition.
